### 2019-03-22 paper-hierarchical.tex
files copied from gitlab.esrf.fr

%------------------------------------------------------------------------------
% Template file for the submission of papers to IUCr journals in LaTeX2e
% using the iucr document class
% Copyright 1999-2013 International Union of Crystallography
% Version 1.6 (28 March 2013)
%------------------------------------------------------------------------------

\documentclass{iucr}              % DO NOT DELETE THIS LINE
% \documentclass[preprint]{iucr}              % DO NOT DELETE THIS LINE

\usepackage{epsfig,color}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}

% \documentclass[a4paper,10pt]{article}
% \usepackage[utf8]{inputenc}
% \usepackage[]{graphicx}
% \usepackage{rotating}
% \usepackage{hyperref}
% \usepackage{amsmath} 
% \usepackage{color}
% \usepackage{listings}
% \usepackage{lscape}
% \usepackage{float}
% \usepackage{verbatim}

% \usepackage{subcaption}
% \usepackage{multirow}

\def\ave#1{{\langle #1\rangle}}
\def\half{{1\over 2}}

\newcommand{\refeq}[1]{Eq.~\ref{#1}}
\newcommand{\refpic}[1]{Fig.~\ref{#1}}
\newcommand{\todo}[1]{{\color{red}[TODO: "#1'']}}
\newcommand{\reftable}[1]{Table.~\ref{#1}}
\newcommand{\inblue}[1]{{\color{blue}#1}}
\newcommand{\inred}[1]{{\color{red}#1}}
\newcommand{\ingreen}[1]{{\color{green}#1}}

\DeclareMathOperator{\sinc}{sinc}
% \lstset{language=Python,
%         keywordstyle=\color{blue}\textbf,
% 	commentstyle=\color[rgb]{0.133,0.545,0.133},
% 	stringstyle=\color[rgb]{0.627,0.126,0.941},
% 	breaklines=true,
% 	showstringspaces=false,
% 	frame=trBL, basicstyle=\scriptsize %\small %\tiny %\footnotesize
%        }

     %-------------------------------------------------------------------------
     % Information about journal to which submitted
     %-------------------------------------------------------------------------
     \journalcode{S}              % Indicate the journal to which submitted
                                  %   A - Acta Crystallographica Section A
                                  %   B - Acta Crystallographica Section B
                                  %   C - Acta Crystallographica Section C
                                  %   D - Acta Crystallographica Section D
                                  %   E - Acta Crystallographica Section E
                                  %   F - Acta Crystallographica Section F
                                  %   J - Journal of Applied Crystallography
                                  %   M - IUCrJ
                                  %   S - Journal of Synchrotron Radiation

\begin{document}                  % DO NOT DELETE THIS LINE

     %-------------------------------------------------------------------------
     % The introductory (header) part of the paper
     %-------------------------------------------------------------------------

     % The title of the paper. Use \shorttitle to indicate an abbreviated title
     % for use in running heads (you will need to uncomment it).

\title{A hierarchical approach for modelling X-ray beamlines. Application to a coherent beamline.}

     % Authors' names and addresses. Use \cauthor for the main (contact) author.
     % Use \author for all other authors. Use \aff for authors' affiliations.
     % Use lower-case letters in square brackets to link authors to their
     % affiliations; if there is only one affiliation address, remove the [a].

\cauthor[a]{Manuel}{Sanchez del Rio}{srio@esrf.eu}{}
\author[a]{Rafael}{Celestre} %{address if different from \aff}
\author[a]{Mark}{Glass} %{address if different from \aff}
\author[a]{Giovanni}{Pirro} %{address if different from \aff}
\author[a]{Juan}{Reyes Herrera} %{address if different from \aff}
\author[a]{Ray}{Barrett} %{address if different from \aff}
\author[a]{Julio Cesar}{da Silva} %{address if different from \aff}
\author[a]{Peter}{Cloetens} %{address if different from \aff}
\author[b]{Xianbo}{Shi}
\author[b]{Luca}{Rebuffi}

\aff[a]{ESRF, 71 Avenue des Martyrs, 38043 Grenoble \country{France}}
\aff[b]{Argonne National Laboratory, 9700 South Cass Avenue, Lemont, Illinois 60439, \country{USA}}

     % Use \shortauthor to indicate an abbreviated author list for use in
     % running heads (you will need to uncomment it).

%\shortauthor{Soape, Author and Doe}

     % Use \vita if required to give biographical details (for authors of
     % invited review papers only). Uncomment it.

%\vita{Author's biography}

     % Keywords (required for Journal of Synchrotron Radiation only)
     % Use the \keyword macro for each word or phrase, e.g. 
     \keyword{Beamline}\keyword{simulations}

%\keyword{keyword}

     % PDB and NDB reference codes for structures referenced in the article and
     % deposited with the Protein Data Bank and Nucleic Acids Database (Acta
     % Crystallographica Section D). Repeat for each separate structure e.g
     % \PDBref[dethiobiotin synthetase]{1byi} \NDBref[d(G$_4$CGC$_4$)]{ad0002}

%\PDBref[optional name]{refcode}
%\NDBref[optional name]{refcode}

\maketitle                        % DO NOT DELETE THIS LINE

\begin{synopsis}
A hierarchical scheme of the calculations of an X-ray beamline for coherent applications is presented. Starting from simple calculations to partial coherence calculations, we show how ray tracing and wave optics software are applied simulating the beamline performances.   
\end{synopsis}

\begin{abstract}
This paper shows different approaches to simulate a modern X-ray beamline. It considers different methodologies with 
increasing complexity as it goes from simple hand-calculations to the most sophisticated available computer simulations based on both ray tracing and waveoptics techniques. A last generation nanofocusing beamline for coherent x-ray applications has been chosen (ID16A at the ESRF) to study in detail the issues 
related to highly demagnifying sychroton sources and exploiting the beam coherence. The performance of the beamline is studied for two storages rings: the old ESRF (emittance 4000 pm) and the new ESRF-EBS (emittance 150 pm). This document gives results in term of flux, intensity distribution, focal size, and partial coherence properties using different software tools. The different algorithms and methodologies are discussed with emphasis in benefits and limitations of each one. 
\end{abstract}

\section{Introduction}

\todo{To be written... 1) Introduction to nanofocus and coherent applications 2) ID16a 3) methods and contents 4) OASYS}

 %-------------------------------------------------------------------------
 % Manuel & Rafael
 %-------------------------------------------------------------------------

\section{Beamline description}

This section presents the reader with the main operational parameters for the ID16A beamline at ESRF. This beamline was built during the Upgrade Phase I at ESRF in \inred{2011??}. It uses one or several undulators as source. During the period 2011-2018 it worked exploiting the ESRF storage ring’s double-bend achromat (DBA) lattice with horizontal emittance about 4000 pm. The replacement of this lattice by a new hybrid multi-bend achromat (HMBA) lattice is being achieved in the ESRF Upgrade Phase II. This new lattice will drastically improve the performances of the ESRF source. The result, the Extreme Brilliant Source (EBS) \cite{ESRF2014} will increase the brightness and coherent fraction by a factor of more than 100, by strongly reducing the horizontal emittance. This is possible using almost the same ring circunference and electron energy $E_e$. As as result, the x-ray sources will be much coherent, specially in the horizontal direction. How this important improvement will afeect a particular beamline is analyzed in this paper.

ID16A \cite{ID16A} is a bemline that provides a nanometric, highly coherent, pink beam. It is dedicated to nanofocusing applications. \todo{please summarize some use and application}. The beamline usually operates at three fixed photon energies: 11200 eV, 17225 eV and 33600 eV. The straight section in the magnet lattice holds three undulators: an in-vacuum undulator of $\lambda_u=26$~mm period, and two other conventional undulators of 18 and 22.4 mm period. The undulator parameters and $K$-values at the resonances for the energies in use are shown in Table \ref{tab:Undulators}.

\begin{table}\label{tab:Undulators}
\centering
\caption{Parameters of the undulators available at the ID16A beamline: period ($\lambda_u$), length ($L$), number of identical undulators ($N$) and deflection parameter ($K$) values for the three main photon energies in use. The valuse are shown for the two ESRF magnetic lattices: the old one (labelled ESRF, with $E_e = 6.04$ GeV) and the new one (EBS, with $E_e = 6.00$ GeV). $n$ is the emission harmonic in use for each particular photon energy.}

\resizebox{\textwidth}{!}{%
\begin{tabular}{lcccccc}
name  & $\lambda_u$ [mm] & $L$ [m] & N & $K$ (11200 eV) & $K$ (17225 eV) & $K$ (33600 eV) \\ 
\input{tableUndulator.txt}
%      & IVU26 & 26.00           & 2.50    & 1      & 0.616 (n=1)    & 1.625 (n=3)    & 0.616 (n=3)    \\
% ESRF & U18.3 & 18.30           & 1.40    & 2      & 1.715 (n=1)    & 0.445 (n=1)    & 1.715 (n=3)    \\
%      & U22.4 & 22.40           & 1.40    & 1      & 0.873 (n=1)    & 1.840 (n=1)    & 0.873 (n=1)    \\ \hline
%      & IVU26 & 26.00           & 2.50    & 1      & 0.590 (n=1)    & 1.606 (n=1)    & 0.590 (n=3)    \\
% EBS  & U18.3 & 18.30           & 1.40    & 2      & 1.156 (n=1)    & 0.411 (n=1)    & 1.156 (n=1)    \\
%      & U22.4 & 22.40           & 1.40    & 1      & 0.852 (n=1)    & 1.821 (n=3)    & 0.852 (n=1)    \\ \hline
\end{tabular}%
}
\end{table}

ID16A is a long beamline (the distance from the center of the straight section to the sample position is 185 m). A strong focusing is performed with a Kirkpatrick-Baez (KB) set of mirrors placed very close (a few cm) upstream the sample position (focal plane), which ensures a high demagnification. In order to match the acceptance of the KB in the vertical and horizontal planes a focusing multilayer monochromator (ML) has
been chosen. It i) monochromatises the beam with a large energy bandwidth (typically $\Delta E/E \approx 10^{-2}$) to allow high (pink) flux, 
and ii) creates a secondary secondary source in horizontal plane 40 m downstream from the undulator, where a horizontal slit (VSS, virtual source slit) is placed. A beamline schematic is shown in Figure~\ref{fig:ID16A}. The positions of the different elements are shown in Table \ref{tab:Positions}. This table also shows the extremely high geometrical demagnification ($\sim 10^3$) needed for focusing the beam to the nanometric range.

\begin{figure}\label{fig:ID16A}
    \centering
    \includegraphics[width=0.9\textwidth,clip=true,trim=50 85 50 85]{GRAPHICS/ID16A.pdf}
    \caption{Schematic of the ID16A beamline showing the horizontal (top) and vertical (bottom) planes and the position of the main optical elements. In the sketch ML stands for the focusing multilayer monochromator; VSS for virtual source slit; and KB for the Kirkpatrick-Baez set of mirrors. The sketch is drawn out of scale.}
\end{figure}

\begin{table}\label{tab:Positions}
    \caption{Position of the main ID16A optical elements with respect to the source (undulator centre). It also shows the values for horizontal and vertical demagnification $M^{-1}=p/q$, where $p$ and $q$ are the the distances object-optics and optics-image, respectively.}
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{rcccccc}
    -                      & Source & Multilayer & Slit  & KB (v) & KB (h) & Focal plane \\ 
    Position {[}m{]}:      & 0.00   & 28.30      & 40.00 & 184.90 & 184.95 & 185.00      \\
    Horizontal $M^{-1}$:   & -      & 2.42       & -     & -      & 2899   & -           \\
    Vertical $M^{-1}$:     & -      & -          & -     & \inblue{1849}   & -      & -           \\ 
    \end{tabular}%
    }%
\end{table}

Simulations were done using the two different storage ring lattices mentioned before: the (present) ESRF DBA where the undulators are in a high-$\beta$ straight section (larger horizontal emittance) and the new EBS (lower horizontal emittance) straight sections. The parameters for the current and the new lattice are shown in Tables \ref{tab:TableSources} and \ref{tab:eBeam}.

\begin{table}\label{tab:TableSources}
    \caption{ESRF and ESRF-EBS main storage ring parameters. Main parameters of the present DBA lattice (ESRF) and futurue HMBA lattice (EBS) storage rings \cite{ESRF2014}.}
        \begin{tabular}{rcc}
                                              & \textbf{ESRF} & \textbf{EBS}      \\ 
        Lattice type:                         & DBA           & HMBA              \\
        Circunference {[}m{]}:                & 844.930       & 843.979           \\
        Beam energy {[}GeV{]}:                & 6.04          & 6.00              \\
        Beam current {[}mA{]}:                & 200           & 200               \\
        Natural emittance {[}pm$\cdot$rad{]}: & 4000          & 147               \\
        Energy spread {[}$\%${]}:             & 0.0011        & 0.00093           \\       
    \end{tabular}
\end{table}

\begin{table}\label{tab:eBeam}
    \centering
    \caption{ESRF and EBS main electron beam parameters. Electron beam parameters for the high-$\beta$ (ESRF) and the EBS straight sections. Values taken at the symmetry point of the straight section, where the insertion devices are placed. Values fir ESRF are taken from \cite{ESRF2014}. For EBS, two (very close) design values are used, from Ref.~\cite{ESRF2014} labelled EBS-S28A, and those coming from further lattice refinement, labelled EBS-S28D. \todo{Remove ESRF-AT?? Check again} }
        \resizebox{\textwidth}{!}{%
        \begin{tabular}{ccccc}
         Lattice & $\sigma_e$ horizontal {[}$\mu$m{]} &  $\sigma_e'$ horizontal {[}$\mu$rad{]} & $\sigma_e$ vertical {[}$\mu$m{]} & $\sigma_e'$ vertical {[}$\mu$rad{]} \\
         \input{tableElectron.txt}

%    \begin{tabular}{rcc}
%                                                & \textbf{ESRF (high-$\beta$)} & \textbf{EBS}     \\ 
%         $\sigma_e$ horizontal {[}$\mu$m{]}:    & 415.00                      & 30.20             \\
%         $\sigma_e'$ horizontal {[}$\mu$rad{]}: & 10.30                       & 4.40              \\
%         $\sigma_e$ vertical {[}$\mu$m{]}:      & 3.40                        & 3.60              \\
%         $\sigma_e'$ vertical {[}$\mu$rad{]}:   & 1.20                        & 1.40              \\ 
    \end{tabular}
    }
\end{table}

\section{Simulations}

For all simulations presented hereafter, we selected the configuration using a single undulator U18.3 placed in the center of the straight section and tuned to its first harmonic at a photon energy of $E=17225$ eV (Table \ref{tab:Undulators}). This section is devided in three lain subsections, that present three main levels of simulations: simple calculations that can be made by hand (section \ref{level0}), ray-tracing simulations (section \ref{level1}) and wave optics simulations (section \ref{level2}).

\subsection{Calculations ``by hand''}
\label{level0}

We first estimate the beam size at different positions at the beamline, in particular at the source plane and at the image plane. A first na{\"{i}}ve estimation of the focal size can be done by considering the source size times the magnification factor $M$. In the paraxial approximation, the magnification factor can be expressed as $M=q/p$, where $p$ is the distance between object and imaging element and $q$ is the distance between imaging element and image. The demagnification is defined, then, as the inverse of the magnification ratio: $M^{-1}=p/q$. Systems where $M>1$ are said to be magnifying, whereas systems with $M<1$ (or $M^{-1}>1$) are said to be demagnifyng. The latter will be our case, thus is mor comfortable to work with demagnification (>1) rather than magnification.

The source size is usually approximated as the convolution of the electron sizes (Table \ref{tab:eBeam}) with undulator natural source size (zero emittance) \cite{elleaume}. Let $\sigma_{e(h,v)}$ and $\sigma_{e(h,v)}'$ be the electron beam size and divergence at a centre of the symmetry of the straight section. The subscripts $h$ and $v$ stand for horizontal and vertical, respectively. The undulator natural source size and divergence are given by:
\begin{align}
    \label{eq:photon small sigmas}
    \sigma_u\approx\frac{2.74}{4\pi}\sqrt{\lambda L} && \sigma_u'\approx0.69\sqrt{\frac{\lambda}{L}},
\end{align}
where $\lambda$ is the emitted wavelength and $L=N\lambda_u$ is the undulator length, with $\lambda_u$ as the magnetic period of the undulator and $N$, the number of magnetic periods. The electron beam usually presents different horizontal and vertical beam emittances, but the undulator natural source size or divergence presents radial symmetry at the central cone (excluding polarization). The photon source size and divergence are given by:
\begin{align}
\label{eq:photon big sigmas}
\Sigma_{h,v}=\sqrt{\sigma_{e(h,v)}^2 + \sigma_u^2} && \Sigma_{h,v}'=\sqrt{\sigma_{e(h,v)}'^2 + \sigma_u'^2}.
\end{align}

% $\Sigma$ is given by the convolution of the electron beam size and the undulator natural source size or equivalently, single electron source size \cite{elleaume}: a storage ring with low emittance will produce small photon source size. It is also convenient to define two other relationships: the first one relates the magnetic field $B$ in an undulator to a dimensionless parameter $K$, the deflection parameter: $K = 93.4\lambda_u[\text{m}]\text{B}[\text{T}]$. The second relationship shows the wavelengths at which the emitted spectral intensity os peaked, or resonance. It is a function of $K$; the harmonic number n; the Lorentz factor $\gamma$ and the observation angle $\phi$:
% \begin{align}
%     \lambda=\frac{\lambda_u}{2\text{n}\gamma^2}\Big(1+\frac{K^2}{2}+\gamma^2\phi^2\Big).
% \end{align}
% The first term is the relativistic contracted wavelength, the second term is the magnetic tuning and the last term is the off-axis wavelength increase \cite{elleaume}.

The focal sizes were estimated, in a first approximation, by multiplying the photon source size by the magnification factors. For predicting the photon beam footprint downstream the beamline and to be able to estimate the needed dimensions of the optical elements to accept the whole beam, it is important to be able to calculate the beam divergence. In an ideal {\'{e}}tendue preserving optical system, the resulting beam divergence transforms inversely as the beam size: it can be calculated as the product between the beam divergence at the source with the demagnification factor $M^{-1}$.

Using the values for U18.3, $n$=1, and  $E=17225$ eV (Table \ref{tab:Undulators}), one obtains the photon source sizes and divergences shown in Table~\ref{tab:HandCalculations}. The focal size would be 139$\times$5 nm$^2$ (H$\times$V) for ESRF and 10$\times$5 nm$^2$ for EBS. The focal spot for ESRF is too large in horizontal for the demandes beamline specification, therefore the VSS could be closed to 50$\mu$m thus reducing the beamsize at the VSS plane by a factor ca. 8 \inred{$\approx$ 403.8/50=8.076}. Therefore the horizontal size at the sample will be affected by the same factor obtaining roughly 17$\times$5 nm$^2$

In general, for grazing incident optics with large demagnifications ($M^{-1} \gg 100$) imaging extended sources, the effect of optical aberrations is very important even using a perfect optical surface that focus point-to-point. As a consequence, the results calculated using only the demagnification factor are usually too optimistic. Moreover imperfections in the optical surface (figure errors, slope errors and microroughness) are often the limiting factors of the real reflective optics. Their effect will studied in Sections~\ref{level1} and \ref{level2}. The smearing of the focal size because of diffractive effects by the mirror clipping will also be discussed.

\begin{table}\label{tab:HandCalculations}
    \centering
    \caption{Photon beam size and divergence for selected ID16A beamline positions. Values below were obtained considering undulator U18.3 tuned to its first harmonic at the photon energy of $E=17225$ eV for both the high-$\beta$ (ESRF) and the EBS straight sections. Values are FWHM. Values in parentheses corresponds to closing the VSS slit to 50$\mu$m (21.3 if approximating by the Gaussian $\sigma$).}
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{r|ccc|ccc}
                            &               & \textbf{ESRF (high-$\beta$)} &   &         & \textbf{EBS}  &    \\ 
    Element:                & Undulator     & VSS          & Sample            & Undulator & VVS   & Sample       \\
    Position [m]:           & 0.00          & 40.00        & 185.00            & 0.00    & 40.00   & 185.00       \\
    $FWHM_h$ [$\mu$m]:      & 977.19        & 404.00 (50.0)& 0.139 (0.017)     & 71.26   & 29.46   & 0.0102       \\
    $FWHM_v$ [$\mu$m]:      & 9.60          & 482          & 0.0052            & 10.02   & 486.5   & 0.0054       \\
    \newline
    $FWHM'_h$ [$\mu$rad]:   & 26.96         & \inred{65.22?}        & 189062   & 15.60   & 37.74   & 109404       \\
    $FWHM'_v$ [$\mu$rad]:   & 12.04         & 12.04                 & 22264    & 12.16   & 12.16   & 22489        \\                                                   
    \end{tabular}%
    }
\end{table}

The next step would be to estimate the number of photons at the sample. For that, we have to know the spectrum at the source. It is not straightforward to get this information ``by hand'' but one can calculate the flux and power using different codes. Perhaps the most performant and well maintained codes spread in the synchrotron community are SPECTRA \cite{codeSPECTRA} and SRW \cite{codeSRW}. In this paper, we use the OASYS suite \cite{codeOASYS} that implements in its XOPPY add-on three codes: US\cite{codeUS}, URGENT: \cite{codeUS} and SRW \cite{codeSRW}. Figure~\ref{fig:FluxU18} shows the spectra for two slit apertures calculated with SRW. The spectrum is calculated over an aperture opened to fully accept the central cone at the resonance peak of the harmonic in use. placed at d=27.1. An aperture size of $S_h=S_v=4$mm  because it is much larger than the Full Width at Half Maximum (FWHM) of the central cone at $d$, that is $\max(\Sigma') d$, with any possible value of $\Sigma'$ in Table~\ref{tab:HandCalculations}. The peak intensity is about  $3.5\cdot10^{14}$ photons/s/0.1{\%}bw (see \ref{fig:FluxU18}). The corresponding spectral peak is very wide and depends mostly on the aperture so it does not change when reducing emittance (from high-$\beta$ (ESRF) to EBS). For smaller apertures (1~mm) the emittance plays an important role in both peak width and peak amplitude. The number of photons at the source at photon energy of 17225 eV \inred{(3.5 $10^{14}$ /17.225 photons/s/eV)} has to be multiplied by the bandpass of the monochromator at that energy \inred{(17225 $10^{-2}$)}, thus resulting $N_{ML}$=3.5 $10^{15}$ photons/s.

\begin{figure}\label{fig:FluxU18}
    \centering
    \includegraphics[width=0.8\textwidth]{GRAPHICS/fluxU18.jpg}
    \caption{Flux of the U18.3 undulator integrated over a square slit placed at 27.1~m from the ID centre with aperture of 1 and 4 mm.}
\end{figure}

Now we have to consider which optical element crop the beam removing photons. This can be analyzed using their angular acceptance or numerical aperture. The multilayer monochromator accepts the whole beam. We consider a VSS closed to $S_h=50 \mu$m in horizontal and fully opened in vertical. For ESRF its transmittivity is 1/8 (as discussed before) and 1 for EBS. In a similar way, the KB mirrors will crop the beam to a size that is the projection of their length to a direction perpendicular to the optical axis. Using mirror length of 60 mm and 26 mm for VFM and HFM, respectively, and incident angle of 15 mrad mrad, the KB is behaving like an aperture of $900 \times 390 \mu m^2$. Dividing this by the distance from the mirror to the VSS in H and to the source in V we get N.A. of6.2$\times$21.6 $mu$rad. Dividing by the beam divergence we have the new transmittivity coefficients 9.5\% (H) and 0.18\% (V)for ESRF and 16.3\% and 0.18\% for EBS. The overall transmission is therefore 0.2\% for ESRF and 2.95\% for EBS, therefore 7.53 $10^{12}$ photons for ESRF and 1.03 $10^{14}$ photons for EBS. \todo{check with referente 6 $10^9$ at 33.6 keV}

One can also discuss some simple consequences about the beam coherence. The coherence fraction $CF$ \cite{arxivCF} of the undulator beam is the occupation of the first coherence mode, as it will be discussed in Section~\ref{level2}. An approximated estimation of it can be obtained using Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas}, for the horizontal and vertical planes we have have
\begin{equation}
 CF_{(h,v)} = \frac{\sigma_u \sigma'_u}{\Sigma_{(h,v)} \Sigma'_{(h,v)}},
\end{equation}
and the total coherent fraction is the product of both ($CF=CF_h \times CF_v$). An estimation of the CF calculated in such a mode gived CF=0.XX\% for the ESRF and XXX for EBS. The CF roughly indicates the ration of the ``coherent photons'' over the total number of photons. A the beamline optics of a beamline target for coherence application must therefore clean the beam to extract only the good coherent photons. A complete cleaning is impossible as there is spatial overlapping of the different coherence modes. However, apertures centered at the optical axis would make the work. If we compare the values transmittivity of the beamline due to geometrical clipping  (0.2\% for ESRF and 2.95\% for EBS) we find them very close to the $CF$ values, meaning that the resulting beam would be highly coherent.

Regarding coherence, one can idealize the beamline in two parts: the ML onochromator and VSS that will prepare a coherent plane wave, as the source is very far from the sample position, and then KB that acts as a focusing element (focal=cm) that also crops the beam like it would be done by a square aperture of $D \times D=$ 400$\times$400 $\mu m^2$ \inred{(we obtained 900 in H, that will be reduced by 0.5 as propagated from 10 cm to 5 cm upstrem from the sample, and 390 in V approximated by 400)}. This aperture has two effects: i) removing a lot of off-axis photons what imply a increase of the coherent fraction to a situation close to a coherent source, and ii) make a broadening of the focal spot because of coherent diffraction. We know from optics that the diffraction by a square aperture of a collapsing spherical wavefront has an intensity distribution of:
\begin{equation}
 I(x,y,z=f) = \frac{4 k^2 D^2}{\pi^2 f^2} \sinc^2\frac{k D x}{2 f} \sinc^2\frac{k D y}{2 f}
\end{equation}
where $k=2\pi/\lambda$ . Considering that the FWHM of $\sinc(x)$ is approximately 2.78, one obtains a FWHM of 7.75$\times$7.75 nm$^2$ nm for the focal spot. This spot size is of the order or smaller than the one obtained by geometrical considerations, therefore one can say that this beamline for the conditions in use is a diffraction limited beamline.

In summary, we show in this section how simple analytical calculations that can be done in its major part ``by hand'' help to estimate the geometrical beam size at focal position (XX$\times$XX nm$^2$) the flux (XX photons/s) and also anticipate that the beam is highly coherent producing a diffraction limited focal spot of  7.75$\times$7.75 nm$^2$. We ignored at this point the effect of real optics including surface errors, the reflectivity of the elements, and other aspects that will be treated in depth in the next sections. This is, however, the first step to be done when analyzing a beamline. A similar work was done at the design level in order to define the working parameters.

 %-------------------------------------------------------------------------
 % Manuel
 %-------------------------------------------------------------------------
\subsection{Ray tracing calculations}
\label{level1}

In this section we perform ray tracing calculations to study the effect of real optics (including aberrations and slope errors) to the beam size and beamline transmittivity. The first subsection deals with the effect of aberrations and transmittivity due to geometrical considerations. The second subsection shows how the effect of slopes erroin the KB mirrors deteriorate the focal spot.

Ray tracing is a simulation method based on tracing some light rays along the optical system from the source to the image and retrieve the beam characteristics at the image (manily size and divergences). For simple systems one can perform ray tracing by hand, selecting some principal rays that will define the location and envelope of the image. Using computers one can trace thousands of rays and calculate the position and divergence distribution of the rays using statistics. Rays are usually generated by Monte Carlo sampling the source characteristics. Ray tracing is based on geometrical optics: a ray is a solution of the Helmholtz equation and travels in vacuum along straight line. Rays are intercepted by the optical elements, which change their trajectory. For reflectors, this change of direction is given by the specular reflection laws. For refractors, the change of direction os defined by the Snell law. Because Snell law uses the refraction index which is wavelength-dependent, it is then interetisting to add additional parameters to each ray, such as the wavelength. This will permit to calculate the refraction index that has to be used by each ray. The information of each ray can also be extended to include some intensity, or better electric field amplitude. This allows to extend the pure geometrical tracing to include physical models that take into account the optical element reflectivity or transmittivity. For mirrors and lenses one should use the Fresnel's equations for reflection and transmission to affect the ray intensity. This combination of using a geometrical model plus a physical model allows to simulate every element used in a synchrotron beamline, such as mirrors and combination of them (e.g., KBs); lenses, CRLs and transfocators; gratings of any type and shape (plane, VLS), crystal systems, etc. It makes ray tracing a simple and extremely powerful technique for calculating the main characteristics of the photon beam (size, divergence, photon distribution) at every point of the beamline. Although very powerful commercial ray tracing packages are available in the market, they are not suitable for synchrotron simulations because they do not cover the particularities of X-ray devices and sources. For that purposes, several packages created in the synchrotron community and are available, like SHADOW \cite{codeSHADOW}, RAY \cite{codeRAY}, McXtrace \cite{codeMCXTRACE} or XRT \cite{codeXRT}. SHADOW is probably the one that is more used and cited. We use SHADOW and its interface ShadowOUI \cite{codeSHADOWOUI}, available as a module of the OASYS suite.

% When analyzing an optical system producing an extremely high demagnification of the source one can think in several limiting factors that could affect the focal sizes calculated by hand in Table~\ref{tab:HandCalculations}.

% \inred{REMOVE!!
% One limiting factor is that the system under consideration is not an {\it imaging system} so an intensity distribution at the source (e.g., a Gaussian distribution) is not necessary imaged into a the same distribution (Gaussian). To test that, we implemented the beamline as three ideal lenses, each of them representing an optical element: ML, KBV and KBH. Ray tracing simulation show that, indeed, the demagnified image of a Gaussian source by ideal lenses shows a vertical distribution that is not Gaussian (see Fig.~\ref{fig:ray tracing ideal elements}a ). The FWHM is not close to what is expected theoretically because of the high numerical aperture. The finite acceptance of the optical surfaces and apertures makes possible a significant reduction of the aberrations. The price to pay is of course a large reduction of the intensity. Defining the multilayer and mirror acceptances as slits right before the ideal focusing elements, we obtain the image sizes shown in Fig.~\ref{fig:ray tracing ideal elements}b. Here, the image dimensions are importantly reduced in horizontal for the ESRF-High$\beta$ because of a great cut in beam dimensions due to the effect of the 50$\mu$m VSS slit and KB acceptance. It can also be seen the complete absence of the tails also due to the effect of beam cut by the apertures.
% }

We performed ray tracing using the elliptical reflectors with finite dimensions as focusing elements. The reflector shape is chosen to be elliptical as this is the theoretical shape producing a point-to-point focusing. For a first calculation  the elliptical elliptical reflectors are considered perfect, and have no slope errors. 
The results in Fig.~\ref{fig:ray tracing}a show two facts. One is that the migration from ESRF lattice to the EBS will significantly improve the horizontal size, and the VSS is not needed. Secontly, one can observe a much larger luminosity in the EBS case. The transimittivity values are 0.165\% for ESRF and 2.28\% for EBS. These values confirm what we calculated by hand (0.2\% for ESRF and and 2.95\% for EBS). As mentioned before, these values are similar to the coherent fraction \cite{coherentfraction} (0.13 and 2.8\%, respectively), meaning that a very high coherence of the beam is expected, therefore some degradation of the beam by clipping the beam by apertures and element acceptance is expected.

% Moreover, the treatment of the slope errors cannot be done without considerent effects of coherence, as discussed in \cite{hybrid}. This will be discussed in the next sections.

\begin{figure}
\label{fig:ray tracing}
\centering
% \includegraphics[width=0.45\textwidth]{GRAPHICS/ideallensESRF.png}
% \includegraphics[width=0.45\textwidth]{GRAPHICS/ideallensEBS.png}
% 
% \includegraphics[width=0.45\textwidth]{GRAPHICS/ideallensaperturesESRF.png}
% \includegraphics[width=0.45\textwidth]{GRAPHICS/ideallensaperturesEBS.png}

\includegraphics[width=0.45\textwidth]{GRAPHICS/idealelementsESRF.png}
\includegraphics[width=0.45\textwidth]{GRAPHICS/idealelementsEBS.png}

\includegraphics[width=0.45\textwidth]{GRAPHICS/hybridnoerrorsESRF.png}
\includegraphics[width=0.45\textwidth]{GRAPHICS/hybridnoerrorsEBS.png}

\includegraphics[width=0.45\textwidth]{GRAPHICS/hybriderrorsESRF.png}
\includegraphics[width=0.45\textwidth]{GRAPHICS/hybriderrorsEBS.png}

\caption{Image produced by different types of ideal focusing for the ESRF (leftcolumn) and EBS (right column). The source with parameters in 
\ref{tab:HandCalculations}. 
The first row has been calculated using the elliptical reflectors (both for ML and KBs) but without slope errors. The focal beam dimensions are written in the plot title.
The second row shows the image calculated the hybrid option (ray tracing plus coherent diffraction) with elliptical reflecting surfaces. It shows the focal spot degradiation duw to the diffraction originated by cropping the beam by the KM mirrors. The calculation in the bottom row also include slope errors using the real mirror profiles measures at the ESRF metrology lab. It can be shown a further degradation and the growing of satellite structures in the vertical intensity distribution. 
}
\end{figure}

The HYBRID method \cite{hybrid} uses geometric ray tracing combined with wavefront propagation. The code computes diffraction effects when the beam is clipped by an aperture or by the finite size of the optics and can also simulate the effect of mirror figure errors when diffraction is present. The method can be applied to an entire beamline by simulating each optical element iteratively under the hypothesis of the validity of far field propagation in all intermediate propagations. If this is not the case, a near field version is available for investigating imaging by individual optics. The interest of the hybrid method is to assess in a fast way the influence of beam coherence because the code is considerably faster than the the full methods for dealing with the partial coherence of the synchrotron beam that are described below. Fig.~\ref{fig:ray tracing}b presents the results using the hybrid model for the beamline with ideal elliptical optical surfaces. When compared with the pure ray tracing in Fig.~\ref{fig:ray tracing}a it can be appreciated that the spot is broadened. This is more evident in the vertical direction, because the smaller focal size. The spot size is about 7-8 nm in V for both storage ring lattices, and in horizontal is about 15$\mu$m for ESRF-High$\beta$ and 11 nm for EBS. It is interesting to note the almost round spot that will be obtained with the new EBS lattice.

The surface finish of the optical elements is another limiting effect for determining the size and intensity of the beam at the image position. It is possible to infere the effect of mirror imperfections and in particular slope errors by only knowing some statistical parameters such as the slope error Root Mean Squate (RMS) and then building synthetic surface profiles to be used in the simulations. This is the typical mode to operate when designing a new beamline where the mirrors do not exist yet. However, in our case the mirrors already exist and have been extensively characterize at the ESRF metrology laboratory. It is therefore highly recommended to use metrology data from real mirrors, when possible. When not available, one can use data from existing mirrors from a database (e.g., \cite{dabam}) and extrapolate some of the effects to the hypothetical new mirrors. The metrology data of the mirrors for the beamline under study are presented in \todo{Appendix 1}. These experimental data are used in all the simulations in this paper. Simulations using the hybrid model including slopes errors are shown in Fig.~\ref{fig:ray tracing}c. It can be observed a small broadening of the peak due to slope errors, but, more importantly, a new satellite structure close to the vertical intensity peak is produced by the slope errors. \todo{may be play with slope errors and use synthetic profiles to show the importance of the low frequencies in the intensity distribution?}

 %-------------------------------------------------------------------------
 % Rafael
 %-------------------------------------------------------------------------

\subsection{Wave optics simulations}
\label{level2}

\subsubsection{Simplified wave optics simulation (coherent case in 1D with point source and ideal elements)}

The characteristics of this particular beamline, a the very long length combined with a small source (in particular for EBS), makes this beamline very adapted for studied requiring coherence. Indeed, it has been shown that the clipping of the beam, mainly due to the acceptance of the KB mirrors produces a significant broadening of the focal spot.  As discussed in Section~\ref{level0}, very simple calculation of the diffraction limit can be sone using a convergent spherical beam clipped by a slit of 400$\times$400$ \mu$m$^2$ placed at 5cm upstream the focus (at the second KB mirror position). The slit vertical size corresponds roughly to the horizontal acceptance of the seconf KB mirror, and it is approximately the same in vertical. A simple simulation of diffraction by a square slit using OASYS is shown in Fig.~\ref{wofry400um}. It reveals a spot of less than 10 nm, thus confirming the analytical predictions, and qualifying this beamline to be a diffraction limited beamline.

\begin{figure}
\label{wofry400um}
\centering
\includegraphics[width=4.25cm]{GRAPHICS/wofry400umslit.png}
\includegraphics[width=4.25cm]{GRAPHICS/wofry400umprofile.png}
\caption{Diffraction spot producedof a spherical wavefront (R=-50mm) by a 400x}
\end{figure}

The first step in the wave optics simulation is to assume a full coherent source and simulate the propagation of such wavefront through the beamline
The wavefront simulation of the beamline include propagation of the beam from element to element (drift spaces) using different possible propagators. Usually a propagator based on the Freskel-Kirchhoff integral is used, as it is valid for both near and far field. The applications of this (and other) operators imply the calculation of an integral (or sum) at each pixel of the image plane. This is costly from the computer point of view, so Fourier Transforms (Fourier Optics) are used to reduce complexity and calculation time. This is practically the only choice for simulating 2D images. The use of Fourier Transforms adds another problem: the result is very dependent on the sampling of the wavefront (i.e., pixel size, number of points, dimension of the window where the wavefront is sample). This means the usually one has to test many configurations before finding one that works reaonably well (even though one can never guarantee the complete accuracy of the sampling used). A full wavefront optics simulation is an iterative process, where the user has to refine the sampling parameters element by element to be sure the wavefront at each position is sampled correctly. This is a time consuming work for performing it with the full 2D simulation. For this reason, a wavefront calculation may be started by a simplified system, using simplified models for the source and focusing elements. Moreover, because the horizontal and vertical coupling is usually small in synchrotron beamlines, it is convinient to separate the simulation in two, one for the horizontal plane and one for the vertical one.

We perform here a simplified wavefront simulation in 1D, thus separating the simulations of the horizontal and vertical planes. The source can be simplified by using a point source (spherical wavefront) and applying a Gaussian intensity profile with $\sigma=\sigma'_u D$ with $\sigma'_u$ given by Eq.~\ref{eq:photon small sigmas} and $D$ the distance from the point source to the observation plane. Each optical element can be simplified by separating its action in two parts: the focusing behaviour implemented as an ideal lens, and the finite size because of the boundaries, implemented as a slit. The WOFRY package in OASYS permits to do this work in a quick way, and it is possible to arrive to a satisfactory result in a few iterations. The beam size at different positions of the beamline are shown in Table~\ref{tablewofry1D}. The final image profiles are in Fig.~\ref{wofry1D}.

\begin{table}[]
    \centering
    \caption{Beam sizes calculated with simplified wave optics propagation (point source and ideal lenses with aperture for modeling the element dimensions). \todo{implement a sort of multielectron system - implement the slope error} }
    \label{tablewofry1D}
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{cccccc} 
    Element:                & ML            & VSS          & KBv               & KBh           & sample        \\
    Position [m]:           & 28.3          & 40.00        & 184.90            & 184.95        & 185.00        \\
    H beam size [$\mu$m]:   & 328           & 1.11         & 4084              & 390           & 0.010         \\
    V beam size [$\mu$m]:   & 328           & 466          & 899               & 450           & 0.010         \\ 
    H scaling factor:       & 1.0           & 0.01         & 220               & 1             & 0.00002       \\
    V scaling factor:       & 1.0           & 0.64         & 5                 & 0.5           & 0.00002       \\
    \end{tabular}%
    }
\end{table}

\begin{figure}
\label{wofry1D}
\centering
\includegraphics[width=4.25cm]{GRAPHICS/wofry1Dh.png}
\includegraphics[width=4.25cm]{GRAPHICS/wofry1Dv.png}
\caption{Intensity distribution at the focal position in horizontal plane (left) and in vertical plane (right).}
\end{figure}

\subsubsection{Full wave optics simulation for coherent case}
% The reduction on the electron beam emittance, by reducing the electron beam horizontal size, leads to an increase on the horizontal lateral coherence of the radiation emitted by an undulator in the symmetry point of a straight section \cite{paper}. Incoherent simulation methods may be insufficient for a full assessment of the beamline design. To accurately account for coherence effects, one has to resort to wave optics simulations (commonly referred to as wave propagation).
% 
The first step in the wave optics simulation is to assume a full coherent source and simulate the propagation of such wavefront through the beamline. The source used on the simulations to follow is the on-axis-single-electron emission from the U18.3 undulator tuned to its first harmonic at $E=17225$ eV, which is inherently fully coherent. Once the source is defined, one has to go around on how to simulate the beamline optical elements. Increasing the complexity of the simulations one can: a) use ideal thin lenses preceded by slits to emulate finite aperture; or b) use grazing incidence focusing optics models \cite{Canestrari2014}. On the top of that, one can add the effect of imperfect optics to the simulations, i.e. slope errors.

The computer code chosen for these calculations is the \textsc{srw} \cite{codeSRW}, which over the past two decades has been extensively benchmarked throughout different synchrotron radiation facilities and has become widely accepted within the X-ray community.

Figure \ref{fig:SingleElectron} shows the horizontal integrated intensity at the focal position for the ID16A. It can been seen that there is little difference between using thin ideal lenses or the elliptical mirror model, confirming the previous ray-tracing results. The elliptical mirror presents a slight reduction in intensity and asymmetry but no significant difference is found, which is in accordance with what is reported in \cite{Canestrari2014}. Notica that the differences are minimal and have to be observed in log plots. It is also important to mention that this equivalency between ideal lenses and elliptical mirrors is only valid at the focal position vicinity: down- or upstream the focal position, differences are not negligible \cite{Canestrari2014}.

It is also possible to add slope errors to the simulations to study the degradation on the focal spot quality (Fig.~\ref{fig:SingleElectron}). For that, the available available metrology data (Appendix 1) are used. For both cases (i.e., simulating the beamline using thin optical elements or elliptical mirrors), the beam sizes at the focal position are little affected by the presence of the slope errors (see Table \ref{tab:SRW_results}). There is a increase in the background level: the radiation that should be on the main lobe (focus) goes to the side lobes as a result of adding the slope errors. This is accompanied by an obvious reduction on the peak intensity. Table \ref{tab:SRW_results} compiles the beam sizes at focal position for mentioned simulations.

\begin{figure}
    \centering
        \includegraphics[width=4.25cm]{GRAPHICS/se_idealTE_idealOE_h.pdf}
        \includegraphics[width=4.25cm]{GRAPHICS/ebs_slopeTE_slopeOE_h.pdf}
    \label{fig:SingleElectron}
    \caption{Wavefront propagation from a single electron emission through the ID16A beamline. }
    %Simulations took less than one minute in a desktop computer Intel Core i7 with 8Gb of RAM memory.}
\end{figure}

\begin{table}\label{tab:SRW_results}
\centering
\caption{Beam dimensions (FWHM) at the focal position for different models of focusing elements.}
\resizebox{\textwidth}{!}{%
\begin{tabular}{ccccc}
Focusing elements                 & Plane      &  Single electron & EBS         & ESRF       \\
ideal lenses/ elliptical mirrors  & Horizontal &  7.94/7.97       & 12.20/12.70 & 17.50/16.80 \\
the same, with errors             & Horizontal &  9.35/10.50       & 12.90/13.65 & 17.70/17.03 \\
ideal lenses/ elliptical mirrors  & Vertical   &  7.11/7.26       & 8.40/9.02   & 8.24/8.77    \\
the same, with errors             & Vertical   &  7.44/8.46       & 8.66/9.87   & 8.52/9.66    \\                                                                 
\end{tabular}
}
\end{table}

\subsubsection{Full wave optics simulation: partial coherent case by means of multielectron Monta Carlo sampling}
The wavefront propagation methodology is based in the sampling of a single wavefront (monochromatic and coherent) in a 2D space (in horizontal and vertical spatial dimensions) at a given point of the beamline. This initial point considered as source cannot be the source physical position, because i) when condidering point or spherical waves all the intensity would come in a single pixel so the wavefront is ill-defined, ii) when using synchrotron sources (i.e., undulators) theory \cite{jackson} permits the calculation of the radiation in positions far from the electron trajectory. The limitation of the method is that one wavefront is not representative for a beam that, in storage rings, is produced by the emission of many electrons circulating through the corresponding magnet structure. Diffraction limited sources would emit radiation that can be described in good approximation by a single wavefront. But even for new ultra low-emittance storage rings (improperly called diffraction-limited sources, see discussion in \cite{coherentfraction}) the contribution of the electron dimensions ($\sigma_{e(h,v)}$ and $\sigma'_{e(h,v)}$) is not negligeable as compared with the natural emission of radiation ($\sigma_u$ and $\sigma'_u$ for undulators as in Eq.~\ref{eq:photon small sigmas}). To account for this effect one should complete or extend the wavefront simulations to take into account that in fact not only a single wavefront contributes to the properties of the beam. There are three approaches currently used: 1) use of convolutions, 2) Monte Carlo multi-electron sampling, and 3) Propagation of coherent modes. The convolution method consist in calculating the emission characteristics and then convolve the resulting intensity with Gaussians describing the electron beam. This method is simple and fast, as only requires simulation of a single wavefront. However, it presents two important limitations: i) it can only be applied to the intensity calculations and not to other parameters of interest for describing coherence, such as degree of coherence, etc., and ii) it can be applied only at the source position: it is usually unknown how the different optical elements will affect the source contribution to the phase space. The next method, the Monte Carlo multi-electron propagation, it is an statistical method based on the idea that one can exactly calculate the emission and its propagation for one electron that is deterministically described by its initial conditions at the entrance of the magnetic structure. It is therefore possible to perform many simulations for different initial conditions of the travelling electron and make the propagation along the beamline until the observation plane. Then one can statistically combine the different wavefronts (one for each electron) to obtain a sampling of the intensity and other magnitudes useful to describe partially coherent beams. This method presents two difficulties. One is convergence. One cannot know a priori how many electrons must be sampled to obtain a given level of accuracy. The second is computing resources. The emission of each electron must be calculated at the source level, propagated along every element of the beamline, and the resulting wavefront must be analyzed to extract the information to be scored and stored. The third method, the propagation of coherent modes wil be discussed in the next section.

The Monte Carlo multielectron method has been proposed by O. Chubar and it is fullt implemented in SRW \cite{codeSRW}. We have used this code for the calculations presented in this section.

\begin{figure}
    \centering
        \includegraphics[width=4.25cm]{GRAPHICS/esrf_idealTE_idealOE_h.pdf}
        \includegraphics[width=4.25cm]{GRAPHICS/esrf_slopeTE_slopeOE_h.pdf}
    \label{fig:FiniteEmittanceA}
    \caption{Wavefront propagation from the ESRF magnetic lattice through the ID16A beamline.}
\end{figure}

\begin{figure}
    \centering
        \includegraphics[width=4.25cm]{GRAPHICS/ebs_idealTE_idealOE_h.pdf}
        \includegraphics[width=4.25cm]{GRAPHICS/ebs_slopeTE_slopeOE_h.pdf}
    \label{fig:FiniteEmittanceB}
    \caption{Wavefront propagation from the EBS magnetic lattice through the ID16A beamline.}
\end{figure}

The source used for simulations is the U18.3 undulator tuned to its first harmonic at $E=17225$ eV with a finite emittance electron beam (see Table \ref{tab:eBeam}). Two set of parameters were used for the electron beam: the current machine parameters at the ESRF (high-$\beta$) and the EBS straight sections (Tables \ref{tab:TableSources} and \ref{tab:eBeam}).

Simulations for both storage ring lattices were carried out modelling the beamline using thin ideal lenses and elliptical mirror models with and without the addition of the slope errors. Results are presented in Figures \ref{fig:FiniteEmittanceA} and \ref{fig:FiniteEmittanceB}; and Table \ref{tab:SRW_results}.

The same observations made on the previous section (full coherent simulations) are pertinent here: there is a reduction on peak intensity when comparing single lens model with the elliptical mirror. The beam sizes and shapes are similar at the focal position, only. Simulations took significantly longer to be completed: few hours on a 56-core computer cluster.

% 
% Synchrotron radiation on the hard X-rays regime is hardly full coherent even for the so-called diffraction limited sources (see Table \ref{tab:Coherence}). To better describe the nature of the emitted radiation, one makes use of the partial coherent simulations using the \textsc{srw}, i.e. incoherent addition of the intensity from the emission from electrons sampled from the source's 6D phase-space \cite{Chubar2011b}. This is a natural extension from the method presented on the previous section.

% 
% \newpage
% 
%  %-------------------------------------------------------------------------
%  % Manuel (?)
%  %-------------------------------------------------------------------------
%  
\subsubsection{Full Wave optics simulation: partial coherence case by means of coherent mode decomposition}
% 
% One of the mentioned drawbacks of the Monte Carlo multielectron methods is the a-priori unability to determine the size of sampling space (i.e., how many electrons) for accurate calculations. 
% Another important limitation is the impossibility with available computer resources to store (although can be computed) the full 4D matrices fully describing the partial coherence properties (cross spectral density) at any point of the beamline.

The cross spectral density (CSD) (sometimes called mutual intensity) defines the correlation of the radiated electric fields between two spatial points $(x_1,y_1)$ and $(x_2,y_2)$ at a given position of the beamline $z$:

\begin{equation}
W(x_1,y_1,x_2,y_2;z,\omega) = <E*(x_1,y_1; z,\omega) E(x_2,y_2;z,\omega)>
\label{CSD}
\end{equation}
where $E$ is the amplitude of the electric field, $\omega$ is the radiation frequency, and $<>$ stands for the ensemble average. The CSD is at the origin of any other information aboute the coherence of the beam, like the spectral density $S$ (or, just intensity): 
\begin{equation}
S(x,y,;z,\omega) = W(x,y,x,y,;z,\omega)
\end{equation}
or the spectral degree of coherence $\mu$:
\begin{equation}
\mu(x_1,y_1,x_2,y_2;z,\omega) = \frac{W(x_1,y_1,x_2,y_2;z,\omega)}{\sqrt{ S(x_1,y_1;z,\omega) S(x_3,y_2;z,\omega}}
\end{equation}
that measures the coherence of the beam, from fully incoherent ($\mu=0$) to fully coherent ($\mu$=1). 
Eq.~\ref{CSD} can in principle be calculated using the Kim's convolution theorem \cite{kim1986} (see \cite{geloni2008} for a full discussion on the validity) but its storage is prohibitif in present computers: supposing sampling one spatial dimension by $N\approx1000$ pixels, a wavefront will contain $N^2$ pizels, and the CSD will be sampled over $N^4$ points. They are complex numbers (16 bits) so the only stockage is of the order of GB or TB. Considering also that the progagation of $W$ from a beamline position $z_1$ to another position $z_2$ will need 10$^8$ to 10$^{12}$ 4D integrals, one can easily imagine the physical impossibility to perform such calculations. It is interesting to remark that although Monte Carlo multielectron method can in principle calculate Eq.~\ref{CSD} the storage at every point is unmanagleable, so usually only some 'cuts' of the CSD are stored, such as $W(x_1,0,x_2,0;z,\omega)$ (horizontal) or $W(0,y_1,0,y_2;z,\omega)$ (vertical).

A solution of this problem is the expansion of the CSD in coherent modes \cite{mandel_wolf}:  
\begin{equation}
 W(x_1,y_1,x_2,y_2;z,\omega) = \sum\limits_{m=0}^{\infty} \lambda_m \Phi_m^{*}(x_1,y_1;z,\omega) \Phi_m(x_2,y_2;z,\omega)
\label{decomposition}
 \end{equation}
where $\lambda_m$ are the eigenvalues or weights, and $\Phi_m$ are the eigenfunctions or coherent modes.

A theoretical study and a numerical algorithm for the coherent mode decomposition of undulator emission in storage rings has been recently proposed  \cite{GlassThesis,GlassEPL} and implemented in the computer package COMSYL \cite{codeCOMSYL}. The innovative concept is the numerical description of the undulator CSD in terms of its coherent modes $\Phi_m(x,y;z,\omega)$ and eigenvalues $\lambda_m $. One can define the mode distribution as: 
\begin{equation}
 d_m = \frac{\lambda_m}{\sum\limits_{i=0}^{\infty} \lambda_i}
\end{equation}
or the cumulated mode distribution as: 
\begin{equation}
 c_m = \sum\limits_{i=0}^{m} d_i
\end{equation}
An important property of this coherent mode decomposition is that it optimizes the representation of the CSD when the series is truncated. In other words, truncating the series in Eq.~\ref{decomposition} up to the term $m$, the CSD obtained approximates better the exact CSD (this with infinite terms) than any other expansion. It also imply that the first term is more important than the second, and so on. The first term (index zero) is the first coherent mode. It relative importance with respect to the full CSD permits to define the coherent fraction (CF) as the occupation of the firts coherence mode: CF=$d_0$. Obviously, a pretty coherent source will require less modes than a mostly incoherent source to arrive up to a certain representation of the CSD.

We performed COMSYL calculations for the U18.3 undulator radiation of ID16A of 1.4 m length using both the EBS and ESRF-High$\beta$ lattices. The obtained CF are 2.8\% for EBS and 0.13\% for ESRF-High$\beta$, indicating that the overall coherent flux emitted by EBS will be about 20 times the existing coherent flux with ESRF-High$\beta$. Fig.~\ref{histomodes}a shows the cumulated occupation of the coherent modes for the full central cone undulator emission. It can be shown how much less number of coherent modes are needed to arrive to a certain degree of cumulated occupation for EBS than for ESRF-High$\beta$. This can also be shown in Fig.\ref{histomodes}b where the number of modes to fill 50\%, 75\% and 95\% CSD occupation are compared.

The coherence fraction of 2.8\% for the EBS source means that for an optimal coherence experiment one should remove 97.2\% of the emitted photons. This task is performed by the beamline. The beamline elements modify the different modesin a different manner. The fist coherence mode is always centered on-axis, and the subsequent ones expand in the space until completely filling the spectral density.   
\begin{figure}
    \centering
        \includegraphics[width=7cm]{GRAPHICS/cumulated_occupation.eps}
        \includegraphics[width=6.5cm]{GRAPHICS/up_to_mode_id16a.eps}
    \label{histomodes}
    \caption{Cumulated occupation of the coherent modes for the EBS and ESRF-High-$\beta$ (top), and number of modes needed to reach a given percent of CSD occupation (bottom). }
\end{figure}

\begin{figure}
    \centering
        \includegraphics[width=\textwidth]{GRAPHICS/ebs_spectral_density.png}
        \includegraphics[width=0.45\textwidth]{GRAPHICS/ebs_mode0.png}
        \includegraphics[width=0.45\textwidth]{GRAPHICS/ebs_mode1.png}
        \includegraphics[width=0.45\textwidth]{GRAPHICS/ebs_mode2.png}
        \includegraphics[width=0.45\textwidth]{GRAPHICS/ebs_mode3.png}
    \label{spectraldensity}
    \caption{XXXX }
\end{figure}

The coherent modes for the EBS have been propagated using a beamline simplified with ideal focusing elements and slits to determine their numerical aperture. Fig.~\ref{fig:mode transmission} shows how the intensity of the first modes is reduced by geometrical considerations. We recall that in the coherent mode expansion, the modes are orthonormal, therefore the integrated intensity for any mode is always one at the source. Different modes will be ``cut'' in a different manner, depending on the slits cropping effect. Fig.~\ref{fig:mode transmission} shows that the lowest mode (zeroth order) transfer is about 25\%, the mode 2 has a 10\% transfer and the 4th mode about 6\%. The other modes have less than 5\% transfer and modes higher than 10 practically do not contribute to the final spectral density. Note that the transmission os mode 2 is higher than mode 1. This is because the mode 3 has its highest intensity at the origin, but mode 2 has a hole in the middle, as shown in Fig. XX. Modes with maximum intensity at origin will be transmitted more than the others. This analysis of the mode transmission is mode dramatic if one looks to the total intensity in the mode, which is given by the eigenvalue. Neglecting modes higher than 10, the intensity transmitted for the source mode zero amounts for 49.5\%, 18.4\% for the mode 2 less than 5\% for the others. As commented before, the source modes cropped by the beamline elements are not longer orthonormal, therefore they are not longer coherent modes of the resulting beam. The new coherent modes can be calculated applying a new coherent mode decomposition on the resulting spectral density. Making this numerical calculation with COMSYL, one obtains a mode occupation spectrum where the lowest mode accounts for the 83\% of the total spectral density (see Fig.~\ref{fig: rediagonalization}). This analysis confirms the quatitative predictions discussed in the previous sections, but give a quantitative figure of the coherence of the beam.

% #########################################################
%            iteration            intensity         intensity/I0   intensity/TotalInt
%                    0               0.23967               1.00000              0.44091 
%                    1              0.022004               0.09181              0.04048 
%                    2               0.10008               0.41757              0.18411 
%                    3              0.028309               0.11812              0.05208 
%                    4              0.061262               0.25561              0.11270 
%                    5              0.026262               0.10958              0.04831 
%                    6               0.03452               0.14403              0.06351 
%                    7               0.01459               0.06087              0.02684 
%                    8              0.011714               0.04887              0.02155 
%                    9             0.0051714               0.02158              0.00951

% #########################################################
%            iteration            intensity         intensity/I0   intensity/TotalInt
%                    0            1.0793e+06               1.00000              0.49537 
%                    1                 93290               0.08644              0.04282 
%                    2            3.9938e+05               0.37004              0.18331 
%                    3            1.0632e+05               0.09851              0.04880 
%                    4            2.1648e+05               0.20058              0.09936 
%                    5                 87305               0.08089              0.04007 
%                    6            1.0794e+05               0.10001              0.04954 
%                    7                 42906               0.03975              0.01969 
%                    8                 32393               0.03001              0.01487 
%                    9                 13446               0.01246              0.00617

\begin{figure}\label{fig:mode transmission}
    \centering
        \includegraphics[width=\textwidth]{GRAPHICS/mode_transmission.png}

    \caption{Transmission of the different modes due to geometrical efffects.}
\end{figure}

\begin{figure}\label{fig:rediagonalization}
    \centering
        \includegraphics[width=\textwidth]{GRAPHICS/rediagonalization.png}

    \caption{Occupation of the coherent modes in the final focused beam.}
\end{figure}

\begin{figure}\label{fig:final modes}
    \centering
        \includegraphics[width=0.45\textwidth]{GRAPHICS/final_mode0.png}
        \includegraphics[width=0.45\textwidth]{GRAPHICS/final_mode1.png}
    \caption{Coherent modes of the focused beam}
\end{figure}

\todo{
\begin{itemize}
 \item propagate modes
 \item calculate coherent fraction at the focal position
 \item show some modes
 \item create some plots with cuts of W
\end{itemize}
}

\section{Discussion}

\todo{
\begin{itemize}
 \item comment on the heat-load deformation 
 \item comment the effect of reflectivity by ML and mirror coating
 \item comment on the main results in terms of intensity and size
 \item EXPERIMENTAL RESULTS?
 \item comment ``where to stop'' in a simulation, depending on the level of accuracy requested
 \item comment on the computer resources used for simulations
 \item comment on software tools: OASYS (and make a graphic)
 \item discuss the particularities of this beamline 
 \item comment on general strategy for beamline simulations and the application of the shown hierarchical model. Make a flowchart
\end{itemize}
}

 %-------------------------------------------------------------------------
 % Manuel & Rafael
 %-------------------------------------------------------------------------
\section{Summary}

Bla bla

\bibliography{paper-hierarchical}
\bibliographystyle{iucr}

\end{document}                    % DO NOT DELETE THIS LINE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-03-24 paper-hierarchical.tex
changed order in bib files
starting text

@@
 \begin{abstract}
-This paper shows different approaches to simulate a modern X-ray beamline. It considers different methodologies with 
-increasing complexity as it goes from simple hand-calculations to the most sophisticated available computer simulations based on both ray tracing and waveoptics techniques.
-A last generation nanofocusing beamline for coherent x-ray applications has been chosen (ID16A at the ESRF) to study in detail the issues
-related to highly demagnifying sychroton sources and exploiting the beam coherence.
+We consider different approaches to simulate a modern X-ray beamline.
+Several methodologies with increasing complexity are applied to discuss the relevant parameters that quantify the beamline performances.
+Parameters such as flux, beam size at the image and coherence properties are obtained from simple hand-calculations to sophisticated computer simulations using ray tracing and waveoptics techniques.
+A last generation x-ray nanofocusing beamline for coherent  applications (ID16A at the ESRF) has been chosen  to study in detail the issues
+related to highly demagnifying sychrotron sources and exploiting the beam coherence.
 The performance of the beamline is studied for two storages rings: the old ESRF (emittance 4000 pm) and the new ESRF-EBS (emittance 150 pm).
@@
 
-\todo{To be written... 1) Introduction to nanofocus and coherent applications 2) ID16a 3) methods and contents 4) OASYS}
+1) Introduction to nanofocus and coherent applications
+
+2) ID16a
+
+3) methods and contents
+
+4) OASYS
 
@@
 
+\bibliographystyle{iucr}
 \bibliography{paper-hierarchical}
-\bibliographystyle{iucr}
 

### 2019-03-24 paper-hierarchical.tex
starting

@@
 \section{Introduction}
+
+Many storage-ring based x-ray synchrotron facilities are building or planning upgrades to increase brilliance and coherent flux by one to three orders of magnitude.
+The first upgrade of a large facility will be the EBS (Extremely Brilliant Source) \cite{orangebook} at the European Synchrotron Radiation Facility (ESRF), aiming to build a storage ring of 150 pm emittance to significantly boost the associated x-ray coherence.
+Applications exploiting beam coherence such as such as x-ray photon correlation spectroscopy, coherent diffraction imaging, propagation-based phase contrast imaging and ptychography will benefit from this update.
+
+Accurate calculation and quantitative evaluation of the parameters related to x-ray coherence is of paramount importance for designing, building and exploiting the new beamlines.
 

### 2019-03-26 paper-hierarchical.tex
working...

--- 2019-03-24 paper-hierarchical.tex
@@
 The performance of the beamline is studied for two storages rings: the old ESRF (emittance 4000 pm) and the new ESRF-EBS (emittance 150 pm).
-This document gives results in term of flux, intensity distribution, focal size, and partial coherence properties using different software tools.
-The different algorithms and methodologies are discussed with emphasis in benefits and limitations of each one.
+In addition to traditional results in term of flux, intensity distribution and focal size, an innovative study on the partial coherence properties based on propagation of coherent modes is presented.
+The different algorithms and methodologies are implemented in the software suite OASYS.
+They are discussed with emphasis in benefits and limitations of each one.
 \end{abstract}
@@
 
-1) Introduction to nanofocus and coherent applications
+Many storage ring based x-ray synchrotron facilities are building or planning upgrades to increase brilliance and coherent flux by one to three orders of magnitude.
+The first upgrade of a large facility will be the EBS (Extremely Brilliant Source) \cite{orangebook} at the European Synchrotron Radiation Facility (ESRF), aiming to build a storage ring of 150 pm emittance to significantly boost the associated x-ray coherence.
+Applications exploiting beam coherence such as such as x-ray photon correlation spectroscopy, coherent diffraction imaging, propagation-based phase contrast imaging and ptychography will strongly benefit from this update.
 
-2) ID16a
+Accurate calculation and quantitative evaluation of the parameters related to x-ray optics is of paramount importance for designing, building and exploiting the new beamlines.
+Every modern beamline follows a procedure of conceptual design and then technical design where detailed simulations are essential.
+The calculations allow the designer to test the design parameters in a virtual computer environment where pros and cons can be studied quantitatively.
+The selection to the best performant optics configuration requires the detailed simulation of the optics imperfections, that in many cases are the limiting factor of the beamline optical configuration.
+Deformation in the optical elements due to heat load need to be controlled, therefore optics simulations have to include results from engineering modeling of the thermal deformation associated to mechanical issues usually dome using finite eelement analysis.
+Traditionally the optics calculations for synchrotron beamlines are done using ray tracing techniques.
+They model the x-ray beam as a collection of different rays that are incoherent one to another.
+This technique is still essential for obtaining information on aberrations, flux propagation, monochromator bandwidth, etc.
+However, the high coherence of the new sources make necessary to complement the ray tracing methods with models including effects of diffraction and scattering from a coherent beam.
+A fully coherent beam is still a dream for the new generation storage rings.
+In spite of a high increate of the coherent fraction of the new beams, the coherent fraction at tipically hard x-ray photon energies is still at few percent for the new sources.
+This makes still important a ray tracing analysis that does not become obsolete, but has to be complemented with an analysis based on wave optics.
+Because of the non perfect coherence the wave optics analysis becomes non trivial and can be done with different levels of approximation, as discussed in this work.
+In addition, wave optics simulations are usually expensive from the computer point of view.
 
-3) methods and contents
+For illustrating the applicability of a battery of methods and solutions for beamline optics simulations we have chosen the ESRF beamline ID16A.
+This beamline provides a high-brilliance beam focused to a few nanometer.
+It shows characteristics increasingly searched in modern beamlines: extremely high demagnification (sub-micrometer beam) and high coherence.
+It combines coherent imaging techniques and X-ray fluorescence microscopy to perform quantitative 3D characterization of the morphology and the elemental composition of specimens in their native state \todo{complete ID16A description}.
 
 4) OASYS
+
+The paper is structures as follows: we first introduce in Section~\ref{Bealine description} the main parameters of the beamline and make the selection of the particular choice of the parameters used in the calculations.
+Section~\label{Simulations} is the central part of the paper, and decribe the different methods used to analyse the beamline with increasing complexity, setting a hierarchy of methodologies that can be exported to the simulations of any beamline.
+In subsection \ref{hand} we calculate with simple methods approximated values of focal size, flux and coherent fraction.
+Then accuracy is improved using geometrical methods in subsection \ref{ray tracing} that give reasonable results as far as some corrections are considered because of the  mentioned characteristics of high demagnification and high coherence of the studied beamline.
+The wave optics...
 
@@
 \section{Beamline description}
+\label{Beamline description}
 

### 2019-03-26 paper-hierarchical.tex
working...

@@
 In addition to traditional results in term of flux, intensity distribution and focal size, an innovative study on the partial coherence properties based on propagation of coherent modes is presented.
-The different algorithms and methodologies are implemented in the software suite OASYS.
+The different algorithms and methodologies are implemented in the software suite OASYS \cite{codeOASYS}.
 They are discussed with emphasis in benefits and limitations of each one.
@@
 
-4) OASYS
+X-ray beamlines are very particular optical systems.
+Some implications of exploiting low wavelengths such as the necessity of using grazing incidence reflectors and the use of perfect crystals as typical hard x-ray monochromators make useless, or at least incomplete, the application of common commercial software packages for the study of synchrotron beamlines.
+For this reason many optics tools have been developed by the synchrotron community.
+Some of them have a long history like SHADOW (more than 40 years old) or SRW (almost 30), for mentioning only codes used in this work.
+These historical tools together with new tools designed ad-hoc packages have been agglutinated in a user friendly software OASYS.
+Many of the calculations presented here have been performed directly in OASYS or with script created with the help of OASYS.
 
-The paper is structures as follows: we first introduce in Section~\ref{Bealine description} the main parameters of the beamline and make the selection of the particular choice of the parameters used in the calculations.
-Section~\label{Simulations} is the central part of the paper, and decribe the different methods used to analyse the beamline with increasing complexity, setting a hierarchy of methodologies that can be exported to the simulations of any beamline.
-In subsection \ref{hand} we calculate with simple methods approximated values of focal size, flux and coherent fraction.
-Then accuracy is improved using geometrical methods in subsection \ref{ray tracing} that give reasonable results as far as some corrections are considered because of the  mentioned characteristics of high demagnification and high coherence of the studied beamline.
-The wave optics...
+The paper is structures as follows: we first introduce in Section~\ref{Beamline description} the main parameters of the beamline and make the selection of the particular choice of the parameters used in the calculations.
+Section~\label{Simulations} is the central part of the paper, and describe the different methods used to analyse the beamline with increasing complexity, setting a hierarchy of methodologies that can be exported to the simulations of any beamline.
+In subsection \ref{level0} we calculate with simple methods approximated values of focal size, flux and coherent fraction.
+Then accuracy is improved using geometrical methods in subsection \ref{level1} that give reasonable results as far as some corrections are considered because of the  mentioned characteristics of high demagnification and high coherence of the studied beamline.
+The wave optics section~]ref{level2} also include different models whose complexity increases when the results expected are more and mode demanding.
+The simulations methods are presented in a hierarchical approach:  they provide more and better results but the price to pay is to deal with more complex and computationally expensive tools.
 
@@
 
-This section presents the reader with the main operational parameters for the ID16A beamline at ESRF.
+This section presents the reader with the main operational parameters for the ID16A beamline at ESRF in two scenarios.
 This beamline was built during the Upgrade Phase I at ESRF in \inred{2011??}.
@@
 This new lattice will drastically improve the performances of the ESRF source.
-The result, the Extreme Brilliant Source (EBS) \cite{ESRF2014} will increase the brightness and coherent fraction by a factor of more than 100, by strongly reducing the horizontal emittance.
+The result, the Extreme Brilliant Source (EBS) \cite{orangebook} will increase the brightness and coherent fraction by a factor of more than 100, by strongly reducing the horizontal emittance.
 This is possible using almost the same ring circunference and electron energy $E_e$.
 As as result, the x-ray sources will be much coherent, specially in the horizontal direction.
-How this important improvement will afeect a particular beamline is analyzed in this paper.
+This work shows some quantitative results on how the storage ring upgrade will affect a particular beamline is analyzed in this paper.
 
@@
 \begin{table}\label{tab:TableSources}
-    \caption{ESRF and ESRF-EBS main storage ring parameters. Main parameters of the present DBA lattice (ESRF) and futurue HMBA lattice (EBS) storage rings \cite{ESRF2014}.}
+    \caption{ESRF and ESRF-EBS main storage ring parameters. Main parameters of the present DBA lattice (ESRF) and future HMBA lattice (EBS) storage rings \cite{orangebook}.}
         \begin{tabular}{rcc}
@@
 For all simulations presented hereafter, we selected the configuration using a single undulator U18.3 placed in the center of the straight section and tuned to its first harmonic at a photon energy of $E=17225$ eV (Table \ref{tab:Undulators}).
-This section is devided in three lain subsections, that present three main levels of simulations: simple calculations that can be made by hand (section \ref{level0}), ray-tracing simulations (section \ref{level1}) and wave optics simulations (section \ref{level2}).
+This section is devided in three subsections, that present three main levels of simulations: simple calculations that can be made by hand (section \ref{level0}), ray-tracing simulations (section \ref{level1}) and wave optics simulations (section \ref{level2}).
 
@@
 In the paraxial approximation, the magnification factor can be expressed as $M=q/p$, where $p$ is the distance between object and imaging element and $q$ is the distance between imaging element and image.
-The demagnification is defined, then, as the inverse of the magnification ratio: $M^{-1}=p/q$.
-Systems where $M>1$ are said to be magnifying, whereas systems with $M<1$ (or $M^{-1}>1$) are said to be demagnifyng.
-The latter will be our case, thus is mor comfortable to work with demagnification (>1) rather than magnification.
+The demagnification is the inverse of the magnification ratio: $M^{-1}=p/q$.
+Systems where $M>1$ are said to be magnifying, whereas systems with $M<1$ (or $M^{-1}>1$) are said to be demagnifying.
+The latter will be our case, thus is more convenient to work with demagnification ($M^{-1}>1$) rather than with magnification.
 
-The source size is usually approximated as the convolution of the electron sizes (Table \ref{tab:eBeam}) with undulator natural source size (zero emittance) \cite{elleaume}.
+The source size is usually approximated supposing a Gaussian distribution of the electron sizes (or divergence) (Table \ref{tab:eBeam}) and undulator natural source size or divergence.
+Although the former is true, the latter is not.
 Let $\sigma_{e(h,v)}$ and $\sigma_{e(h,v)}'$ be the electron beam size and divergence at a centre of the symmetry of the straight section.
 The subscripts $h$ and $v$ stand for horizontal and vertical, respectively.
-The undulator natural source size and divergence are given by:
+The undulator natural source size and divergence are given by \cite{elleaume}:
 \begin{align}
     \label{eq:photon small sigmas}
-    \sigma_u\approx\frac{2.74}{4\pi}\sqrt{\lambda L} && \sigma_u'\approx0.69\sqrt{\frac{\lambda}{L}},
+    \sigma_u=\frac{2.74}{4\pi}\sqrt{\lambda L}\approx \sqrt{\frac{\lambda L}{2 \pi^2}}  && \sigma_u' = 0.69\sqrt{\frac{\lambda}{L}}\approx \sqrt{\frac{\lambda}{2 L}},
 \end{align}
@@
 The electron beam usually presents different horizontal and vertical beam emittances, but the undulator natural source size or divergence presents radial symmetry at the central cone (excluding polarization).
-The photon source size and divergence are given by:
+The photon source size and divergence are given by the convolution of the electron sizes and the undulator natural sizes.
+If we consider Gaussian distributions, the photon source is
 \begin{align}
@@
 
-The focal sizes were estimated, in a first approximation, by multiplying the photon source size by the magnification factors.
-For predicting the photon beam footprint downstream the beamline and to be able to estimate the needed dimensions of the optical elements to accept the whole beam, it is important to be able to calculate the beam divergence.
-In an ideal {\'{e}}tendue preserving optical system, the resulting beam divergence transforms inversely as the beam size: it can be calculated as the product between the beam divergence at the source with the demagnification factor $M^{-1}$.
+The beam sizes at different positions at the beamline were estimated in a different way depending if the beam is focused or not.
+If the beam is focused, its dimensions in a first approximation are the photon source size multiplied by the corresponding magnification factor.
+For predicting the photon beam in a non-focal position the beam divergence is used.
+In an ideal {\'{e}}tendue preserving optical system, where the conservation of the Smith-Helmholz invariante applies, the resulting beam divergence transforms inversely as the beam size: it can be calculated as the product between the beam divergence at the source with the demagnification factor $M^{-1}$.
 
-Using the values for U18.3, $n$=1, and  $E=17225$ eV (Table \ref{tab:Undulators}), one obtains the photon source sizes and divergences shown in Table~\ref{tab:HandCalculations}.
-The focal size would be 139$\times$5 nm$^2$ (H$\times$V) for ESRF and 10$\times$5 nm$^2$ for EBS.
-The focal spot for ESRF is too large in horizontal for the demandes beamline specification, therefore the VSS could be closed to 50$\mu$m thus reducing the beamsize at the VSS plane by a factor ca.
+Using the values for U18.3, $n$=1, and  $E=17225$ eV (Table \ref{tab:Undulators}), one obtains a focal size 139$\times$5 nm$^2$ (H$\times$V) for ESRF and 10$\times$5 nm$^2$ for EBS (Table~\ref{tab:HandCalculations}).
+The focal spot for ESRF lattice is too large in horizontal for the beamline specification, therefore the VSS could be closed to 50$\mu$m thus reducing the beamsize at the VSS plane by a factor ca.
 8 \inred{$\approx$ 403.8/50=8.076}.
@@
 The next step would be to estimate the number of photons at the sample.
-For that, we have to know the spectrum at the source.
+For that, we need the spectrum at the source.
 It is not straightforward to get this information ``by hand'' but one can calculate the flux and power using different codes.
 Perhaps the most performant and well maintained codes spread in the synchrotron community are SPECTRA \cite{codeSPECTRA} and SRW \cite{codeSRW}.
-In this paper, we use the OASYS suite \cite{codeOASYS} that implements in its XOPPY add-on three codes: US\cite{codeUS}, URGENT: \cite{codeUS} and SRW \cite{codeSRW}.
+The XOPPY package in the OASYS suite implements three codes: US\cite{codeUS}, URGENT: \cite{codeUS} and SRW \cite{codeSRW}.
 Figure~\ref{fig:FluxU18} shows the spectra for two slit apertures calculated with SRW.
 The spectrum is calculated over an aperture opened to fully accept the central cone at the resonance peak of the harmonic in use.
-placed at d=27.1.
-An aperture size of $S_h=S_v=4$mm  because it is much larger than the Full Width at Half Maximum (FWHM) of the central cone at $d$, that is $\max(\Sigma') d$, with any possible value of $\Sigma'$ in Table~\ref{tab:HandCalculations}.
+placed at $d$=27.1m.
+An aperture size of $S_h=S_v=4$ mm it is much larger than the Full Width at Half Maximum (FWHM) of the central cone at $d$, that is $\max(\Sigma') d$ \todo{HOW MUCH?}, with any possible value of $\Sigma'$ in Table~\ref{tab:HandCalculations}.
 The peak intensity is about  $3.5\cdot10^{14}$ photons/s/0.1{\%}bw (see \ref{fig:FluxU18}).
@@
 
-Now we have to consider which optical element crop the beam removing photons.
-This can be analyzed using their angular acceptance or numerical aperture.
+\inred{Next we would like to estimate} the geometrical transmitivity of the beamline, therefore we have to consider which optical element crop the beam and remove photons.
+This can be analysed using their angular acceptance or numerical aperture.
 The multilayer monochromator accepts the whole beam.

### 2019-03-28 paper-hierarchical.tex
more work

@@
 Perhaps the most performant and well maintained codes spread in the synchrotron community are SPECTRA \cite{codeSPECTRA} and SRW \cite{codeSRW}.
-The XOPPY package in the OASYS suite implements three codes: US\cite{codeUS}, URGENT: \cite{codeUS} and SRW \cite{codeSRW}.
-Figure~\ref{fig:FluxU18} shows the spectra for two slit apertures calculated with SRW.
-The spectrum is calculated over an aperture opened to fully accept the central cone at the resonance peak of the harmonic in use.
-placed at $d$=27.1m.
-An aperture size of $S_h=S_v=4$ mm it is much larger than the Full Width at Half Maximum (FWHM) of the central cone at $d$, that is $\max(\Sigma') d$ \todo{HOW MUCH?}, with any possible value of $\Sigma'$ in Table~\ref{tab:HandCalculations}.
-The peak intensity is about  $3.5\cdot10^{14}$ photons/s/0.1{\%}bw (see \ref{fig:FluxU18}).
-The corresponding spectral peak is very wide and depends mostly on the aperture so it does not change when reducing emittance (from high-$\beta$ (ESRF) to EBS).
-For smaller apertures (1~mm) the emittance plays an important role in both peak width and peak amplitude.
-The number of photons at the source at photon energy of 17225 eV \inred{(3.5 $10^{14}$ /17.225 photons/s/eV)} has to be multiplied by the bandpass of the monochromator at that energy \inred{(17225 $10^{-2}$)}, thus resulting $N_{ML}$=3.5 $10^{15}$ photons/s.
+The XOPPY package in the OASYS suite implements three codes: US\cite{codeUS}, URGENT: \cite{codeURGENT} and SRW \cite{codeSRW}.
+Figure~\ref{fig:FluxU18} shows the spectra calculated over an aperture if 1 $\times$ 1 mm$^2$ placed at $d$=27.1 m.
+This aperture opening is set to fully accept the central cone at the resonance peak of the harmonic in use ($\max(\Sigma') \times d = 0.73$ mm), with $\Sigma'$ values from in Table~\ref{tab:HandCalculations}.
+The peak intensity is about  $2.9\cdot10^{14}$ photons/s/0.1{\%}bw for EBS and 1.9 10$^{14}$ for ESRF (see \ref{fig:FluxU18}).
+One can appreciate how the emittance plays an important role in both peak width and specially in peak amplitude.
+The number of photons per second after the multilayer monochromator is 2.9 10$^{14}$ for EBS and 1.9 10$^{14}$ for ESRF, considering that the energy bandwidth transmitted by a multilater monochromator is $\Delta E/E \approx 10^{-2}$ and the bandwidth in \ref{fig:FluxU18} is 10$^{-3}$, and supposing at this point ideal reflectivity (100\%) for the multilayers.
 
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{GRAPHICS/fluxU18.jpg}
+    \includegraphics[width=0.8\textwidth]{GRAPHICS/fluxU18.png}
     \caption{Flux of the U18.3 undulator integrated over a square slit placed at 27.1~m from the ID centre with aperture of 1 and 4 mm.}
@@
 
-\inred{Next we would like to estimate} the geometrical transmitivity of the beamline, therefore we have to consider which optical element crop the beam and remove photons.
+Next we would like to estimate the geometrical transmitivity of the beamline, therefore we have to consider which optical element crop the beam and remove photons.
 This can be analysed using their angular acceptance or numerical aperture.
 The multilayer monochromator accepts the whole beam.
-We consider a VSS closed to $S_h=50 \mu$m in horizontal and fully opened in vertical.
+We consider a VSS closed to $S_h=50~\mu$m in horizontal and fully opened in vertical.
 For ESRF its transmittivity is 1/8 (as discussed before) and 1 for EBS.
 In a similar way, the KB mirrors will crop the beam to a size that is the projection of their length to a direction perpendicular to the optical axis.
-Using mirror length of 60 mm and 26 mm for VFM and HFM, respectively, and incident angle of 15 mrad mrad, the KB is behaving like an aperture of $900 \times 390 \mu m^2$.
-Dividing this by the distance from the mirror to the VSS in H and to the source in V we get N.A.
-of6.2$\times$21.6 $mu$rad.
-Dividing by the beam divergence we have the new transmittivity coefficients 9.5\% (H) and 0.18\% (V)for ESRF and 16.3\% and 0.18\% for EBS.
-The overall transmission is therefore 0.2\% for ESRF and 2.95\% for EBS, therefore 7.53 $10^{12}$ photons for ESRF and 1.03 $10^{14}$ photons for EBS.
-\todo{check with referente 6 $10^9$ at 33.6 keV}
+Using mirror length of $L_v$=60 mm and $L_h$=26 mm for VFM and HFM, respectively, and incident angle of 15 mrad, the KB is behaving like an aperture of $900 \times 390 \mu m^2$.
+We get N.A.
+values of 6.2$\times$21.6 $mu$rad (aperture size divided by the distance from the mirror to the VSS in H and to the source in V).
+The transmittivity coefficient is the ratio of the accepted N.A.
+over the beam divergence, resulting 9.5\% (H) and 0.18\% (V)for ESRF and 16.3\% and 0.18\% for EBS.
+The overall transmission is the product of H and V ratios, therefore 0.2\% for ESRF and 2.95\% for EBS, therefore 3.8 $10^{11}$ photons/s for ESRF and 8.5 $10^{12}$ photons/s for EBS.
+\todo{check with references $10^9$ at 33.6 keV. I get: 1.6e+15 * 0.00295 * 0.7**3 = 1.62e+12. How closed are the slits? }
 
 One can also discuss some simple consequences about the beam coherence.
-The coherence fraction $CF$ \cite{arxivCF} of the undulator beam is the occupation of the first coherence mode, as it will be discussed in Section~\ref{level2}.
-An approximated estimation of it can be obtained using Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas}, for the horizontal and vertical planes we have have
+The coherence fraction (CF) \cite{arxivCF} of the undulator beam is the occupation of the lowest coherent mode, as it will be discussed in Section~\ref{level2}.
+An approximated estimation of it can be obtained using Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas}.
+The CF in its the horizontal and vertical planes are
 \begin{equation}
@@
 \end{equation}
-and the total coherent fraction is the product of both ($CF=CF_h \times CF_v$).
-An estimation of the CF calculated in such a mode gived CF=0.XX\% for the ESRF and XXX for EBS.
-The CF roughly indicates the ration of the ``coherent photons'' over the total number of photons.
-A the beamline optics of a beamline target for coherence application must therefore clean the beam to extract only the good coherent photons.
-A complete cleaning is impossible as there is spatial overlapping of the different coherence modes.
+and the total coherent fraction is the product of both (CF=CF$_h \times $CF$_v$).
+An estimation of the CF calculated in such a mode give CF=0.12\% for the ESRF and 2.73\% for EBS at 17 keV photon energy.
+The CF roughly indicates the ratio of the ``coherent photons'' over the total number of photons.
+It is a good indicator of the quality of a coherent beam (a fully coherent beam has CF=1 and an incoherent beam CF=0).
+The beamline optics acts as a ``coherence filter'' by removing the ``bad'' or ``incoherent'' photons and in consequence increasing the CF.
+A complete cleaning is impossible because of the spatial overlapping of the different coherence modes.
 However, apertures centered at the optical axis would make the work.
-If we compare the values transmittivity of the beamline due to geometrical clipping  (0.2\% for ESRF and 2.95\% for EBS) we find them very close to the $CF$ values, meaning that the resulting beam would be highly coherent.
+If we compare the values transmittivity of the beamline due to geometrical clipping  (0.2\% for ESRF and 2.95\% for EBS) with the the CF values, we remark that they are very close.
+This implies that the beam at the focal position would be highly coherent because the fraction of the photons removed from the beam matches the CF value that quantify the fraction of ``good'' or ``coherent'' photons.
 
-Regarding coherence, one can idealize the beamline in two parts: the ML onochromator and VSS that will prepare a coherent plane wave, as the source is very far from the sample position, and then KB that acts as a focusing element (focal=cm) that also crops the beam like it would be done by a square aperture of $D \times D=$ 400$\times$400 $\mu m^2$ \inred{(we obtained 900 in H, that will be reduced by 0.5 as propagated from 10 cm to 5 cm upstrem from the sample, and 390 in V approximated by 400)}.
-This aperture has two effects: i) removing a lot of off-axis photons what imply a increase of the coherent fraction to a situation close to a coherent source, and ii) make a broadening of the focal spot because of coherent diffraction.
-We know from optics that the diffraction by a square aperture of a collapsing spherical wavefront has an intensity distribution of:
-\begin{equation}
- I(x,y,z=f) = \frac{4 k^2 D^2}{\pi^2 f^2} \sinc^2\frac{k D x}{2 f} \sinc^2\frac{k D y}{2 f}
-\end{equation}
-where $k=2\pi/\lambda$ .
-Considering that the FWHM of $\sinc(x)$ is approximately 2.78, one obtains a FWHM of 7.75$\times$7.75 nm$^2$ nm for the focal spot.
-This spot size is of the order or smaller than the one obtained by geometrical considerations, therefore one can say that this beamline for the conditions in use is a diffraction limited beamline.
+To assess the focused beam size, one can idealize the beamline in two parts.
+First, the ML monochromator and VSS that will prepare a coherent plane wave as the source is very far from the sample position.
+Second, the KB has a double role, acting as a focusing element of focal distance $f$=5 cm, but also crops the beam as as it would be done by a square aperture of $D \times D=$ 400$\times$400 $\mu m^2$ (the length of the mirror projected in a plane perpendicular to the beam propagation) \inred{(we obtained 900 in H, that will be reduced by 0.5 as propagated from 10 cm to 5 cm upstrem from the sample, and 390 in V approximated by 400)}.
+This aperture has two effects: i) as discussed before, it removes a lot of off-axis photons increasing the CF to a value close than one, and ii) make a broadening of the focal spot because of coherent diffraction.
+We know from optics that the diffraction by a square aperture of a collapsing spherical wavefront has an intensity distribution proportional to $\sinc^2(k D x/2 f) \sinc^2(k D y / 2 f)$, where $k=2\pi/\lambda$, and $x$ ($y$) is the horizontal (vertical) coordinate .
+Considering that the FWHM of $\sinc(x)$ is approximately 2.78, one obtains a FWHM of 7.75$\times$7.75 nm$^2$ for the focal spot.
+This spot size is of the order or larger than the one obtained by geometrical considerations, therefore contributing significantly to the size of the focused beam.
+One can conclude that for the conditions in use this is a diffraction limited beamline.
 
-In summary, we show in this section how simple analytical calculations that can be done in its major part ``by hand'' help to estimate the geometrical beam size at focal position (XX$\times$XX nm$^2$) the flux (XX photons/s) and also anticipate that the beam is highly coherent producing a diffraction limited focal spot of  7.75$\times$7.75 nm$^2$.
-We ignored at this point the effect of real optics including surface errors, the reflectivity of the elements, and other aspects that will be treated in depth in the next sections.
+In summary, we show in this section how simple analytical calculations that can be done in its major part ``by hand'' help to estimate the geometrical beam size at focal position (17$\times$5 nm$^2$ for ESRF, 10$\times$5 nm$^2$ for the EBS) the flux (3.8 $10^{11}$ photons/s for ESRF and 8.5 $10^{12}$ photons/s for EBS) and also anticipate that the beam is highly coherent producing a diffraction limited focal spot of  7.75$\times$7.75 nm$^2$.
+We ignored at this point the effect of real optics containing surface errors, the reflectivity of the elements, and other aspects that will be treated in depth in the next sections.
 This is, however, the first step to be done when analyzing a beamline.
-A similar work was done at the design level in order to define the working parameters.
+It also helps in having a critical eye or sanity check on the numerical calcuations that will provide more precised values, but must remain the this order of magnitude.
 
@@
 The first subsection deals with the effect of aberrations and transmittivity due to geometrical considerations.
-The second subsection shows how the effect of slopes erroin the KB mirrors deteriorate the focal spot.
+The second subsection shows how the effect of slopes errors the KB mirrors degrade the focal spot.
 
-Ray tracing is a simulation method based on tracing some light rays along the optical system from the source to the image and retrieve the beam characteristics at the image (manily size and divergences).
+Ray tracing is a simulation method based on tracing some light rays along the optical system from the source to the image and retrieve the statistics of rays at the image to measure beam characteristics.
 For simple systems one can perform ray tracing by hand, selecting some principal rays that will define the location and envelope of the image.
@@
 This allows to extend the pure geometrical tracing to include physical models that take into account the optical element reflectivity or transmittivity.
-For mirrors and lenses one should use the Fresnel's equations for reflection and transmission to affect the ray intensity.
-This combination of using a geometrical model plus a physical model allows to simulate every element used in a synchrotron beamline, such as mirrors and combination of them (e.g., KBs); lenses, CRLs and transfocators; gratings of any type and shape (plane, VLS), crystal systems, etc.
-It makes ray tracing a simple and extremely powerful technique for calculating the main characteristics of the photon beam (size, divergence, photon distribution) at every point of the beamline.
+This combination of using a geometrical model plus a physical model allows to simulate every element used in a synchrotron beamline, such as mirrors and combination of them (e.g., KBs); lenses, compound refractive lenses and transfocators; gratings of any type and shape, crystal systems, etc.
+Ray tracing is a simple and extremely powerful technique for calculating the main characteristics of the photon beam (size, divergence, photon distribution) at every point of the beamline.
 Although very powerful commercial ray tracing packages are available in the market, they are not suitable for synchrotron simulations because they do not cover the particularities of X-ray devices and sources.
@@
 One is that the migration from ESRF lattice to the EBS will significantly improve the horizontal size, and the VSS is not needed.
-Secontly, one can observe a much larger luminosity in the EBS case.
+For EBS, the whole beam passes though the VSS, producing a final spot with almost Gaussian horizontal intensity profile instead of the step distribution for ESRF.
+Secondly, one can observe a much larger luminosity in the EBS case.
 The transimittivity values are 0.165\% for ESRF and 2.28\% for EBS.
 These values confirm what we calculated by hand (0.2\% for ESRF and and 2.95\% for EBS).
-As mentioned before, these values are similar to the coherent fraction \cite{coherentfraction} (0.13 and 2.8\%, respectively), meaning that a very high coherence of the beam is expected, therefore some degradation of the beam by clipping the beam by apertures and element acceptance is expected.
+As mentioned before, these values are similar to the coherent fraction \cite{coherentfraction} (0.12 for ESRF and 2.73\% for EBS), meaning that a very high coherence of the beam is expected, therefore some degradation of the beam by clipping the beam by apertures and element acceptance is expected.
 
@@
 It can be observed a small broadening of the peak due to slope errors, but, more importantly, a new satellite structure close to the vertical intensity peak is produced by the slope errors.
-\todo{may be play with slope errors and use synthetic profiles to show the importance of the low frequencies in the intensity distribution?}
+
+\todo{
+\begin{itemize}
+ \item play with slope errors and use synthetic profiles to show the importance of the low frequencies in the intensity distribution?
+ \item explain why KB errors are treated with hybrid and not with shadow. In Discussion section? 
+\end{itemize}
+}
 
@@
 
+In this section the coherence properties of the beamline are analysed using, again, different methodologies classified by complexity and computer resources.
+We first analyze the beamline supposing a fully coherent beam.
+Subection~\ref{wofry} shows a very simpified 1D model.
+Subsection~\ref{srw_se} used a full 2D model with accurate simulation of the undulator emission but without considering storage ring emittance.
+These simplifications permit to address diffraction effects by apertures and slope errors.
+Then we introduce methods to treat the partial coherence originated by the storage ring emittance.
+Two variants are shown, a Monte Carlo approach (subsection~\ref{srw_me}) that permits us to calculate in great detail the intensity dostribution of the focused beam, and a decomposition of the radiation in coherent modes (subsection~\ref{comsyl}) that allows to evaluate quantitatively the quality of the coherent beam.
+
 \subsubsection{Simplified wave optics simulation (coherent case in 1D with point source and ideal elements)}
+\label{wofry}
 
-The characteristics of this particular beamline, a the very long length combined with a small source (in particular for EBS), makes this beamline very adapted for studied requiring coherence.
+The very long distance source-sample combined with a small source (in particular for EBS) makes this beamline very adapted for studied requiring coherence.
 Indeed, it has been shown that the clipping of the beam, mainly due to the acceptance of the KB mirrors produces a significant broadening of the focal spot.
@@
 \centering
-\includegraphics[width=4.25cm]{GRAPHICS/wofry400umslit.png}
-\includegraphics[width=4.25cm]{GRAPHICS/wofry400umprofile.png}
+\includegraphics[width=0.45\textwidth]{GRAPHICS/wofry400umslit.png}
+\includegraphics[width=0.45\textwidth]{GRAPHICS/wofry400umprofile.png}
 \caption{Diffraction spot producedof a spherical wavefront (R=-50mm) by a 400x}
@@
 
-The first step in the wave optics simulation is to assume a full coherent source and simulate the propagation of such wavefront through the beamline
-The wavefront simulation of the beamline include propagation of the beam from element to element (drift spaces) using different possible propagators.
-Usually a propagator based on the Freskel-Kirchhoff integral is used, as it is valid for both near and far field.
-The applications of this (and other) operators imply the calculation of an integral (or sum) at each pixel of the image plane.
-This is costly from the computer point of view, so Fourier Transforms (Fourier Optics) are used to reduce complexity and calculation time.
-This is practically the only choice for simulating 2D images.
-The use of Fourier Transforms adds another problem: the result is very dependent on the sampling of the wavefront (i.e., pixel size, number of points, dimension of the window where the wavefront is sample).
-This means the usually one has to test many configurations before finding one that works reaonably well (even though one can never guarantee the complete accuracy of the sampling used).
-A full wavefront optics simulation is an iterative process, where the user has to refine the sampling parameters element by element to be sure the wavefront at each position is sampled correctly.
-This is a time consuming work for performing it with the full 2D simulation.
-For this reason, a wavefront calculation may be started by a simplified system, using simplified models for the source and focusing elements.
-Moreover, because the horizontal and vertical coupling is usually small in synchrotron beamlines, it is convinient to separate the simulation in two, one for the horizontal plane and one for the vertical one.
-
 We perform here a simplified wavefront simulation in 1D, thus separating the simulations of the horizontal and vertical planes.
-The source can be simplified by using a point source (spherical wavefront) and applying a Gaussian intensity profile with $\sigma=\sigma'_u D$ with $\sigma'_u$ given by Eq.~\ref{eq:photon small sigmas} and $D$ the distance from the point source to the observation plane.
+The source can be simplified by using a point source (spherical wavefront) and applying a Gaussian intensity profile with $\sigma=\sigma'_u d$ with $\sigma'_u$ given by Eq.~\ref{eq:photon small sigmas} and $d$ the distance from the point source to the observation plane.
 Each optical element can be simplified by separating its action in two parts: the focusing behaviour implemented as an ideal lens, and the finite size because of the boundaries, implemented as a slit.
-The WOFRY package in OASYS permits to do this work in a quick way, and it is possible to arrive to a satisfactory result in a few iterations.
+The WOFRY package in OASYS permits to do this work, and it is possible to arrive to a satisfactory result in a few iterations.
 The beam size at different positions of the beamline are shown in Table~\ref{tablewofry1D}.
@@
     \caption{Beam sizes calculated with simplified wave optics propagation (point source and ideal lenses with aperture for modeling the element dimensions). 
-\todo{implement a sort of multielectron system - implement the slope error} }
+    }
     \label{tablewofry1D}
@@
     \begin{tabular}{cccccc} 
+    % using WAVEOPTICS WOFRY workspace
     Element:                & ML            & VSS          & KBv               & KBh           & sample        \\
     Position [m]:           & 28.3          & 40.00        & 184.90            & 184.95        & 185.00        \\
-    H beam size [$\mu$m]:   & 328           & 1.11         & 4084              & 390           & 0.010         \\
-    V beam size [$\mu$m]:   & 328           & 466          & 899               & 450           & 0.010         \\ 
-    H scaling factor:       & 1.0           & 0.01         & 220               & 1             & 0.00002       \\
+    H beam size [$\mu$m]:   & 329           & 1.11         & 4084              & 390           & 0.008         \\
+    V beam size [$\mu$m]:   & 329           & 466          & 899               & 450           & 0.007         \\ 
+    H scaling factor:       & 1.0           & 0.01         & 440               & 1             & 0.00002       \\
     V scaling factor:       & 1.0           & 0.64         & 5                 & 0.5           & 0.00002       \\
@@
 \centering
-\includegraphics[width=4.25cm]{GRAPHICS/wofry1Dh.png}
-\includegraphics[width=4.25cm]{GRAPHICS/wofry1Dv.png}
-\caption{Intensity distribution at the focal position in horizontal plane (left) and in vertical plane (right).}
+\includegraphics[width=0.45\textwidth]{GRAPHICS/wofry1Dh.png}
+\includegraphics[width=0.45\textwidth]{GRAPHICS/wofry1Dv.png}
+\caption{Intensity distribution at the focal position in horizontal plane (left, FWHM=8.1 nm) and in vertical plane (right, FWHM=7.1 nm).
+\todo{implement a sort of multielectron system - implement the slope error. In Discussion? } 
+}
 \end{figure}
@@
 \subsubsection{Full wave optics simulation for coherent case}
+\label{srw_se}
 % The reduction on the electron beam emittance, by reducing the electron beam horizontal size, leads to an increase on the horizontal lateral coherence of the radiation emitted by an undulator in the symmetry point of a straight section \cite{paper}.
@@
 % 
-The first step in the wave optics simulation is to assume a full coherent source and simulate the propagation of such wavefront through the beamline.
-The source used on the simulations to follow is the on-axis-single-electron emission from the U18.3 undulator tuned to its first harmonic at $E=17225$ eV, which is inherently fully coherent.
+A next step in the wave optics simulation is, maintaining the full coherence, to generate a 2D source wavefront that matches in detail the real source (an undulator in a storage ring) and then simulate the propagation of such a wavefront through the optical elements with models that match well the characteristics of the real elements.
+The source simulated is the on-axis-single-electron emission from the U18.3 undulator tuned to its first harmonic at $E=17225$ eV.
+The emission by a single-electron is fully coherent.
 Once the source is defined, one has to go around on how to simulate the beamline optical elements.
-Increasing the complexity of the simulations one can: a) use ideal thin lenses preceded by slits to emulate finite aperture; or b) use grazing incidence focusing optics models \cite{Canestrari2014}.
+Increasing the complexity of the simulations one can: a) use ideal thin lenses preceded by slits to emulate finite aperture (as done in the 1D model presented in Section~\ref{wofry}); or b) use grazing incidence focusing optics models \cite{Canestrari2014}.
 On the top of that, one can add the effect of imperfect optics to the simulations, i.e.
@@
 
-The computer code chosen for these calculations is the \textsc{srw} \cite{codeSRW}, which over the past two decades has been extensively benchmarked throughout different synchrotron radiation facilities and has become widely accepted within the X-ray community.
+The computer code chosen for these calculations is Synchrotron Radiation Workshop (\textsc{srw}) \cite{codeSRW}, which over the past two decades has been extensively benchmarked throughout different synchrotron radiation facilities and has become widely accepted within the X-ray community.
 
 Figure \ref{fig:SingleElectron} shows the horizontal integrated intensity at the focal position for the ID16A.
-It can been seen that there is little difference between using thin ideal lenses or the elliptical mirror model, confirming the previous ray-tracing results.
+It can been seen that there is little difference between using thin ideal lenses or the elliptical mirror model.
 The elliptical mirror presents a slight reduction in intensity and asymmetry but no significant difference is found, which is in accordance with what is reported in \cite{Canestrari2014}.
-Notica that the differences are minimal and have to be observed in log plots.
+Notice that the differences are minimal and to remark them we have chosen log plots to represent the intensity distributions at the focal position.
 It is also important to mention that this equivalency between ideal lenses and elliptical mirrors is only valid at the focal position vicinity: down- or upstream the focal position, differences are not negligible \cite{Canestrari2014}.
@@
 It is also possible to add slope errors to the simulations to study the degradation on the focal spot quality (Fig.~\ref{fig:SingleElectron}).
-For that, the available available metrology data (Appendix 1) are used.
-For both cases (i.e., simulating the beamline using thin optical elements or elliptical mirrors), the beam sizes at the focal position are little affected by the presence of the slope errors (see Table \ref{tab:SRW_results}).
-There is a increase in the background level: the radiation that should be on the main lobe (focus) goes to the side lobes as a result of adding the slope errors.
+For that, the available available metrology data \inred{(Appendix 1)} are used.
+For both cases (i.e., simulating the beamline using thin optical elements or elliptical mirrors), the beam sizes calculated as the FWHM at the focal position are little affected by the presence of the slope errors (see Table \ref{tab:SRW_results}).
+There is a significant reduction and shuffling of the interference fringes observed with no slope error.
+They are almost cancelled, with is an important raise of the background level: the radiation found on the main lobe (focus) shifts to the side lobes when the slope errors are considered.
 This is accompanied by an obvious reduction on the peak intensity.
@@
     \label{fig:SingleElectron}
-    \caption{Wavefront propagation from a single electron emission through the ID16A beamline. }
+    \caption{Wavefront propagation from a single electron emission through the ID16A beamline.
+Left: the perfect focusing with ideal lenses (black) is compared with the focusing using elliptical mirrors (red).
+Right: the same as in the previous plot, but adding the effect of the mirrors slope errors in bothe cases.}
     %Simulations took less than one minute in a desktop computer Intel Core i7 with 8Gb of RAM memory.}
@@
 \subsubsection{Full wave optics simulation: partial coherent case by means of multielectron Monta Carlo sampling}
+\label{srw_me}
 The wavefront propagation methodology is based in the sampling of a single wavefront (monochromatic and coherent) in a 2D space (in horizontal and vertical spatial dimensions) at a given point of the beamline.
@@
 The beam sizes and shapes are similar at the focal position, only.
-Simulations took significantly longer to be completed: few hours on a 56-core computer cluster.
 
@@
 \subsubsection{Full Wave optics simulation: partial coherence case by means of coherent mode decomposition}
+\label{comsyl}
 % 
@@
 
-The cross spectral density (CSD) (sometimes called mutual intensity) defines the correlation of the radiated electric fields between two spatial points $(x_1,y_1)$ and $(x_2,y_2)$ at a given position of the beamline $z$:
+The cross spectral density (CSD) (sometimes called mutual intensity) completely describes the coherence of a beam. 
+At a given plane at a distance $z$ from the source, for a photon beam of frequency $\omega$, the CSD is the correlation of the radiated electric fields between two spatial points $(x_1,y_1)$ and $(x_2,y_2)$:
 
-\begin{equation}
+\begin{equation}\label{CSD}
 W(x_1,y_1,x_2,y_2;z,\omega) = <E*(x_1,y_1; z,\omega) E(x_2,y_2;z,\omega)>
-\label{CSD}
 \end{equation}
-where $E$ is the amplitude of the electric field, $\omega$ is the radiation frequency, and $<>$ stands for the ensemble average.
-The CSD is at the origin of any other information aboute the coherence of the beam, like the spectral density $S$ (or, just intensity):
-\begin{equation}
+where $E$ is the amplitude of the electric field, and $<>$ stands for the ensemble average.
+The CSD is at the origin of any other information about the coherence of the beam, like the spectral density $S$ (or, spatial distribution of the intensity):
+\begin{equation}\label{SD}
 S(x,y,;z,\omega) = W(x,y,x,y,;z,\omega)
@@
 or the spectral degree of coherence $\mu$:
-\begin{equation}
+\begin{equation}\label{SDC}
 \mu(x_1,y_1,x_2,y_2;z,\omega) = \frac{W(x_1,y_1,x_2,y_2;z,\omega)}{\sqrt{ S(x_1,y_1;z,\omega) S(x_3,y_2;z,\omega}}
@@
 Considering also that the progagation of $W$ from a beamline position $z_1$ to another position $z_2$ will need 10$^8$ to 10$^{12}$ 4D integrals, one can easily imagine the physical impossibility to perform such calculations.
-It is interesting to remark that although Monte Carlo multielectron method can in principle calculate Eq.~\ref{CSD} the storage at every point is unmanagleable, so usually only some 'cuts' of the CSD are stored, such as $W(x_1,0,x_2,0;z,\omega)$ (horizontal) or $W(0,y_1,0,y_2;z,\omega)$ (vertical).
 
-A solution of this problem is the expansion of the CSD in coherent modes \cite{mandel_wolf}:  
+A solution to this problem is the expansion of the CSD in coherent modes \cite{mandel_wolf}:  
 \begin{equation}
@@
  \end{equation}
-where $\lambda_m$ are the eigenvalues or weights, and $\Phi_m$ are the eigenfunctions or coherent modes.
+where $\lambda_m$ are the eigenvalues or weights, and $\Phi_m$ are the eigenfunctions or (orthonormal) coherent modes.
 
-A theoretical study and a numerical algorithm for the coherent mode decomposition of undulator emission in storage rings has been recently proposed  \cite{GlassThesis,GlassEPL} and implemented in the computer package COMSYL \cite{codeCOMSYL}.
+A theoretical study and a numerical algorithm for the coherent mode decomposition of undulator emission in storage rings has been recently proposed  \cite{GlassThesis,GlassEPL} and is implemented in the computer package COMSYL \cite{codeCOMSYL}.
 The innovative concept is the numerical description of the undulator CSD in terms of its coherent modes $\Phi_m(x,y;z,\omega)$ and eigenvalues $\lambda_m $.
-One can define the mode distribution as:
+One can define the mode distribution as
 \begin{equation}
- d_m = \frac{\lambda_m}{\sum\limits_{i=0}^{\infty} \lambda_i}
+ d_m = \frac{\lambda_m}{\sum\limits_{i=0}^{\infty} \lambda_i},
 \end{equation}
-or the cumulated mode distribution as: 
+or the cumulated mode distribution like
 \begin{equation}
- c_m = \sum\limits_{i=0}^{m} d_i
+ c_m = \sum\limits_{i=0}^{m} d_i.
 \end{equation}
-An important property of this coherent mode decomposition is that it optimizes the representation of the CSD when the series is truncated.
+An important property of the coherent mode decomposition is that it optimizes the representation of the CSD when the series is truncated.
 In other words, truncating the series in Eq.~\ref{decomposition} up to the term $m$, the CSD obtained approximates better the exact CSD (this with infinite terms) than any other expansion.
 It also imply that the first term is more important than the second, and so on.
-The first term (index zero) is the first coherent mode.
-It relative importance with respect to the full CSD permits to define the coherent fraction (CF) as the occupation of the firts coherence mode: CF=$d_0$.
+The first or lower term (index zero) is the most important coherent mode.
+It permits to define the coherent fraction (CF) as the occupation of the lower coherent mode: CF=$d_0$.
 Obviously, a pretty coherent source will require less modes than a mostly incoherent source to arrive up to a certain representation of the CSD.
@@
 The obtained CF are 2.8\% for EBS and 0.13\% for ESRF-High$\beta$, indicating that the overall coherent flux emitted by EBS will be about 20 times the existing coherent flux with ESRF-High$\beta$.
-Fig.~\ref{histomodes}a shows the cumulated occupation of the coherent modes for the full central cone undulator emission.
-It can be shown how much less number of coherent modes are needed to arrive to a certain degree of cumulated occupation for EBS than for ESRF-High$\beta$.
-This can also be shown in Fig.\ref{histomodes}b where the number of modes to fill 50\%, 75\% and 95\% CSD occupation are compared.
+Fig.~\ref{fig:histomodes}a shows the cumulated occupation of the coherent modes for the full central cone undulator emission.
+It shows that the EBS cumulated occupation grows very steply with the number of coherent modes, whereas for ESRF the number of coherent modes needed to attain a given value of cumulated occupation much higher.
+This can also be shown in Fig.\ref{fig:histomodes}b, where the number of modes to fill 50\%, 75\% and 95\% of spectral density are compared for EBS and ESRF.
 
@@
 This task is performed by the beamline.
-The beamline elements modify the different modesin a different manner.
+The beamline elements modify the different modes in a different manner.
 The fist coherence mode is always centered on-axis, and the subsequent ones expand in the space until completely filling the spectral density.
-\begin{figure}
+\begin{figure}\label{fig:histomodes}
     \centering
@@
         \includegraphics[width=6.5cm]{GRAPHICS/up_to_mode_id16a.eps}
-    \label{histomodes}
     \caption{Cumulated occupation of the coherent modes for the EBS and ESRF-High-$\beta$ (top), and number of modes needed to reach a given percent of CSD occupation (bottom). }
@@
 
-\begin{figure}
+\begin{figure}\label{fig:spectraldensity}
     \centering
@@
         \includegraphics[width=0.45\textwidth]{GRAPHICS/ebs_mode3.png}
-    \label{spectraldensity}
     \caption{XXXX }
@@
 
+The first step in the wave optics simulation is to assume a full coherent source and simulate the propagation of such wavefront through the beamline
+The wavefront simulation of the beamline include propagation of the beam from element to element (drift spaces) using different possible propagators.
+Usually a propagator based on the Freskel-Kirchhoff integral is used, as it is valid for both near and far field.
+The applications of this (and other) operators imply the calculation of an integral (or sum) at each pixel of the image plane.
+This is costly from the computer point of view, so Fourier Transforms (Fourier Optics) are used to reduce complexity and calculation time.
+This is practically the only choice for simulating 2D images.
+The use of Fourier Transforms adds another problem: the result is very dependent on the sampling of the wavefront (i.e., pixel size, number of points, dimension of the window where the wavefront is sample).
+This means the usually one has to test many configurations before finding one that works reaonably well (even though one can never guarantee the complete accuracy of the sampling used).
+A full wavefront optics simulation is an iterative process, where the user has to refine the sampling parameters element by element to be sure the wavefront at each position is sampled correctly.
+This is a time consuming work for performing it with the full 2D simulation.
+For this reason, a wavefront calculation may be started by a simplified system, using simplified models for the source and focusing elements.
+Moreover, because the horizontal and vertical coupling is usually small in synchrotron beamlines, it is convinient to separate the simulation in two, one for the horizontal plane and one for the vertical one
+
+Simulations took significantly longer to be completed for the multi e case: few hours on a 56-core computer cluster.
+
+It is interesting to remark that although Monte Carlo multielectron method can in principle calculate Eq.~\ref{CSD} the storage at every point is unmanagleable, so usually only some 'cuts' of the CSD are stored, such as $W(x_1,0,x_2,0;z,\omega)$ (horizontal) or $W(0,y_1,0,y_2;z,\omega)$ (vertical) \todo{move this to the discussion?}.
+
 \todo{

### 2019-03-28 paper-hierarchical.tex
update
+

@@
 Section~\label{Simulations} is the central part of the paper, and describe the different methods used to analyse the beamline with increasing complexity, setting a hierarchy of methodologies that can be exported to the simulations of any beamline.
-In subsection \ref{level0} we calculate with simple methods approximated values of focal size, flux and coherent fraction.
-Then accuracy is improved using geometrical methods in subsection \ref{level1} that give reasonable results as far as some corrections are considered because of the  mentioned characteristics of high demagnification and high coherence of the studied beamline.
+In Section \ref{level0} we calculate with simple methods approximated values of focal size, flux and coherent fraction.
+Then accuracy is improved using geometrical methods in Section \ref{level1} that give reasonable results as far as some corrections are considered because of the  mentioned characteristics of high demagnification and high coherence of the studied beamline.
 The wave optics section~]ref{level2} also include different models whose complexity increases when the results expected are more and mode demanding.
@@
 The surface finish of the optical elements is another limiting effect for determining the size and intensity of the beam at the image position.
-It is possible to infere the effect of mirror imperfections and in particular slope errors by only knowing some statistical parameters such as the slope error Root Mean Squate (RMS) and then building synthetic surface profiles to be used in the simulations.
+It is possible to infere the effect of mirror imperfections and in particular slope errors by only knowing some statistical parameters such as the slope error Root Mean Square (RMS) and then building synthetic surface profiles to be used in the simulations.
 This is the typical mode to operate when designing a new beamline where the mirrors do not exist yet.
@@
 
-The very long distance source-sample combined with a small source (in particular for EBS) makes this beamline very adapted for studied requiring coherence.
+The very long distance source-sample combined with a small source (in particular for EBS) makes this beamline suitable for experimental techniques requiring a highly coherent beam.
 Indeed, it has been shown that the clipping of the beam, mainly due to the acceptance of the KB mirrors produces a significant broadening of the focal spot.
@@
 
-The Monte Carlo multielectron method has been proposed by O. Chubar and it is fullt implemented in SRW \cite{codeSRW}. We have used this code for the calculations presented in this section.
+The Monte Carlo multielectron method has been proposed by O. Chubar and it is fully implemented in SRW \cite{codeSRW}. We have used this code for the calculations presented in this section.
 
@@
 A solution to this problem is the expansion of the CSD in coherent modes \cite{mandel_wolf}:  
-\begin{equation}
+\begin{equation}\label{decomposition}
  W(x_1,y_1,x_2,y_2;z,\omega) = \sum\limits_{m=0}^{\infty} \lambda_m \Phi_m^{*}(x_1,y_1;z,\omega) \Phi_m(x_2,y_2;z,\omega)
-\label{decomposition}
  \end{equation}
@@
 
-The coherence fraction of 2.8\% for the EBS source means that for an optimal coherence experiment one should remove 97.2\% of the emitted photons.
-This task is performed by the beamline.
-The beamline elements modify the different modes in a different manner.
-The fist coherence mode is always centered on-axis, and the subsequent ones expand in the space until completely filling the spectral density.
 \begin{figure}\label{fig:histomodes}
@@
 \end{figure}
+
+The spectral density at the EBS photon source is represented in Fig.~\ref{fig:spectraldensity}a.
+It shows an elliptical shape with horizontal size larger than the vertical.
+The spectral density is reproduced by the addition of the coherent modes weighted by the eigenvectors (from Eq.~\ref{decomposition})
+\begin{equation}
+S(x,y) = \sum_{m=0}^{\infty} \lambda_m |\Phi_m(x,y)|^2.
+\end{equation}
+Figs.~\ref{fig:spectraldensity}b-e represent the coherent modes with $m$=0 to $m$=3.
+It can be shown that the first mode is centered on the axis but extends over an spatial region smaller than the spectral density.
+Next mode ($m$=1) contributes to extend the spectral density in horizontal, but presents zero intensity at the origin.
+Successive modes extend  the spectral density in horizontal, sometimes presenting a lobe centerer at zero and sometimes not (e.g., $m$=3).
+Modes with several lobes along the vertical will also appear, but not so often than in the horizontal, because of the elliptical shape of the spectral density.
+The coherent modes are orthonormal, thus verifying
+\begin{equation}
+S(x,y) = \int_{-\infty}^{\infty} \Phi^*_m(x,y) \Phi_n(x,y)~dx~dy = \delta_{mn}
+\end{equation}
+with $\delta_{mn}$ the Kronecker delta.
+The coherence fraction of 2.8\% for the EBS source means that for an optimal coherence experiment one should remove 97.2\% of the emitted photons.
+This task is performed by the beamline.
+The beamline elements modify the different modes in a different manner, but it will never be possible to remove all the photons except from the lowest coherent mode, because there is an overlapping of the intensity of the lowest mode with the other ones.
+Therefore it is impossible to obtain a fully coherent beam that must contain only a single mode.
 
@@
         \includegraphics[width=0.45\textwidth]{GRAPHICS/ebs_mode3.png}
-    \caption{XXXX }
+    \caption{a) Spectral density $S(x,y)$ at the source position for EBS.  }
 \end{figure}
 
-The coherent modes for the EBS have been propagated using a beamline simplified with ideal focusing elements and slits to determine their numerical aperture.
+The coherent modes for the EBS have been propagated using a beamline simplified with ideal focusing elements and slits to determine their numerical aperture (like in Section~\ref{wofry}).
 Fig.~\ref{fig:mode transmission} shows how the intensity of the first modes is reduced by geometrical considerations.
-We recall that in the coherent mode expansion, the modes are orthonormal, therefore the integrated intensity for any mode is always one at the source.
-Different modes will be ``cut'' in a different manner, depending on the slits cropping effect.
-Fig.~\ref{fig:mode transmission} shows that the lowest mode (zeroth order) transfer is about 25\%, the mode 2 has a 10\% transfer and the 4th mode about 6\%.
-The other modes have less than 5\% transfer and modes higher than 10 practically do not contribute to the final spectral density.
-Note that the transmission os mode 2 is higher than mode 1.
-This is because the mode 3 has its highest intensity at the origin, but mode 2 has a hole in the middle, as shown in Fig.
-XX.
-Modes with maximum intensity at origin will be transmitted more than the others.
-This analysis of the mode transmission is mode dramatic if one looks to the total intensity in the mode, which is given by the eigenvalue.
-Neglecting modes higher than 10, the intensity transmitted for the source mode zero amounts for 49.5\%, 18.4\% for the mode 2 less than 5\% for the others.
-As commented before, the source modes cropped by the beamline elements are not longer orthonormal, therefore they are not longer coherent modes of the resulting beam.
-The new coherent modes can be calculated applying a new coherent mode decomposition on the resulting spectral density.
-Making this numerical calculation with COMSYL, one obtains a mode occupation spectrum where the lowest mode accounts for the 83\% of the total spectral density (see Fig.~\ref{fig: rediagonalization}).
-This analysis confirms the quatitative predictions discussed in the previous sections, but give a quantitative figure of the coherence of the beam.
+%We recall that in the coherent mode expansion, the modes are orthonormal, therefore the integrated intensity for any mode is always one at the source. 
+As discussed, coherent modes will be ``cut'' in a different manner, depending on the slits cropping effect.
+Fig.~\ref{fig:mode transmission} shows that the lowest mode ($m$=0) transfer is about 25\%, the mode $m$=2 has a 10\% transfer and the mode $m$=4 about 6\%.
+The other modes have less than 5\% transfer and modes higher than $m$=10 practically do not contribute to the final spectral density.
+Note that the transmission of mode $m$=2 is higher than mode $m$=1, because this one does not contain a central lobe (Fig.~\ref{fig:spactraldensity}c).
+
+\begin{figure}\label{fig:mode transmission}
+    \centering
+        \includegraphics[width=\textwidth]{GRAPHICS/mode_transmission.png}
+    \caption{Transmission of the different modes due to geometrical effects.}
+\end{figure}
+
+This analysis of the mode transmission is mode dramatic if one looks to the total intensity in the mode, which is given by its eigenvalue.
+Neglecting modes higher than 10, the lowest coherent mode ($m$=0) carries the 49.5\% of the beam intensity, 18.4\% for the mode $m$=2, and less than 5\% for the others.
+As commented before, the coherent modes at the source, propagated and cropped by the beamline, are not longer orthonormal, therefore they cannot be called coherent modes of the resulting beam.
+At this point, the new coherent modes can be calculated applying a new coherent mode decomposition on the cross spectral density of the transmitted beam, that is known from source coherent mode expansion.
+Making this numerical calculation with COMSYL, one obtains a mode occupation spectrum (see Fig.~\ref{fig:rediagonalization}) where the lowest mode accounts for the 83\% of the total spectral density.
+The next mode accounts for only XX\% and the rest are practically noise.
+The intensity of these two modes is shown in Fig.~\ref{fig:final modes}.
+The first mode has dimensions of XXxXX nm$^2$ and the spectral density, that is practically given by the summ of the intensity of these two modes, is XXxXX nm$^2$.
+
+This analysis confirms the quantitative predictions discussed in the previous sections, but give a quantitative figure of the coherence of the beam, a  coherent fraction of CF=83\% that measures the "coherence purity".
 
@@
 
-\begin{figure}\label{fig:mode transmission}
-    \centering
-        \includegraphics[width=\textwidth]{GRAPHICS/mode_transmission.png}
-
-    \caption{Transmission of the different modes due to geometrical efffects.}
-\end{figure}
-
 \begin{figure}\label{fig:rediagonalization}
@@
 \begin{itemize}
- \item propagate modes
- \item calculate coherent fraction at the focal position
- \item show some modes
  \item create some plots with cuts of W
@@
 
-Simulations took significantly longer to be completed for the multi e case: few hours on a 56-core computer cluster.
+Simulations took significantly longer to be completed for the multi electron case: few hours on a 56-core computer cluster.
 
 It is interesting to remark that although Monte Carlo multielectron method can in principle calculate Eq.~\ref{CSD} the storage at every point is unmanagleable, so usually only some 'cuts' of the CSD are stored, such as $W(x_1,0,x_2,0;z,\omega)$ (horizontal) or $W(0,y_1,0,y_2;z,\omega)$ (vertical) \todo{move this to the discussion?}.
+
+The fist coherence mode is always centered on-axis (see, e.g, Fig.~\ref{fig:spectraldensity}b) and will always be centered when propagated until a position where an optical element (an aperture or mirror with finite dimensions) will crop it removing part of its intensity but letting the central part to go throught.
+The subsequent modes will be more absorbed as they extend more.
+In particular the modes that do not have a central lobe will lose more intensity.
 

### 2019-03-29 paper-hierarchical.tex
almost there

@@
 Several methodologies with increasing complexity are applied to discuss the relevant parameters that quantify the beamline performances.
-Parameters such as flux, beam size at the image and coherence properties are obtained from simple hand-calculations to sophisticated computer simulations using ray tracing and waveoptics techniques.
+Parameters such as flux, beam size at the image and coherence properties are obtained from simple analytical calculations to sophisticated computer simulations using ray tracing and waveoptics techniques.
 A last generation x-ray nanofocusing beamline for coherent  applications (ID16A at the ESRF) has been chosen  to study in detail the issues
@@
 
-\subsection{Calculations ``by hand''}
+\subsection{Analytical calculations}
 \label{level0}
@@
 For that, we need the spectrum at the source.
-It is not straightforward to get this information ``by hand'' but one can calculate the flux and power using different codes.
+It is not straightforward to get this information analytically but one can calculate the flux and power using different codes.
 Perhaps the most performant and well maintained codes spread in the synchrotron community are SPECTRA \cite{codeSPECTRA} and SRW \cite{codeSRW}.
@@
 The CF in its the horizontal and vertical planes are
-\begin{equation}
+\begin{equation}\label{eq:coherent fraction}
  CF_{(h,v)} = \frac{\sigma_u \sigma'_u}{\Sigma_{(h,v)} \Sigma'_{(h,v)}},
@@
 Ray tracing is a simulation method based on tracing some light rays along the optical system from the source to the image and retrieve the statistics of rays at the image to measure beam characteristics.
-For simple systems one can perform ray tracing by hand, selecting some principal rays that will define the location and envelope of the image.
+For simple systems one can perform ray tracing even by hand, selecting some principal rays that will define the location and envelope of the image.
 Using computers one can trace thousands of rays and calculate the position and divergence distribution of the rays using statistics.
@@
 The transimittivity values are 0.165\% for ESRF and 2.28\% for EBS.
-These values confirm what we calculated by hand (0.2\% for ESRF and and 2.95\% for EBS).
+These values confirm the analytical estimations (0.2\% for ESRF and and 2.95\% for EBS).
 As mentioned before, these values are similar to the coherent fraction \cite{coherentfraction} (0.12 for ESRF and 2.73\% for EBS), meaning that a very high coherence of the beam is expected, therefore some degradation of the beam by clipping the beam by apertures and element acceptance is expected.
@@
 The coherent modes are orthonormal, thus verifying
-\begin{equation}
-S(x,y) = \int_{-\infty}^{\infty} \Phi^*_m(x,y) \Phi_n(x,y)~dx~dy = \delta_{mn}
+\begin{equation}\label{eq:mode normalization}
+M_{mn} = \int_{-\infty}^{\infty} \Phi^*_m(x,y) \Phi_n(x,y)~dx~dy = \delta_{mn}
 \end{equation}
@@
 
-The coherent modes for the EBS have been propagated using a beamline simplified with ideal focusing elements and slits to determine their numerical aperture (like in Section~\ref{wofry}).
+The coherent modes calculated at the EBS source have been propagated using a beamline simplified with ideal focusing elements and slits to determine their numerical aperture (like in Section~\ref{wofry}).
 Fig.~\ref{fig:mode transmission} shows how the intensity of the first modes is reduced by geometrical considerations.
@@
 As discussed, coherent modes will be ``cut'' in a different manner, depending on the slits cropping effect.
-Fig.~\ref{fig:mode transmission} shows that the lowest mode ($m$=0) transfer is about 25\%, the mode $m$=2 has a 10\% transfer and the mode $m$=4 about 6\%.
-The other modes have less than 5\% transfer and modes higher than $m$=10 practically do not contribute to the final spectral density.
-Note that the transmission of mode $m$=2 is higher than mode $m$=1, because this one does not contain a central lobe (Fig.~\ref{fig:spactraldensity}c).
-
-\begin{figure}\label{fig:mode transmission}
-    \centering
-        \includegraphics[width=\textwidth]{GRAPHICS/mode_transmission.png}
-    \caption{Transmission of the different modes due to geometrical effects.}
-\end{figure}
-
-This analysis of the mode transmission is mode dramatic if one looks to the total intensity in the mode, which is given by its eigenvalue.
-Neglecting modes higher than 10, the lowest coherent mode ($m$=0) carries the 49.5\% of the beam intensity, 18.4\% for the mode $m$=2, and less than 5\% for the others.
-As commented before, the coherent modes at the source, propagated and cropped by the beamline, are not longer orthonormal, therefore they cannot be called coherent modes of the resulting beam.
-At this point, the new coherent modes can be calculated applying a new coherent mode decomposition on the cross spectral density of the transmitted beam, that is known from source coherent mode expansion.
-Making this numerical calculation with COMSYL, one obtains a mode occupation spectrum (see Fig.~\ref{fig:rediagonalization}) where the lowest mode accounts for the 83\% of the total spectral density.
-The next mode accounts for only XX\% and the rest are practically noise.
-The intensity of these two modes is shown in Fig.~\ref{fig:final modes}.
-The first mode has dimensions of XXxXX nm$^2$ and the spectral density, that is practically given by the summ of the intensity of these two modes, is XXxXX nm$^2$.
-
-This analysis confirms the quantitative predictions discussed in the previous sections, but give a quantitative figure of the coherence of the beam, a  coherent fraction of CF=83\% that measures the "coherence purity".
+This imply that the Eq.~\ref{eq:mode normalization} is not longer verified, and for the propagated modes $M_{mm}\le$1 and $M_{mn}\ne$0 for $m\ne n$.
+Therefore, the coherent modes of the source after propagation and cropping by the beamline are no longer coherent modes of the resulting beam.
+Fig.~\ref{fig:mode transmission} shows that the most important transfer ratio ($M_{mm}$)transfer for the lowest mode is about $M_{00}$=25\%, then $M_{22}$=10\% and $M_{44}$=6\%.
+The other modes shown transfer less than 5\% transfer, and modes higher than $m$=10 practically do not contribute to the final spectral density.
+Note that, as discussed before, the transmission of mode $m$=2 is higher than mode $m$=1, because this one does not contain a central lobe (Fig.~\ref{fig:spectraldensity}c).
 
@@
 
+\begin{figure}\label{fig:mode transmission}
+    \centering
+        \includegraphics[width=\textwidth]{GRAPHICS/propagated_modes.eps}
+    \caption{Transmission $M_{mm}$ (see Eq.~\ref{eq:mode normalization}) of the different modes of the EBS source  due to the effect of propagation and cropping by the beamline optics.}
+\end{figure}
+
+This analysis of the mode transmission is mode dramatic if one looks to the total intensity in the mode, which is given by its eigenvalue.
+Neglecting modes higher than 10, the lowest coherent mode ($m$=0) carries the 49.5\% of the beam intensity, 18.4\% for the mode $m$=2, and less than 5\% for the others.
+
+% 0    8.283396938003417809e-01
+% 1    1.334192926516760969e-01
+% 2    1.726189402927864533e-02
+% 3    7.157008449213129661e-03
+% 4    6.639480916324663404e-03
+% 5    3.894724551090737296e-03
+% 6    2.436986236753659139e-03
+% 7    4.732687905858104635e-04
+% 8    2.758652782324634112e-04
+% 9    7.761681183345831123e-05
+% 10   1.934547528193778418e-05
+
 \begin{figure}\label{fig:rediagonalization}
     \centering
-        \includegraphics[width=\textwidth]{GRAPHICS/rediagonalization.png}
+        \includegraphics[width=\textwidth]{GRAPHICS/rediagonalized_modes.eps}
+    \caption{Occupation of the coherent modes of the final focused beam.}
+\end{figure}
 
-    \caption{Occupation of the coherent modes in the final focused beam.}
-\end{figure}
+As commented before, the coherent modes at the source, propagated and cropped by the beamline, are not longer orthonormal, therefore they cannot be called coherent modes of the resulting beam.
+At this point, the new coherent modes can be calculated applying a new coherent mode decomposition on the cross spectral density of the transmitted beam, that is known from source coherent mode expansion.
+Making this numerical calculation with COMSYL, one obtains a mode occupation spectrum (see Fig.~\ref{fig:rediagonalization}) where the lowest mode accounts for the 83\% of the total spectral density.
+The next mode accounts for only 13\% and the rest are practically negligeable.
+The intensity of these two coherent modes is shown in Fig.~\ref{fig:final modes}.
+The first mode has FWHM dimensions of 7.3$\times$6.7 nm$^2$.
+The spectral density, that is practically given by the contribution of the first two modes has a FWHM equal to the $m$=0 mode because, in horizontal the $m$=1 mode does contribute only to the tails and in vertical is identical to mode $m$=0.
+
+This analysis confirms the quantitative predictions discussed in the previous sections, but give a quantitative figure of the coherence of the beam, a  coherent fraction of CF=83\% that measures the "coherence purity".
 
@@
     \centering
-        \includegraphics[width=0.45\textwidth]{GRAPHICS/final_mode0.png}
-        \includegraphics[width=0.45\textwidth]{GRAPHICS/final_mode1.png}
-    \caption{Coherent modes of the focused beam}
+        \includegraphics[width=0.95\textwidth]{GRAPHICS/final_mode0.png}
+        \includegraphics[width=0.95\textwidth]{GRAPHICS/final_mode1.png}
+    \caption{Coherent modes of the focused beam.
+Top: lower mode ($m$=0), with 83\% occupation and FWHM of 7.3$\times$6.7 nm$^2$.
+Botton: mode $m$=1, with 13\% occupation and FWHM of 19.0$\times$6.7 nm$^2$}
 \end{figure}
@@
 \begin{itemize}
+ \item introduce slope errors in WOFRY propagation
+ \item propagate modes with SRW
  \item create some plots with cuts of W
@@
 \section{Discussion} 
+\label{discussion}
 
-The first step in the wave optics simulation is to assume a full coherent source and simulate the propagation of such wavefront through the beamline
+After the analysis of how the ID16A optics modify the characteristics of the x-ray beam done in the previous section, when one looks back to the numbers obtained by simple analytical canculations, one can verify that these analytical previsions were a quite good approximation of the values then obtained using more sophisticated methods.
+The size of the focused beam and the transmission calculated using ray tracing are in good agreement.
+
+When comparing beam sizes, one usually assume that the intensity distributions are Gaussian and therefore well characterised by the FWHM value.
+From Fig.~\ref{fig:ray tracing} one can see that the horizontal intensity distribution is not Gaussian for the ESRF, as a result of cropping the beam with the VSS opened to 50$\mu$m.
+Although the electron beam statistics is in a very good approximations Gaussian, the undulator emission is not.
+In our ray tracing simulations we always suppose that we work at the resonance, then we applied the Gaussian approximation used in Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas} to compute the geometrical characteristics of the source.
+This is certainly acceptable at the resonance but will fail out of resonance.
+Indeed, these lack of Gaussianity is relevant in the calculations of the coherent fraction.
+We approximate the coherence fraction by Eq.~\ref{eq:coherent fraction}, which for zero emittance ($\sigma_{x,y}=\sigma_{x',y'}$=0) becomes CF=1, as expected because a zero emittance electron beam, that is a single electron or a filament beam affected by an magnetic field emmits a fully coherent wavefront.
+However, the photon source for a filament beam would have an emittance of $\sigma_u \sigma_{u'} \approx \lambda / 2 \pi$ (using Eqs.~\ref{eq:photon small sigmas}), which is in contradiction with theory, that says that a pure coherent state has emittance of $\lambda / 4 \pi$, as happens for a Gaussian photon beam.
+This is justified considering that Eqs.~\ref{eq:photon small sigmas} are approximations obtained by fitting with Gaussians the theoretical non-Gaussian distributions (see Ref.~\cite{elleaume}).
+The correct definition of CF as the occupation of the lowest coherent mode does not present these incongruences.
+However, its calculation requires to perform the coherent mode decomposition, as done in Section~\ref{comsyl} which is very costly.
+It is therefore the case to find the compromise between accuracy and cost, or where in the hierarchical series of methodologies proposed one should stop.
+
+The combination of ray tracing and wave optics methods used indicate that the focalized beam size due to  demagnification with mirrors (giving for EBS 10x5 nm2 analytically, 8.4x4.8 nm2 with ray tracing) has to be corrected with diffraction effect (8.4x7.8 hybrid, 12.7x9.2 SRW) and then add the slope error effect.
+The slope errors are very small, necessary not to blow up the size and kit it in the specifications range (10.8x8.2 hybrid 13.6x9.9 SRW).
+It is important to treat the the slope errors correctly.
+Slope errors can be included in the ray tracing using the data in Appendix A and applying specular reflection on the optical surface with errors.
+This is the standard mode of including slope errors in SHADOW, using measured profiles or synthetic ones.
+It works well for the case of incoherent beams, but in our case, because of the high coherence of the beam the ray tracing method (specular reflection) overestimates the blowing of the focal spot conducting to wrong results.
+Therefore the slope errors of the KB system have been considered using a ``wave optics approach'', as explained in Ref.~\cite{hybrid}.
+Here, the mirror heigh errors $z(x)$ are projected to the plane perpendicular to the propagation ($z(x) \sin \theta$, with $\theta$ the grazing angle) and originate a phase shift $\psi$ in the wavefront phase proportional to the optical path $\psi = - k z(x) \sin \theta $.
+This is the algorithm used in the simulations in Sections~\ref{level1}, \ref{srw_se} and \ref{srw_me}.
+
+% The first step in the wave optics simulation is to assume a full coherent source and simulate the propagation of such wavefront through the beamline.
+
 The wavefront simulation of the beamline include propagation of the beam from element to element (drift spaces) using different possible propagators.
-Usually a propagator based on the Freskel-Kirchhoff integral is used, as it is valid for both near and far field.
+A propagator based on the Freskel-Kirchhoff integral is used.
 The applications of this (and other) operators imply the calculation of an integral (or sum) at each pixel of the image plane.
@@
 This is a time consuming work for performing it with the full 2D simulation.
-For this reason, a wavefront calculation may be started by a simplified system, using simplified models for the source and focusing elements.
-Moreover, because the horizontal and vertical coupling is usually small in synchrotron beamlines, it is convinient to separate the simulation in two, one for the horizontal plane and one for the vertical one
+For this reason, we found very useful to start a wavefront calculation in 1D, separating the horizontal and vertical simulations, and using a simplified system with ideal sources and ideal focusing elements.
+This is ofted a good choice because the horizontal and vertical coupling is usually small in synchrotron beamlines.
+This 1D simulations help in refining the choice of propagator and its parameters (resampling, zoom factors) that will be valid if a, more costly, full 2D simulation is envisaged.
 
-Simulations took significantly longer to be completed for the multi electron case: few hours on a 56-core computer cluster.
+Many of the simulations presented here (ray tracing and coherent optics in Sections XXXXX) can be done interactively in a normal laptop.
+However, when one one to obtain information on the partial coherence, very long calculations have to be done.
+For example, Monte Carlo calculations in Section~\ref{srw_me} took significantly longer to be completed for the multi electron case: a few hours on a 56-core computer cluster.
+For performing the coherent mode decomposition the calculation time depends very much on the dimensioning on the problem.
+The EBS case is reasonable: we sampled the wavefronts using 1007x335 pixels and calculated 1103 modes that represent 98\% of the spectral density.
+This represents 5 MB per wavefront and a matrix of 1695 GB to diagonalize.
+It took less that four hours in the a cluster of 28 cores.
+The calculation is more expensive if one wants to decompose a source that is pretty incoherent like our ESRF-High-$\beta$.
+This is a limit case, as the wavefields have to be vey large XXxXX to accomodate the large espatial extension of the source.
+The calculation presented here took XX days in a large XX cluster.
+To give an idea of the real cost, a similar work send to the AWS cluster costed about 2000 USD.
+This was done as a test case, to evaluate the limits of the calculations, but it is certainly not recommended to do similar things, not only for the cost but also for the lack of interest to treat with partial coherence methods a source that is quite incoherent.
+In Appendix B we present for the ESRF High$\beta$ similar results than those in Section~\ref{comsyl}.
 
-It is interesting to remark that although Monte Carlo multielectron method can in principle calculate Eq.~\ref{CSD} the storage at every point is unmanagleable, so usually only some 'cuts' of the CSD are stored, such as $W(x_1,0,x_2,0;z,\omega)$ (horizontal) or $W(0,y_1,0,y_2;z,\omega)$ (vertical) \todo{move this to the discussion?}.
+There are some effects that are typically studied when designing and simulating a beamline that have been ignores here.
+For example the study of the power emitted by the source and its heat load effect in the first (or firsts) slits and optical elements (here the ML monochromator).
+Also reflectivity of the elements are not considered.
+Thir effect can also be fully simulated, but we preferred here to integrate them in the simulations for the sake of clarity.
+Again an analytical estimation would give a reflectivity about 0.7 for multilayers and 0.9 for mirrors, therefore a reduction factor of (0.7)**3 ML for the beamline flux due to reflectivity of mirror coatings.
 
-The fist coherence mode is always centered on-axis (see, e.g, Fig.~\ref{fig:spectraldensity}b) and will always be centered when propagated until a position where an optical element (an aperture or mirror with finite dimensions) will crop it removing part of its intensity but letting the central part to go throught.
+In conclusion, we have shown how to obtain the relevant parameters necessary to quantify flux, transmission, fucused beam size, and coherent fraction for a modern beamline with already exceptional performances in the old ESRF DBA lattices, and advanced the improvements to be expected when the EBS will start in 2020.
+In this discussion we obtained these parameters using a hierarchical flowchart, with increasing complexity.
+The phase of analytical estimation of the main parameters is essential, not only to start with some values but also because these values help to perform sanity check of the computer simulations.
+A ray tracing simulation is inexpensive and always recommended.
+The comparison of the CF beamline transmittivity would give an idea of how coherent is the beam, or how many modes will be present in your final beam.
+For high coherence (a few modes), like for ID16A, the wave optics calculations are necessary.
+If the final beam has a very high coherence, like in our case, coherent optics (1D simplifications, and single electron calculations) may give good results.
+We may have stopped there, as in our case coherence is very high.
+The COMSYL calculations presented here introduce a new way of quantitatively assess the quality of the partial coherent beam via the calculation of the coherent fraction.
+
+% It is interesting to remark that although Monte Carlo multielectron method can in principle calculate Eq.~\ref{CSD} the storage at every point is unmanagleable, so usually only some 'cuts' of the CSD are stored, such as $W(x_1,0,x_2,0;z,\omega)$ (horizontal) or $W(0,y_1,0,y_2;z,\omega)$ (vertical) \todo{move this to the discussion?}.
+% 
+% 
+% The fist coherence mode is always centered on-axis (see, e.g, Fig.~\ref{fig:spectraldensity}b) and will always be centered when propagated until a position where an optical element (an aperture or mirror with finite dimensions) will crop it removing part of its intensity but letting the central part to go throught.
 The subsequent modes will be more absorbed as they extend more.
@@
 
+\ack{\todo{someone to thank?}}
+
+\appendix
+
+\section{Metrology of the KB mirrors}
+
+\section{COMSYL results for the ESRF-H$\beta$}
+
 \bibliographystyle{iucr}

### 2019-03-29 paper-hierarchical.tex
Updates from Overleaf

@@
 
-After the analysis of how the ID16A optics modify the characteristics of the x-ray beam done in the previous section, when one looks back to the numbers obtained by simple analytical canculations, one can verify that these analytical previsions were a quite good approximation of the values then obtained using more sophisticated methods.
-The size of the focused beam and the transmission calculated using ray tracing are in good agreement.
-
-When comparing beam sizes, one usually assume that the intensity distributions are Gaussian and therefore well characterised by the FWHM value.
+After the complete analysis done in the previous section on how the ID16A optics modify the characteristics of the x-ray beam, one can verify that the analytical previsions were a good estimator of the values then obtained using more sophisticated methods.
+The size of the focused beam (e.g, 10x5 for EBS) and the transmission (3\%) calculated analytically are in good agreement with geometric ray tracing (8.4x4.8, and 2.28\%).
+When characterising beam sizes, one usually assume that the intensity distributions are Gaussian and therefore they are well characterised by the FWHM value.
 From Fig.~\ref{fig:ray tracing} one can see that the horizontal intensity distribution is not Gaussian for the ESRF, as a result of cropping the beam with the VSS opened to 50$\mu$m.
@@
 Indeed, these lack of Gaussianity is relevant in the calculations of the coherent fraction.
-We approximate the coherence fraction by Eq.~\ref{eq:coherent fraction}, which for zero emittance ($\sigma_{x,y}=\sigma_{x',y'}$=0) becomes CF=1, as expected because a zero emittance electron beam, that is a single electron or a filament beam affected by an magnetic field emmits a fully coherent wavefront.
+We approximate the coherence fraction by Eq.~\ref{eq:coherent fraction}, which for zero emittance ($\sigma_{x,y}=\sigma_{x',y'}$=0) becomes CF=1, as expected because a zero emittance electron beam (a single electron or a filament beam) that travel in an magnetic field emits a fully coherent wavefront.
 However, the photon source for a filament beam would have an emittance of $\sigma_u \sigma_{u'} \approx \lambda / 2 \pi$ (using Eqs.~\ref{eq:photon small sigmas}), which is in contradiction with theory, that says that a pure coherent state has emittance of $\lambda / 4 \pi$, as happens for a Gaussian photon beam.
@@
 Therefore the slope errors of the KB system have been considered using a ``wave optics approach'', as explained in Ref.~\cite{hybrid}.
-Here, the mirror heigh errors $z(x)$ are projected to the plane perpendicular to the propagation ($z(x) \sin \theta$, with $\theta$ the grazing angle) and originate a phase shift $\psi$ in the wavefront phase proportional to the optical path $\psi = - k z(x) \sin \theta $.
+Here, the mirror height errors $z(x)$ are projected to the plane perpendicular to the propagation ($z(x) \sin \theta$, with $\theta$ the grazing angle).
+They introduce a phase shift $\psi$ in the wavefront phase proportional to the optical path $\psi = - k z(x) \sin \theta $.
 This is the algorithm used in the simulations in Sections~\ref{level1}, \ref{srw_se} and \ref{srw_me}.
@@
 The applications of this (and other) operators imply the calculation of an integral (or sum) at each pixel of the image plane.
-This is costly from the computer point of view, so Fourier Transforms (Fourier Optics) are used to reduce complexity and calculation time.
-This is practically the only choice for simulating 2D images.
-The use of Fourier Transforms adds another problem: the result is very dependent on the sampling of the wavefront (i.e., pixel size, number of points, dimension of the window where the wavefront is sample).
-This means the usually one has to test many configurations before finding one that works reaonably well (even though one can never guarantee the complete accuracy of the sampling used).
+This is costly from the computer point of view, so Fourier Transforms (Fourier Optics) and its Fast Fourier Transform implementation are used to reduce amount of operations and the calculation time.
+This is practically the only choice for propagating 2D wavefronts.
+The use of Fourier Transforms adds another problem: the result is very dependent on the sampling of the wavefront (i.e., pixel size, number of points, dimension of the window where the wavefront is defined).
+This means the usually one has to test many configurations before finding one that works reasonably well (even though one can never guarantee the complete accuracy of the sampling used).
 A full wavefront optics simulation is an iterative process, where the user has to refine the sampling parameters element by element to be sure the wavefront at each position is sampled correctly.
-This is a time consuming work for performing it with the full 2D simulation.
-For this reason, we found very useful to start a wavefront calculation in 1D, separating the horizontal and vertical simulations, and using a simplified system with ideal sources and ideal focusing elements.
-This is ofted a good choice because the horizontal and vertical coupling is usually small in synchrotron beamlines.
-This 1D simulations help in refining the choice of propagator and its parameters (resampling, zoom factors) that will be valid if a, more costly, full 2D simulation is envisaged.
-
-Many of the simulations presented here (ray tracing and coherent optics in Sections XXXXX) can be done interactively in a normal laptop.
-However, when one one to obtain information on the partial coherence, very long calculations have to be done.
-For example, Monte Carlo calculations in Section~\ref{srw_me} took significantly longer to be completed for the multi electron case: a few hours on a 56-core computer cluster.
-For performing the coherent mode decomposition the calculation time depends very much on the dimensioning on the problem.
-The EBS case is reasonable: we sampled the wavefronts using 1007x335 pixels and calculated 1103 modes that represent 98\% of the spectral density.
+This is a time consuming work in particular for 2D simulation.
+We found very useful to start a wavefront simulation in 1D, separating the horizontal and vertical simulations, using a simplified system with ideal sources and ideal focusing elements.
+The use of 1D propagation is justified because in in synchrotron beamlines the horizontal and vertical coupling is usually small.
+This 1D simulations help in refining the choice of propagator and its parameters (resampling, zoom factors) that can be reused in a full 2D simulations.
+
+Regarding computational effort, simulations concerning ray tracing (Section~\ref{level1}) and coherent optics (Sections~\ref{wofry} and \ref{srw_se}) can be done interactively in a normal laptop.
+However, for those concerning partial coherence, very long calculations are needed.
+For example, Monte Carlo calculations in Section~\ref{srw_me} for the multi electron case \inred{HOW MANY ELECTRONS?} spent a few hours on a 56-core computer cluster.
+For performing the COMSYL coherent mode decomposition the calculation time depends very much on the dimensions on the problem.
+For EBS the wavefronts were sampled using 1007x335 pixels, and 1103 modes were computed representing 98\% of the spectral density.
 This represents 5 MB per wavefront and a matrix of 1695 GB to diagonalize.
@@
 The calculation is more expensive if one wants to decompose a source that is pretty incoherent like our ESRF-High-$\beta$.
-This is a limit case, as the wavefields have to be vey large XXxXX to accomodate the large espatial extension of the source.
-The calculation presented here took XX days in a large XX cluster.
-To give an idea of the real cost, a similar work send to the AWS cluster costed about 2000 USD.
-This was done as a test case, to evaluate the limits of the calculations, but it is certainly not recommended to do similar things, not only for the cost but also for the lack of interest to treat with partial coherence methods a source that is quite incoherent.
+This is a limit case, it was used wavefields of XxX pixels to accommodate the large spatial extension of the source.
+The calculation took XX days in a large XX cluster.
+%To give an idea of the real cost, a similar work send to the AWS cluster costed about 2000 USD. 
+This was done as a test case, to demonstrate the feasibility of the calculations.
+It is not recommended to do similar calculations that imply a the cost because a source that is quite incoherent was treated with partial coherence methods.
 In Appendix B we present for the ESRF High$\beta$ similar results than those in Section~\ref{comsyl}.
 
-There are some effects that are typically studied when designing and simulating a beamline that have been ignores here.
+Some effects that are typically studied when designing and simulating a beamline have not been discussed here.
 For example the study of the power emitted by the source and its heat load effect in the first (or firsts) slits and optical elements (here the ML monochromator).
-Also reflectivity of the elements are not considered.
-Thir effect can also be fully simulated, but we preferred here to integrate them in the simulations for the sake of clarity.
-Again an analytical estimation would give a reflectivity about 0.7 for multilayers and 0.9 for mirrors, therefore a reduction factor of (0.7)**3 ML for the beamline flux due to reflectivity of mirror coatings.
-
-In conclusion, we have shown how to obtain the relevant parameters necessary to quantify flux, transmission, fucused beam size, and coherent fraction for a modern beamline with already exceptional performances in the old ESRF DBA lattices, and advanced the improvements to be expected when the EBS will start in 2020.
-In this discussion we obtained these parameters using a hierarchical flowchart, with increasing complexity.
-The phase of analytical estimation of the main parameters is essential, not only to start with some values but also because these values help to perform sanity check of the computer simulations.
-A ray tracing simulation is inexpensive and always recommended.
-The comparison of the CF beamline transmittivity would give an idea of how coherent is the beam, or how many modes will be present in your final beam.
-For high coherence (a few modes), like for ID16A, the wave optics calculations are necessary.
+Also, the reflectivity of the elements are not considered.
+The simulation of this was not included in the simulations, because it can be estimated analytically considering an average reflectivity about \inred{0.7 for multilayers and 0.9 for mirrors, therefore a reduction factor of (0.7)**3 ML for the beamline flux due to reflectivity of mirror coatings}.
+
+\todo{discuss coherence length}
+
+\section{Summary and conclusions}
+
+We have obtained quantitative values of flux, transmission, size of the focused beam, and coherent fraction for ID16A, a modern beamline with already exceptional performances in terms of nanofucusing and coherence.
+We predicted advances the improvements in these parameters when the EBS will start in 2020.
+We obtained these parameters using a hierarchical flowchart with, applying methods with increasing complexity and requiring higher computational effort.
+
+The analytical estimation of the main parameters is mandatory first step in the beamline conception and analysis.
+We applied concepts based on optical magnification and numerical aperture acceptance and calculated approximately the coherent fraction at the source.
+These values help to perform sanity check of the computer simulations.
+A ray tracing simulation is inexpensive and provides important information.
+The comparison of the CF with the beamline transmitivity would give an idea of the beam coherence, reflected in the number modes present in the final beam.
+For high coherence, like for ID16A, the wave optics calculations are necessary.
 If the final beam has a very high coherence, like in our case, coherent optics (1D simplifications, and single electron calculations) may give good results.
-We may have stopped there, as in our case coherence is very high.
-The COMSYL calculations presented here introduce a new way of quantitatively assess the quality of the partial coherent beam via the calculation of the coherent fraction.
+Full calculations based on coherent mode decomposition Were introduced here.
+It is a new way of quantitatively assess the quality of the partial coherent beam via the calculation of the coherent modes of the source, propagation of the modes and recalculating the coherent mode expansion on the final beam from which one obtains the CF, a parameter that valutated the coherence quality.
 
@@
 
- %-------------------------------------------------------------------------
- % Manuel & Rafael
- %-------------------------------------------------------------------------
-\section{Summary}
-
-Bla bla
-
 \ack{\todo{someone to thank?}}

### 2019-04-03 paper-hierarchical.tex
Updates from Overleaf

@@
 The calculation is more expensive if one wants to decompose a source that is pretty incoherent like our ESRF-High-$\beta$.
-This is a limit case, it was used wavefields of XxX pixels to accommodate the large spatial extension of the source.
-The calculation took XX days in a large XX cluster.
+This is a limit case, it was used wavefields of 10075x201 pixels to accommodate the large spatial extension of the source.
+We obtained 4016 modes totalizing only XX\% of the spectral density.
+The calculation took more than 4 days in a 84 cores cluster.
 %To give an idea of the real cost, a similar work send to the AWS cluster costed about 2000 USD. 
@@
 We predicted advances the improvements in these parameters when the EBS will start in 2020.
-We obtained these parameters using a hierarchical flowchart with, applying methods with increasing complexity and requiring higher computational effort.
+We obtained these parameters using a hierarchical flowchart applying methods with increasing complexity and requiring higher computational effort.
 
@@
 For high coherence, like for ID16A, the wave optics calculations are necessary.
-If the final beam has a very high coherence, like in our case, coherent optics (1D simplifications, and single electron calculations) may give good results.
+If the final beam has a very high coherence coherent optics (1D simplifications, and single electron calculations) may give good results.
 Full calculations based on coherent mode decomposition Were introduced here.
-It is a new way of quantitatively assess the quality of the partial coherent beam via the calculation of the coherent modes of the source, propagation of the modes and recalculating the coherent mode expansion on the final beam from which one obtains the CF, a parameter that valutated the coherence quality.
+It is a new way of quantitatively assess the quality of the partial coherent beam via the calculation of the coherent modes of the source, propagation of the modes and recalculation the coherent mode expansion on the final beam.
+The CF or occupations of the first mode is the parameter used to measure the coherence quality.
 
@@
 \section{Metrology of the KB mirrors}
+The residual height error of the multilayer monochromator and the KB mirrors, after removing the main ellitical shape, are shown in Fig.~\ref{fog:metrology}.
+These profiles are used along the text in ray tracing and wave optics simulations.
+
+\begin{figure}\label{fig:metrology}
+\includegraphics[width=0.8\textwidth]{GRAPHICS/ID16A_ML_W-B4C_120Strp.png}
+\includegraphics[width=0.8\textwidth]{GRAPHICS/ID16A_KB1_VF.png}
+\includegraphics[width=0.8\textwidth]{GRAPHICS/ID16A_KB1_HF.png}
+\caption{Error height profiles used for the simulations for a) the ML monochromator, b) the vertical focusing mirror os the KB, and c) T=the horozontal focusing mirror of the KB.}
+\end{figure}
 

### 2019-04-03 paper-hierarchical.tex
version sent to coauthors

@@
 Several methodologies with increasing complexity are applied to discuss the relevant parameters that quantify the beamline performances.
-Parameters such as flux, beam size at the image and coherence properties are obtained from simple analytical calculations to sophisticated computer simulations using ray tracing and waveoptics techniques.
-A last generation x-ray nanofocusing beamline for coherent  applications (ID16A at the ESRF) has been chosen  to study in detail the issues
-related to highly demagnifying sychrotron sources and exploiting the beam coherence.
+Parameters such as flux, dimensions and intensity distribution of the focused beam and coherence properties are obtained from simple analytical calculations to sophisticated computer simulations using ray tracing and waveoptics techniques.
+A last generation x-ray nanofocusing beamline for coherent  applications (ID16A at the ESRF) has been chosen  to study in detail the issues related to highly demagnifying sychrotron sources and exploiting the beam coherence.
 The performance of the beamline is studied for two storages rings: the old ESRF (emittance 4000 pm) and the new ESRF-EBS (emittance 150 pm).
-In addition to traditional results in term of flux, intensity distribution and focal size, an innovative study on the partial coherence properties based on propagation of coherent modes is presented.
+In addition to traditional results in term of flux and beam sizes, an innovative study on the partial coherence properties based on propagation of coherent modes is presented.
 The different algorithms and methodologies are implemented in the software suite OASYS \cite{codeOASYS}.
@@
 Many storage ring based x-ray synchrotron facilities are building or planning upgrades to increase brilliance and coherent flux by one to three orders of magnitude.
-The first upgrade of a large facility will be the EBS (Extremely Brilliant Source) \cite{orangebook} at the European Synchrotron Radiation Facility (ESRF), aiming to build a storage ring of 150 pm emittance to significantly boost the associated x-ray coherence.
-Applications exploiting beam coherence such as such as x-ray photon correlation spectroscopy, coherent diffraction imaging, propagation-based phase contrast imaging and ptychography will strongly benefit from this update.
+The first upgrade of a large facility will be the EBS (Extremely Brilliant Source) \cite{orangebook} at the European Synchrotron Radiation Facility (ESRF).
+The new storage ring of 150 pm emittance that will be commissioned in 2020 will significantly boost the x-ray coherence of the synchrotron beams.
+Applications exploiting beam coherence such as x-ray photon correlation spectroscopy, coherent diffraction imaging, propagation-based phase contrast imaging and ptychography will strongly benefit from this update.
 
 Accurate calculation and quantitative evaluation of the parameters related to x-ray optics is of paramount importance for designing, building and exploiting the new beamlines.
-Every modern beamline follows a procedure of conceptual design and then technical design where detailed simulations are essential.
-The calculations allow the designer to test the design parameters in a virtual computer environment where pros and cons can be studied quantitatively.
+Every modern beamline follows a procedure of conceptual design plus technical design where detailed simulations are essential.
+These calculations permit to test the design parameters in a virtual computer environment where advantages and limitations can be studied quantitatively.
 The selection to the best performant optics configuration requires the detailed simulation of the optics imperfections, that in many cases are the limiting factor of the beamline optical configuration.
@@
 It shows characteristics increasingly searched in modern beamlines: extremely high demagnification (sub-micrometer beam) and high coherence.
-It combines coherent imaging techniques and X-ray fluorescence microscopy to perform quantitative 3D characterization of the morphology and the elemental composition of specimens in their native state \todo{complete ID16A description}.
+It combines coherent imaging techniques and X-ray fluorescence microscopy to perform quantitative 3D characterization of the morphology and the elemental composition of specimens in their native state.
 
 X-ray beamlines are very particular optical systems.
-Some implications of exploiting low wavelengths such as the necessity of using grazing incidence reflectors and the use of perfect crystals as typical hard x-ray monochromators make useless, or at least incomplete, the application of common commercial software packages for the study of synchrotron beamlines.
+Some implications of exploiting the low wavelengths such as the necessity of using grazing incidence reflectors and the use of perfect crystals as typical hard x-ray monochromators make useless, or at least incomplete, the application of common commercial software packages for the study of synchrotron beamlines.
 For this reason many optics tools have been developed by the synchrotron community.
 Some of them have a long history like SHADOW (more than 40 years old) or SRW (almost 30), for mentioning only codes used in this work.
-These historical tools together with new tools designed ad-hoc packages have been agglutinated in a user friendly software OASYS.
+These historical tools together with other new applications have been agglutinated in a user friendly software OASYS.
 Many of the calculations presented here have been performed directly in OASYS or with script created with the help of OASYS.
 
-The paper is structures as follows: we first introduce in Section~\ref{Beamline description} the main parameters of the beamline and make the selection of the particular choice of the parameters used in the calculations.
-Section~\label{Simulations} is the central part of the paper, and describe the different methods used to analyse the beamline with increasing complexity, setting a hierarchy of methodologies that can be exported to the simulations of any beamline.
+The paper is structured as follows: we first introduce in Section~\ref{Beamline description} the main parameters of the beamline and make the selection of the particular choice of the parameters used in the calculations.
+Section~\ref{Simulations} is the central part of the paper, and describe the different methods used to analyse the beamline with increasing complexity, setting a hierarchy of methodologies that can be exported to the simulations of any beamline.
 In Section \ref{level0} we calculate with simple methods approximated values of focal size, flux and coherent fraction.
 Then accuracy is improved using geometrical methods in Section \ref{level1} that give reasonable results as far as some corrections are considered because of the  mentioned characteristics of high demagnification and high coherence of the studied beamline.
-The wave optics section~]ref{level2} also include different models whose complexity increases when the results expected are more and mode demanding.
+The wave optics section~\ref{level2} also include different models whose complexity increases when the results expected are more and mode demanding.
 The simulations methods are presented in a hierarchical approach:  they provide more and better results but the price to pay is to deal with more complex and computationally expensive tools.
@@
  %-------------------------------------------------------------------------
- % Manuel & Rafael
+ %
  %-------------------------------------------------------------------------
@@
 \input{tableUndulator.txt}
-%      & IVU26 & 26.00           & 2.50    & 1      & 0.616 (n=1)    & 1.625 (n=3)    & 0.616 (n=3)    \\
-% ESRF & U18.3 & 18.30           & 1.40    & 2      & 1.715 (n=1)    & 0.445 (n=1)    & 1.715 (n=3)    \\
-%      & U22.4 & 22.40           & 1.40    & 1      & 0.873 (n=1)    & 1.840 (n=1)    & 0.873 (n=1)    \\ \hline
-%      & IVU26 & 26.00           & 2.50    & 1      & 0.590 (n=1)    & 1.606 (n=1)    & 0.590 (n=3)    \\
-% EBS  & U18.3 & 18.30           & 1.40    & 2      & 1.156 (n=1)    & 0.411 (n=1)    & 1.156 (n=1)    \\
-%      & U22.4 & 22.40           & 1.40    & 1      & 0.852 (n=1)    & 1.821 (n=3)    & 0.852 (n=1)    \\ \hline
 \end{tabular}%
@@
     Horizontal $M^{-1}$:   & -      & 2.42       & -     & -      & 2899   & -           \\
-    Vertical $M^{-1}$:     & -      & -          & -     & \inblue{1849}   & -      & -           \\ 
+    Vertical $M^{-1}$:     & -      & -          & -     & 1849   & -      & -           \\ 
     \end{tabular}%
@@
 
-Simulations were done using the two different storage ring lattices mentioned before: the (present) ESRF DBA where the undulators are in a high-$\beta$ straight section (larger horizontal emittance) and the new EBS (lower horizontal emittance) straight sections.
+Simulations were done using the two different storage ring lattices mentioned before: the (already dismantled) ESRF DBA where the undulators are in a high-$\beta$ straight section (larger horizontal emittance) and the new EBS (lower horizontal emittance) straight sections.
 The parameters for the current and the new lattice are shown in Tables \ref{tab:TableSources} and \ref{tab:eBeam}.
@@
 Values taken at the symmetry point of the straight section, where the insertion devices are placed.
-Values fir ESRF are taken from \cite{ESRF2014}.
+Values for ESRF are taken from \cite{ESRF2014}.
 For EBS, two (very close) design values are used, from Ref.~\cite{ESRF2014} labelled EBS-S28A, and those coming from further lattice refinement, labelled EBS-S28D.
-\todo{Remove ESRF-AT??
-Check again} }
+\todo{Remove ESRF-AT and EBS-S28A} }
         \resizebox{\textwidth}{!}{%
@@
 \section{Simulations}
+\label{Simulations}
 
@@
 The source size is usually approximated supposing a Gaussian distribution of the electron sizes (or divergence) (Table \ref{tab:eBeam}) and undulator natural source size or divergence.
-Although the former is true, the latter is not.
-Let $\sigma_{e(h,v)}$ and $\sigma_{e(h,v)}'$ be the electron beam size and divergence at a centre of the symmetry of the straight section.
+Fo some cases, as the undulator emission, the Gaussian approximation may not be accurate and have some implications that will be discussed later.
+Let $\sigma_{e(h,v)}$ and $\sigma_{e(h,v)}'$ be the electron beam size and divergence standard deviation of the electron distributions at a centre of the symmetry of the straight section in a plane perpendicular to the main propagation direction.
 The subscripts $h$ and $v$ stand for horizontal and vertical, respectively.
+
 The undulator natural source size and divergence are given by \cite{elleaume}:
@@
 \end{align}
-where $\lambda$ is the emitted wavelength and $L=N\lambda_u$ is the undulator length, with $\lambda_u$ as the magnetic period of the undulator and $N$, the number of magnetic periods.
+where $\lambda$ is the emitted wavelength and $L=N\Lambda_u$ is the undulator length, with $\Lambda_u$ as the magnetic period of the undulator and $N$, the number of magnetic periods.
 The electron beam usually presents different horizontal and vertical beam emittances, but the undulator natural source size or divergence presents radial symmetry at the central cone (excluding polarization).
-The photon source size and divergence are given by the convolution of the electron sizes and the undulator natural sizes.
+The photon source size and divergence are approximately given by the convolution of the electron sizes and the undulator natural sizes.
 If we consider Gaussian distributions, the photon source is
@@
 
-% $\Sigma$ is given by the convolution of the electron beam size and the undulator natural source size or equivalently, single electron source size \cite{elleaume}: a storage ring with low emittance will produce small photon source size.
-It is also convenient to define two other relationships: the first one relates the magnetic field $B$ in an undulator to a dimensionless parameter $K$, the deflection parameter: $K = 93.4\lambda_u[\text{m}]\text{B}[\text{T}]$.
-The second relationship shows the wavelengths at which the emitted spectral intensity os peaked, or resonance.
-It is a function of $K$; the harmonic number n; the Lorentz factor $\gamma$ and the observation angle $\phi$:
-% \begin{align}
-%     \lambda=\frac{\lambda_u}{2\text{n}\gamma^2}\Big(1+\frac{K^2}{2}+\gamma^2\phi^2\Big).
-% \end{align}
-% The first term is the relativistic contracted wavelength, the second term is the magnetic tuning and the last term is the off-axis wavelength increase \cite{elleaume}.
-
 The beam sizes at different positions at the beamline were estimated in a different way depending if the beam is focused or not.
 If the beam is focused, its dimensions in a first approximation are the photon source size multiplied by the corresponding magnification factor.
-For predicting the photon beam in a non-focal position the beam divergence is used.
+For predicting the photon beam sizes in a position out of the focus the beam divergence is used.
 In an ideal {\'{e}}tendue preserving optical system, where the conservation of the Smith-Helmholz invariante applies, the resulting beam divergence transforms inversely as the beam size: it can be calculated as the product between the beam divergence at the source with the demagnification factor $M^{-1}$.
@@
 8 \inred{$\approx$ 403.8/50=8.076}.
-Therefore the horizontal size at the sample will be affected by the same factor obtaining roughly 17$\times$5 nm$^2$
-
-In general, for grazing incident optics with large demagnifications ($M^{-1} \gg 100$) imaging extended sources, the effect of optical aberrations is very important even using a perfect optical surface that focus point-to-point.
+Therefore, for ESRF, the horizontal size at the sample will be affected by the same factor obtaining roughly 17$\times$5 nm$^2$
+
+In general, for grazing incident optics imaging extended sources with large demagnifications ($M^{-1} \gg 100$), the effect of optical aberrations is very important even using a perfect optical surface that focus point-to-point.
 As a consequence, the results calculated using only the demagnification factor are usually too optimistic.
@@
 Their effect will studied in Sections~\ref{level1} and \ref{level2}.
-The smearing of the focal size because of diffractive effects by the mirror clipping will also be discussed.
+%The smearing of the focal size because of diffractive effects by the mirror clipping will also be discussed.
 
@@
 
-The next step would be to estimate the number of photons at the sample.
+Next, we want to estimate the number of photons at the sample.
 For that, we need the spectrum at the source.
@@
 Perhaps the most performant and well maintained codes spread in the synchrotron community are SPECTRA \cite{codeSPECTRA} and SRW \cite{codeSRW}.
-The XOPPY package in the OASYS suite implements three codes: US\cite{codeUS}, URGENT: \cite{codeURGENT} and SRW \cite{codeSRW}.
+The XOPPY package in the OASYS suite implements three codes: US \cite{codeUS}, URGENT \cite{codeURGENT} and SRW.
 Figure~\ref{fig:FluxU18} shows the spectra calculated over an aperture if 1 $\times$ 1 mm$^2$ placed at $d$=27.1 m.
@@
 The peak intensity is about  $2.9\cdot10^{14}$ photons/s/0.1{\%}bw for EBS and 1.9 10$^{14}$ for ESRF (see \ref{fig:FluxU18}).
-One can appreciate how the emittance plays an important role in both peak width and specially in peak amplitude.
-The number of photons per second after the multilayer monochromator is 2.9 10$^{14}$ for EBS and 1.9 10$^{14}$ for ESRF, considering that the energy bandwidth transmitted by a multilater monochromator is $\Delta E/E \approx 10^{-2}$ and the bandwidth in \ref{fig:FluxU18} is 10$^{-3}$, and supposing at this point ideal reflectivity (100\%) for the multilayers.
+One can appreciate how the emittance plays an important role in both peak width and specially in peak value.
+The number of photons per second after the multilayer monochromator is 2.9 10$^{15}$ for EBS and 1.9 10$^{15}$ for ESRF, considering that the energy bandwidth transmitted by a multilater monochromator is $\Delta E/E \approx 10^{-2}$ and the bandwidth in Fig.~\ref{fig:FluxU18} is 10$^{-3}$, and supposing at this point ideal reflectivity (100\%) for the multilayers.
 
@@
 
-Next we would like to estimate the geometrical transmitivity of the beamline, therefore we have to consider which optical element crop the beam and remove photons.
+To estimate the geometrical transmitivity of the beamline we have to consider which optical elements crop the beam and remove photons.
 This can be analysed using their angular acceptance or numerical aperture.
@@
 over the beam divergence, resulting 9.5\% (H) and 0.18\% (V)for ESRF and 16.3\% and 0.18\% for EBS.
-The overall transmission is the product of H and V ratios, therefore 0.2\% for ESRF and 2.95\% for EBS, therefore 3.8 $10^{11}$ photons/s for ESRF and 8.5 $10^{12}$ photons/s for EBS.
-\todo{check with references $10^9$ at 33.6 keV. I get: 1.6e+15 * 0.00295 * 0.7**3 = 1.62e+12. How closed are the slits? }
+The overall transmission is the product of H and V ratios, therefore 0.2\% for ESRF and 2.95\% for EBS, therefore 3.8 $10^{12}$ photons/s for ESRF and 8.5 $10^{13}$ photons/s for EBS.
+\todo{check with references $10^9$ at 33.6 keV.
+I get: 1.6e+15 * 0.00295 * 0.7**3 = 1.62e+12.
+How closed are the slits?
+We used the undulator peak value over the whole ML bandwidth, that overestimates the integrated flux.}
 
@@
 An approximated estimation of it can be obtained using Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas}.
+The coherent fraction is the ratio of the phase space for the coherent beam (zero emittance) over the actual beam.
 The CF in its the horizontal and vertical planes are
@@
 A complete cleaning is impossible because of the spatial overlapping of the different coherence modes.
-However, apertures centered at the optical axis would make the work.
+However, apertures centered at the optical axis would clean the beam from these undesired photons.
 If we compare the values transmittivity of the beamline due to geometrical clipping  (0.2\% for ESRF and 2.95\% for EBS) with the the CF values, we remark that they are very close.
-This implies that the beam at the focal position would be highly coherent because the fraction of the photons removed from the beam matches the CF value that quantify the fraction of ``good'' or ``coherent'' photons.
-
-To assess the focused beam size, one can idealize the beamline in two parts.
+This suggests that the beam at the focal position will be highly coherent because the fraction of the photons removed from the beam matches the CF value that quantify the fraction of ``good'' or ``coherent'' photons.
+
+To assess the focused beam size given by a coherent beam, one can idealize the beamline in two parts.
 First, the ML monochromator and VSS that will prepare a coherent plane wave as the source is very far from the sample position.
-Second, the KB has a double role, acting as a focusing element of focal distance $f$=5 cm, but also crops the beam as as it would be done by a square aperture of $D \times D=$ 400$\times$400 $\mu m^2$ (the length of the mirror projected in a plane perpendicular to the beam propagation) \inred{(we obtained 900 in H, that will be reduced by 0.5 as propagated from 10 cm to 5 cm upstrem from the sample, and 390 in V approximated by 400)}.
+Second, the KB has a double role, acting as a focusing element of focal distance $f$=5 cm, but also crops the beam as as it would be done by a square aperture of $D \times D=$ 400$\times$400 $\mu m^2$ (the length of the mirror projected in a plane perpendicular to the beam propagation)
+\todo{REMOVE: we obtained 900 in H, that will be reduced by 0.5 as propagated from 10 cm to 5 cm upstrem from the sample, and 390 in V approximated by 400}.
 This aperture has two effects: i) as discussed before, it removes a lot of off-axis photons increasing the CF to a value close than one, and ii) make a broadening of the focal spot because of coherent diffraction.
@@
 
-In summary, we show in this section how simple analytical calculations that can be done in its major part ``by hand'' help to estimate the geometrical beam size at focal position (17$\times$5 nm$^2$ for ESRF, 10$\times$5 nm$^2$ for the EBS) the flux (3.8 $10^{11}$ photons/s for ESRF and 8.5 $10^{12}$ photons/s for EBS) and also anticipate that the beam is highly coherent producing a diffraction limited focal spot of  7.75$\times$7.75 nm$^2$.
+In summary, we show in this section how simple analytical calculations that can be done in its major part ``by hand'' help to estimate the geometrical beam size at focal position (17$\times$5 nm$^2$ for ESRF, 10$\times$5 nm$^2$ for the EBS) the flux (3.8 $10^{12}$ photons/s for ESRF and 8.5 $10^{13}$ photons/s for EBS) and also anticipate that the beam is highly coherent producing a diffraction limited focal spot of  7.75$\times$7.75 nm$^2$.
 We ignored at this point the effect of real optics containing surface errors, the reflectivity of the elements, and other aspects that will be treated in depth in the next sections.
@@
 %-------------------------------------------------------------------------
- % Manuel
+% 
 %-------------------------------------------------------------------------
@@
 
-% When analyzing an optical system producing an extremely high demagnification of the source one can think in several limiting factors that could affect the focal sizes calculated by hand in Table~\ref{tab:HandCalculations}.
-
-% \inred{REMOVE!!
-% One limiting factor is that the system under consideration is not an {\it imaging system} so an intensity distribution at the source (e.g., a Gaussian distribution) is not necessary imaged into a the same distribution (Gaussian).
-To test that, we implemented the beamline as three ideal lenses, each of them representing an optical element: ML, KBV and KBH.
-Ray tracing simulation show that, indeed, the demagnified image of a Gaussian source by ideal lenses shows a vertical distribution that is not Gaussian (see Fig.~\ref{fig:ray tracing ideal elements}a ).
-The FWHM is not close to what is expected theoretically because of the high numerical aperture.
-The finite acceptance of the optical surfaces and apertures makes possible a significant reduction of the aberrations.
-The price to pay is of course a large reduction of the intensity.
-Defining the multilayer and mirror acceptances as slits right before the ideal focusing elements, we obtain the image sizes shown in Fig.~\ref{fig:ray tracing ideal elements}b.
-Here, the image dimensions are importantly reduced in horizontal for the ESRF-High$\beta$ because of a great cut in beam dimensions due to the effect of the 50$\mu$m VSS slit and KB acceptance.
-It can also be seen the complete absence of the tails also due to the effect of beam cut by the apertures.
-% }
-
 We performed ray tracing using the elliptical reflectors with finite dimensions as focusing elements.
@@
 These values confirm the analytical estimations (0.2\% for ESRF and and 2.95\% for EBS).
-As mentioned before, these values are similar to the coherent fraction \cite{coherentfraction} (0.12 for ESRF and 2.73\% for EBS), meaning that a very high coherence of the beam is expected, therefore some degradation of the beam by clipping the beam by apertures and element acceptance is expected.
-
-% Moreover, the treatment of the slope errors cannot be done without considerent effects of coherence, as discussed in \cite{hybrid}. This will be discussed in the next sections.
+As mentioned before, these values are similar to the coherent fraction \cite{arxivCF} (0.12 for ESRF and 2.73\% for EBS), meaning that a very high coherence of the beam is expected, therefore some degradation of the beam by clipping the beam by apertures and element acceptance is expected.
 
@@
 \centering
-% \includegraphics[width=0.45\textwidth]{GRAPHICS/ideallensESRF.png}
-% \includegraphics[width=0.45\textwidth]{GRAPHICS/ideallensEBS.png}
-% 
-% \includegraphics[width=0.45\textwidth]{GRAPHICS/ideallensaperturesESRF.png}
-% \includegraphics[width=0.45\textwidth]{GRAPHICS/ideallensaperturesEBS.png}
-
 \includegraphics[width=0.45\textwidth]{GRAPHICS/idealelementsESRF.png}
@@
 \includegraphics[width=0.45\textwidth]{GRAPHICS/hybriderrorsEBS.png}
-
-\caption{Image produced by different types of ideal focusing for the ESRF (leftcolumn) and EBS (right column). The source with parameters in 
+\caption{Image produced by different types of ideal focusing for the ESRF (left column) and EBS (right column). The source with parameters in 
 \ref{tab:HandCalculations}. 
@@
 The second row shows the image calculated the hybrid option (ray tracing plus coherent diffraction) with elliptical reflecting surfaces.
-It shows the focal spot degradiation duw to the diffraction originated by cropping the beam by the KM mirrors.
+It shows the focal spot degradiation duw to the diffraction originated by cropping the beam by the KB mirrors.
 The calculation in the bottom row also include slope errors using the real mirror profiles measures at the ESRF metrology lab.
 It can be shown a further degradation and the growing of satellite structures in the vertical intensity distribution.
+Values of the histograms FWHM are displayed in the graphic titles.
 }
@@
 The surface finish of the optical elements is another limiting effect for determining the size and intensity of the beam at the image position.
-It is possible to infere the effect of mirror imperfections and in particular slope errors by only knowing some statistical parameters such as the slope error Root Mean Square (RMS) and then building synthetic surface profiles to be used in the simulations.
+It is possible to infere the effect of mirror imperfections and in particular slope errors by only knowing some statistical parameters such as the slope error RMS and then building synthetic surface profiles to be used in the simulations.
 This is the typical mode to operate when designing a new beamline where the mirrors do not exist yet.
@@
 When not available, one can use data from existing mirrors from a database (e.g., \cite{dabam}) and extrapolate some of the effects to the hypothetical new mirrors.
-The metrology data of the mirrors for the beamline under study are presented in \todo{Appendix 1}.
+The metrology data of the mirrors for the beamline under study are presented in Appendix~\ref{appendix:metrology}.
 These experimental data are used in all the simulations in this paper.
@@
 
-\todo{
-\begin{itemize}
- \item play with slope errors and use synthetic profiles to show the importance of the low frequencies in the intensity distribution?
- \item explain why KB errors are treated with hybrid and not with shadow. In Discussion section? 
-\end{itemize}
-}
-
 %-------------------------------------------------------------------------
- % Rafael
+% 
 %-------------------------------------------------------------------------
@@
 \centering
-\includegraphics[width=0.45\textwidth]{GRAPHICS/wofry400umslit.png}
-\includegraphics[width=0.45\textwidth]{GRAPHICS/wofry400umprofile.png}
-\caption{Diffraction spot producedof a spherical wavefront (R=-50mm) by a 400x}
+\includegraphics[width=0.48\textwidth]{GRAPHICS/wofry400umslit.png}
+\includegraphics[width=0.48\textwidth]{GRAPHICS/wofry400umprofile.png}
+\caption{Diffraction spot at focal position produced an aperture (400$\times$400 $\mu$m$^2$) intercepting an spherical collapsing wavefront (R=-50mm).
+\todo{Remove this figure?}}
 \end{figure}
@@
     \caption{Beam sizes calculated with simplified wave optics propagation (point source and ideal lenses with aperture for modeling the element dimensions).
+The zoom or scaling factor used for the propagation is also displayed.
     }
@@
     \begin{tabular}{cccccc} 
-    % using WAVEOPTICS WOFRY workspace
     Element:                & ML            & VSS          & KBv               & KBh           & sample        \\
@@
 \caption{Intensity distribution at the focal position in horizontal plane (left, FWHM=8.1 nm) and in vertical plane (right, FWHM=7.1 nm). 
-\todo{implement a sort of multielectron system - implement the slope error. In Discussion? } 
 }
@@
 \label{srw_se}
-% The reduction on the electron beam emittance, by reducing the electron beam horizontal size, leads to an increase on the horizontal lateral coherence of the radiation emitted by an undulator in the symmetry point of a straight section \cite{paper}.
-Incoherent simulation methods may be insufficient for a full assessment of the beamline design.
-To accurately account for coherence effects, one has to resort to wave optics simulations (commonly referred to as wave propagation).
-% 
+
 A next step in the wave optics simulation is, maintaining the full coherence, to generate a 2D source wavefront that matches in detail the real source (an undulator in a storage ring) and then simulate the propagation of such a wavefront through the optical elements with models that match well the characteristics of the real elements.
@@
 The computer code chosen for these calculations is Synchrotron Radiation Workshop (\textsc{srw}) \cite{codeSRW}, which over the past two decades has been extensively benchmarked throughout different synchrotron radiation facilities and has become widely accepted within the X-ray community.
-
 Figure \ref{fig:SingleElectron} shows the horizontal integrated intensity at the focal position for the ID16A.
@@
 It is also possible to add slope errors to the simulations to study the degradation on the focal spot quality (Fig.~\ref{fig:SingleElectron}).
-For that, the available available metrology data \inred{(Appendix 1)} are used.
+For that, the available available metrology data (see Appendix \ref{appendix:metrology})  are used.
 For both cases (i.e., simulating the beamline using thin optical elements or elliptical mirrors), the beam sizes calculated as the FWHM at the focal position are little affected by the presence of the slope errors (see Table \ref{tab:SRW_results}).
@@
 \centering
-\caption{Beam dimensions (FWHM) at the focal position for different models of focusing elements.}
+\caption{Beam dimensions (FWHM, in nm) at the focal position for different models of focusing elements.}
 \resizebox{\textwidth}{!}{%
@@
 Diffraction limited sources would emit radiation that can be described in good approximation by a single wavefront.
-But even for new ultra low-emittance storage rings (improperly called diffraction-limited sources, see discussion in \cite{coherentfraction}) the contribution of the electron dimensions ($\sigma_{e(h,v)}$ and $\sigma'_{e(h,v)}$) is not negligeable as compared with the natural emission of radiation ($\sigma_u$ and $\sigma'_u$ for undulators as in Eq.~\ref{eq:photon small sigmas}).
+But even for new ultra low-emittance storage rings (improperly called diffraction-limited sources, see discussion in \cite{arxivCF}) the contribution of the electron dimensions ($\sigma_{e(h,v)}$ and $\sigma'_{e(h,v)}$) is not negligeable as compared with the natural emission of radiation ($\sigma_u$ and $\sigma'_u$ for undulators as in Eq.~\ref{eq:photon small sigmas}).
 To account for this effect one should complete or extend the wavefront simulations to take into account that in fact not only a single wavefront contributes to the properties of the beam.
@@
 The innovative concept is the numerical description of the undulator CSD in terms of its coherent modes $\Phi_m(x,y;z,\omega)$ and eigenvalues $\lambda_m $.
-One can define the mode distribution as
+One can define the occupation of the coherent modes as
 \begin{equation}
@@
 \end{equation}
-or the cumulated mode distribution like
+or the cumulated occupation like
 \begin{equation}
@@
         \includegraphics[width=0.45\textwidth]{GRAPHICS/ebs_mode3.png}
-    \caption{a) Spectral density $S(x,y)$ at the source position for EBS.  }
+    \caption{a) Spectral density $S(x,y)$ at the source position for EBS. b) Intensity distribution of the first coherent modes ($m$=0 to $m$=3) at the source position.
+    \todo{label plots}
+    }
 \end{figure}
@@
 
-After the complete analysis done in the previous section on how the ID16A optics modify the characteristics of the x-ray beam, one can verify that the analytical previsions were a good estimator of the values then obtained using more sophisticated methods.
+After the complete analysis done in the previous section on how the ID16A optics modify the characteristics of the x-ray beam, one can verify that the analytical previsions (Section~\ref{level0}) were a good estimator of the values then obtained using more sophisticated methods.
 The size of the focused beam (e.g, 10x5 for EBS) and the transmission (3\%) calculated analytically are in good agreement with geometric ray tracing (8.4x4.8, and 2.28\%).
@@
 
-The combination of ray tracing and wave optics methods used indicate that the focalized beam size due to  demagnification with mirrors (giving for EBS 10x5 nm2 analytically, 8.4x4.8 nm2 with ray tracing) has to be corrected with diffraction effect (8.4x7.8 hybrid, 12.7x9.2 SRW) and then add the slope error effect.
-The slope errors are very small, necessary not to blow up the size and kit it in the specifications range (10.8x8.2 hybrid 13.6x9.9 SRW).
+The combination of ray tracing and wave optics methods used indicate that the focalized beam size due to  demagnification with mirrors (giving for EBS 10$\times$5 nm$^2$ analytically, 8.4$\times$4.8 nm$^2$ with ray tracing) has to be corrected with diffraction effect (8.4$\times$7.8 nm$^2$ hybrid, 12.7$\times$9.2 nm$^2$ SRW) and then add the slope error effect.
+The slope errors do not to blow up the size and kit it in the specifications range (10.8$\times$8.2 nm$^2$ hybrid 13.6$\times$9.9 nm$^2$ SRW).
 It is important to treat the the slope errors correctly.
-Slope errors can be included in the ray tracing using the data in Appendix A and applying specular reflection on the optical surface with errors.
+Slope errors can be included in the ray tracing using the data in Appendix~\ref{appendix:metrology} and applying specular reflection on the optical surface with errors.
 This is the standard mode of including slope errors in SHADOW, using measured profiles or synthetic ones.
@@
 Here, the mirror height errors $z(x)$ are projected to the plane perpendicular to the propagation ($z(x) \sin \theta$, with $\theta$ the grazing angle).
-They introduce a phase shift $\psi$ in the wavefront phase proportional to the optical path $\psi = - k z(x) \sin \theta $.
+They introduce a phase shift $\psi$ in the wavefront phase proportional to the optical path $\psi = k z(x) \sin \theta $.
 This is the algorithm used in the simulations in Sections~\ref{level1}, \ref{srw_se} and \ref{srw_me}.
@@
 However, for those concerning partial coherence, very long calculations are needed.
-For example, Monte Carlo calculations in Section~\ref{srw_me} for the multi electron case \inred{HOW MANY ELECTRONS?} spent a few hours on a 56-core computer cluster.
+For example, Monte Carlo calculations in Section~\ref{srw_me} for the multi electron case spent a few hours on a 56-core computer cluster for simulating \inred{NN} electrons.
 For performing the COMSYL coherent mode decomposition the calculation time depends very much on the dimensions on the problem.
-For EBS the wavefronts were sampled using 1007x335 pixels, and 1103 modes were computed representing 98\% of the spectral density.
+For EBS the wavefronts were sampled using 1007$\times$335 pixels, and 1103 modes were computed representing 98\% of the spectral density.
 This represents 5 MB per wavefront and a matrix of 1695 GB to diagonalize.
@@
 The calculation is more expensive if one wants to decompose a source that is pretty incoherent like our ESRF-High-$\beta$.
-This is a limit case, it was used wavefields of 10075x201 pixels to accommodate the large spatial extension of the source.
-We obtained 4016 modes totalizing only XX\% of the spectral density.
+This is a limit case, it was used wavefields of 10075$\times$201 pixels to accommodate the large spatial extension of the source.
+We obtained 4016 modes totalizing only 86\% of the spectral density.
 The calculation took more than 4 days in a 84 cores cluster.
-%To give an idea of the real cost, a similar work send to the AWS cluster costed about 2000 USD. 
 This was done as a test case, to demonstrate the feasibility of the calculations.
 It is not recommended to do similar calculations that imply a the cost because a source that is quite incoherent was treated with partial coherence methods.
-In Appendix B we present for the ESRF High$\beta$ similar results than those in Section~\ref{comsyl}.
+\inred{In Appendix B we present for the ESRF High$\beta$ similar results than those in Section~\ref{comsyl}}.
 
@@
 
-\todo{discuss coherence length}
+\todo{
+Complete discussion. Some ideas:
+\begin{itemize}
+ \item DISCUSS EXPERIMENTAL RESULTS?
+ \item Discuss coherence length?
+ \item Comment on the heat-load deformation 
+ \item Extend the effect of reflectivity by ML and mirror coating
+ \item Talk a bit more on Oasys?
+ \item comment on general strategy for beamline simulations and the application of the shown hierarchical model. Make a flowchart
+\end{itemize}
+ }
 
@@
 %
-% 
-% The fist coherence mode is always centered on-axis (see, e.g, Fig.~\ref{fig:spectraldensity}b) and will always be centered when propagated until a position where an optical element (an aperture or mirror with finite dimensions) will crop it removing part of its intensity but letting the central part to go throught.
-The subsequent modes will be more absorbed as they extend more.
-In particular the modes that do not have a central lobe will lose more intensity.
-
-\todo{
-\begin{itemize}
- \item comment on the heat-load deformation 
- \item comment the effect of reflectivity by ML and mirror coating
- \item comment on the main results in terms of intensity and size
- \item EXPERIMENTAL RESULTS?
- \item comment ``where to stop'' in a simulation, depending on the level of accuracy requested
- \item comment on the computer resources used for simulations
- \item comment on software tools: OASYS (and make a graphic)
- \item discuss the particularities of this beamline 
- \item comment on general strategy for beamline simulations and the application of the shown hierarchical model. Make a flowchart
-\end{itemize}
-}
 
@@
 \section{Metrology of the KB mirrors}
-The residual height error of the multilayer monochromator and the KB mirrors, after removing the main ellitical shape, are shown in Fig.~\ref{fog:metrology}.
+\label{appendix:metrology}
+The residual height error of the multilayer monochromator and the KB mirrors, after removing the main ellitical shape, are shown in Fig.~\ref{fig:metrology}.
 These profiles are used along the text in ray tracing and wave optics simulations.
@@
 \includegraphics[width=0.8\textwidth]{GRAPHICS/ID16A_KB1_HF.png}
-\caption{Error height profiles used for the simulations for a) the ML monochromator, b) the vertical focusing mirror os the KB, and c) T=the horozontal focusing mirror of the KB.}
+\caption{Error height profiles used for the simulations for a) the ML monochromator, b) the KB vertical focusing mirror, and c) the KB horizontal focusing mirror.}
 \end{figure}
@@
 \section{COMSYL results for the ESRF-H$\beta$}
+\todo{Under development}
 

### 2019-05-07 paper-hierarchical.tex
Updates from Overleaf

@@
 
-\usepackage{epsfig,color}
+\usepackage{epsfig}
 \usepackage{graphicx}
 \usepackage{amsmath,amssymb}
+\usepackage{color,xcolor}
+\usepackage{siunitx}
+\usepackage[normalem]{ulem}
 
@@
 
+%-------------------------------------------------------------
+%%%%%%%%%%% to ignore the \sout{}, uncomment the line below
+%\renewcommand{\sout}[1]{\unskip}
+%-------------------------------------------------------------
+%%%%%%%%%%% to change color from red to black
+%ctrl+f, put 'red' in 'Search for' and 'black' in 'Replace with', then press 'All'
+%-------------------------------------------------------------
+
      %-------------------------------------------------------------------------
@@
 We consider different approaches to simulate a modern X-ray beamline.
-Several methodologies with increasing complexity are applied to discuss the relevant parameters that quantify the beamline performances.
-Parameters such as flux, dimensions and intensity distribution of the focused beam and coherence properties are obtained from simple analytical calculations to sophisticated computer simulations using ray tracing and waveoptics techniques.
-A last generation x-ray nanofocusing beamline for coherent  applications (ID16A at the ESRF) has been chosen  to study in detail the issues related to highly demagnifying sychrotron sources and exploiting the beam coherence.
-The performance of the beamline is studied for two storages rings: the old ESRF (emittance 4000 pm) and the new ESRF-EBS (emittance 150 pm).
+Several methodologies with increasing complexity are applied to discuss the relevant parameters that quantify the beamline {\color{green}\sout{performances}performance}.
+Parameters such as flux, dimensions and intensity distribution of the focused beam and coherence properties are obtained from simple analytical calculations to sophisticated computer simulations using ray tracing and {\color{red}\sout{waveoptics}wave optics} techniques.
+A {\color{green}\sout{last}latest} generation x-ray nanofocusing beamline for coherent  applications (ID16A at the ESRF) has been chosen  to study in detail the issues related to highly demagnifying  {\color{green}\sout{sychrotron}synchrotron} sources and exploiting the beam coherence.
+The performance of the beamline is studied for two storage{\color{green}\sout{s}} rings: the old ESRF (emittance 4000 pm) and the new ESRF-EBS (emittance 150 pm).
 In addition to traditional results in term of flux and beam sizes, an innovative study on the partial coherence properties based on propagation of coherent modes is presented.
 The different algorithms and methodologies are implemented in the software suite OASYS \cite{codeOASYS}.
-They are discussed with emphasis in benefits and limitations of each one.
+They are discussed with emphasis beamline {\color{green}\sout{in}placed upon the } benefits and limitations of each.
 \end{abstract}
@@
 The first upgrade of a large facility will be the EBS (Extremely Brilliant Source) \cite{orangebook} at the European Synchrotron Radiation Facility (ESRF).
-The new storage ring of 150 pm emittance that will be commissioned in 2020 will significantly boost the x-ray coherence of the synchrotron beams.
+The new storage ring of 150 pm emittance that will be commissioned in 2020 will significantly boost the  {\color{green}\sout{x-ray}} coherence of the  {\color{green}\sout{synchrotron} x-ray} beams.
 Applications exploiting beam coherence such as x-ray photon correlation spectroscopy, coherent diffraction imaging, propagation-based phase contrast imaging and ptychography will strongly benefit from this update.
 
-Accurate calculation and quantitative evaluation of the parameters related to x-ray optics is of paramount importance for designing, building and exploiting the new beamlines.
+Accurate calculation and quantitative evaluation of the parameters related to x-ray optics is of {\color{green}\sout{paramount} great} importance for designing, building and exploiting {\color{green}\sout{the}} new beamlines.
 Every modern beamline follows a procedure of conceptual design plus technical design where detailed simulations are essential.
-These calculations permit to test the design parameters in a virtual computer environment where advantages and limitations can be studied quantitatively.
-The selection to the best performant optics configuration requires the detailed simulation of the optics imperfections, that in many cases are the limiting factor of the beamline optical configuration.
-Deformation in the optical elements due to heat load need to be controlled, therefore optics simulations have to include results from engineering modeling of the thermal deformation associated to mechanical issues usually dome using finite eelement analysis.
+These calculations  {\color{green}\sout{permit to test} allowing testing of} the design parameters in a virtual computer environment where advantages and limitations can be studied quantitatively.
+{\color{green}\sout{The selection to best performant optics configuration} Selection of the optimal optical layout}  requires the detailed simulation of the optics imperfections, that in many cases are the limiting factor of the beamline optical configuration.
+{\color{green}\sout{Deformation in} Deformations of} the optical elements due to heat load need to be controlled, therefore optics simulations have to include results from engineering modeling of the thermal deformations {\color{green}\sout{associated to mechanical issues}} usually {\color{red}\sout{dome}done} using finite {\color{red}\sout{e}}element analysis.
 Traditionally the optics calculations for synchrotron beamlines are done using ray tracing techniques.
-They model the x-ray beam as a collection of different rays that are incoherent one to another.
-This technique is still essential for obtaining information on aberrations, flux propagation, monochromator bandwidth, etc.
-However, the high coherence of the new sources make necessary to complement the ray tracing methods with models including effects of diffraction and scattering from a coherent beam.
-A fully coherent beam is still a dream for the new generation storage rings.
-In spite of a high increate of the coherent fraction of the new beams, the coherent fraction at tipically hard x-ray photon energies is still at few percent for the new sources.
-This makes still important a ray tracing analysis that does not become obsolete, but has to be complemented with an analysis based on wave optics.
-Because of the non perfect coherence the wave optics analysis becomes non trivial and can be done with different levels of approximation, as discussed in this work.
-In addition, wave optics simulations are usually expensive from the computer point of view.
-
-For illustrating the applicability of a battery of methods and solutions for beamline optics simulations we have chosen the ESRF beamline ID16A.
-This beamline provides a high-brilliance beam focused to a few nanometer.
-It shows characteristics increasingly searched in modern beamlines: extremely high demagnification (sub-micrometer beam) and high coherence.
-It combines coherent imaging techniques and X-ray fluorescence microscopy to perform quantitative 3D characterization of the morphology and the elemental composition of specimens in their native state.
+They model the x-ray beam as a collection of {\color{green}\sout{different}incoherent} rays.
+{\color{green}\sout{that are incoherent one to another}}  This technique is still essential for obtaining information on aberrations, flux propagation, monochromator bandwidth, etc.
+However, the high coherence of the new sources {\color{green}\sout{make necessary}can introduce the necessity} to complement the ray tracing methods with models including effects of diffraction and scattering from a coherent beam.
+A fully coherent {\color{green} hard x-ray}beam is still a dream for the new generation storage rings.
+In spite of a {\color{green}\sout{high increate of }considerable increase in}  the coherent fraction of the new beams, the coherent fraction at {\color{red}typically\sout{tipically}} hard x-ray photon energies is still a few percent for the new sources.
+{\color{green}\sout{This makes still important a } Consequently} ray tracing analysis {\color{green}\sout{that}} does not become obsolete, but has to be complemented with an analysis based on wave optics.
+Because of the {\color{green}\sout{non perfect} partial} coherence the wave optics analysis becomes non trivial and can be done with different levels of approximation, as discussed in this work.
+In addition, wave optics simulations are usually computationally expensive.\todo{I noticed that there is very few citations to references in the introduction.
+This is not so nice, as we don't knowledge any previous work.
+Maybe one could add a few citations in the introduction.}
+
+To illustrate the applicability of a variety of methods and solutions for beamline optical simulations we have chosen the ESRF beamline ID16A~{\color{red}\cite{ID16A}}.
+This beamline provides a high-brilliance beam focused to a few nanometers.
+It displays characteristics increasingly aimed for in modern beamlines: extremely high demagnification (sub-micrometer focused beam size) and high coherence.
+It combines coherent imaging techniques and X-ray fluorescence microscopy to perform quantitative 3D characterization of the morphology and the elemental composition of specimens in their native state~{\color{red}\cite{XRFID16A}}.
 
 X-ray beamlines are very particular optical systems.
-Some implications of exploiting the low wavelengths such as the necessity of using grazing incidence reflectors and the use of perfect crystals as typical hard x-ray monochromators make useless, or at least incomplete, the application of common commercial software packages for the study of synchrotron beamlines.
+Some implications of exploiting the {\color{red}\sout{low}short} wavelengths such as the use of grazing incidence reflectors and the use of perfect crystals as typical hard x-ray monochromators render common commercial software packages ill-adapted for the study of synchrotron beamlines.
 For this reason many optics tools have been developed by the synchrotron community.
-Some of them have a long history like SHADOW (more than 40 years old) or SRW (almost 30), for mentioning only codes used in this work.
-These historical tools together with other new applications have been agglutinated in a user friendly software OASYS.
-Many of the calculations presented here have been performed directly in OASYS or with script created with the help of OASYS.
-
-The paper is structured as follows: we first introduce in Section~\ref{Beamline description} the main parameters of the beamline and make the selection of the particular choice of the parameters used in the calculations.
-Section~\ref{Simulations} is the central part of the paper, and describe the different methods used to analyse the beamline with increasing complexity, setting a hierarchy of methodologies that can be exported to the simulations of any beamline.
-In Section \ref{level0} we calculate with simple methods approximated values of focal size, flux and coherent fraction.
-Then accuracy is improved using geometrical methods in Section \ref{level1} that give reasonable results as far as some corrections are considered because of the  mentioned characteristics of high demagnification and high coherence of the studied beamline.
-The wave optics section~\ref{level2} also include different models whose complexity increases when the results expected are more and mode demanding.
-The simulations methods are presented in a hierarchical approach:  they provide more and better results but the price to pay is to deal with more complex and computationally expensive tools.
+Some of them have a long history like SHADOW (more than 40 years old) or SRW (almost 30), to mention only codes used in this work.
+These historical tools together with other new applications have been collected together in a user friendly software package: OASYS.
+Many of the calculations presented here have been performed directly in OASYS or with scripts created with the help of OASYS.
+
+The paper is structured as follows: we first introduce in Section~\ref{Beamline description} the main parameters of the beamline and explain the particular choice of the parameters used in the calculations.
+Section~\ref{Simulations} is the central part of the paper, and describes the different methods used to analyse the beamline with increasing complexity, setting a hierarchy of methodologies that can be exported to the simulations of any beamline.
+In Section \ref{level0} we calculate with simple methods approximate values of focal size, flux and coherent fraction.
+Accuracy can improved using geometrical methods in Section \ref{level1} that allow a more complete simulation of the beamline optics within certain limits which are related, in this example, to high demagnification and high coherence of the studied beamline.
+The wave optics section~\ref{level2} includes different models of increasing complexity as the required accuracy of the results becomes more demanding.
+The simulation methods are presented in a hierarchical approach:  they provide more and better results but the price to pay is to deal with more complex and computationally expensive tools.
 
@@
 It uses one or several undulators as source.
-During the period 2011-2018 it worked exploiting the ESRF storage ring’s double-bend achromat (DBA) lattice with horizontal emittance about 4000 pm.
-The replacement of this lattice by a new hybrid multi-bend achromat (HMBA) lattice is being achieved in the ESRF Upgrade Phase II.
+During the period 2011-2018 it operated exploiting the ESRF storage ring’s double-bend achromat (DBA) lattice with horizontal emittance about 4000 pm.
+In the second phase of the ESRF Upgrade the original storage ring is being replaced by a new hybrid multi-bend achromat (HMBA) lattice.
 This new lattice will drastically improve the performances of the ESRF source.
-The result, the Extreme Brilliant Source (EBS) \cite{orangebook} will increase the brightness and coherent fraction by a factor of more than 100, by strongly reducing the horizontal emittance.
-This is possible using almost the same ring circunference and electron energy $E_e$.
-As as result, the x-ray sources will be much coherent, specially in the horizontal direction.
-This work shows some quantitative results on how the storage ring upgrade will affect a particular beamline is analyzed in this paper.
-
-ID16A \cite{ID16A} is a bemline that provides a nanometric, highly coherent, pink beam.
+The resulting Extreme Brilliant Source (EBS) \cite{orangebook} will increase the brightness and coherent fraction by a factor of more than 100, by strongly reducing the horizontal emittance.
+This is possible using almost the same ring circumference and electron energy $E_e$.
+As a result, the x-ray sources will be much coherent, specially in the horizontal direction.
+In this paper, some quantitative results on how the storage ring upgrade will affect a particular beamline are presented.
+%RC190404: @PhaseI -> from the orange book: On the basis of the Purple Book, the ESRF Council decided in June 2008 to launch the ESRF UP by establishing a two-phase process.
+The Upgrade Programme Phase I (UP PI) for the period 2009-2015, and the Upgrade Programme Phase II (UP PII) to be prepared in the period 2012-2014, and to be presented to the Council for consideration in 2014 and launch in 2015.
+% The  Upgrade  Programme  Phase  I  has  been  on  the  ESFRI  Roadmap  since  its  fi rst publication  in  2006,  and  is  today  heading  towards  its  successful  completion  by  the end of 2015.
+
+ID16A \cite{ID16A} is a beamline that provides a nanometric, highly coherent beam with broad energy-bandpass.
 It is dedicated to nanofocusing applications.
 \todo{please summarize some use and application}.
-The beamline usually operates at three fixed photon energies: 11200 eV, 17225 eV and 33600 eV.
-The straight section in the magnet lattice holds three undulators: an in-vacuum undulator of $\lambda_u=26$~mm period, and two other conventional undulators of 18 and 22.4 mm period.
+The beamline can operate at three fixed photon energies: 11200 eV, 17225 eV and 33600 eV.
+The straight section in the magnet lattice houses three undulators: an in-vacuum undulator of $\lambda_u=26$~mm period, and two other conventional undulators of 18 and 22.4 mm period.
 The undulator parameters and $K$-values at the resonances for the energies in use are shown in Table \ref{tab:Undulators}.
+
+%RC190404: From the ID16D page -> The beamline provides a high-brilliance beam focused down to nanometer size, allowing quantitative 3D characterization of the morphology and the elemental composition of specimens in their native state by combining coherent imaging techniques and X-ray fluorescence microscopy.
+% The 185 m long beamline provides nano-focused beams for analytical imaging.
+The nano-imaging beamline addresses problems in biology, biomedicine and nano-technology using X-ray fluorescence microscopy and nano-tomography.
+It is optimised for ultimate hard X-ray focusing of a beam with a large energy bandwidth at specific energies (17keV or 33.6 keV).
+Cryogenic sample preservation is available for 2D and 3D imaging.
+% Techniques: coherent imaging; phase-contrast imaging; near-field and far-field ptychography; X-ray fluorescence microscopy;  scanning transmission X-ray microscopy; nanotomography
+% http://www.esrf.eu/home/UsersAndScience/Experiments/XNP/ID16A/over.html
 
@@
 ID16A is a long beamline (the distance from the center of the straight section to the sample position is 185 m).
-A strong focusing is performed with a Kirkpatrick-Baez (KB) set of mirrors placed very close (a few cm) upstream the sample position (focal plane), which ensures a high demagnification.
+Strong focusing is performed using a Kirkpatrick-Baez (KB) mirror system placed a few cm upstream of the sample position (focal plane), which ensures a high demagnification.
 In order to match the acceptance of the KB in the vertical and horizontal planes a focusing multilayer monochromator (ML) has
-been chosen. It i) monochromatises the beam with a large energy bandwidth (typically $\Delta E/E \approx 10^{-2}$) to allow high (pink) flux, 
-and ii) creates a secondary secondary source in horizontal plane 40 m downstream from the undulator, where a horizontal slit (VSS, virtual source slit) is placed.
+been chosen. It i) monochromatises the beam with a large energy bandwidth (typically $\Delta E/E \approx 10^{-2}$) to allow high flux, 
+and ii) creates a secondary secondary source in the horizontal plane 40 m downstream of the undulator, where a horizontal slit (VSS, virtual source slit) is placed.
 A beamline schematic is shown in Figure~\ref{fig:ID16A}.
@@
 For all simulations presented hereafter, we selected the configuration using a single undulator U18.3 placed in the center of the straight section and tuned to its first harmonic at a photon energy of $E=17225$ eV (Table \ref{tab:Undulators}).
-This section is devided in three subsections, that present three main levels of simulations: simple calculations that can be made by hand (section \ref{level0}), ray-tracing simulations (section \ref{level1}) and wave optics simulations (section \ref{level2}).
+This section is divided in three subsections, that present three main levels of simulations: simple calculations that can be made by hand (section \ref{level0}), ray-tracing simulations (section \ref{level1}) and wave optics simulations (section \ref{level2}).
 
@@
 
-We first estimate the beam size at different positions at the beamline, in particular at the source plane and at the image plane.
-A first na{\"{i}}ve estimation of the focal size can be done by considering the source size times the magnification factor $M$.
+We first estimate the beam size at different positions in the beamline, in particular at the source plane and at the image plane.
+A first na{\"{i}}ve estimation of the focal size can be made by considering the source size times the magnification factor $M$.
 In the paraxial approximation, the magnification factor can be expressed as $M=q/p$, where $p$ is the distance between object and imaging element and $q$ is the distance between imaging element and image.
@@
 
-The source size is usually approximated supposing a Gaussian distribution of the electron sizes (or divergence) (Table \ref{tab:eBeam}) and undulator natural source size or divergence.
-Fo some cases, as the undulator emission, the Gaussian approximation may not be accurate and have some implications that will be discussed later.
-Let $\sigma_{e(h,v)}$ and $\sigma_{e(h,v)}'$ be the electron beam size and divergence standard deviation of the electron distributions at a centre of the symmetry of the straight section in a plane perpendicular to the main propagation direction.
+The source size (or divergence) is usually approximated supposing a Gaussian distribution of the electron {\color{green}\sout{sizes}positions} (or divergence) (Table \ref{tab:eBeam}) and {\color{green}\sout{undulator natural}single electron photon} source size or divergence.
+In some cases, such as for undulator emission, this Gaussian approximation may not be accurate and have some implications that will be discussed later.
+Let $\sigma_{e(h,v)}$ and $\sigma_{e(h,v)}'$ be the standard deviations of electron beam size and divergence  at the centre of  symmetry of the straight section in a plane perpendicular to the main propagation direction.
 The subscripts $h$ and $v$ stand for horizontal and vertical, respectively.
 
-The undulator natural source size and divergence are given by \cite{elleaume}:
+The undulator natural photon source size and divergence are given by \cite{elleaume}:
 \begin{align}
@@
 The electron beam usually presents different horizontal and vertical beam emittances, but the undulator natural source size or divergence presents radial symmetry at the central cone (excluding polarization).
-The photon source size and divergence are approximately given by the convolution of the electron sizes and the undulator natural sizes.
+The photon source size and divergence are approximately given by the convolution of the electron beam sizes and the undulator natural sizes.
 If we consider Gaussian distributions, the photon source is
@@
 
-The beam sizes at different positions at the beamline were estimated in a different way depending if the beam is focused or not.
-If the beam is focused, its dimensions in a first approximation are the photon source size multiplied by the corresponding magnification factor.
+The beam sizes at different positions in the beamline are estimated in different ways depending upon whether the beam is focused or not.
+If the beam is focused, its dimensions, to a first approximation, are the photon source size multiplied by the corresponding magnification factor.
 For predicting the photon beam sizes in a position out of the focus the beam divergence is used.
-In an ideal {\'{e}}tendue preserving optical system, where the conservation of the Smith-Helmholz invariante applies, the resulting beam divergence transforms inversely as the beam size: it can be calculated as the product between the beam divergence at the source with the demagnification factor $M^{-1}$.
-
-Using the values for U18.3, $n$=1, and  $E=17225$ eV (Table \ref{tab:Undulators}), one obtains a focal size 139$\times$5 nm$^2$ (H$\times$V) for ESRF and 10$\times$5 nm$^2$ for EBS (Table~\ref{tab:HandCalculations}).
-The focal spot for ESRF lattice is too large in horizontal for the beamline specification, therefore the VSS could be closed to 50$\mu$m thus reducing the beamsize at the VSS plane by a factor ca.
+In an ideal {\'{e}}tendue conserving optical system, where the Smith-Helmholz invariant applies, the resulting beam divergence transforms inversely as the beam size: it can be calculated as the product between the beam divergence at the source with the demagnification factor $M^{-1}$.
+
+Using the values for U18.3, $n$=1, and  $E=17225$ eV (Table \ref{tab:Undulators}) and the optical configuration described in Table \ref{tab:Positions}, one obtains a focal size 139$\times$5 nm$^2$ (H$\times$V) for ESRF and 10$\times$5 nm$^2$ for EBS (Table~\ref{tab:HandCalculations}).
+The horizontal focal spot for the ESRF lattice is larger than the specified target value for the beamline, therefore the VSS could be closed to 50$\mu$m thus reducing the beamsize at the VSS plane by a factor ca.
 8 \inred{$\approx$ 403.8/50=8.076}.
@@
 
-In general, for grazing incident optics imaging extended sources with large demagnifications ($M^{-1} \gg 100$), the effect of optical aberrations is very important even using a perfect optical surface that focus point-to-point.
-As a consequence, the results calculated using only the demagnification factor are usually too optimistic.
-Moreover imperfections in the optical surface (figure errors, slope errors and microroughness) are often the limiting factors of the real reflective optics.
-Their effect will studied in Sections~\ref{level1} and \ref{level2}.
+In general, for simple elliptical cylinder figured grazing incident optics imaging extended sources with large demagnifications ($M^{-1} \gg 100$), the effect of optical aberrations is very important even using a perfect optical surface.
+As a consequence, the results calculated using only the demagnification factor are usually optimistic.
+Moreover imperfections in the optical surface (figure errors, slope errors and microroughness) are often the limiting factors of real reflective optics.
+Their effects will studied in Sections~\ref{level1} and \ref{level2}.
+%Effects of slope errors and diffraction limits can be simply estimated by hand calculations as well...
 %The smearing of the focal size because of diffractive effects by the mirror clipping will also be discussed.
@@
 Values are FWHM.
-Values in parentheses corresponds to closing the VSS slit to 50$\mu$m (21.3 if approximating by the Gaussian $\sigma$).}
+Values in parentheses correspond to closing the VSS slit to 50$\mu$m (21.3 if approximating by the Gaussian $\sigma$).}
     \resizebox{\textwidth}{!}{%
@@
 It is not straightforward to get this information analytically but one can calculate the flux and power using different codes.
-Perhaps the most performant and well maintained codes spread in the synchrotron community are SPECTRA \cite{codeSPECTRA} and SRW \cite{codeSRW}.
+Perhaps the most advanced and well maintained codes used in the synchrotron community are SPECTRA \cite{codeSPECTRA} and SRW \cite{codeSRW}.
 The XOPPY package in the OASYS suite implements three codes: US \cite{codeUS}, URGENT \cite{codeURGENT} and SRW.
-Figure~\ref{fig:FluxU18} shows the spectra calculated over an aperture if 1 $\times$ 1 mm$^2$ placed at $d$=27.1 m.
+Figure~\ref{fig:FluxU18} shows the spectra calculated \todo{indicate which code used} over an aperture of 1 $\times$ 1 mm$^2$ placed at $d$=27.1 m.
 This aperture opening is set to fully accept the central cone at the resonance peak of the harmonic in use ($\max(\Sigma') \times d = 0.73$ mm), with $\Sigma'$ values from in Table~\ref{tab:HandCalculations}.
@@
 One can appreciate how the emittance plays an important role in both peak width and specially in peak value.
-The number of photons per second after the multilayer monochromator is 2.9 10$^{15}$ for EBS and 1.9 10$^{15}$ for ESRF, considering that the energy bandwidth transmitted by a multilater monochromator is $\Delta E/E \approx 10^{-2}$ and the bandwidth in Fig.~\ref{fig:FluxU18} is 10$^{-3}$, and supposing at this point ideal reflectivity (100\%) for the multilayers.
+The number of photons per second after the multilayer monochromator is 2.9 10$^{15}$ for EBS and 1.9 10$^{15}$ for ESRF, considering that the energy bandwidth transmitted by a multilayer monochromator is $\Delta E/E \approx 10^{-2}$ and the bandwidth in Fig.~\ref{fig:FluxU18} is 10$^{-3}$, and supposing, at this point, ideal reflectivity (100\%) for the multilayers.
 
@@
 
-To estimate the geometrical transmitivity of the beamline we have to consider which optical elements crop the beam and remove photons.
+To estimate the geometrical {\color{red}\sout{transmitivity}transmission?} of the beamline we have to consider which optical elements crop the beam and remove photons.
 This can be analysed using their angular acceptance or numerical aperture.
@@
 We consider a VSS closed to $S_h=50~\mu$m in horizontal and fully opened in vertical.
-For ESRF its transmittivity is 1/8 (as discussed before) and 1 for EBS.
-In a similar way, the KB mirrors will crop the beam to a size that is the projection of their length to a direction perpendicular to the optical axis.
-Using mirror length of $L_v$=60 mm and $L_h$=26 mm for VFM and HFM, respectively, and incident angle of 15 mrad, the KB is behaving like an aperture of $900 \times 390 \mu m^2$.
+For ESRF its {\color{red}\sout{transmittivity}transmission?} is 1/8 (as discussed before) and 1 for EBS.
+In a similar way, the KB mirrors will crop the beam to a size that is the projection of their {\color{red}useful optical} length onto a direction perpendicular to the optical axis.
+Using a mirror length of $L_v$=60 mm and $L_h$=26 mm for VFM and HFM, respectively, and incident angle of 15 mrad, the KB behaves as an aperture of $900 \times 390 \mu m^2$.
 We get N.A.
-values of 6.2$\times$21.6 $mu$rad (aperture size divided by the distance from the mirror to the VSS in H and to the source in V).
-The transmittivity coefficient is the ratio of the accepted N.A.
+values of 6.2$\times$21.6 $\mu rad^2$ (aperture size divided by the distance from the mirror to the VSS in H and to the source in V).
+The {\color{red}\sout{transmittivity}transmission?} coefficient is the ratio of the accepted N.A.
 over the beam divergence, resulting 9.5\% (H) and 0.18\% (V)for ESRF and 16.3\% and 0.18\% for EBS.
@@
 We used the undulator peak value over the whole ML bandwidth, that overestimates the integrated flux.}
-
-One can also discuss some simple consequences about the beam coherence.
-The coherence fraction (CF) \cite{arxivCF} of the undulator beam is the occupation of the lowest coherent mode, as it will be discussed in Section~\ref{level2}.
-An approximated estimation of it can be obtained using Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas}.
-The coherent fraction is the ratio of the phase space for the coherent beam (zero emittance) over the actual beam.
-The CF in its the horizontal and vertical planes are
+\todo{need to consider ML reflectivity for the 3 reflections, also undulator is not perfect, horizontal slit may be closed to less than $50 \mu m$, also KB mirrors may not be fully illuminated and slope error losses due to initial ML mono are neglected}
+
+One can also make some simple inferences about the beam coherence.
+The coherent fraction (CF) \cite{arxivCF} of the undulator beam is the occupation of the lowest coherent mode, and will be discussed further in Section~\ref{level2}.
+An approximated estimation of CF can be obtained using Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas}.
+The coherent fraction is the ratio of the phase space {\color{green}\sout{of the fully coherent beam (zero emittance) over the actual beam} of the true beam relative to that of the fully coherent (zero emittance) beam}.
+The CF in the horizontal and vertical planes are
 \begin{equation}\label{eq:coherent fraction}
@@
 and the total coherent fraction is the product of both (CF=CF$_h \times $CF$_v$).
-An estimation of the CF calculated in such a mode give CF=0.12\% for the ESRF and 2.73\% for EBS at 17 keV photon energy.
+An estimation of the CF calculated in this way gives CF=0.12\% for the ESRF and 2.73\% for EBS at 17 keV photon energy.
 The CF roughly indicates the ratio of the ``coherent photons'' over the total number of photons.
 It is a good indicator of the quality of a coherent beam (a fully coherent beam has CF=1 and an incoherent beam CF=0).
-The beamline optics acts as a ``coherence filter'' by removing the ``bad'' or ``incoherent'' photons and in consequence increasing the CF.
+The beamline optics act as a ``coherence filter'' by removing the ``bad'' or ``incoherent'' photons and in consequence increasing the CF.
 A complete cleaning is impossible because of the spatial overlapping of the different coherence modes.
 However, apertures centered at the optical axis would clean the beam from these undesired photons.
-If we compare the values transmittivity of the beamline due to geometrical clipping  (0.2\% for ESRF and 2.95\% for EBS) with the the CF values, we remark that they are very close.
+If we compare the values {\color{green}\sout{transmittivity} transmission}of the beamline due to geometrical aperturing  (0.2\% for ESRF and 2.95\% for EBS) with the the CF values, we remark that they are very close.
 This suggests that the beam at the focal position will be highly coherent because the fraction of the photons removed from the beam matches the CF value that quantify the fraction of ``good'' or ``coherent'' photons.
 
-To assess the focused beam size given by a coherent beam, one can idealize the beamline in two parts.
-First, the ML monochromator and VSS that will prepare a coherent plane wave as the source is very far from the sample position.
-Second, the KB has a double role, acting as a focusing element of focal distance $f$=5 cm, but also crops the beam as as it would be done by a square aperture of $D \times D=$ 400$\times$400 $\mu m^2$ (the length of the mirror projected in a plane perpendicular to the beam propagation)
-\todo{REMOVE: we obtained 900 in H, that will be reduced by 0.5 as propagated from 10 cm to 5 cm upstrem from the sample, and 390 in V approximated by 400}.
-This aperture has two effects: i) as discussed before, it removes a lot of off-axis photons increasing the CF to a value close than one, and ii) make a broadening of the focal spot because of coherent diffraction.
-We know from optics that the diffraction by a square aperture of a collapsing spherical wavefront has an intensity distribution proportional to $\sinc^2(k D x/2 f) \sinc^2(k D y / 2 f)$, where $k=2\pi/\lambda$, and $x$ ($y$) is the horizontal (vertical) coordinate .
+To assess the focused beam size given by a coherent beam, one can consider the beamline in two sections: First, the ML monochromator and VSS that produce a coherent plane wave as the source (vertical) and VSS (horizontal) are both very far from the {\color{green}\sout{sample position} entrance aperture of the focusing optics}; Second, the KB-system playing a dual role, acting as a focusing element of focal distance $f$=5 cm, but also cropping the beam by acting as a square aperture of $D \times D=$ 400$\times$400 $\mu m^2$ (the length of the mirror projected in a plane perpendicular to the beam propagation)
+\todo{REMOVE: we obtained 900 in H, that will be reduced by 0.5 as propagated from 10 cm to 5 cm upstream of the sample, and 390 in V approximated by 400}.
+This aperture has two effects: i) as discussed before, it removes a lot of off-axis photons increasing the CF to a value close to one, and ii) broadens of the focal spot because of coherent diffraction.
+The diffraction by a square aperture of a collapsing spherical wavefront has an intensity distribution proportional to $\sinc^2(k D x/2 f) \sinc^2(k D y / 2 f)$, \todo{include citation?} where $k=2\pi/\lambda$, and $x$ ($y$) is the horizontal (vertical) coordinate .
 Considering that the FWHM of $\sinc(x)$ is approximately 2.78, one obtains a FWHM of 7.75$\times$7.75 nm$^2$ for the focal spot.
-This spot size is of the order or larger than the one obtained by geometrical considerations, therefore contributing significantly to the size of the focused beam.
+This spot size is of the order or larger than that calculated geometrical considerations alone, and therefore contributes significantly to the overall size of the focused beam.
 One can conclude that for the conditions in use this is a diffraction limited beamline.
 
-In summary, we show in this section how simple analytical calculations that can be done in its major part ``by hand'' help to estimate the geometrical beam size at focal position (17$\times$5 nm$^2$ for ESRF, 10$\times$5 nm$^2$ for the EBS) the flux (3.8 $10^{12}$ photons/s for ESRF and 8.5 $10^{13}$ photons/s for EBS) and also anticipate that the beam is highly coherent producing a diffraction limited focal spot of  7.75$\times$7.75 nm$^2$.
-We ignored at this point the effect of real optics containing surface errors, the reflectivity of the elements, and other aspects that will be treated in depth in the next sections.
+In summary, we have shown in this section how simple analytical calculations that can be done mostly ``by hand'' help to estimate the geometrical beam size at focal position (17$\times$5 nm$^2$ for ESRF, 10$\times$5 nm$^2$ for the EBS) the flux (3.8 $10^{12}$ photons/s for ESRF and 8.5 $10^{13}$ photons/s for EBS) and also anticipate that the beam is highly coherent producing a diffraction limited focal spot of  7.75$\times$7.75 nm$^2$.
+Thus far we ignored  the effect of real optics exhibiting surface errors, the reflectivity of the elements, and other aspects that will be treated in depth in the next sections.
 This is, however, the first step to be done when analyzing a beamline.
-It also helps in having a critical eye or sanity check on the numerical calcuations that will provide more precised values, but must remain the this order of magnitude.
+It also helps in providing an initial benchmark estimate which should, to a large extent, provide a check of the validity of the numerical calculations that will provide more precise values, but should remain at the same order of magnitude.
 
@@
 \label{level1}
-
-In this section we perform ray tracing calculations to study the effect of real optics (including aberrations and slope errors) to the beam size and beamline transmittivity.
+\todo{transmittivity or transmission}
+In this section we perform ray tracing calculations to study the effect of real optics (including aberrations and slope errors) upon the beam size and beamline transmittivity.
 The first subsection deals with the effect of aberrations and transmittivity due to geometrical considerations.
-The second subsection shows how the effect of slopes errors the KB mirrors degrade the focal spot.
-
-Ray tracing is a simulation method based on tracing some light rays along the optical system from the source to the image and retrieve the statistics of rays at the image to measure beam characteristics.
+The second subsection shows how the effect of slope errors of the KB mirrors degrade the focal spot.
+
+Ray tracing is a simulation method based on tracing some light rays along the optical system from the source to the image and retrieving the statistics of rays at the image to measure beam characteristics.
 For simple systems one can perform ray tracing even by hand, selecting some principal rays that will define the location and envelope of the image.
@@
 For reflectors, this change of direction is given by the specular reflection laws.
-For refractors, the change of direction os defined by the Snell law.
-Because Snell law uses the refraction index which is wavelength-dependent, it is then interetisting to add additional parameters to each ray, such as the wavelength.
-This will permit to calculate the refraction index that has to be used by each ray.
-The information of each ray can also be extended to include some intensity, or better electric field amplitude.
+For refractors, the change of direction is defined by  Snell's law.
+Because Snell's law uses the refractive index which is wavelength-dependent, it is then interesting to assign additional attributes to each ray, such as the wavelength.
+This will permit to calculate the refractive index that should  be used for each ray.
+The characteristics of each ray can also be extended to include some intensity, or better electric field amplitude.
 This allows to extend the pure geometrical tracing to include physical models that take into account the optical element reflectivity or transmittivity.
-This combination of using a geometrical model plus a physical model allows to simulate every element used in a synchrotron beamline, such as mirrors and combination of them (e.g., KBs); lenses, compound refractive lenses and transfocators; gratings of any type and shape, crystal systems, etc.
+This combination of using a geometrical model plus a physical model allows to simulate every element used in a synchrotron beamline, such as mirrors and combinations of them (e.g., KBs); lenses, compound refractive lenses and transfocators; gratings of any type and shape, crystal systems, etc.
 Ray tracing is a simple and extremely powerful technique for calculating the main characteristics of the photon beam (size, divergence, photon distribution) at every point of the beamline.
-Although very powerful commercial ray tracing packages are available in the market, they are not suitable for synchrotron simulations because they do not cover the particularities of X-ray devices and sources.
+Although very powerful commercial ray tracing packages are available in the market, they are generally ill adapted for synchrotron simulations because they do not cover the particularities of X-ray devices and sources.
 For that purposes, several packages created in the synchrotron community and are available, like SHADOW \cite{codeSHADOW}, RAY \cite{codeRAY}, McXtrace \cite{codeMCXTRACE} or XRT \cite{codeXRT}.
-SHADOW is probably the one that is more used and cited.
-We use SHADOW and its interface ShadowOUI \cite{codeSHADOWOUI}, available as a module of the OASYS suite.
-
-We performed ray tracing using the elliptical reflectors with finite dimensions as focusing elements.
+SHADOW is probably the one that is most widely used and cited.
+For this study we use SHADOW and its interface ShadowOUI \cite{codeSHADOWOUI}, available as a module of the OASYS suite.
+
+We performed ray tracing \todo{ it is currently unclear to me whether the ML mono was also simulated - and whether slope errors of this mono were included.
+The ML mono is a spherical reflector} using elliptical cylinder reflectors with finite dimensions as the KB focusing elements.
 The reflector shape is chosen to be elliptical as this is the theoretical shape producing a point-to-point focusing.
-For a first calculation  the elliptical elliptical reflectors are considered perfect, and have no slope errors.
-The results in Fig.~\ref{fig:ray tracing}a show two facts.
+For a first calculation  the elliptical  reflectors are considered perfect, and have no slope errors.
+The results in Fig.~\ref{fig:ray tracing}a show two phenomena.
 One is that the migration from ESRF lattice to the EBS will significantly improve the horizontal size, and the VSS is not needed.
@@
 Secondly, one can observe a much larger luminosity in the EBS case.
-The transimittivity values are 0.165\% for ESRF and 2.28\% for EBS.
+The transmittivity values are 0.165\% for ESRF and 2.28\% for EBS.
 These values confirm the analytical estimations (0.2\% for ESRF and and 2.95\% for EBS).
-As mentioned before, these values are similar to the coherent fraction \cite{arxivCF} (0.12 for ESRF and 2.73\% for EBS), meaning that a very high coherence of the beam is expected, therefore some degradation of the beam by clipping the beam by apertures and element acceptance is expected.
+As mentioned before, these values are similar to the coherent fraction \cite{arxivCF} (0.12 for ESRF and 2.73\% for EBS), meaning that a very high coherence of the beam is expected, therefore some broadening of the beam {\color{green}\sout{by clipping the beam by apertures and element acceptance} due to aperturing by the mirrors.}
 
@@
 \ref{tab:HandCalculations}. 
-The first row has been calculated using the elliptical reflectors (both for ML and KBs) but without slope errors. The focal beam dimensions are written in the plot title.
+The first row has been calculated using the elliptical reflectors (both for ML \todo{ML is not elliptical - it is a sphere} and KBs) but without slope errors.
+The focal beam dimensions are written in the plot title.
 The second row shows the image calculated the hybrid option (ray tracing plus coherent diffraction) with elliptical reflecting surfaces.
-It shows the focal spot degradiation duw to the diffraction originated by cropping the beam by the KB mirrors.
-The calculation in the bottom row also include slope errors using the real mirror profiles measures at the ESRF metrology lab.
-It can be shown a further degradation and the growing of satellite structures in the vertical intensity distribution.
-Values of the histograms FWHM are displayed in the graphic titles.
+It shows the focal spot degradation due to the diffraction arising from the beam aperturing by the KB mirrors.
+The calculations in the bottom row also include slope errors using the real mirror profiles measures at the ESRF metrology lab.
+Further beam broadening and the growth of satellite structures in the vertical intensity distribution can be observed.
+Values of the histograms FWHM are displayed in the graph titles \todo{improve legibility of titles - too small}.
 }
@@
 
-\ack{\todo{someone to thank?}}
+\ack{\todo{someone to thank?
+- RC: before the PHANGS in Trieste I was contacting Oleg about the elliptical mirrors.
+There was a bug there and he corrected it.
+Only after fixed it we were able to simulate the ID16A with the mirror elements.
+So maybe something mentioning some disucssions on the early stages of the SRW simulations.}}
 
@@
 These profiles are used along the text in ray tracing and wave optics simulations.
-
+\todo{
+RC: maybe plot the PSD for the three mirrors? Put them on the same scale to show how nice the KB's are?
+ }
 \begin{figure}\label{fig:metrology}

### 2019-05-07 paper-hierarchical.tex
Updates from Overleaf

@@
 The wave optics section~\ref{level2} includes different models of increasing complexity as the required accuracy of the results becomes more demanding.
-The simulation methods are presented in a hierarchical approach:  they provide more and better results but the price to pay is to deal with more complex and computationally expensive tools.
+\inblue{In the Discussion (Section \ref{discussion}) the results of the different simulation methods are commented.
+The hierarchical approach allows more accurate and better results but the price to pay is to deal with more complex and computationally expensive tools.
+Finally Section \ref{summary} summarizes the results.}
 
@@
 \todo{please summarize some use and application}.
-The beamline can operate at three fixed photon energies: 11200 eV, 17225 eV and 33600 eV.
-The straight section in the magnet lattice houses three undulators: an in-vacuum undulator of $\lambda_u=26$~mm period, and two other conventional undulators of 18 and 22.4 mm period.
+The beamline can operate at two fixed photon energies: 17050 eV and 33600 eV.
+The straight section in the magnet lattice houses four undulators, two 18 mm and another two of 22.4 mm period.
 The undulator parameters and $K$-values at the resonances for the energies in use are shown in Table \ref{tab:Undulators}.
@@
 \begin{tabular}{lcccccc}
-name  & $\lambda_u$ [mm] & $L$ [m] & N & $K$ (11200 eV) & $K$ (17225 eV) & $K$ (33600 eV) \\ 
+name  & $\lambda_u$ [mm] & $L$ [m] & N & $K$ (11200 eV) & $K$ (17050 eV) & $K$ (33600 eV) \\ 
 \input{tableUndulator.txt}
@@
 
-For all simulations presented hereafter, we selected the configuration using a single undulator U18.3 placed in the center of the straight section and tuned to its first harmonic at a photon energy of $E=17225$ eV (Table \ref{tab:Undulators}).
+For all simulations presented hereafter, we selected the configuration using a single undulator U18.3 placed in the center of the straight section and tuned to its first harmonic at a photon energy of $E=17225$ eV (\inblue{this value is not identical but close to the usual operation value shown in }Table \ref{tab:Undulators}).
 This section is divided in three subsections, that present three main levels of simulations: simple calculations that can be made by hand (section \ref{level0}), ray-tracing simulations (section \ref{level1}) and wave optics simulations (section \ref{level2}).
@@
     \label{eq:photon small sigmas}
-    \sigma_u=\frac{2.74}{4\pi}\sqrt{\lambda L}\approx \sqrt{\frac{\lambda L}{2 \pi^2}}  && \sigma_u' = 0.69\sqrt{\frac{\lambda}{L}}\approx \sqrt{\frac{\lambda}{2 L}},
+    \sigma_u=\frac{2.74}{4\pi}\sqrt{\lambda L}\approx \frac{1}{2  \pi}\sqrt{2 \lambda L}  && \sigma_u' = 0.69\sqrt{\frac{\lambda}{L}}\approx \sqrt{\frac{\lambda}{2 L}},
 \end{align}
@@
 The HYBRID method \cite{hybrid} uses geometric ray tracing combined with wavefront propagation.
-The code computes diffraction effects when the beam is clipped by an aperture or by the finite size of the optics and can also simulate the effect of mirror figure errors when diffraction is present.
+The code computes diffraction effects when the beam is apertured and can also simulate the effect of mirror figure errors when diffraction is present.
 The method can be applied to an entire beamline by simulating each optical element iteratively under the hypothesis of the validity of far field propagation in all intermediate propagations.
@@
 When compared with the pure ray tracing in Fig.~\ref{fig:ray tracing}a it can be appreciated that the spot is broadened.
-This is more evident in the vertical direction, because the smaller focal size.
+This is more evident in the vertical direction, because of the smaller focal size.
 The spot size is about 7-8 nm in V for both storage ring lattices, and in horizontal is about 15$\mu$m for ESRF-High$\beta$ and 11 nm for EBS.
@@
 
-The surface finish of the optical elements is another limiting effect for determining the size and intensity of the beam at the image position.
-It is possible to infere the effect of mirror imperfections and in particular slope errors by only knowing some statistical parameters such as the slope error RMS and then building synthetic surface profiles to be used in the simulations.
-This is the typical mode to operate when designing a new beamline where the mirrors do not exist yet.
-However, in our case the mirrors already exist and have been extensively characterize at the ESRF metrology laboratory.
+The surface figure of the optical elements is another limiting effect for determining the size and intensity of the beam at the image position.
+It is possible to infer the effect of mirror imperfections and in particular slope errors by only knowing some statistical parameters such as the RMS slope error  and then building synthetic surface profiles to be used in the simulations.
+This is the typical approach used when designing a new beamline where the mirrors do not exist yet.
+However, in our case the mirrors already exist and have been extensively characterized at the ESRF optical metrology laboratory.
 It is therefore highly recommended to use metrology data from real mirrors, when possible.
@@
 These experimental data are used in all the simulations in this paper.
-Simulations using the hybrid model including slopes errors are shown in Fig.~\ref{fig:ray tracing}c.
-It can be observed a small broadening of the peak due to slope errors, but, more importantly, a new satellite structure close to the vertical intensity peak is produced by the slope errors.
+Simulations using the hybrid model including slope errors are shown in Fig.~\ref{fig:ray tracing}c.
+A small broadening of the peak due to slope errors can be observed, but, more importantly, a new satellite structure close to the vertical intensity peak is produced by the slope errors.
 
@@
 
-In this section the coherence properties of the beamline are analysed using, again, different methodologies classified by complexity and computer resources.
+In this section the coherence properties of the beamline are analysed using, again, different approaches  distinguished by their complexity and necessary computer resources.
 We first analyze the beamline supposing a fully coherent beam.
-Subection~\ref{wofry} shows a very simpified 1D model.
-Subsection~\ref{srw_se} used a full 2D model with accurate simulation of the undulator emission but without considering storage ring emittance.
-These simplifications permit to address diffraction effects by apertures and slope errors.
-Then we introduce methods to treat the partial coherence originated by the storage ring emittance.
-Two variants are shown, a Monte Carlo approach (subsection~\ref{srw_me}) that permits us to calculate in great detail the intensity dostribution of the focused beam, and a decomposition of the radiation in coherent modes (subsection~\ref{comsyl}) that allows to evaluate quantitatively the quality of the coherent beam.
+Subection~\ref{wofry} shows a very simplified 1D model.
+Subsection~\ref{srw_se} used a full 2D model with accurate simulation of the undulator emission but without considering storage ring emittance (i.e.
+again using a fully coherent beam).
+These simplifications permit the simulation of diffraction effects by apertures and slope/figure errors.
+Subsequently we introduce methods to treat the partial coherence originated by the storage ring emittance.
+Two variants are shown, a Monte Carlo approach (subsection~\ref{srw_me}) that permits us to calculate in great detail the intensity distribution of the focused beam, and a decomposition of the radiation in coherent modes (subsection~\ref{comsyl}) that allows   quantitative evaluation of the coherent beam characteristics.
 
@@
 
-The very long distance source-sample combined with a small source (in particular for EBS) makes this beamline suitable for experimental techniques requiring a highly coherent beam.
+The very long  source-{\color{green}\sout{sample} optics} distance combined with a small source (in particular for EBS) makes this beamline suitable for experimental techniques requiring a highly coherent beam.
 Indeed, it has been shown that the clipping of the beam, mainly due to the acceptance of the KB mirrors produces a significant broadening of the focal spot.
-As discussed in Section~\ref{level0}, very simple calculation of the diffraction limit can be sone using a convergent spherical beam clipped by a slit of 400$\times$400$ \mu$m$^2$ placed at 5cm upstream the focus (at the second KB mirror position).
+As discussed in Section~\ref{level0}, very simple calculation of the diffraction limit can be done using a convergent spherical beam clipped by a slit of 400$\times$400$ \mu$m$^2$ placed  5cm upstream the focus (at the second KB mirror position).
 The slit vertical size corresponds roughly to the horizontal acceptance of the seconf KB mirror, and it is approximately the same in vertical.
@@
 \section{Summary and conclusions}
+\label{summary}
 

### 2019-05-07 paper-hierarchical.tex
Updates from Overleaf

@@
 Indeed, it has been shown that the clipping of the beam, mainly due to the acceptance of the KB mirrors produces a significant broadening of the focal spot.
-As discussed in Section~\ref{level0}, very simple calculation of the diffraction limit can be done using a convergent spherical beam clipped by a slit of 400$\times$400$ \mu$m$^2$ placed  5cm upstream the focus (at the second KB mirror position).
-The slit vertical size corresponds roughly to the horizontal acceptance of the seconf KB mirror, and it is approximately the same in vertical.
+As discussed in Section~\ref{level0}, very simple calculation of the diffraction limit can be done using a convergent spherical beam clipped by a slit of 400$\times$400$ \mu$m$^2$ placed  5cm upstream of the focus (at the second KB mirror position).
+The slit vertical size corresponds roughly to the horizontal acceptance of the second KB mirror, and it is approximately the same in vertical.
 A simple simulation of diffraction by a square slit using OASYS is shown in Fig.~\ref{wofry400um}.
-It reveals a spot of less than 10 nm, thus confirming the analytical predictions, and qualifying this beamline to be a diffraction limited beamline.
+It reveals a spot of less than 10 nm, thus confirming the analytical predictions, and qualifying this beamline to be a diffraction-limited beamline.
 
@@
 Each optical element can be simplified by separating its action in two parts: the focusing behaviour implemented as an ideal lens, and the finite size because of the boundaries, implemented as a slit.
-The WOFRY package in OASYS permits to do this work, and it is possible to arrive to a satisfactory result in a few iterations.
+The WOFRY package in OASYS permits such simulations, and it is possible to achieve satisfactory results in a few caluclation iterations.
 The beam size at different positions of the beamline are shown in Table~\ref{tablewofry1D}.
@@
 
-A next step in the wave optics simulation is, maintaining the full coherence, to generate a 2D source wavefront that matches in detail the real source (an undulator in a storage ring) and then simulate the propagation of such a wavefront through the optical elements with models that match well the characteristics of the real elements.
+A next step in the wave optics simulation is, {\color{green}\sout{maintaining the} assuming} full coherence, to generate a 2D source wavefront that matches in detail the real source (an undulator in a storage ring) and then simulate the propagation of such a wavefront through the optical elements with models that match well the characteristics of the real elements.
 The source simulated is the on-axis-single-electron emission from the U18.3 undulator tuned to its first harmonic at $E=17225$ eV.
 The emission by a single-electron is fully coherent.
-Once the source is defined, one has to go around on how to simulate the beamline optical elements.
-Increasing the complexity of the simulations one can: a) use ideal thin lenses preceded by slits to emulate finite aperture (as done in the 1D model presented in Section~\ref{wofry}); or b) use grazing incidence focusing optics models \cite{Canestrari2014}.
-On the top of that, one can add the effect of imperfect optics to the simulations, i.e.
-slope errors.
+Once the source is defined, an approach of how best to simulate the beamline optical elements.
+Increasing the complexity of the simulations one can: a) use ideal thin lenses preceded by slits to emulate finite apertures (as done in the 1D model presented in Section~\ref{wofry}); or b) use grazing incidence focusing optics models \cite{Canestrari2014}.
+Furthermore, one can introduce the effect of imperfect optics to the simulations, i.e.
+slope/figure errors.
 
@@
 It can been seen that there is little difference between using thin ideal lenses or the elliptical mirror model.
-The elliptical mirror presents a slight reduction in intensity and asymmetry but no significant difference is found, which is in accordance with what is reported in \cite{Canestrari2014}.
-Notice that the differences are minimal and to remark them we have chosen log plots to represent the intensity distributions at the focal position.
-It is also important to mention that this equivalency between ideal lenses and elliptical mirrors is only valid at the focal position vicinity: down- or upstream the focal position, differences are not negligible \cite{Canestrari2014}.
-
-It is also possible to add slope errors to the simulations to study the degradation on the focal spot quality (Fig.~\ref{fig:SingleElectron}).
-For that, the available available metrology data (see Appendix \ref{appendix:metrology})  are used.
+The elliptical mirror presents a slight reduction in intensity and asymmetry but no significant differences are observed, which is in accordance with  the findings of \cite{Canestrari2014}.
+Notice that the differences are minimal and to accentuate them we have chosen logarithmically scaled plots to represent the intensity distributions at the focal position.
+It is also important to mention that this equivalency between ideal lenses and elliptical mirrors is only valid close to the focal position: down- or upstream the focal position, differences are no longer negligible \cite{Canestrari2014}.
+
+It is also possible to add slope errors to the simulations to study the resulting impact upon the focal spot quality (Fig.~\ref{fig:SingleElectron}).
+For that, the available surface metrology data of the reflective elements (see Appendix \ref{appendix:metrology})  are used.
 For both cases (i.e., simulating the beamline using thin optical elements or elliptical mirrors), the beam sizes calculated as the FWHM at the focal position are little affected by the presence of the slope errors (see Table \ref{tab:SRW_results}).
-There is a significant reduction and shuffling of the interference fringes observed with no slope error.
-They are almost cancelled, with is an important raise of the background level: the radiation found on the main lobe (focus) shifts to the side lobes when the slope errors are considered.
+There is a significant reduction and {\color{green}\sout{shuffling} blurring} of the interference fringes compared to observations with no slope error.
+They are almost completely removed and associated with a large increase of the background level: the radiation found on the main lobe (focus) shifts to the side lobes when the slope errors are considered.
 This is accompanied by an obvious reduction on the peak intensity.
@@
 Left: the perfect focusing with ideal lenses (black) is compared with the focusing using elliptical mirrors (red).
-Right: the same as in the previous plot, but adding the effect of the mirrors slope errors in bothe cases.}
-    %Simulations took less than one minute in a desktop computer Intel Core i7 with 8Gb of RAM memory.}
+Right: the same as in the previous plot, but adding the effect of the mirrors slope errors in both cases.}
+    %Simulations took less than one minute using a desktop computer Intel Core i7 with 8Gb of RAM memory.}
 \end{figure}
@@
 
-\subsubsection{Full wave optics simulation: partial coherent case by means of multielectron Monta Carlo sampling}
+\subsubsection{Full wave optics simulation: partial coherent case by means of multielectron Monte Carlo sampling}
 \label{srw_me}
 The wavefront propagation methodology is based in the sampling of a single wavefront (monochromatic and coherent) in a 2D space (in horizontal and vertical spatial dimensions) at a given point of the beamline.
-This initial point considered as source cannot be the source physical position, because i) when condidering point or spherical waves all the intensity would come in a single pixel so the wavefront is ill-defined, ii) when using synchrotron sources (i.e., undulators) theory \cite{jackson} permits the calculation of the radiation in positions far from the electron trajectory.
+This initial point considered as source cannot be the source physical position, because i) when considering point or spherical waves all the intensity would be concentrated in a single pixel so the wavefront is ill-defined, ii) when using synchrotron sources (i.e., undulators) theory \cite{jackson} permits the calculation of the radiation in positions far from the electron trajectory.
 The limitation of the method is that one wavefront is not representative for a beam that, in storage rings, is produced by the emission of many electrons circulating through the corresponding magnet structure.
@@
 There are three approaches currently used: 1) use of convolutions, 2) Monte Carlo multi-electron sampling, and 3) Propagation of coherent modes.
-The convolution method consist in calculating the emission characteristics and then convolve the resulting intensity with Gaussians describing the electron beam.
+The convolution method consists of calculating the emission characteristics and then convolving the resulting intensity with Gaussians describing the electron beam.
 This method is simple and fast, as only requires simulation of a single wavefront.
 However, it presents two important limitations: i) it can only be applied to the intensity calculations and not to other parameters of interest for describing coherence, such as degree of coherence, etc., and ii) it can be applied only at the source position: it is usually unknown how the different optical elements will affect the source contribution to the phase space.
-The next method, the Monte Carlo multi-electron propagation, it is an statistical method based on the idea that one can exactly calculate the emission and its propagation for one electron that is deterministically described by its initial conditions at the entrance of the magnetic structure.
+The next method, the Monte Carlo multi-electron propagation,  is a statistical method based on the idea that one can exactly calculate the emission and its propagation for one electron that is deterministically described by its initial conditions at the entrance of the magnetic structure.
 It is therefore possible to perform many simulations for different initial conditions of the travelling electron and make the propagation along the beamline until the observation plane.
@@
 
-The same observations made on the previous section (full coherent simulations) are pertinent here: there is a reduction on peak intensity when comparing single lens model with the elliptical mirror.
+The same observations made in the previous section (fully coherent simulations) are pertinent here: there is a reduction of the peak intensity when comparing simple lens model with the elliptical mirror.
 The beam sizes and shapes are similar at the focal position, only.
+{\todo{more detailed comments on comparisons with filament beam results? For comparison also information on calculation time, number of electrons, ...}}
 
@@
 In other words, truncating the series in Eq.~\ref{decomposition} up to the term $m$, the CSD obtained approximates better the exact CSD (this with infinite terms) than any other expansion.
-It also imply that the first term is more important than the second, and so on.
+It also implies that the first term is more important than the second, and so on.
 The first or lower term (index zero) is the most important coherent mode.
-It permits to define the coherent fraction (CF) as the occupation of the lower coherent mode: CF=$d_0$.
-Obviously, a pretty coherent source will require less modes than a mostly incoherent source to arrive up to a certain representation of the CSD.
+It permits definition of the coherent fraction (CF) as the occupation of the lower coherent mode: CF=$d_0$.
+Obviously, a more coherent source will require less modes than a mostly incoherent source to approximate the CSD to a given precision.
 
@@
 Fig.~\ref{fig:histomodes}a shows the cumulated occupation of the coherent modes for the full central cone undulator emission.
-It shows that the EBS cumulated occupation grows very steply with the number of coherent modes, whereas for ESRF the number of coherent modes needed to attain a given value of cumulated occupation much higher.
+It shows that the EBS cumulated occupation grows very steeply with the number of coherent modes, whereas for ESRF the number of coherent modes needed to attain a given value of cumulated occupation much higher.
 This can also be shown in Fig.\ref{fig:histomodes}b, where the number of modes to fill 50\%, 75\% and 95\% of spectral density are compared for EBS and ESRF.
@@
 Figs.~\ref{fig:spectraldensity}b-e represent the coherent modes with $m$=0 to $m$=3.
-It can be shown that the first mode is centered on the axis but extends over an spatial region smaller than the spectral density.
-Next mode ($m$=1) contributes to extend the spectral density in horizontal, but presents zero intensity at the origin.
-Successive modes extend  the spectral density in horizontal, sometimes presenting a lobe centerer at zero and sometimes not (e.g., $m$=3).
+It can be shown that the first mode is centered on the axis but extends over a spatial region smaller than the spectral density.
+The next mode ($m$=1) contributes to extend the spectral density in horizontal, but presents zero intensity at the origin.
+Successive modes extend  the spectral density in horizontal, sometimes presenting a lobe center at zero and sometimes not (e.g., $m$=3).
 Modes with several lobes along the vertical will also appear, but not so often than in the horizontal, because of the elliptical shape of the spectral density.
@@
 The beamline elements modify the different modes in a different manner, but it will never be possible to remove all the photons except from the lowest coherent mode, because there is an overlapping of the intensity of the lowest mode with the other ones.
-Therefore it is impossible to obtain a fully coherent beam that must contain only a single mode.
+Therefore it is impossible to obtain a fully coherent beam containing only a single mode.
 
@@
 
-The coherent modes calculated at the EBS source have been propagated using a beamline simplified with ideal focusing elements and slits to determine their numerical aperture (like in Section~\ref{wofry}).
+The coherent modes calculated at the EBS source have been propagated \todo{Using which propagation code - is this all in COMSYL?} using a simplified beamline  with ideal focusing elements and slits to determine their numerical aperture (like in Section~\ref{wofry}).
 Fig.~\ref{fig:mode transmission} shows how the intensity of the first modes is reduced by geometrical considerations.
 %We recall that in the coherent mode expansion, the modes are orthonormal, therefore the integrated intensity for any mode is always one at the source. 
-As discussed, coherent modes will be ``cut'' in a different manner, depending on the slits cropping effect.
-This imply that the Eq.~\ref{eq:mode normalization} is not longer verified, and for the propagated modes $M_{mm}\le$1 and $M_{mn}\ne$0 for $m\ne n$.
+As discussed, coherent modes will be ``cut'' in a different manner, depending on the slits aperturing effect.
+This implies that the Eq.~\ref{eq:mode normalization} is not longer {\color{green}\sout{verfified} satisfied}, and for the propagated modes $M_{mm}\le$1 and $M_{mn}\ne$0 for $m\ne n$.
 Therefore, the coherent modes of the source after propagation and cropping by the beamline are no longer coherent modes of the resulting beam.
@@
 The other modes shown transfer less than 5\% transfer, and modes higher than $m$=10 practically do not contribute to the final spectral density.
-Note that, as discussed before, the transmission of mode $m$=2 is higher than mode $m$=1, because this one does not contain a central lobe (Fig.~\ref{fig:spectraldensity}c).
+Note that, as discussed before, the transmission of mode $m$=2 is higher than mode $m$=1, because this one does not display a central lobe (Fig.~\ref{fig:spectraldensity}c).
 
@@
 
-This analysis of the mode transmission is mode dramatic if one looks to the total intensity in the mode, which is given by its eigenvalue.
+This analysis of the mode transmission is more dramatic if one looks at the total intensity in the mode, which is given by its eigenvalue.
 Neglecting modes higher than 10, the lowest coherent mode ($m$=0) carries the 49.5\% of the beam intensity, 18.4\% for the mode $m$=2, and less than 5\% for the others.

### 2019-06-11 paper-hierarchical.tex
updated Ray's corrections

@@
 We consider different approaches to simulate a modern X-ray beamline.
-Several methodologies with increasing complexity are applied to discuss the relevant parameters that quantify the beamline {\color{green}\sout{performances}performance}.
-Parameters such as flux, dimensions and intensity distribution of the focused beam and coherence properties are obtained from simple analytical calculations to sophisticated computer simulations using ray tracing and {\color{red}\sout{waveoptics}wave optics} techniques.
-A {\color{green}\sout{last}latest} generation x-ray nanofocusing beamline for coherent  applications (ID16A at the ESRF) has been chosen  to study in detail the issues related to highly demagnifying  {\color{green}\sout{sychrotron}synchrotron} sources and exploiting the beam coherence.
-The performance of the beamline is studied for two storage{\color{green}\sout{s}} rings: the old ESRF (emittance 4000 pm) and the new ESRF-EBS (emittance 150 pm).
+Several methodologies with increasing complexity are applied to discuss the relevant parameters that quantify the beamline performance.
+Parameters such as flux, dimensions and intensity distribution of the focused beam and coherence properties are obtained from simple analytical calculations to sophisticated computer simulations using ray tracing and wave optics techniques.
+A latest generation x-ray nanofocusing beamline for coherent  applications (ID16A at the ESRF) has been chosen  to study in detail the issues related to highly demagnifying  synchrotron sources and exploiting the beam coherence.
+The performance of the beamline is studied for two storage rings: the old ESRF-1 (emittance 4000 pm) and the new ESRF-EBS (emittance 150 pm).
 In addition to traditional results in term of flux and beam sizes, an innovative study on the partial coherence properties based on propagation of coherent modes is presented.
 The different algorithms and methodologies are implemented in the software suite OASYS \cite{codeOASYS}.
-They are discussed with emphasis beamline {\color{green}\sout{in}placed upon the } benefits and limitations of each.
+They are discussed with emphasis beamline placed upon the benefits and limitations of each.
 \end{abstract}
@@
 The first upgrade of a large facility will be the EBS (Extremely Brilliant Source) \cite{orangebook} at the European Synchrotron Radiation Facility (ESRF).
-The new storage ring of 150 pm emittance that will be commissioned in 2020 will significantly boost the  {\color{green}\sout{x-ray}} coherence of the  {\color{green}\sout{synchrotron} x-ray} beams.
+The new storage ring of 150 pm emittance that will be commissioned in 2020 will significantly boost the coherence of the x-ray beams.
 Applications exploiting beam coherence such as x-ray photon correlation spectroscopy, coherent diffraction imaging, propagation-based phase contrast imaging and ptychography will strongly benefit from this update.
 
-Accurate calculation and quantitative evaluation of the parameters related to x-ray optics is of {\color{green}\sout{paramount} great} importance for designing, building and exploiting {\color{green}\sout{the}} new beamlines.
+Accurate calculation and quantitative evaluation of the parameters related to x-ray optics is of great importance for designing, building and exploiting new beamlines.
 Every modern beamline follows a procedure of conceptual design plus technical design where detailed simulations are essential.
-These calculations  {\color{green}\sout{permit to test} allowing testing of} the design parameters in a virtual computer environment where advantages and limitations can be studied quantitatively.
-{\color{green}\sout{The selection to best performant optics configuration} Selection of the optimal optical layout}  requires the detailed simulation of the optics imperfections, that in many cases are the limiting factor of the beamline optical configuration.
-{\color{green}\sout{Deformation in} Deformations of} the optical elements due to heat load need to be controlled, therefore optics simulations have to include results from engineering modeling of the thermal deformations {\color{green}\sout{associated to mechanical issues}} usually {\color{red}\sout{dome}done} using finite {\color{red}\sout{e}}element analysis.
+These calculations  allowing testing of the design parameters in a virtual computer environment where advantages and limitations can be studied quantitatively.
+Selection of the optimal optical layout requires the detailed simulation of the optics imperfections, that in many cases are the limiting factor of the beamline optical configuration.
+Deformations of the optical elements due to heat load need to be controlled, therefore optics simulations have to include results from engineering modeling of the thermal deformations usually done using finite element analysis.
 Traditionally the optics calculations for synchrotron beamlines are done using ray tracing techniques.
-They model the x-ray beam as a collection of {\color{green}\sout{different}incoherent} rays.
-{\color{green}\sout{that are incoherent one to another}}  This technique is still essential for obtaining information on aberrations, flux propagation, monochromator bandwidth, etc.
-However, the high coherence of the new sources {\color{green}\sout{make necessary}can introduce the necessity} to complement the ray tracing methods with models including effects of diffraction and scattering from a coherent beam.
-A fully coherent {\color{green} hard x-ray}beam is still a dream for the new generation storage rings.
-In spite of a {\color{green}\sout{high increate of }considerable increase in}  the coherent fraction of the new beams, the coherent fraction at {\color{red}typically\sout{tipically}} hard x-ray photon energies is still a few percent for the new sources.
-{\color{green}\sout{This makes still important a } Consequently} ray tracing analysis {\color{green}\sout{that}} does not become obsolete, but has to be complemented with an analysis based on wave optics.
-Because of the {\color{green}\sout{non perfect} partial} coherence the wave optics analysis becomes non trivial and can be done with different levels of approximation, as discussed in this work.
-In addition, wave optics simulations are usually computationally expensive.\todo{I noticed that there is very few citations to references in the introduction.
-This is not so nice, as we don't knowledge any previous work.
-Maybe one could add a few citations in the introduction.}
-
-To illustrate the applicability of a variety of methods and solutions for beamline optical simulations we have chosen the ESRF beamline ID16A~{\color{red}\cite{ID16A}}.
+They model the x-ray beam as a collection of incoherent rays.
+This technique is still essential for obtaining information on aberrations, flux propagation, monochromator bandwidth, etc.
+However, the high coherence of the new sources can introduce the necessity to complement the ray tracing methods with models including effects of diffraction and scattering from a coherent beam.
+A fully coherent hard x-ray beam is still a dream for the new generation storage rings.
+In spite of a considerable increase in the coherent fraction of the new beams, the coherent fraction at typically hard x-ray photon energies is still a few percent for the new sources.
+Consequently ray tracing analysis does not become obsolete, but has to be complemented with an analysis based on wave optics.
+Because of the partial coherence the wave optics analysis becomes non trivial and can be done with different levels of approximation, as discussed in this work.
+In addition, wave optics simulations are usually computationally expensive.
+
+To illustrate the applicability of a variety of methods and solutions for beamline optical simulations we have chosen the ESRF beamline ID16A~\cite{ID16A}.
 This beamline provides a high-brilliance beam focused to a few nanometers.
 It displays characteristics increasingly aimed for in modern beamlines: extremely high demagnification (sub-micrometer focused beam size) and high coherence.
-It combines coherent imaging techniques and X-ray fluorescence microscopy to perform quantitative 3D characterization of the morphology and the elemental composition of specimens in their native state~{\color{red}\cite{XRFID16A}}.
+It combines coherent imaging techniques and X-ray fluorescence microscopy to perform quantitative 3D characterization of the morphology and the elemental composition of specimens in their native state~\ingreen{\cite{XRFID16A}}.
 
 X-ray beamlines are very particular optical systems.
-Some implications of exploiting the {\color{red}\sout{low}short} wavelengths such as the use of grazing incidence reflectors and the use of perfect crystals as typical hard x-ray monochromators render common commercial software packages ill-adapted for the study of synchrotron beamlines.
+Some implications of exploiting the short wavelengths such as the use of grazing incidence reflectors and the use of perfect crystals as typical hard x-ray monochromators render common commercial software packages ill-adapted for the study of synchrotron beamlines.
 For this reason many optics tools have been developed by the synchrotron community.
@@
 The wave optics section~\ref{level2} includes different models of increasing complexity as the required accuracy of the results becomes more demanding.
-\inblue{In the Discussion (Section \ref{discussion}) the results of the different simulation methods are commented.
+In the Discussion (Section \ref{discussion}) the results of the different simulation methods are commented.
 The hierarchical approach allows more accurate and better results but the price to pay is to deal with more complex and computationally expensive tools.
-Finally Section \ref{summary} summarizes the results.}
+Finally Section \ref{summary} summarizes the results.
 
@@
 This section presents the reader with the main operational parameters for the ID16A beamline at ESRF in two scenarios.
-This beamline was built during the Upgrade Phase I at ESRF in \inred{2011??}.
+This beamline was built during the Upgrade Phase I at ESRF (2009-2015).
 It uses one or several undulators as source.
-During the period 2011-2018 it operated exploiting the ESRF storage ring’s double-bend achromat (DBA) lattice with horizontal emittance about 4000 pm.
+During the period 2011-2018 it operated exploiting the ESRF-1 storage ring’s double-bend achromat (DBA) lattice with horizontal emittance about 4000 pm.
 In the second phase of the ESRF Upgrade the original storage ring is being replaced by a new hybrid multi-bend achromat (HMBA) lattice.
@@
 It is dedicated to nanofocusing applications.
-\todo{please summarize some use and application}.
 The beamline can operate at two fixed photon energies: 17050 eV and 33600 eV.
@@
 \caption{Parameters of the undulators available at the ID16A beamline: period ($\lambda_u$), length ($L$), number of identical undulators ($N$) and deflection parameter ($K$) values for the three main photon energies in use.
-The valuse are shown for the two ESRF magnetic lattices: the old one (labelled ESRF, with $E_e = 6.04$ GeV) and the new one (EBS, with $E_e = 6.00$ GeV).
+The valuse are shown for the two ESRF magnetic lattices: the old ESRF-1 (labelled ESRF, with $E_e = 6.04$ GeV) and the new ESRF-EBS (labelled EBS, with $E_e = 6.00$ GeV).
 $n$ is the emission harmonic in use for each particular photon energy.}
@@
 
-Simulations were done using the two different storage ring lattices mentioned before: the (already dismantled) ESRF DBA where the undulators are in a high-$\beta$ straight section (larger horizontal emittance) and the new EBS (lower horizontal emittance) straight sections.
+Simulations were done using the two different storage ring lattices mentioned before: the (already dismantled) ESRF-1 where the undulators are in a high-$\beta$ straight section (larger horizontal emittance) and the new EBS (lower horizontal emittance) straight sections.
 The parameters for the current and the new lattice are shown in Tables \ref{tab:TableSources} and \ref{tab:eBeam}.
@@
 \begin{table}\label{tab:TableSources}
-    \caption{ESRF and ESRF-EBS main storage ring parameters. Main parameters of the present DBA lattice (ESRF) and future HMBA lattice (EBS) storage rings \cite{orangebook}.}
+    \caption{ESRF and ESRF-EBS main storage ring parameters. Main parameters of the present DBA lattice (ESRF-1) and future HMBA lattice (EBS) storage rings \cite{orangebook}.}
         \begin{tabular}{rcc}
-                                              & \textbf{ESRF} & \textbf{EBS}      \\ 
+                                              & \textbf{ESRF-1} & \textbf{EBS}      \\ 
         Lattice type:                         & DBA           & HMBA              \\
@@
     \centering
-    \caption{ESRF and EBS main electron beam parameters.
-Electron beam parameters for the high-$\beta$ (ESRF) and the EBS straight sections.
+    \caption{ESRF-1 and EBS main electron beam parameters.
+Electron beam parameters for the high-$\beta$ (labelled ESRF) and the EBS straight sections.
 Values taken at the symmetry point of the straight section, where the insertion devices are placed.
@@
 
-For all simulations presented hereafter, we selected the configuration using a single undulator U18.3 placed in the center of the straight section and tuned to its first harmonic at a photon energy of $E=17225$ eV (\inblue{this value is not identical but close to the usual operation value shown in }Table \ref{tab:Undulators}).
+For all simulations presented hereafter, we selected the configuration using a single undulator U18.3 placed in the center of the straight section and tuned to its first harmonic at a photon energy of $E=17225$ eV (a value close to the usual operation value shown in Table \ref{tab:Undulators}).
 This section is divided in three subsections, that present three main levels of simulations: simple calculations that can be made by hand (section \ref{level0}), ray-tracing simulations (section \ref{level1}) and wave optics simulations (section \ref{level2}).
@@
 
-The source size (or divergence) is usually approximated supposing a Gaussian distribution of the electron {\color{green}\sout{sizes}positions} (or divergence) (Table \ref{tab:eBeam}) and {\color{green}\sout{undulator natural}single electron photon} source size or divergence.
+The source size (or divergence) is usually approximated supposing a Gaussian distribution of the electron positions (or divergence) (Table \ref{tab:eBeam}) and single-electron photon source size or divergence.
 In some cases, such as for undulator emission, this Gaussian approximation may not be accurate and have some implications that will be discussed later.
@@
 
-Using the values for U18.3, $n$=1, and  $E=17225$ eV (Table \ref{tab:Undulators}) and the optical configuration described in Table \ref{tab:Positions}, one obtains a focal size 139$\times$5 nm$^2$ (H$\times$V) for ESRF and 10$\times$5 nm$^2$ for EBS (Table~\ref{tab:HandCalculations}).
-The horizontal focal spot for the ESRF lattice is larger than the specified target value for the beamline, therefore the VSS could be closed to 50$\mu$m thus reducing the beamsize at the VSS plane by a factor ca.
-8 \inred{$\approx$ 403.8/50=8.076}.
-Therefore, for ESRF, the horizontal size at the sample will be affected by the same factor obtaining roughly 17$\times$5 nm$^2$
+Using the values for U18.3, $n$=1, and  $E=17225$ eV (Table \ref{tab:Undulators}) and the optical configuration described in Table \ref{tab:Positions}, one obtains a focal size 139$\times$5 nm$^2$ (H$\times$V) for ESRF-1 and 10$\times$5 nm$^2$ for EBS (Table~\ref{tab:HandCalculations}).
+The horizontal focal spot for the ESRF-1 lattice is larger than the specified target value for the beamline, therefore the VSS could be closed to 50$\mu$m thus reducing the beamsize at the VSS plane by a factor ca.
+8.
+%\inred{8 $\approx$ 403.8/50=8.076}
+Therefore, for ESRF-1, the horizontal size at the sample will be affected by the same factor obtaining roughly 17$\times$5 nm$^2$
 
@@
     \newline
-    $FWHM'_h$ [$\mu$rad]:   & 26.96         & \inred{65.22?}        & 189062   & 15.60   & 37.74   & 109404       \\
+    $FWHM'_h$ [$\mu$rad]:   & 26.96         & 65.22        & 189062   & 15.60   & 37.74   & 109404       \\
     $FWHM'_v$ [$\mu$rad]:   & 12.04         & 12.04                 & 22264    & 12.16   & 12.16   & 22489        \\                                                   
@@
 The XOPPY package in the OASYS suite implements three codes: US \cite{codeUS}, URGENT \cite{codeURGENT} and SRW.
-Figure~\ref{fig:FluxU18} shows the spectra calculated \todo{indicate which code used} over an aperture of 1 $\times$ 1 mm$^2$ placed at $d$=27.1 m.
+Figure~\ref{fig:FluxU18} shows the spectra calculated using SRW over an aperture of 1 $\times$ 1 mm$^2$ placed at $d$=27.1 m.
 This aperture opening is set to fully accept the central cone at the resonance peak of the harmonic in use ($\max(\Sigma') \times d = 0.73$ mm), with $\Sigma'$ values from in Table~\ref{tab:HandCalculations}.
-The peak intensity is about  $2.9\cdot10^{14}$ photons/s/0.1{\%}bw for EBS and 1.9 10$^{14}$ for ESRF (see \ref{fig:FluxU18}).
+The peak intensity is about  $2.9\cdot10^{14}$ photons/s/0.1{\%}bw for EBS and 1.9 10$^{14}$ for ESRF-1 (see \ref{fig:FluxU18}).
 One can appreciate how the emittance plays an important role in both peak width and specially in peak value.
-The number of photons per second after the multilayer monochromator is 2.9 10$^{15}$ for EBS and 1.9 10$^{15}$ for ESRF, considering that the energy bandwidth transmitted by a multilayer monochromator is $\Delta E/E \approx 10^{-2}$ and the bandwidth in Fig.~\ref{fig:FluxU18} is 10$^{-3}$, and supposing, at this point, ideal reflectivity (100\%) for the multilayers.
+The number of photons per second after the multilayer monochromator is 2.9 10$^{15}$ for EBS and 1.9 10$^{15}$ for ESRF-1, considering that the energy bandwidth transmitted by a multilayer monochromator is $\Delta E/E \approx 10^{-2}$ and the bandwidth in Fig.~\ref{fig:FluxU18} is 10$^{-3}$, and supposing, at this point, ideal reflectivity (100\%) for the multilayers.
 
@@
 
-To estimate the geometrical {\color{red}\sout{transmitivity}transmission?} of the beamline we have to consider which optical elements crop the beam and remove photons.
+To estimate the geometrical transmission of the beamline we have to consider which optical elements crop the beam and remove photons.
 This can be analysed using their angular acceptance or numerical aperture.
@@
 We consider a VSS closed to $S_h=50~\mu$m in horizontal and fully opened in vertical.
-For ESRF its {\color{red}\sout{transmittivity}transmission?} is 1/8 (as discussed before) and 1 for EBS.
-In a similar way, the KB mirrors will crop the beam to a size that is the projection of their {\color{red}useful optical} length onto a direction perpendicular to the optical axis.
+For ESRF its transmission is 1/8 (as discussed before) and 1 for EBS.
+In a similar way, the KB mirrors will crop the beam to a size that is the projection of their useful optical length onto a direction perpendicular to the optical axis.
 Using a mirror length of $L_v$=60 mm and $L_h$=26 mm for VFM and HFM, respectively, and incident angle of 15 mrad, the KB behaves as an aperture of $900 \times 390 \mu m^2$.
@@
 values of 6.2$\times$21.6 $\mu rad^2$ (aperture size divided by the distance from the mirror to the VSS in H and to the source in V).
-The {\color{red}\sout{transmittivity}transmission?} coefficient is the ratio of the accepted N.A.
+The transmission coefficient is the ratio of the accepted N.A.
 over the beam divergence, resulting 9.5\% (H) and 0.18\% (V)for ESRF and 16.3\% and 0.18\% for EBS.
 The overall transmission is the product of H and V ratios, therefore 0.2\% for ESRF and 2.95\% for EBS, therefore 3.8 $10^{12}$ photons/s for ESRF and 8.5 $10^{13}$ photons/s for EBS.
-\todo{check with references $10^9$ at 33.6 keV.
+%\todo{check with references $10^9$ at 33.6 keV.
 I get: 1.6e+15 * 0.00295 * 0.7**3 = 1.62e+12.
@@
 We used the undulator peak value over the whole ML bandwidth, that overestimates the integrated flux.}
-\todo{need to consider ML reflectivity for the 3 reflections, also undulator is not perfect, horizontal slit may be closed to less than $50 \mu m$, also KB mirrors may not be fully illuminated and slope error losses due to initial ML mono are neglected}
+%\todo{need to consider ML reflectivity for the 3 reflections, also undulator is not perfect, horizontal slit may be closed to less than $50 \mu m$, also KB mirrors may not be fully illuminated and slope error losses due to initial ML mono are neglected}
 
@@
 An approximated estimation of CF can be obtained using Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas}.
-The coherent fraction is the ratio of the phase space {\color{green}\sout{of the fully coherent beam (zero emittance) over the actual beam} of the true beam relative to that of the fully coherent (zero emittance) beam}.
+The coherent fraction is the ratio of the phase space of the fully coherent (zero emittance) beam raletive to the true beam.
 The CF in the horizontal and vertical planes are
@@
 However, apertures centered at the optical axis would clean the beam from these undesired photons.
-If we compare the values {\color{green}\sout{transmittivity} transmission}of the beamline due to geometrical aperturing  (0.2\% for ESRF and 2.95\% for EBS) with the the CF values, we remark that they are very close.
+If we compare the values of the beamline transmission due to geometrical aperturing  (0.2\% for ESRF-1 and 2.95\% for EBS) with the the CF values, we remark that they are very close.
 This suggests that the beam at the focal position will be highly coherent because the fraction of the photons removed from the beam matches the CF value that quantify the fraction of ``good'' or ``coherent'' photons.
 
-To assess the focused beam size given by a coherent beam, one can consider the beamline in two sections: First, the ML monochromator and VSS that produce a coherent plane wave as the source (vertical) and VSS (horizontal) are both very far from the {\color{green}\sout{sample position} entrance aperture of the focusing optics}; Second, the KB-system playing a dual role, acting as a focusing element of focal distance $f$=5 cm, but also cropping the beam by acting as a square aperture of $D \times D=$ 400$\times$400 $\mu m^2$ (the length of the mirror projected in a plane perpendicular to the beam propagation)
-\todo{REMOVE: we obtained 900 in H, that will be reduced by 0.5 as propagated from 10 cm to 5 cm upstream of the sample, and 390 in V approximated by 400}.
+To assess the focused beam size given by a coherent beam, one can consider the beamline in two sections: First, the ML monochromator and VSS that produce a coherent plane wave as the source (vertical) and VSS (horizontal) are both very far from the entrance aperture of the focusing optics.
+Second, the KB-system playing a dual role, acting as a focusing element of focal distance $f$=5 cm, but also cropping the beam by acting as a square aperture of $D \times D=$ 400$\times$400 $\mu m^2$ (the length of the mirror projected in a plane perpendicular to the beam propagation).
+%\todo{REMOVE: we obtained 900 in H, that will be reduced by 0.5 as propagated from 10 cm to 5 cm upstream of the sample, and 390 in V approximated by 400}.
 This aperture has two effects: i) as discussed before, it removes a lot of off-axis photons increasing the CF to a value close to one, and ii) broadens of the focal spot because of coherent diffraction.
-The diffraction by a square aperture of a collapsing spherical wavefront has an intensity distribution proportional to $\sinc^2(k D x/2 f) \sinc^2(k D y / 2 f)$, \todo{include citation?} where $k=2\pi/\lambda$, and $x$ ($y$) is the horizontal (vertical) coordinate .
+The diffraction by a square aperture of a collapsing spherical wavefront has an intensity distribution proportional to $\sinc^2(k D x/2 f) \sinc^2(k D y / 2 f)$, \todo{include citation?} where $k=2\pi/\lambda$, and $x$ ($y$) is the horizontal (vertical) coordinate.
 Considering that the FWHM of $\sinc(x)$ is approximately 2.78, one obtains a FWHM of 7.75$\times$7.75 nm$^2$ for the focal spot.
@@
 \label{level1}
-\todo{transmittivity or transmission}
-In this section we perform ray tracing calculations to study the effect of real optics (including aberrations and slope errors) upon the beam size and beamline transmittivity.
-The first subsection deals with the effect of aberrations and transmittivity due to geometrical considerations.
+In this section we perform ray tracing calculations to study the effect of real optics (including aberrations and slope errors) upon the beam size and beamline transmission.
+The first subsection deals with the effect of aberrations and transmission due to geometrical considerations.
 The second subsection shows how the effect of slope errors of the KB mirrors degrade the focal spot.
@@
 The characteristics of each ray can also be extended to include some intensity, or better electric field amplitude.
-This allows to extend the pure geometrical tracing to include physical models that take into account the optical element reflectivity or transmittivity.
+This allows to extend the pure geometrical tracing to include physical models that take into account the optical element reflectivity or transmission.
 This combination of using a geometrical model plus a physical model allows to simulate every element used in a synchrotron beamline, such as mirrors and combinations of them (e.g., KBs); lenses, compound refractive lenses and transfocators; gratings of any type and shape, crystal systems, etc.
@@
 The results in Fig.~\ref{fig:ray tracing}a show two phenomena.
-One is that the migration from ESRF lattice to the EBS will significantly improve the horizontal size, and the VSS is not needed.
-For EBS, the whole beam passes though the VSS, producing a final spot with almost Gaussian horizontal intensity profile instead of the step distribution for ESRF.
+One is that the migration from ESRF-1 lattice to the EBS will significantly improve the horizontal size, and the VSS is not needed.
+For EBS, the whole beam passes though the VSS, producing a final spot with almost Gaussian horizontal intensity profile instead of the step distribution for ESRF-1.
 Secondly, one can observe a much larger luminosity in the EBS case.
-The transmittivity values are 0.165\% for ESRF and 2.28\% for EBS.
-These values confirm the analytical estimations (0.2\% for ESRF and and 2.95\% for EBS).
-As mentioned before, these values are similar to the coherent fraction \cite{arxivCF} (0.12 for ESRF and 2.73\% for EBS), meaning that a very high coherence of the beam is expected, therefore some broadening of the beam {\color{green}\sout{by clipping the beam by apertures and element acceptance} due to aperturing by the mirrors.}
+The transmission values are 0.165\% for ESRF-1 and 2.28\% for EBS.
+These values confirm the analytical estimations (0.2\% for ESRF-1 and and 2.95\% for EBS).
+As mentioned before, these values are similar to the coherent fraction \cite{arxivCF} (0.12 for ESRF and 2.73\% for EBS), meaning that a very high coherence of the beam is expected, therefore it is expected some broadening of the beam due to diffraction effects originated by the mirrors acceptance.
 
@@
 This is more evident in the vertical direction, because of the smaller focal size.
-The spot size is about 7-8 nm in V for both storage ring lattices, and in horizontal is about 15$\mu$m for ESRF-High$\beta$ and 11 nm for EBS.
+The spot size is about 7-8 nm in V for both storage ring lattices, and in horizontal is about 15$\mu$m for ESRF-1 High$\beta$ and 11 nm for EBS.
 It is interesting to note the almost round spot that will be obtained with the new EBS lattice.
@@
 
-The very long  source-{\color{green}\sout{sample} optics} distance combined with a small source (in particular for EBS) makes this beamline suitable for experimental techniques requiring a highly coherent beam.
+The very long  source-optics distance combined with a small source (in particular for EBS) makes this beamline suitable for experimental techniques requiring a highly coherent beam.
 Indeed, it has been shown that the clipping of the beam, mainly due to the acceptance of the KB mirrors produces a significant broadening of the focal spot.
@@
 
-A next step in the wave optics simulation is, {\color{green}\sout{maintaining the} assuming} full coherence, to generate a 2D source wavefront that matches in detail the real source (an undulator in a storage ring) and then simulate the propagation of such a wavefront through the optical elements with models that match well the characteristics of the real elements.
+A next step in the wave optics simulation is, assuming full coherence, to generate a 2D source wavefront that matches in detail the real source (an undulator in a storage ring) and then simulate the propagation of such a wavefront through the optical elements with models that match well the characteristics of the real elements.
 The source simulated is the on-axis-single-electron emission from the U18.3 undulator tuned to its first harmonic at $E=17225$ eV.
@@
 For both cases (i.e., simulating the beamline using thin optical elements or elliptical mirrors), the beam sizes calculated as the FWHM at the focal position are little affected by the presence of the slope errors (see Table \ref{tab:SRW_results}).
-There is a significant reduction and {\color{green}\sout{shuffling} blurring} of the interference fringes compared to observations with no slope error.
+There is a significant reduction and blurring of the interference fringes compared to observations with no slope error.
 They are almost completely removed and associated with a large increase of the background level: the radiation found on the main lobe (focus) shifts to the side lobes when the slope errors are considered.
@@
 
-The coherent modes calculated at the EBS source have been propagated \todo{Using which propagation code - is this all in COMSYL?} using a simplified beamline  with ideal focusing elements and slits to determine their numerical aperture (like in Section~\ref{wofry}).
+The coherent modes calculated by COMSYL at the EBS source have been propagated using the WOFRY package implementing a simplified beamline with ideal focusing elements and slits to determine their numerical aperture (like in Section~\ref{wofry}).
 Fig.~\ref{fig:mode transmission} shows how the intensity of the first modes is reduced by geometrical considerations.
@@
 As discussed, coherent modes will be ``cut'' in a different manner, depending on the slits aperturing effect.
-This implies that the Eq.~\ref{eq:mode normalization} is not longer {\color{green}\sout{verfified} satisfied}, and for the propagated modes $M_{mm}\le$1 and $M_{mn}\ne$0 for $m\ne n$.
+This implies that the Eq.~\ref{eq:mode normalization} is not longer satisfied, and for the propagated modes $M_{mm}\le$1 and $M_{mn}\ne$0 for $m\ne n$.
 Therefore, the coherent modes of the source after propagation and cropping by the beamline are no longer coherent modes of the resulting beam.

### 2019-06-12 paper-hierarchical.tex
upgrade

@@
 Strong focusing is performed using a Kirkpatrick-Baez (KB) mirror system placed a few cm upstream of the sample position (focal plane), which ensures a high demagnification.
-In order to match the acceptance of the KB in the vertical and horizontal planes a focusing multilayer monochromator (ML) has
+In order to match the acceptance of the KB in the vertical and horizontal planes a focusing (cylindrically bent) multilayer monochromator (ML) has
 been chosen. It i) monochromatises the beam with a large energy bandwidth (typically $\Delta E/E \approx 10^{-2}$) to allow high flux, 
-and ii) creates a secondary secondary source in the horizontal plane 40 m downstream of the undulator, where a horizontal slit (VSS, virtual source slit) is placed.
+and ii) it focus tangentially the beam  in the horizontal plane to create a secondary secondary source at 40 m downstream of the undulator, where a horizontal slit (VSS) is placed.
 A beamline schematic is shown in Figure~\ref{fig:ID16A}.
@@
     \begin{tabular}{rcccccc}
-    -                      & Source & Multilayer & Slit  & KB (v) & KB (h) & Focal plane \\ 
+    -                      & Source & Multilayer (ML) & Slit (VSS) & KB (v) & KB (h) & Focal plane \\ 
     Position {[}m{]}:      & 0.00   & 28.30      & 40.00 & 184.90 & 184.95 & 185.00      \\
@@
 This aperture has two effects: i) as discussed before, it removes a lot of off-axis photons increasing the CF to a value close to one, and ii) broadens of the focal spot because of coherent diffraction.
-The diffraction by a square aperture of a collapsing spherical wavefront has an intensity distribution proportional to $\sinc^2(k D x/2 f) \sinc^2(k D y / 2 f)$, \todo{include citation?} where $k=2\pi/\lambda$, and $x$ ($y$) is the horizontal (vertical) coordinate.
+The diffraction by a square aperture of a collapsing spherical wavefront has an intensity distribution proportional to $\sinc^2(k D x/2 f) \sinc^2(k D y / 2 f)$,
+%\todo{include citation?}
+where $k=2\pi/\lambda$, and $x$ ($y$) is the horizontal (vertical) coordinate.
 Considering that the FWHM of $\sinc(x)$ is approximately 2.78, one obtains a FWHM of 7.75$\times$7.75 nm$^2$ for the focal spot.
@@
 
-We performed ray tracing \todo{ it is currently unclear to me whether the ML mono was also simulated - and whether slope errors of this mono were included.
-The ML mono is a spherical reflector} using elliptical cylinder reflectors with finite dimensions as the KB focusing elements.
+We performed ray tracing of the ID16A beamline using undulator source tuned ti have its resonance at 17225 eV, considering the ML as a cylindrical mirror, and using elliptical cylinder reflectors with finite dimensions as the KB focusing elements.
 The reflector shape is chosen to be elliptical as this is the theoretical shape producing a point-to-point focusing.
-For a first calculation  the elliptical  reflectors are considered perfect, and have no slope errors.
+For a first calculation the reflectors (ML and KBs) are considered perfect, and have no slope errors.
 The results in Fig.~\ref{fig:ray tracing}a show two phenomena.
@@
 Secondly, one can observe a much larger luminosity in the EBS case.
-The transmission values are 0.165\% for ESRF-1 and 2.28\% for EBS.
+The transmission values are 0.162\% for ESRF-1 and 2.22\% for EBS.
 These values confirm the analytical estimations (0.2\% for ESRF-1 and and 2.95\% for EBS).
-As mentioned before, these values are similar to the coherent fraction \cite{arxivCF} (0.12 for ESRF and 2.73\% for EBS), meaning that a very high coherence of the beam is expected, therefore it is expected some broadening of the beam due to diffraction effects originated by the mirrors acceptance.
+As mentioned before, these values are similar to the coherent fraction \cite{arxivCF} (0.12 for ESRF-1 and 2.73\% for EBS), meaning that a very high coherence of the beam is expected, therefore it is expected some broadening of the beam due to diffraction effects originated by the mirrors acceptance.
 
@@
 \centering
+~~~~~ESRF-1~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~EBS
+\newline
 \includegraphics[width=0.45\textwidth]{GRAPHICS/idealelementsESRF.png}
@@
 \includegraphics[width=0.45\textwidth]{GRAPHICS/hybriderrorsEBS.png}
-\caption{Image produced by different types of ideal focusing for the ESRF (left column) and EBS (right column). The source with parameters in 
-\ref{tab:HandCalculations}. 
-The first row has been calculated using the elliptical reflectors (both for ML \todo{ML is not elliptical - it is a sphere} and KBs) but without slope errors.
-The focal beam dimensions are written in the plot title.
-The second row shows the image calculated the hybrid option (ray tracing plus coherent diffraction) with elliptical reflecting surfaces.
+\caption{Image produced by different types of ideal focusing for the ESRF (left column) and EBS (right column). The source with parameters in Table~\ref{tab:HandCalculations}. 
+The first row has been calculated using the ideal reflectors (cylindrical for ML, elliptical for the BK mirrors) without slope errors. 
+The second row shows the image calculated the hybrid option (ray tracing plus coherent diffraction) also for ideal reflectors with no slope errors.
 It shows the focal spot degradation due to the diffraction arising from the beam aperturing by the KB mirrors.
-The calculations in the bottom row also include slope errors using the real mirror profiles measures at the ESRF metrology lab.
+The calculations in the bottom row also include slope errors using the real mirror profiles measures at the ESRF metrology lab (see Appendix A).
 Further beam broadening and the growth of satellite structures in the vertical intensity distribution can be observed.
-Values of the histograms FWHM are displayed in the graph titles \todo{improve legibility of titles - too small}.
+The focal beam dimensions (FWHM of the histograms) are written in the plot title.
 }
@@
 We first analyze the beamline supposing a fully coherent beam.
-Subection~\ref{wofry} shows a very simplified 1D model.
-Subsection~\ref{srw_se} used a full 2D model with accurate simulation of the undulator emission but without considering storage ring emittance (i.e.
+Subection~\ref{wofry} shows a simplified 1D model.
+Subsection~\ref{srw_se} uses a full 2D model with accurate simulation of the undulator emission but without considering storage ring emittance (i.e.
 again using a fully coherent beam).
@@
 Indeed, it has been shown that the clipping of the beam, mainly due to the acceptance of the KB mirrors produces a significant broadening of the focal spot.
-As discussed in Section~\ref{level0}, very simple calculation of the diffraction limit can be done using a convergent spherical beam clipped by a slit of 400$\times$400$ \mu$m$^2$ placed  5cm upstream of the focus (at the second KB mirror position).
-The slit vertical size corresponds roughly to the horizontal acceptance of the second KB mirror, and it is approximately the same in vertical.
-A simple simulation of diffraction by a square slit using OASYS is shown in Fig.~\ref{wofry400um}.
-It reveals a spot of less than 10 nm, thus confirming the analytical predictions, and qualifying this beamline to be a diffraction-limited beamline.
-
-\begin{figure}
-\label{wofry400um}
-\centering
-\includegraphics[width=0.48\textwidth]{GRAPHICS/wofry400umslit.png}
-\includegraphics[width=0.48\textwidth]{GRAPHICS/wofry400umprofile.png}
-\caption{Diffraction spot at focal position produced an aperture (400$\times$400 $\mu$m$^2$) intercepting an spherical collapsing wavefront (R=-50mm).
-\todo{Remove this figure?}}
-\end{figure}
-
+As discussed in Section~\ref{level0}, very simple calculation of the diffraction limit can be done using a convergent spherical beam clipped by a slit of 400$\times$400$ \mu$m$^2$ placed 5cm upstream of the focus (at the second KB mirror position), producing a spot of about 7.75$\times$7.75 nm$^2$, therefore qualifying this beamline to be a diffraction-limited beamline.
 We perform here a simplified wavefront simulation in 1D, thus separating the simulations of the horizontal and vertical planes.
@@
 \centering
-\includegraphics[width=0.45\textwidth]{GRAPHICS/wofry1Dh.png}
-\includegraphics[width=0.45\textwidth]{GRAPHICS/wofry1Dv.png}
-\caption{Intensity distribution at the focal position in horizontal plane (left, FWHM=8.1 nm) and in vertical plane (right, FWHM=7.1 nm). 
+\includegraphics[width=0.95\textwidth]{GRAPHICS/wofry1D.png}
+\caption{Intensity distributions at the focal position (horizontal and vertical cuts) as calculated by a simplified 1D waveoptics model. 
 }
@@
     \centering
-        \includegraphics[width=4.25cm]{GRAPHICS/se_idealTE_idealOE_h.pdf}
-        \includegraphics[width=4.25cm]{GRAPHICS/ebs_slopeTE_slopeOE_h.pdf}
+        \includegraphics[width=0.95\textwidth]{GRAPHICS/se_idealTE_idealOE_h.pdf}
+        \includegraphics[width=0.95\textwidth]{GRAPHICS/ebs_slopeTE_slopeOE_h.pdf}
     \label{fig:SingleElectron}
-    \caption{Wavefront propagation from a single electron emission through the ID16A beamline.
-Left: the perfect focusing with ideal lenses (black) is compared with the focusing using elliptical mirrors (red).
-Right: the same as in the previous plot, but adding the effect of the mirrors slope errors in both cases.}
+    \caption{Intensity profile at the sample position produced by the single electron emission propagated through the ID16A beamline.
+Top: the perfect focusing with ideal lenses (black) is compared with the focusing using elliptical mirrors (red).
+Bottom: the same as in the previous plot, but adding the effect of the mirrors slope errors in both cases.}
     %Simulations took less than one minute using a desktop computer Intel Core i7 with 8Gb of RAM memory.}
@@
 The limitation of the method is that one wavefront is not representative for a beam that, in storage rings, is produced by the emission of many electrons circulating through the corresponding magnet structure.
+
 Diffraction limited sources would emit radiation that can be described in good approximation by a single wavefront.
@@
 To account for this effect one should complete or extend the wavefront simulations to take into account that in fact not only a single wavefront contributes to the properties of the beam.
-There are three approaches currently used: 1) use of convolutions, 2) Monte Carlo multi-electron sampling, and 3) Propagation of coherent modes.
+
+\begin{figure}
+    \centering
+        \includegraphics[width=0.95\textwidth]{GRAPHICS/esrf_idealTE_idealOE_h.pdf}
+        \includegraphics[width=0.95\textwidth]{GRAPHICS/esrf_slopeTE_slopeOE_h.pdf}
+    \label{fig:FiniteEmittanceA}
+    \caption{
+    Intensity profile at the sample position produced by multi-electron method sampled with parameters of the ESRF-1 lattice, using ideal focusing elements (top), and including the effect of the mirrors slope errors (bottom).
+    }
+\end{figure}
+
+\begin{figure}
+    \centering
+        \includegraphics[width=0.95\textwidth]{GRAPHICS/ebs_idealTE_idealOE_h.pdf}
+        \includegraphics[width=0.95\textwidth]{GRAPHICS/ebs_slopeTE_slopeOE_h.pdf}
+    \label{fig:FiniteEmittanceB}
+    \caption{
+    Intensity profile at the sample position produced by multi-electron method sampled with parameters of the EBS lattice, using ideal focusing elements (top), and including the effect of the mirrors slope errors (bottom).
+    }
+\end{figure}
+
+There are three approaches currently used for including the electron emittance effects in the waveoptics simulations: 1) use of convolutions, 2) Monte Carlo multi-electron sampling, and 3) Propagation of coherent modes.
 The convolution method consists of calculating the emission characteristics and then convolving the resulting intensity with Gaussians describing the electron beam.
@@
 
-\begin{figure}
-    \centering
-        \includegraphics[width=4.25cm]{GRAPHICS/esrf_idealTE_idealOE_h.pdf}
-        \includegraphics[width=4.25cm]{GRAPHICS/esrf_slopeTE_slopeOE_h.pdf}
-    \label{fig:FiniteEmittanceA}
-    \caption{Wavefront propagation from the ESRF magnetic lattice through the ID16A beamline.}
-\end{figure}
-
-\begin{figure}
-    \centering
-        \includegraphics[width=4.25cm]{GRAPHICS/ebs_idealTE_idealOE_h.pdf}
-        \includegraphics[width=4.25cm]{GRAPHICS/ebs_slopeTE_slopeOE_h.pdf}
-    \label{fig:FiniteEmittanceB}
-    \caption{Wavefront propagation from the EBS magnetic lattice through the ID16A beamline.}
-\end{figure}
-
 The source used for simulations is the U18.3 undulator tuned to its first harmonic at $E=17225$ eV with a finite emittance electron beam (see Table \ref{tab:eBeam}).
-Two set of parameters were used for the electron beam: the current machine parameters at the ESRF (high-$\beta$) and the EBS straight sections (Tables \ref{tab:TableSources} and \ref{tab:eBeam}).
-
+Two set of parameters were used for the electron beam: the machine parameters of the ESRF-1 lattice (high-$\beta$) and the new EBS straight sections (Tables \ref{tab:TableSources} and \ref{tab:eBeam}).
 Simulations for both storage ring lattices were carried out modelling the beamline using thin ideal lenses and elliptical mirror models with and without the addition of the slope errors.
@@
 The beam sizes and shapes are similar at the focal position, only.
-{\todo{more detailed comments on comparisons with filament beam results? For comparison also information on calculation time, number of electrons, ...}}
+%{\todo{more detailed comments on comparisons with filament beam results? For comparison also information on calculation time, number of electrons, ...}
 
@@
 \end{equation}
-that measures the coherence of the beam, from fully incoherent ($\mu=0$) to fully coherent ($\mu$=1). 
+that measures the coherence of the beam between two points, and ranges from fully incoherent ($\mu=0$) to fully coherent ($\mu$=1). 
 Eq.~\ref{CSD} can in principle be calculated using the Kim's convolution theorem \cite{kim1986} (see \cite{geloni2008} for a full discussion on the validity) but its storage is prohibitif in present computers: supposing sampling one spatial dimension by $N\approx1000$ pixels, a wavefront will contain $N^2$ pizels, and the CSD will be sampled over $N^4$ points.
@@
 
-We performed COMSYL calculations for the U18.3 undulator radiation of ID16A of 1.4 m length using both the EBS and ESRF-High$\beta$ lattices.
-The obtained CF are 2.8\% for EBS and 0.13\% for ESRF-High$\beta$, indicating that the overall coherent flux emitted by EBS will be about 20 times the existing coherent flux with ESRF-High$\beta$.
+We performed COMSYL calculations for the U18.3 undulator radiation of ID16A of 1.4 m length using both the EBS and ESRF-1 lattices.
+The obtained CF are 2.8\% for EBS and 0.13\% for ESRF-1 High$\beta$, indicating that the overall coherent flux emitted by EBS will be about 20 times the existing coherent flux with ESRF-High$\beta$.
 Fig.~\ref{fig:histomodes}a shows the cumulated occupation of the coherent modes for the full central cone undulator emission.
@@
     \centering
-        \includegraphics[width=7cm]{GRAPHICS/cumulated_occupation.eps}
-        \includegraphics[width=6.5cm]{GRAPHICS/up_to_mode_id16a.eps}
+        \includegraphics[width=0.95\textwidth]{GRAPHICS/cumulated_occupation.eps}
+        \includegraphics[width=0.95\textwidth]{GRAPHICS/up_to_mode_id16a.eps}
     \caption{Cumulated occupation of the coherent modes for the EBS and ESRF-High-$\beta$ (top), and number of modes needed to reach a given percent of CSD occupation (bottom). }
@@
     \caption{a) Spectral density $S(x,y)$ at the source position for EBS. b) Intensity distribution of the first coherent modes ($m$=0 to $m$=3) at the source position.
-    \todo{label plots}
     }

### 2019-06-13 paper-hierarchical.tex
Updates from Overleaf

@@
      %-------------------------------------------------------------------------
-     \journalcode{S}              % Indicate the journal to which submitted
+     \journalcode{K}              % Indicate the journal to which submitted
                                   %   A - Acta Crystallographica Section A
@@
 An approximated estimation of CF can be obtained using Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas}.
-The coherent fraction is the ratio of the phase space of the fully coherent (zero emittance) beam raletive to the true beam.
+The coherent fraction is the ratio of the phase space of the fully coherent (zero emittance) beam relative to the true beam.
 The CF in the horizontal and vertical planes are
@@
 It shows the focal spot degradation due to the diffraction arising from the beam aperturing by the KB mirrors.
-The calculations in the bottom row also include slope errors using the real mirror profiles measures at the ESRF metrology lab (see Appendix A).
+The calculations in the bottom row also include slope errors using the real mirror profiles measures at the ESRF metrology lab (see Appendix~\ref{appendix:metrology}).
 Further beam broadening and the growth of satellite structures in the vertical intensity distribution can be observed.
@@
 It is also possible to add slope errors to the simulations to study the resulting impact upon the focal spot quality (Fig.~\ref{fig:SingleElectron}).
-For that, the available surface metrology data of the reflective elements (see Appendix \ref{appendix:metrology})  are used.
+For that, the available surface metrology data of the reflective elements (see Appendix~\ref{appendix:metrology})  are used.
 For both cases (i.e., simulating the beamline using thin optical elements or elliptical mirrors), the beam sizes calculated as the FWHM at the focal position are little affected by the presence of the slope errors (see Table \ref{tab:SRW_results}).
@@
 \end{figure}
-
-\todo{
-\begin{itemize}
- \item introduce slope errors in WOFRY propagation
- \item propagate modes with SRW
- \item create some plots with cuts of W
-\end{itemize}
-}
+%\todo{
+%\begin{itemize}
+% \item introduce slope errors in WOFRY propagation
+% \item propagate modes with SRW
+% \item create some plots with cuts of W
+%\end{itemize}
+%}
 
@@
 It is not recommended to do similar calculations that imply a the cost because a source that is quite incoherent was treated with partial coherence methods.
-\inred{In Appendix B we present for the ESRF High$\beta$ similar results than those in Section~\ref{comsyl}}.
+%\inred{In Appendix B we present for the ESRF High$\beta$ similar results than those in Section~\ref{comsyl}}.
 
@@
 Also, the reflectivity of the elements are not considered.
-The simulation of this was not included in the simulations, because it can be estimated analytically considering an average reflectivity about \inred{0.7 for multilayers and 0.9 for mirrors, therefore a reduction factor of (0.7)**3 ML for the beamline flux due to reflectivity of mirror coatings}.
-
-\todo{
-Complete discussion. Some ideas:
-\begin{itemize}
- \item DISCUSS EXPERIMENTAL RESULTS?
- \item Discuss coherence length?
- \item Comment on the heat-load deformation 
- \item Extend the effect of reflectivity by ML and mirror coating
- \item Talk a bit more on Oasys?
- \item comment on general strategy for beamline simulations and the application of the shown hierarchical model. Make a flowchart
-\end{itemize}
- }
+The simulation of this was not included in the simulations, because it can be estimated analytically considering an average reflectivity about 0.7 for multilayers and 0.9 for mirrors, therefore a transmission factor of 0.44 for the beamline flux due to reflectivity of mirror coatings.
+
+%\todo{
+%Complete discussion. Some ideas:
+%\begin{itemize}
+% \item DISCUSS EXPERIMENTAL RESULTS?
+% \item Discuss coherence length?
+% \item Comment on the heat-load deformation 
+% \item Extend the effect of reflectivity by ML and mirror %coating
+% \item Talk a bit more on Oasys?
+% \item comment on general strategy for beamline %simulations and the application of the shown hierarchical %model. Make a flowchart
+%\end{itemize}
+% }
 
@@
 
-\ack{\todo{someone to thank?
+%\ack{\todo{someone to thank?
 - RC: before the PHANGS in Trieste I was contacting Oleg about the elliptical mirrors.
 There was a bug there and he corrected it.
-Only after fixed it we were able to simulate the ID16A with the mirror elements.
-So maybe something mentioning some disucssions on the early stages of the SRW simulations.}}
+%Only after fixed it we were able to simulate the ID16A with the mirror elements.
+So maybe something mentioning some disucssions on the early stages of the SRW %simulations.}}
 
@@
 These profiles are used along the text in ray tracing and wave optics simulations.
-\todo{
-RC: maybe plot the PSD for the three mirrors? Put them on the same scale to show how nice the KB's are?
- }
+
 \begin{figure}\label{fig:metrology}
-\includegraphics[width=0.8\textwidth]{GRAPHICS/ID16A_ML_W-B4C_120Strp.png}
-\includegraphics[width=0.8\textwidth]{GRAPHICS/ID16A_KB1_VF.png}
-\includegraphics[width=0.8\textwidth]{GRAPHICS/ID16A_KB1_HF.png}
-\caption{Error height profiles used for the simulations for a) the ML monochromator, b) the KB vertical focusing mirror, and c) the KB horizontal focusing mirror.}
-\end{figure}
-
-\section{COMSYL results for the ESRF-H$\beta$}
-\todo{Under development}
+\includegraphics[width=0.95\textwidth]{GRAPHICS/metrology.eps}
+\caption{Error height profiles used for the simulations (left) and Power Spectral Density (right) for the ML monochromator (top row), the KB vertical focusing mirror (central row), and the KB horizontal focusing mirror (bottom row).}
+\end{figure}
+
+%\section{COMSYL results for the ESRF-H$\beta$}
+%\todo{Under development}
 

### 2019-06-14 paper-hierarchical.tex
Rafael changes

@@
 A hierarchical scheme of the calculations of an X-ray beamline for coherent applications is presented.
-Starting from simple calculations to partial coherence calculations, we show how ray tracing and wave optics software are applied simulating the beamline performances.
+Starting from simple calculations to partial coherence calculations, we show how ray-tracing and wave optics software are applied simulating the beamline performances.
 \end{synopsis}
@@
 \begin{abstract}
-We consider different approaches to simulate a modern X-ray beamline.
+We consider different approaches to simulating a modern X-ray beamline.
 Several methodologies with increasing complexity are applied to discuss the relevant parameters that quantify the beamline performance.
-Parameters such as flux, dimensions and intensity distribution of the focused beam and coherence properties are obtained from simple analytical calculations to sophisticated computer simulations using ray tracing and wave optics techniques.
-A latest generation x-ray nanofocusing beamline for coherent  applications (ID16A at the ESRF) has been chosen  to study in detail the issues related to highly demagnifying  synchrotron sources and exploiting the beam coherence.
-The performance of the beamline is studied for two storage rings: the old ESRF-1 (emittance 4000 pm) and the new ESRF-EBS (emittance 150 pm).
-In addition to traditional results in term of flux and beam sizes, an innovative study on the partial coherence properties based on propagation of coherent modes is presented.
+Parameters such as flux, dimensions and intensity distribution of the focused beam and coherence properties are obtained from simple analytical calculations to sophisticated computer simulations using ray-tracing and wave optics techniques.
+A latest-generation X-ray nanofocusing beamline for coherent applications (ID16A at the ESRF) has been chosen to study in detail the issues related to highly demagnifying synchrotron sources and exploiting the beam coherence.
+The performance of the beamline is studied for two storage rings: the old ESRF-1 (emittance 4000~pm) and the new ESRF-EBS (emittance 150~pm).
+In addition to traditional results in terms of flux and beam sizes, an innovative study on the partial coherence properties based on the propagation of coherent modes is presented.
 The different algorithms and methodologies are implemented in the software suite OASYS \cite{codeOASYS}.
-They are discussed with emphasis beamline placed upon the benefits and limitations of each.
+Those are discussed with emphasis placed upon the their benefits and limitations.
 \end{abstract}
@@
 
-Many storage ring based x-ray synchrotron facilities are building or planning upgrades to increase brilliance and coherent flux by one to three orders of magnitude.
+Many storage-ring-based X-ray synchrotron facilities are building or planning upgrades to increase brilliance and coherent flux by one to three orders of magnitude.
 The first upgrade of a large facility will be the EBS (Extremely Brilliant Source) \cite{orangebook} at the European Synchrotron Radiation Facility (ESRF).
-The new storage ring of 150 pm emittance that will be commissioned in 2020 will significantly boost the coherence of the x-ray beams.
-Applications exploiting beam coherence such as x-ray photon correlation spectroscopy, coherent diffraction imaging, propagation-based phase contrast imaging and ptychography will strongly benefit from this update.
-
-Accurate calculation and quantitative evaluation of the parameters related to x-ray optics is of great importance for designing, building and exploiting new beamlines.
+The new storage ring of 150~pm emittance will be commissioned in 2020 and will significantly boost the coherence of the X-ray beams.
+Applications exploiting beam coherence such as X-ray photon correlation spectroscopy, coherent diffraction imaging, propagation-based phase-contrast imaging, and ptychography will strongly benefit from this update.
+
+Accurate calculation and quantitative evaluation of the parameters related to X-ray optics are of great importance for designing, building and exploiting new beamlines.
 Every modern beamline follows a procedure of conceptual design plus technical design where detailed simulations are essential.
@@
 Deformations of the optical elements due to heat load need to be controlled, therefore optics simulations have to include results from engineering modeling of the thermal deformations usually done using finite element analysis.
-Traditionally the optics calculations for synchrotron beamlines are done using ray tracing techniques.
-They model the x-ray beam as a collection of incoherent rays.
+Traditionally the optics calculations for synchrotron beamlines are done using ray-tracing techniques.
+They model the X-ray beam as a collection of incoherent rays.
 This technique is still essential for obtaining information on aberrations, flux propagation, monochromator bandwidth, etc.
-However, the high coherence of the new sources can introduce the necessity to complement the ray tracing methods with models including effects of diffraction and scattering from a coherent beam.
-A fully coherent hard x-ray beam is still a dream for the new generation storage rings.
-In spite of a considerable increase in the coherent fraction of the new beams, the coherent fraction at typically hard x-ray photon energies is still a few percent for the new sources.
-Consequently ray tracing analysis does not become obsolete, but has to be complemented with an analysis based on wave optics.
-Because of the partial coherence the wave optics analysis becomes non trivial and can be done with different levels of approximation, as discussed in this work.
+However, the high coherence of the new sources can introduce the necessity to complement the ray-tracing methods with models including effects of diffraction and scattering from a coherent beam.
+A fully coherent hard X-ray beam is still a dream for the new generation storage rings.
+Despite a considerable increase in the coherent fraction of the new beams, the coherent fraction at typically hard X-ray photon energies is still a few percents for the new sources.
+Ray-tracing analysis does not become obsolete but has to be complemented with an analysis based on wave optics.
+Because of the partial coherence, the wave optics analysis becomes nontrivial and can be done with different levels of approximation, as discussed in this work.
 In addition, wave optics simulations are usually computationally expensive.
@@
 X-ray beamlines are very particular optical systems.
-Some implications of exploiting the short wavelengths such as the use of grazing incidence reflectors and the use of perfect crystals as typical hard x-ray monochromators render common commercial software packages ill-adapted for the study of synchrotron beamlines.
-For this reason many optics tools have been developed by the synchrotron community.
-Some of them have a long history like SHADOW (more than 40 years old) or SRW (almost 30), to mention only codes used in this work.
-These historical tools together with other new applications have been collected together in a user friendly software package: OASYS.
+Some implications of exploiting the short wavelengths such as the use of grazing incidence reflectors and the use of perfect crystals as typical hard X-ray monochromators render common commercial software packages ill-adapted for the study of synchrotron beamlines.
+For this reason, many optics tools have been developed by the synchrotron community.
+Some of them have a long history like SHADOW - from 1984 \cite{firstShadow}; or SRW - from 1998 \cite{codeSRW}; to mention only codes used in this work.
+These historical tools together with other new applications have been collected together in a user-friendly software package: OASYS \cite{codeOASYS}.
 Many of the calculations presented here have been performed directly in OASYS or with scripts created with the help of OASYS.
 
-The paper is structured as follows: we first introduce in Section~\ref{Beamline description} the main parameters of the beamline and explain the particular choice of the parameters used in the calculations.
-Section~\ref{Simulations} is the central part of the paper, and describes the different methods used to analyse the beamline with increasing complexity, setting a hierarchy of methodologies that can be exported to the simulations of any beamline.
+ In this paper, some quantitative results on how the storage ring upgrade will affect a particular beamline are presented and it is structured as follows: we first introduce in Section~\ref{Beamline description} the main parameters of the beamline and explain the particular choice of the parameters used in the calculations.
+Section~\ref{Simulations} is the central part of the paper, and describes the different methods used to analyze the beamline with increasing complexity, setting a hierarchy of methodologies that can be exported to the simulations of any beamline.
 In Section \ref{level0} we calculate with simple methods approximate values of focal size, flux and coherent fraction.
-Accuracy can improved using geometrical methods in Section \ref{level1} that allow a more complete simulation of the beamline optics within certain limits which are related, in this example, to high demagnification and high coherence of the studied beamline.
+Accuracy can be improved using geometrical methods in Section \ref{level1} that allow a more complete simulation of the beamline optics within certain limits which are related, in this example, to high demagnification and high coherence of the studied beamline.
 The wave optics section~\ref{level2} includes different models of increasing complexity as the required accuracy of the results becomes more demanding.
@@
 The hierarchical approach allows more accurate and better results but the price to pay is to deal with more complex and computationally expensive tools.
-Finally Section \ref{summary} summarizes the results.
+Finally, Section \ref{summary} summarizes the results.
 
@@
 
-This section presents the reader with the main operational parameters for the ID16A beamline at ESRF in two scenarios.
-This beamline was built during the Upgrade Phase I at ESRF (2009-2015).
-It uses one or several undulators as source.
-During the period 2011-2018 it operated exploiting the ESRF-1 storage ring’s double-bend achromat (DBA) lattice with horizontal emittance about 4000 pm.
-In the second phase of the ESRF Upgrade the original storage ring is being replaced by a new hybrid multi-bend achromat (HMBA) lattice.
+This section presents the reader with the main operational parameters for the ID16A beamline at ESRF \cite{ID16A} in two scenarios.
+This beamline was built during the Upgrade Phase~I at ESRF (2009-2015).
+It uses one or several undulators as the source.
+During the period 2011-2018, it operated exploiting the ESRF-1 storage ring’s double-bend achromat (DBA) lattice with horizontal emittance about 4000~pm.
+In the second phase of the ESRF Upgrade, the original storage ring is being replaced by a new hybrid multi-bend achromat (HMBA) lattice.
 This new lattice will drastically improve the performances of the ESRF source.
@@
 This is possible using almost the same ring circumference and electron energy $E_e$.
-As a result, the x-ray sources will be much coherent, specially in the horizontal direction.
-In this paper, some quantitative results on how the storage ring upgrade will affect a particular beamline are presented.
-%RC190404: @PhaseI -> from the orange book: On the basis of the Purple Book, the ESRF Council decided in June 2008 to launch the ESRF UP by establishing a two-phase process.
-The Upgrade Programme Phase I (UP PI) for the period 2009-2015, and the Upgrade Programme Phase II (UP PII) to be prepared in the period 2012-2014, and to be presented to the Council for consideration in 2014 and launch in 2015.
-% The  Upgrade  Programme  Phase  I  has  been  on  the  ESFRI  Roadmap  since  its  fi rst publication  in  2006,  and  is  today  heading  towards  its  successful  completion  by  the end of 2015.
-
-ID16A \cite{ID16A} is a beamline that provides a nanometric, highly coherent beam with broad energy-bandpass.
+As a result, the X-ray sources will be much more coherent, especially in the horizontal direction.
+
+ID16A is a beamline that provides a nanometric, highly coherent beam with broad energy-bandpass.
 It is dedicated to nanofocusing applications.
-The beamline can operate at two fixed photon energies: 17050 eV and 33600 eV.
-The straight section in the magnet lattice houses four undulators, two 18 mm and another two of 22.4 mm period.
+The beamline can operate at two fixed photon energies: 17050~eV and 33600~eV.
+The straight section in the magnet lattice houses four undulators, two 18~mm and another two of 22.4~mm period.
 The undulator parameters and $K$-values at the resonances for the energies in use are shown in Table \ref{tab:Undulators}.
 
-%RC190404: From the ID16D page -> The beamline provides a high-brilliance beam focused down to nanometer size, allowing quantitative 3D characterization of the morphology and the elemental composition of specimens in their native state by combining coherent imaging techniques and X-ray fluorescence microscopy.
-% The 185 m long beamline provides nano-focused beams for analytical imaging.
-The nano-imaging beamline addresses problems in biology, biomedicine and nano-technology using X-ray fluorescence microscopy and nano-tomography.
-It is optimised for ultimate hard X-ray focusing of a beam with a large energy bandwidth at specific energies (17keV or 33.6 keV).
-Cryogenic sample preservation is available for 2D and 3D imaging.
-% Techniques: coherent imaging; phase-contrast imaging; near-field and far-field ptychography; X-ray fluorescence microscopy;  scanning transmission X-ray microscopy; nanotomography
-% http://www.esrf.eu/home/UsersAndScience/Experiments/XNP/ID16A/over.html
-
 \begin{table}\label{tab:Undulators}
@@
 \caption{Parameters of the undulators available at the ID16A beamline: period ($\lambda_u$), length ($L$), number of identical undulators ($N$) and deflection parameter ($K$) values for the three main photon energies in use.
-The valuse are shown for the two ESRF magnetic lattices: the old ESRF-1 (labelled ESRF, with $E_e = 6.04$ GeV) and the new ESRF-EBS (labelled EBS, with $E_e = 6.00$ GeV).
+The valuse are shown for the two ESRF magnetic lattices: the old ESRF-1 (labelled ESRF, with $E_e = 6.04$~GeV) and the new ESRF-EBS (labelled EBS, with $E_e = 6.00$~GeV).
 $n$ is the emission harmonic in use for each particular photon energy.}
@@
 \begin{tabular}{lcccccc}
-name  & $\lambda_u$ [mm] & $L$ [m] & N & $K$ (11200 eV) & $K$ (17050 eV) & $K$ (33600 eV) \\ 
+name  & $\lambda_u$ [mm] & $L$ [m] & N & $K$ (11200~eV) & $K$ (17050~eV) & $K$ (33600~eV) \\ 
 \input{tableUndulator.txt}
@@
 
-ID16A is a long beamline (the distance from the center of the straight section to the sample position is 185 m).
+ID16A is a long beamline: the distance from the center of the straight section to the sample position is 185~m.
 Strong focusing is performed using a Kirkpatrick-Baez (KB) mirror system placed a few cm upstream of the sample position (focal plane), which ensures a high demagnification.
-In order to match the acceptance of the KB in the vertical and horizontal planes a focusing (cylindrically bent) multilayer monochromator (ML) has
-been chosen. It i) monochromatises the beam with a large energy bandwidth (typically $\Delta E/E \approx 10^{-2}$) to allow high flux, 
-and ii) it focus tangentially the beam  in the horizontal plane to create a secondary secondary source at 40 m downstream of the undulator, where a horizontal slit (VSS) is placed.
+In order to match the acceptance of the KB in the vertical and horizontal planes, a focusing (cylindrically bent) multilayer monochromator (ML) has
+been chosen. It i) monochromatizes the beam with a large energy bandwidth (typically $\Delta E/E \approx 10^{-2}$) to allow high flux, 
+and ii) it focus tangentially the beam in the horizontal plane to create a secondary source at 40~m downstream of the undulator, where a horizontal slit (VSS) is placed.
 A beamline schematic is shown in Figure~\ref{fig:ID16A}.
@@
 
-Simulations were done using the two different storage ring lattices mentioned before: the (already dismantled) ESRF-1 where the undulators are in a high-$\beta$ straight section (larger horizontal emittance) and the new EBS (lower horizontal emittance) straight sections.
+% Simulations were done using the two different storage ring lattices mentioned before: the (already dismantled) ESRF-1 where the undulators are in a high-$\beta$ straight section (larger horizontal emittance) and the new EBS (lower horizontal emittance) straight sections.
 The parameters for the current and the new lattice are shown in Tables \ref{tab:TableSources} and \ref{tab:eBeam}.
@@
 \begin{table}\label{tab:TableSources}
-    \caption{ESRF and ESRF-EBS main storage ring parameters. Main parameters of the present DBA lattice (ESRF-1) and future HMBA lattice (EBS) storage rings \cite{orangebook}.}
+    \caption{Main parameters of the present DBA lattice (ESRF-1) and future HMBA lattice (EBS) storage rings \cite{orangebook}.}
         \begin{tabular}{rcc}
@@
 Values taken at the symmetry point of the straight section, where the insertion devices are placed.
-Values for ESRF are taken from \cite{ESRF2014}.
+Values for ESRF-1 are taken from \cite{ESRF2014}.
 For EBS, two (very close) design values are used, from Ref.~\cite{ESRF2014} labelled EBS-S28A, and those coming from further lattice refinement, labelled EBS-S28D.
@@
 
-For all simulations presented hereafter, we selected the configuration using a single undulator U18.3 placed in the center of the straight section and tuned to its first harmonic at a photon energy of $E=17225$ eV (a value close to the usual operation value shown in Table \ref{tab:Undulators}).
-This section is divided in three subsections, that present three main levels of simulations: simple calculations that can be made by hand (section \ref{level0}), ray-tracing simulations (section \ref{level1}) and wave optics simulations (section \ref{level2}).
+For all simulations presented hereafter, we selected the configuration using a single undulator U18.3 placed in the center of the straight section and tuned to its first harmonic at a photon energy of $E=17225$~eV: a value close to the usual operation value shown in Table \ref{tab:Undulators}.
+Calculations were done using the two different storage ring lattices mentioned before: the (already dismantled) ESRF-1 where the undulators are in a high-$\beta$ straight section (larger horizontal emittance) and the new EBS (lower horizontal emittance) straight sections.
+The parameters for the current and the new lattice are shown in Tables \ref{tab:TableSources} and \ref{tab:eBeam}.
+This section is divided into three subsections, that present three main levels of simulations: simple calculations that can be made by hand, ray-tracing simulations and wave optics simulations: sections  \ref{level0}, \ref{level1} section \ref{level2} respectively.
 
@@
 A first na{\"{i}}ve estimation of the focal size can be made by considering the source size times the magnification factor $M$.
-In the paraxial approximation, the magnification factor can be expressed as $M=q/p$, where $p$ is the distance between object and imaging element and $q$ is the distance between imaging element and image.
+In the paraxial approximation, the magnification factor can be expressed as $M=q/p$, where $p$ is the distance between object and imaging element and $q$ is the distance between the imaging element and image.
 The demagnification is the inverse of the magnification ratio: $M^{-1}=p/q$.
 Systems where $M>1$ are said to be magnifying, whereas systems with $M<1$ (or $M^{-1}>1$) are said to be demagnifying.
-The latter will be our case, thus is more convenient to work with demagnification ($M^{-1}>1$) rather than with magnification.
-
-The source size (or divergence) is usually approximated supposing a Gaussian distribution of the electron positions (or divergence) (Table \ref{tab:eBeam}) and single-electron photon source size or divergence.
+The latter will be our case: it is, consequently, more convenient to work with demagnification ($M^{-1}>1$) rather than with magnification.
+
+The source size and divergence are usually approximated supposing a Gaussian distribution of the electron positions and divergences (Table \ref{tab:eBeam}) and single-electron photon source size and divergence.
 In some cases, such as for undulator emission, this Gaussian approximation may not be accurate and have some implications that will be discussed later.
-Let $\sigma_{e(h,v)}$ and $\sigma_{e(h,v)}'$ be the standard deviations of electron beam size and divergence  at the centre of  symmetry of the straight section in a plane perpendicular to the main propagation direction.
+Let $\sigma_{e(h,v)}$ and $\sigma_{e(h,v)}'$ be the standard deviations of electron beam size and divergence at the center of symmetry of the straight section in a plane perpendicular to the main propagation direction.
 The subscripts $h$ and $v$ stand for horizontal and vertical, respectively.
@@
 The photon source size and divergence are approximately given by the convolution of the electron beam sizes and the undulator natural sizes.
-If we consider Gaussian distributions, the photon source is
+If we consider Gaussian distributions, the photon source is:
 \begin{align}
@@
 For predicting the photon beam sizes in a position out of the focus the beam divergence is used.
-In an ideal {\'{e}}tendue conserving optical system, where the Smith-Helmholz invariant applies, the resulting beam divergence transforms inversely as the beam size: it can be calculated as the product between the beam divergence at the source with the demagnification factor $M^{-1}$.
-
-Using the values for U18.3, $n$=1, and  $E=17225$ eV (Table \ref{tab:Undulators}) and the optical configuration described in Table \ref{tab:Positions}, one obtains a focal size 139$\times$5 nm$^2$ (H$\times$V) for ESRF-1 and 10$\times$5 nm$^2$ for EBS (Table~\ref{tab:HandCalculations}).
-The horizontal focal spot for the ESRF-1 lattice is larger than the specified target value for the beamline, therefore the VSS could be closed to 50$\mu$m thus reducing the beamsize at the VSS plane by a factor ca.
+In an ideal {\'{e}}tendue-conserving optical system, where the Smith-Helmholz invariant applies, the resulting beam divergence transforms inversely as the beam size: it can be calculated as the product between the beam divergence at the source with the demagnification factor $M^{-1}$.
+
+Using the values for U18.3, $n=1$, and  $E=17225$~eV (Table \ref{tab:Undulators}) and the optical configuration described in Table \ref{tab:Positions}, one obtains a focal size 139$\times$5~nm$^2$ (H$\times$V) for ESRF-1 and $10\times5$~nm$^2$ for EBS (Table~\ref{tab:HandCalculations}).
+The horizontal focal spot for the ESRF-1 lattice is larger than the specified target value for the beamline, therefore the VSS could be closed to 50~$\mu$m thus reducing the beam size at the VSS plane by a factor ca.
 8.
 %\inred{8 $\approx$ 403.8/50=8.076}
-Therefore, for ESRF-1, the horizontal size at the sample will be affected by the same factor obtaining roughly 17$\times$5 nm$^2$
+Therefore, for ESRF-1, the horizontal size at the sample will be affected by the same factor obtaining roughly 17$\times$5~nm$^2$.
 
@@
 As a consequence, the results calculated using only the demagnification factor are usually optimistic.
-Moreover imperfections in the optical surface (figure errors, slope errors and microroughness) are often the limiting factors of real reflective optics.
-Their effects will studied in Sections~\ref{level1} and \ref{level2}.
+Moreover, imperfections in the optical surface (figure errors, slope errors and microroughness) are often the limiting factors of real reflective optics.
+Their effects will be studied in Sections~\ref{level1} and \ref{level2}.
+
 %Effects of slope errors and diffraction limits can be simply estimated by hand calculations as well...
@@
     \caption{Photon beam size and divergence for selected ID16A beamline positions.
-Values below were obtained considering undulator U18.3 tuned to its first harmonic at the photon energy of $E=17225$ eV for both the high-$\beta$ (ESRF) and the EBS straight sections.
+Values below were obtained considering undulator U18.3 tuned to its first harmonic at the photon energy of $E=17225$~eV for both the high-$\beta$ (ESRF-1) and the EBS straight sections.
 Values are FWHM.
-Values in parentheses correspond to closing the VSS slit to 50$\mu$m (21.3 if approximating by the Gaussian $\sigma$).}
+Values in parentheses correspond to closing the VSS slit to 50~$\mu$m (21.3 if approximating by the Gaussian $\sigma$).}
     \resizebox{\textwidth}{!}{%
@@
     Position [m]:           & 0.00          & 40.00        & 185.00            & 0.00    & 40.00   & 185.00       \\
-    $FWHM_h$ [$\mu$m]:      & 977.19        & 404.00 (50.0)& 0.139 (0.017)     & 71.26   & 29.46   & 0.0102       \\
-    $FWHM_v$ [$\mu$m]:      & 9.60          & 482          & 0.0052            & 10.02   & 486.5   & 0.0054       \\
+    FWHM$_h$ [$\mu$m]:      & 977.19        & 404.00 (50.0)& 0.139 (0.017)     & 71.26   & 29.46   & 0.0102       \\
+    FWHM$_v$ [$\mu$m]:      & 9.60          & 482          & 0.0052            & 10.02   & 486.5   & 0.0054       \\
     \newline
-    $FWHM'_h$ [$\mu$rad]:   & 26.96         & 65.22        & 189062   & 15.60   & 37.74   & 109404       \\
-    $FWHM'_v$ [$\mu$rad]:   & 12.04         & 12.04                 & 22264    & 12.16   & 12.16   & 22489        \\                                                   
+    FWHM'$_h$ [$\mu$rad]:   & 26.96         & 65.22        & 189062   & 15.60   & 37.74   & 109404       \\
+    FWHM'$_v$ [$\mu$rad]:   & 12.04         & 12.04                 & 22264    & 12.16   & 12.16   & 22489        \\                                                   
     \end{tabular}%
@@
 
-Next, we want to estimate the number of photons at the sample.
+Next, we want to estimate the number of photons in the sample.
 For that, we need the spectrum at the source.
 It is not straightforward to get this information analytically but one can calculate the flux and power using different codes.
-Perhaps the most advanced and well maintained codes used in the synchrotron community are SPECTRA \cite{codeSPECTRA} and SRW \cite{codeSRW}.
+Perhaps the most advanced and well-maintained codes used in the synchrotron community are SPECTRA \cite{codeSPECTRA} and SRW \cite{codeSRW}.
 The XOPPY package in the OASYS suite implements three codes: US \cite{codeUS}, URGENT \cite{codeURGENT} and SRW.
-Figure~\ref{fig:FluxU18} shows the spectra calculated using SRW over an aperture of 1 $\times$ 1 mm$^2$ placed at $d$=27.1 m.
-This aperture opening is set to fully accept the central cone at the resonance peak of the harmonic in use ($\max(\Sigma') \times d = 0.73$ mm), with $\Sigma'$ values from in Table~\ref{tab:HandCalculations}.
-The peak intensity is about  $2.9\cdot10^{14}$ photons/s/0.1{\%}bw for EBS and 1.9 10$^{14}$ for ESRF-1 (see \ref{fig:FluxU18}).
-One can appreciate how the emittance plays an important role in both peak width and specially in peak value.
-The number of photons per second after the multilayer monochromator is 2.9 10$^{15}$ for EBS and 1.9 10$^{15}$ for ESRF-1, considering that the energy bandwidth transmitted by a multilayer monochromator is $\Delta E/E \approx 10^{-2}$ and the bandwidth in Fig.~\ref{fig:FluxU18} is 10$^{-3}$, and supposing, at this point, ideal reflectivity (100\%) for the multilayers.
+Figure~\ref{fig:FluxU18} shows the spectra calculated using SRW over an aperture of $1 \times 1$~mm$^2$ placed at $d=27.1$~m.
+This aperture opening is set to fully accept the central cone at the resonance peak of the harmonic in use ($\max(\Sigma') \times d = 0.73$~mm), with $\Sigma'$ values from in Table~\ref{tab:HandCalculations}.
+The peak intensity is about  $2.9\cdot10^{14}$ photons/s/0.1{\%}bw for EBS and $1.9\cdot10^{14}$ for ESRF-1 (see \ref{fig:FluxU18}).
+One can appreciate how the emittance plays an important role in both peak width and especially in peak value.
+The number of photons per second after the multilayer monochromator is $2.9\cdot10^{15}$ for EBS and $1.9\cdot10^{15}$ for ESRF-1, considering that the energy bandwidth transmitted by a multilayer monochromator is $\Delta E/E \approx 10^{-2}$ and the bandwidth in Fig.~\ref{fig:FluxU18} is 10$^{-3}$, and supposing, at this point, ideal reflectivity (100~\%) for the multilayers.
 
@@
 To estimate the geometrical transmission of the beamline we have to consider which optical elements crop the beam and remove photons.
-This can be analysed using their angular acceptance or numerical aperture.
+This can be analyzed using their angular acceptance or numerical aperture.
 The multilayer monochromator accepts the whole beam.
 We consider a VSS closed to $S_h=50~\mu$m in horizontal and fully opened in vertical.
-For ESRF its transmission is 1/8 (as discussed before) and 1 for EBS.
+For ESRF-1 its transmission is 1/8 (as discussed before) and 1 for EBS.
 In a similar way, the KB mirrors will crop the beam to a size that is the projection of their useful optical length onto a direction perpendicular to the optical axis.
-Using a mirror length of $L_v$=60 mm and $L_h$=26 mm for VFM and HFM, respectively, and incident angle of 15 mrad, the KB behaves as an aperture of $900 \times 390 \mu m^2$.
+Using a mirror length of $L_v=60$~mm and $L_h=26$~mm for VFM and HFM, respectively, and incident angle of 15~mrad, the KB behaves as an aperture of $900 \times 390~\mu$m$^2$.
 We get N.A.
-values of 6.2$\times$21.6 $\mu rad^2$ (aperture size divided by the distance from the mirror to the VSS in H and to the source in V).
+values of $6.2\times21.6~\mu$rad$^2$ (aperture size divided by the distance from the mirror to the VSS in H and to the source in V).
 The transmission coefficient is the ratio of the accepted N.A.
-over the beam divergence, resulting 9.5\% (H) and 0.18\% (V)for ESRF and 16.3\% and 0.18\% for EBS.
-The overall transmission is the product of H and V ratios, therefore 0.2\% for ESRF and 2.95\% for EBS, therefore 3.8 $10^{12}$ photons/s for ESRF and 8.5 $10^{13}$ photons/s for EBS.
+over the beam divergence, resulting in 9.5\% (H) and 0.18\% (V)for ESRF-1 and 16.3\% and 0.18\% for EBS.
+The overall transmission is the product of H and V ratios, therefore 0.2\% for ESRF-1 and 2.95\% for EBS, therefore $3.8\cdot10^{12}$ photons/s for ESRF-1 and $8.5\cdot10^{13}$ photons/s for EBS.
 %\todo{check with references $10^9$ at 33.6 keV.
@@
 We used the undulator peak value over the whole ML bandwidth, that overestimates the integrated flux.}
-%\todo{need to consider ML reflectivity for the 3 reflections, also undulator is not perfect, horizontal slit may be closed to less than $50 \mu m$, also KB mirrors may not be fully illuminated and slope error losses due to initial ML mono are neglected}
+%\todo{need to consider ML reflectivity for the 3 reflections, also undulator is not perfect, the horizontal slit may be closed to less than $50 \mu m$, also KB mirrors may not be fully illuminated and slope error losses due to initial ML mono are neglected}
 
 One can also make some simple inferences about the beam coherence.
-The coherent fraction (CF) \cite{arxivCF} of the undulator beam is the occupation of the lowest coherent mode, and will be discussed further in Section~\ref{level2}.
+The coherent fraction (CF) \cite{arxivCF} of the undulator beam is the occupation of the lowest coherent mode and will be discussed further in Section~\ref{level2}.
 An approximated estimation of CF can be obtained using Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas}.
@@
 and the total coherent fraction is the product of both (CF=CF$_h \times $CF$_v$).
-An estimation of the CF calculated in this way gives CF=0.12\% for the ESRF and 2.73\% for EBS at 17 keV photon energy.
+An estimation of the CF calculated in this way gives CF=0.12\% for the ESRF-1 and 2.73\% for EBS at 17 keV photon energy.
 The CF roughly indicates the ratio of the ``coherent photons'' over the total number of photons.
@@
 However, apertures centered at the optical axis would clean the beam from these undesired photons.
-If we compare the values of the beamline transmission due to geometrical aperturing  (0.2\% for ESRF-1 and 2.95\% for EBS) with the the CF values, we remark that they are very close.
-This suggests that the beam at the focal position will be highly coherent because the fraction of the photons removed from the beam matches the CF value that quantify the fraction of ``good'' or ``coherent'' photons.
+If we compare the values of the beamline transmission due to the geometrical aperture  (0.2\% for ESRF-1 and 2.95\% for EBS) with the CF values, we remark that they are very close.
+This suggests that the beam at the focal position will be highly coherent because the fraction of the photons removed from the beam matches the CF value that quantifies the fraction of ``good'' or ``coherent'' photons.
 
 To assess the focused beam size given by a coherent beam, one can consider the beamline in two sections: First, the ML monochromator and VSS that produce a coherent plane wave as the source (vertical) and VSS (horizontal) are both very far from the entrance aperture of the focusing optics.
-Second, the KB-system playing a dual role, acting as a focusing element of focal distance $f$=5 cm, but also cropping the beam by acting as a square aperture of $D \times D=$ 400$\times$400 $\mu m^2$ (the length of the mirror projected in a plane perpendicular to the beam propagation).
+Second, the KB-system playing a dual role, acting as a focusing element of focal distance $f=5$~cm, but also cropping the beam by acting as a square aperture of $D \times D=$ 400$\times$400~$\mu$m$^2$ (the length of the mirror projected in a plane perpendicular to the beam propagation).
 %\todo{REMOVE: we obtained 900 in H, that will be reduced by 0.5 as propagated from 10 cm to 5 cm upstream of the sample, and 390 in V approximated by 400}.
@@
 where $k=2\pi/\lambda$, and $x$ ($y$) is the horizontal (vertical) coordinate.
-Considering that the FWHM of $\sinc(x)$ is approximately 2.78, one obtains a FWHM of 7.75$\times$7.75 nm$^2$ for the focal spot.
-This spot size is of the order or larger than that calculated geometrical considerations alone, and therefore contributes significantly to the overall size of the focused beam.
-One can conclude that for the conditions in use this is a diffraction limited beamline.
-
-In summary, we have shown in this section how simple analytical calculations that can be done mostly ``by hand'' help to estimate the geometrical beam size at focal position (17$\times$5 nm$^2$ for ESRF, 10$\times$5 nm$^2$ for the EBS) the flux (3.8 $10^{12}$ photons/s for ESRF and 8.5 $10^{13}$ photons/s for EBS) and also anticipate that the beam is highly coherent producing a diffraction limited focal spot of  7.75$\times$7.75 nm$^2$.
+Considering that the FWHM of $\sinc(x)$ is approximately 2.78, one obtains a FWHM of 7.75$\times$7.75~nm$^2$ for the focal spot.
+This spot size is of the order or larger than that calculated geometrical considerations alone and therefore contributes significantly to the overall size of the focused beam.
+One can conclude that for the conditions in use this is a diffraction-limited beamline.
+
+In summary, we have shown in this section how simple analytical calculations that can be done mostly ``by hand'' help to estimate the geometrical beam size at focal position (17$\times$5~nm$^2$ for ESRF-1, 10$\times$5~nm$^2$ for the EBS) the flux ($3.8\cdot10^{12}$ photons/s for ESRF-1 and $8.5\cdot10^{13}$ photons/s for EBS) and also anticipate that the beam is highly coherent producing a diffraction-limited focal spot of  7.75$\times$7.75~nm$^2$.
 Thus far we ignored the effect of real optics exhibiting surface errors, the reflectivity of the elements, and other aspects that will be treated in depth in the next sections.
@@
 %-------------------------------------------------------------------------
-\subsection{Ray tracing calculations}
+\subsection{Ray-tracing calculations}
 \label{level1}
-In this section we perform ray tracing calculations to study the effect of real optics (including aberrations and slope errors) upon the beam size and beamline transmission.
+In this section, we perform ray-tracing calculations to study the effect of real optics (including aberrations and slope errors) upon the beam size and beamline transmission.
 The first subsection deals with the effect of aberrations and transmission due to geometrical considerations.
@@
 
-Ray tracing is a simulation method based on tracing some light rays along the optical system from the source to the image and retrieving the statistics of rays at the image to measure beam characteristics.
-For simple systems one can perform ray tracing even by hand, selecting some principal rays that will define the location and envelope of the image.
+Ray-tracing is a simulation method based on tracing some light rays along the optical system from the source to the image and retrieving the statistics of rays at the image to measure beam characteristics.
+For simple systems one can perform ray-tracing even by hand, selecting some principal rays that will define the location and envelope of the image.
 Using computers one can trace thousands of rays and calculate the position and divergence distribution of the rays using statistics.
 Rays are usually generated by Monte Carlo sampling the source characteristics.
-Ray tracing is based on geometrical optics: a ray is a solution of the Helmholtz equation and travels in vacuum along straight line.
+Ray-tracing is based on geometrical optics: a ray is a solution of the Helmholtz equation and travels in vacuum along a straight line.
 Rays are intercepted by the optical elements, which change their trajectory.
@@
 Because Snell's law uses the refractive index which is wavelength-dependent, it is then interesting to assign additional attributes to each ray, such as the wavelength.
-This will permit to calculate the refractive index that should  be used for each ray.
-The characteristics of each ray can also be extended to include some intensity, or better electric field amplitude.
-This allows to extend the pure geometrical tracing to include physical models that take into account the optical element reflectivity or transmission.
-This combination of using a geometrical model plus a physical model allows to simulate every element used in a synchrotron beamline, such as mirrors and combinations of them (e.g., KBs); lenses, compound refractive lenses and transfocators; gratings of any type and shape, crystal systems, etc.
-Ray tracing is a simple and extremely powerful technique for calculating the main characteristics of the photon beam (size, divergence, photon distribution) at every point of the beamline.
-Although very powerful commercial ray tracing packages are available in the market, they are generally ill adapted for synchrotron simulations because they do not cover the particularities of X-ray devices and sources.
+This will permit us to calculate the refractive index that should be used for each ray.
+The characteristics of each ray can also be extended to include the electric field amplitude.
+This allows extending the pure geometrical tracing to include physical models that take into account the optical element reflectivity or transmission.
+This combination of using a geometrical model plus a physical model allows to simulate every element used in a synchrotron beamline, such as mirrors and combinations of them (e.g., KBs); lenses, compound refractive lenses, and transfocators; gratings of any type and shape, crystal systems, etc.
+Ray-tracing is a simple and extremely powerful technique for calculating the main characteristics of the photon beam (size, divergence, photon distribution) at every point of the beamline.
+Although very powerful commercial ray-tracing packages are available in the market, they are generally ill-adapted for synchrotron simulations because they do not cover the particularities of X-ray devices and sources.
 For that purposes, several packages created in the synchrotron community and are available, like SHADOW \cite{codeSHADOW}, RAY \cite{codeRAY}, McXtrace \cite{codeMCXTRACE} or XRT \cite{codeXRT}.
-SHADOW is probably the one that is most widely used and cited.
-For this study we use SHADOW and its interface ShadowOUI \cite{codeSHADOWOUI}, available as a module of the OASYS suite.
-
-We performed ray tracing of the ID16A beamline using undulator source tuned ti have its resonance at 17225 eV, considering the ML as a cylindrical mirror, and using elliptical cylinder reflectors with finite dimensions as the KB focusing elements.
+For this study, we use SHADOW and its interface ShadowOUI \cite{codeSHADOWOUI}, available as a module of the OASYS suite.
+
+We performed ray-tracing of the ID16A beamline using an undulator source tuned to have its resonance at 17225~eV, considering the ML as a cylindrical mirror, and using elliptical cylinder reflectors with finite dimensions as the KB focusing elements.
 The reflector shape is chosen to be elliptical as this is the theoretical shape producing a point-to-point focusing.
-For a first calculation the reflectors (ML and KBs) are considered perfect, and have no slope errors.
-The results in Fig.~\ref{fig:ray tracing}a show two phenomena.
+For a first calculation, the reflectors (ML and KBs) are considered perfect and have no slope errors.
+The results in Fig.~\ref{fig:ray-tracing}a show two phenomena.
 One is that the migration from ESRF-1 lattice to the EBS will significantly improve the horizontal size, and the VSS is not needed.
@@
 The transmission values are 0.162\% for ESRF-1 and 2.22\% for EBS.
-These values confirm the analytical estimations (0.2\% for ESRF-1 and and 2.95\% for EBS).
+These values confirm the analytical estimations (0.2\% for ESRF-1 and 2.95\% for EBS).
 As mentioned before, these values are similar to the coherent fraction \cite{arxivCF} (0.12 for ESRF-1 and 2.73\% for EBS), meaning that a very high coherence of the beam is expected, therefore it is expected some broadening of the beam due to diffraction effects originated by the mirrors acceptance.
@@
 \begin{figure}
-\label{fig:ray tracing}
+\label{fig:ray-tracing}
 \centering
@@
 The first row has been calculated using the ideal reflectors (cylindrical for ML, elliptical for the BK mirrors) without slope errors. 
-The second row shows the image calculated the hybrid option (ray tracing plus coherent diffraction) also for ideal reflectors with no slope errors.
+The second row shows the image calculated the hybrid option (ray-tracing plus coherent diffraction) also for ideal reflectors with no slope errors.
 It shows the focal spot degradation due to the diffraction arising from the beam aperturing by the KB mirrors.
@@
 
-The HYBRID method \cite{hybrid} uses geometric ray tracing combined with wavefront propagation.
-The code computes diffraction effects when the beam is apertured and can also simulate the effect of mirror figure errors when diffraction is present.
-The method can be applied to an entire beamline by simulating each optical element iteratively under the hypothesis of the validity of far field propagation in all intermediate propagations.
+The hybrid method \cite{hybrid} uses geometric ray-tracing combined with wavefront propagation.
+The code computes diffraction effects when the beam is cropped and can also simulate the effect of mirror figure errors when diffraction is present.
+The method can be applied to an entire beamline by simulating each optical element iteratively under the hypothesis of the validity of far-field propagation in all intermediate propagations.
 If this is not the case, a near field version is available for investigating imaging by individual optics.
-The interest of the hybrid method is to assess in a fast way the influence of beam coherence because the code is considerably faster than the the full methods for dealing with the partial coherence of the synchrotron beam that are described below.
-Fig.~\ref{fig:ray tracing}b presents the results using the hybrid model for the beamline with ideal elliptical optical surfaces.
-When compared with the pure ray tracing in Fig.~\ref{fig:ray tracing}a it can be appreciated that the spot is broadened.
+The interest of the hybrid method is to assess in a fast way the influence of beam coherence because the code is considerably faster than the full methods for dealing with the partial coherence of the synchrotron beam that are described below.
+Fig.~\ref{fig:ray-tracing}b presents the results using the hybrid model for the beamline with ideal elliptical optical surfaces.
+When compared with the pure ray-tracing in Fig.~\ref{fig:ray-tracing}a it can be appreciated that the spot is broadened.
 This is more evident in the vertical direction, because of the smaller focal size.
-The spot size is about 7-8 nm in V for both storage ring lattices, and in horizontal is about 15$\mu$m for ESRF-1 High$\beta$ and 11 nm for EBS.
+The spot size is about 7-8~nm in V for both storage ring lattices, and in horizontal is about 15~nm for ESRF-1 High$\beta$ and 11~nm for EBS.
 It is interesting to note the almost round spot that will be obtained with the new EBS lattice.
@@
 It is possible to infer the effect of mirror imperfections and in particular slope errors by only knowing some statistical parameters such as the RMS slope error and then building synthetic surface profiles to be used in the simulations.
+It is, though, highly recommended to use metrology data from real mirrors, when possible.
+When not available, one can use data from existing mirrors from a database (e.g., \cite{dabam}) and extrapolate some of the effects to the hypothetical new mirrors.
 This is the typical approach used when designing a new beamline where the mirrors do not exist yet.
-However, in our case the mirrors already exist and have been extensively characterized at the ESRF optical metrology laboratory.
-It is therefore highly recommended to use metrology data from real mirrors, when possible.
-When not available, one can use data from existing mirrors from a database (e.g., \cite{dabam}) and extrapolate some of the effects to the hypothetical new mirrors.
+However, in our case, the mirrors already exist and have been extensively characterized at the ESRF optical metrology laboratory.
 The metrology data of the mirrors for the beamline under study are presented in Appendix~\ref{appendix:metrology}.
 These experimental data are used in all the simulations in this paper.
-Simulations using the hybrid model including slope errors are shown in Fig.~\ref{fig:ray tracing}c.
-A small broadening of the peak due to slope errors can be observed, but, more importantly, a new satellite structure close to the vertical intensity peak is produced by the slope errors.
+Simulations using the hybrid model including slope errors are shown in Fig.~\ref{fig:ray-tracing}c.
+A small broadening of the peak due to slope errors can be observed, but, more importantly, a new satellite structure close to the vertical intensity peak is produced.
 
@@
 
-In this section the coherence properties of the beamline are analysed using, again, different approaches  distinguished by their complexity and necessary computer resources.
+In this section the coherence properties of the beamline are analyzed using, again, different approaches distinguished by their complexity and necessary computer resources.
 We first analyze the beamline supposing a fully coherent beam.
 Subection~\ref{wofry} shows a simplified 1D model.
-Subsection~\ref{srw_se} uses a full 2D model with accurate simulation of the undulator emission but without considering storage ring emittance (i.e.
-again using a fully coherent beam).
-These simplifications permit the simulation of diffraction effects by apertures and slope/figure errors.
-Subsequently we introduce methods to treat the partial coherence originated by the storage ring emittance.
+Subsection~\ref{srw_se} uses a full 2D model with accurate simulation of the undulator emission but without considering storage ring emittance.
+These simplifications allow the simulation of diffraction effects by apertures and slope/figure errors.
+Subsequently, we introduce methods to treat the partial coherence originated by the storage ring emittance.
 Two variants are shown, a Monte Carlo approach (subsection~\ref{srw_me}) that permits us to calculate in great detail the intensity distribution of the focused beam, and a decomposition of the radiation in coherent modes (subsection~\ref{comsyl}) that allows quantitative evaluation of the coherent beam characteristics.
@@
 
-The very long  source-optics distance combined with a small source (in particular for EBS) makes this beamline suitable for experimental techniques requiring a highly coherent beam.
-Indeed, it has been shown that the clipping of the beam, mainly due to the acceptance of the KB mirrors produces a significant broadening of the focal spot.
-As discussed in Section~\ref{level0}, very simple calculation of the diffraction limit can be done using a convergent spherical beam clipped by a slit of 400$\times$400$ \mu$m$^2$ placed 5cm upstream of the focus (at the second KB mirror position), producing a spot of about 7.75$\times$7.75 nm$^2$, therefore qualifying this beamline to be a diffraction-limited beamline.
-We perform here a simplified wavefront simulation in 1D, thus separating the simulations of the horizontal and vertical planes.
+As discussed in Section~\ref{level0}, the clipping of the beam, mainly due to the acceptance of the KB mirrors produces a significant broadening of the focal spot.
+Very simple calculation of the diffraction limit can be done using a convergent spherical beam clipped by a slit of 400$\times$400$~\mu$m$^2$ placed 5~cm upstream of the focus (at the second KB mirror position), producing a spot of about 7.75$\times$7.75~nm$^2$, therefore qualifying this beamline to be a diffraction-limited beamline.
+We perform here a simplified wavefront simulation in 1D, thus decoupling the simulations of the horizontal and vertical planes.
 The source can be simplified by using a point source (spherical wavefront) and applying a Gaussian intensity profile with $\sigma=\sigma'_u d$ with $\sigma'_u$ given by Eq.~\ref{eq:photon small sigmas} and $d$ the distance from the point source to the observation plane.
-Each optical element can be simplified by separating its action in two parts: the focusing behaviour implemented as an ideal lens, and the finite size because of the boundaries, implemented as a slit.
-The WOFRY package in OASYS permits such simulations, and it is possible to achieve satisfactory results in a few caluclation iterations.
-The beam size at different positions of the beamline are shown in Table~\ref{tablewofry1D}.
+Each optical element can be simplified by separating its action into two parts: the focusing behavior implemented as an ideal lens, and the finite size because of the boundaries, implemented as a slit.
+The WOFRY package \cite{codeWOFRY} in OASYS permits such simulations, and it is possible to achieve satisfactory results in a few calculation iterations.
+The beam size at different positions of the beamline is shown in Table~\ref{tablewofry1D}.
 The final image profiles are in Fig.~\ref{wofry1D}.
@@
 A next step in the wave optics simulation is, assuming full coherence, to generate a 2D source wavefront that matches in detail the real source (an undulator in a storage ring) and then simulate the propagation of such a wavefront through the optical elements with models that match well the characteristics of the real elements.
-The source simulated is the on-axis-single-electron emission from the U18.3 undulator tuned to its first harmonic at $E=17225$ eV.
-The emission by a single-electron is fully coherent.
+The emission by a filament beam is fully coherent.
 Once the source is defined, an approach of how best to simulate the beamline optical elements.
 Increasing the complexity of the simulations one can: a) use ideal thin lenses preceded by slits to emulate finite apertures (as done in the 1D model presented in Section~\ref{wofry}); or b) use grazing incidence focusing optics models \cite{Canestrari2014}.
-Furthermore, one can introduce the effect of imperfect optics to the simulations, i.e.
-slope/figure errors.
-
-The computer code chosen for these calculations is Synchrotron Radiation Workshop (\textsc{srw}) \cite{codeSRW}, which over the past two decades has been extensively benchmarked throughout different synchrotron radiation facilities and has become widely accepted within the X-ray community.
+Furthermore, one can introduce the effect of imperfect optics on the simulations, i.e.
+slope/figure errors to both ideal lenses and mirrors.
+
+The computer code chosen for these calculations is ``Synchrotron Radiation Workshop'' (SRW) \cite{codeSRW}, which over the past two decades has been extensively benchmarked throughout different synchrotron radiation facilities and has become widely accepted within the X-ray community.
 Figure \ref{fig:SingleElectron} shows the horizontal integrated intensity at the focal position for the ID16A.
-It can been seen that there is little difference between using thin ideal lenses or the elliptical mirror model.
+It can be seen that there is little difference between using thin ideal lenses or the elliptical mirror model.
 The elliptical mirror presents a slight reduction in intensity and asymmetry but no significant differences are observed, which is in accordance with the findings of \cite{Canestrari2014}.
 Notice that the differences are minimal and to accentuate them we have chosen logarithmically scaled plots to represent the intensity distributions at the focal position.
-It is also important to mention that this equivalency between ideal lenses and elliptical mirrors is only valid close to the focal position: down- or upstream the focal position, differences are no longer negligible \cite{Canestrari2014}.
+It is also important to mention that this equivalency between ideal lenses and elliptical mirrors is only valid close to the focal position: down- or upstream the focal position, differences are no longer negligible - c.f.
+Fig.~5 in \cite{Canestrari2014}.
 
@@
 They are almost completely removed and associated with a large increase of the background level: the radiation found on the main lobe (focus) shifts to the side lobes when the slope errors are considered.
-This is accompanied by an obvious reduction on the peak intensity.
+This is accompanied by an obvious reduction in the peak intensity.
 Table \ref{tab:SRW_results} compiles the beam sizes at focal position for mentioned simulations.
@@
     \label{fig:SingleElectron}
-    \caption{Intensity profile at the sample position produced by the single electron emission propagated through the ID16A beamline.
+    \caption{Intensity profile at the sample position produced by the single-electron emission propagated through the ID16A beamline.
 Top: the perfect focusing with ideal lenses (black) is compared with the focusing using elliptical mirrors (red).
@@
 \begin{tabular}{ccccc}
-Focusing elements                 & Plane      &  Single electron & EBS         & ESRF       \\
+Focusing elements                 & Plane      &  Single electron & EBS         & ESRF-1       \\
 ideal lenses/ elliptical mirrors  & Horizontal &  7.94/7.97       & 12.20/12.70 & 17.50/16.80 \\
@@
 
-\subsubsection{Full wave optics simulation: partial coherent case by means of multielectron Monte Carlo sampling}
+\subsubsection{Full wave optics simulation: partial coherent case by means of multi-electron Monte Carlo sampling}
 \label{srw_me}
-The wavefront propagation methodology is based in the sampling of a single wavefront (monochromatic and coherent) in a 2D space (in horizontal and vertical spatial dimensions) at a given point of the beamline.
+The wavefront propagation methodology is based on the sampling of a single wavefront (monochromatic and coherent) in a 2D space (in horizontal and vertical spatial dimensions) at a given point of the beamline.
 This initial point considered as source cannot be the source physical position, because i) when considering point or spherical waves all the intensity would be concentrated in a single pixel so the wavefront is ill-defined, ii) when using synchrotron sources (i.e., undulators) theory \cite{jackson} permits the calculation of the radiation in positions far from the electron trajectory.
@@
 But even for new ultra low-emittance storage rings (improperly called diffraction-limited sources, see discussion in \cite{arxivCF}) the contribution of the electron dimensions ($\sigma_{e(h,v)}$ and $\sigma'_{e(h,v)}$) is not negligeable as compared with the natural emission of radiation ($\sigma_u$ and $\sigma'_u$ for undulators as in Eq.~\ref{eq:photon small sigmas}).
-To account for this effect one should complete or extend the wavefront simulations to take into account that in fact not only a single wavefront contributes to the properties of the beam.
+To account for this effect one should complete or extend the wavefront simulations to take into account that in fact, not only a single wavefront contributes to the properties of the beam.
 
@@
 
-There are three approaches currently used for including the electron emittance effects in the waveoptics simulations: 1) use of convolutions, 2) Monte Carlo multi-electron sampling, and 3) Propagation of coherent modes.
+There are three approaches currently used for including the electron emittance effects in the wave optics simulations: 1) use of convolutions, 2) Monte Carlo multi-electron sampling, and 3) Propagation of coherent modes.
 The convolution method consists of calculating the emission characteristics and then convolving the resulting intensity with Gaussians describing the electron beam.
 This method is simple and fast, as only requires simulation of a single wavefront.
-However, it presents two important limitations: i) it can only be applied to the intensity calculations and not to other parameters of interest for describing coherence, such as degree of coherence, etc., and ii) it can be applied only at the source position: it is usually unknown how the different optical elements will affect the source contribution to the phase space.
+However, it presents two important limitations: i) it can only be applied to the intensity calculations and not to other parameters of interest for describing coherence, such as the degree of coherence, etc., and ii) it can be applied only at the source position: it is usually unknown how the different optical elements will affect the source contribution to the phase space.
 The next method, the Monte Carlo multi-electron propagation,  is a statistical method based on the idea that one can exactly calculate the emission and its propagation for one electron that is deterministically described by its initial conditions at the entrance of the magnetic structure.
-It is therefore possible to perform many simulations for different initial conditions of the travelling electron and make the propagation along the beamline until the observation plane.
+It is, therefore, possible to perform many simulations for different initial conditions of the traveling electron and make the propagation along the beamline until the observation plane.
 Then one can statistically combine the different wavefronts (one for each electron) to obtain a sampling of the intensity and other magnitudes useful to describe partially coherent beams.
@@
 The second is computing resources.
-The emission of each electron must be calculated at the source level, propagated along every element of the beamline, and the resulting wavefront must be analyzed to extract the information to be scored and stored.
-The third method, the propagation of coherent modes wil be discussed in the next section.
-
-The Monte Carlo multielectron method has been proposed by O. Chubar and it is fully implemented in SRW \cite{codeSRW}. We have used this code for the calculations presented in this section.
-
-The source used for simulations is the U18.3 undulator tuned to its first harmonic at $E=17225$ eV with a finite emittance electron beam (see Table \ref{tab:eBeam}).
+The emission of each electron must be calculated at the source level, propagated through every element of the beamline, and the resulting wavefront must be analyzed to extract the information to be scored and stored.
+In the third method, the propagation of coherent modes will be discussed in the next section.
+The Monte Carlo multi-electron method has been proposed by O. Chubar and it is fully implemented in SRW \cite{multiE}. We have used this code for the calculations presented in this section.
+
+The source used for simulations is the U18.3 undulator tuned to its first harmonic at $E=17225$~eV with a finite emittance electron beam (see Table \ref{tab:eBeam}).
 Two set of parameters were used for the electron beam: the machine parameters of the ESRF-1 lattice (high-$\beta$) and the new EBS straight sections (Tables \ref{tab:TableSources} and \ref{tab:eBeam}).
-Simulations for both storage ring lattices were carried out modelling the beamline using thin ideal lenses and elliptical mirror models with and without the addition of the slope errors.
+Simulations for both storage ring lattices were carried out modeling the beamline using thin ideal lenses and elliptical mirror models with and without the addition of the slope errors.
 Results are presented in Figures \ref{fig:FiniteEmittanceA} and \ref{fig:FiniteEmittanceB}; and Table \ref{tab:SRW_results}.
 
-The same observations made in the previous section (fully coherent simulations) are pertinent here: there is a reduction of the peak intensity when comparing simple lens model with the elliptical mirror.
+The same observations made in the previous section (fully coherent simulations) are pertinent here: there is a reduction of the peak intensity when comparing the simple lens model with the elliptical mirror.
 The beam sizes and shapes are similar at the focal position, only.
@@
 % 
-% One of the mentioned drawbacks of the Monte Carlo multielectron methods is the a-priori unability to determine the size of sampling space (i.e., how many electrons) for accurate calculations. 
+% One of the mentioned drawbacks of the Monte Carlo multi-electron methods is the a-priori unability to determine the size of sampling space (i.e., how many electrons) for accurate calculations. 
 % Another important limitation is the impossibility with available computer resources to store (although can be computed) the full 4D matrices fully describing the partial coherence properties (cross spectral density) at any point of the beamline.
 
-The cross spectral density (CSD) (sometimes called mutual intensity) completely describes the coherence of a beam. 
+The cross-spectral density (CSD) (sometimes called mutual intensity) completely describes the coherence of a beam. 
 At a given plane at a distance $z$ from the source, for a photon beam of frequency $\omega$, the CSD is the correlation of the radiated electric fields between two spatial points $(x_1,y_1)$ and $(x_2,y_2)$:
@@
 that measures the coherence of the beam between two points, and ranges from fully incoherent ($\mu=0$) to fully coherent ($\mu$=1). 
-Eq.~\ref{CSD} can in principle be calculated using the Kim's convolution theorem \cite{kim1986} (see \cite{geloni2008} for a full discussion on the validity) but its storage is prohibitif in present computers: supposing sampling one spatial dimension by $N\approx1000$ pixels, a wavefront will contain $N^2$ pizels, and the CSD will be sampled over $N^4$ points.
-They are complex numbers (16 bits) so the only stockage is of the order of GB or TB.
-Considering also that the progagation of $W$ from a beamline position $z_1$ to another position $z_2$ will need 10$^8$ to 10$^{12}$ 4D integrals, one can easily imagine the physical impossibility to perform such calculations.
+Eq.~\ref{CSD} can in principle be calculated using Kim's convolution theorem \cite{kim1986} (see \cite{geloni2008} for a full discussion on the validity) but its storage is prohibitive in present computers: supposing sampling one spatial dimension by $N\approx1000$ pixels, a wavefront will contain $N^2$ pixels, and the CSD will be sampled over $N^4$ points.
+They are complex numbers (16 bits) so the only storage is of the order of GB or TB.
+Considering also that the propagation of $W$ from a beamline position $z_1$ to another position $z_2$ will need 10$^8$ to 10$^{12}$ 4D integrals, one can easily imagine the physical impossibility to perform such calculations.
 
@@
 The innovative concept is the numerical description of the undulator CSD in terms of its coherent modes $\Phi_m(x,y;z,\omega)$ and eigenvalues $\lambda_m $.
-One can define the occupation of the coherent modes as
+One can define the occupation of coherent modes as
 \begin{equation}
@@
 The first or lower term (index zero) is the most important coherent mode.
-It permits definition of the coherent fraction (CF) as the occupation of the lower coherent mode: CF=$d_0$.
-Obviously, a more coherent source will require less modes than a mostly incoherent source to approximate the CSD to a given precision.
-
-We performed COMSYL calculations for the U18.3 undulator radiation of ID16A of 1.4 m length using both the EBS and ESRF-1 lattices.
-The obtained CF are 2.8\% for EBS and 0.13\% for ESRF-1 High$\beta$, indicating that the overall coherent flux emitted by EBS will be about 20 times the existing coherent flux with ESRF-High$\beta$.
+It permits the definition of the coherent fraction (CF) as the occupation of the lower coherent mode: CF=$d_0$.
+Obviously, a more coherent source will require fewer modes than a mostly incoherent source to approximate the CSD to a given precision.
+
+We performed COMSYL calculations for the U18.3 undulator radiation of ID16A of 1.4~m length using both the EBS and ESRF-1 lattices.
+The obtained CF is 2.8\% for EBS and 0.13\% for ESRF-1 High$\beta$, indicating that the overall coherent flux emitted by EBS will be about 20 times the existing coherent flux with ESRF-High$\beta$.
 Fig.~\ref{fig:histomodes}a shows the cumulated occupation of the coherent modes for the full central cone undulator emission.
-It shows that the EBS cumulated occupation grows very steeply with the number of coherent modes, whereas for ESRF the number of coherent modes needed to attain a given value of cumulated occupation much higher.
-This can also be shown in Fig.\ref{fig:histomodes}b, where the number of modes to fill 50\%, 75\% and 95\% of spectral density are compared for EBS and ESRF.
+It shows that the EBS cumulated occupation grows very steeply with the number of coherent modes, whereas for ESRF-1 the number of coherent modes needed to attain a given value of cumulated occupation much higher.
+This can also be shown in Fig.\ref{fig:histomodes}b, where the number of modes to fill 50\%, 75\% and 95\% of spectral density is compared for EBS and ESRF-1.
 
@@
 The spectral density at the EBS photon source is represented in Fig.~\ref{fig:spectraldensity}a.
-It shows an elliptical shape with horizontal size larger than the vertical.
+It shows an elliptical shape with a horizontal size larger than the vertical.
 The spectral density is reproduced by the addition of the coherent modes weighted by the eigenvectors (from Eq.~\ref{decomposition})
@@
 It can be shown that the first mode is centered on the axis but extends over a spatial region smaller than the spectral density.
-The next mode ($m$=1) contributes to extend the spectral density in horizontal, but presents zero intensity at the origin.
+The next mode ($m$=1) contributes to extend the spectral density in horizontal but presents zero intensity at the origin.
 Successive modes extend the spectral density in horizontal, sometimes presenting a lobe center at zero and sometimes not (e.g., $m$=3).
@@
 This task is performed by the beamline.
-The beamline elements modify the different modes in a different manner, but it will never be possible to remove all the photons except from the lowest coherent mode, because there is an overlapping of the intensity of the lowest mode with the other ones.
+The beamline elements modify the different modes in a different manner, but it will never be possible to remove all the photons except from the lowest coherent mode because there is an overlapping of the intensity of the lowest mode with the other ones.
 Therefore it is impossible to obtain a fully coherent beam containing only a single mode.
@@
 %We recall that in the coherent mode expansion, the modes are orthonormal, therefore the integrated intensity for any mode is always one at the source. 
-As discussed, coherent modes will be ``cut'' in a different manner, depending on the slits aperturing effect.
-This implies that the Eq.~\ref{eq:mode normalization} is not longer satisfied, and for the propagated modes $M_{mm}\le$1 and $M_{mn}\ne$0 for $m\ne n$.
+As discussed, coherent modes will be ``cut'' in a different manner, depending on the effect of the slit.
+This implies that the Eq.~\ref{eq:mode normalization} is no longer satisfied, and for the propagated modes $M_{mm}\le$1 and $M_{mn}\ne$0 for $m\ne n$.
 Therefore, the coherent modes of the source after propagation and cropping by the beamline are no longer coherent modes of the resulting beam.
@@
 As commented before, the coherent modes at the source, propagated and cropped by the beamline, are not longer orthonormal, therefore they cannot be called coherent modes of the resulting beam.
-At this point, the new coherent modes can be calculated applying a new coherent mode decomposition on the cross spectral density of the transmitted beam, that is known from source coherent mode expansion.
-Making this numerical calculation with COMSYL, one obtains a mode occupation spectrum (see Fig.~\ref{fig:rediagonalization}) where the lowest mode accounts for the 83\% of the total spectral density.
-The next mode accounts for only 13\% and the rest are practically negligeable.
+At this point, the new coherent modes can be calculated by applying a new coherent mode decomposition on the cross-spectral density of the transmitted beam, which is known from source coherent mode expansion.
+Making this numerical calculation with COMSYL, one obtains a mode occupation spectrum (see Fig.~\ref{fig:rediagonalization}) where the lowest mode accounts for 83\% of the total spectral density.
+The next mode accounts for only 13\% and the rest is practically negligible.
 The intensity of these two coherent modes is shown in Fig.~\ref{fig:final modes}.
-The first mode has FWHM dimensions of 7.3$\times$6.7 nm$^2$.
-The spectral density, that is practically given by the contribution of the first two modes has a FWHM equal to the $m$=0 mode because, in horizontal the $m$=1 mode does contribute only to the tails and in vertical is identical to mode $m$=0.
-
-This analysis confirms the quantitative predictions discussed in the previous sections, but give a quantitative figure of the coherence of the beam, a  coherent fraction of CF=83\% that measures the "coherence purity".
+The first mode has FWHM dimensions of 7.3$\times$6.7~nm$^2$.
+The spectral density, that is practically given by the contribution of the first two modes has a FWHM equal to the $m$=0 mode because in horizontal the $m$=1 mode does contribute only to the tails and in vertical is identical to mode $m$=0.
+
+This analysis confirms the quantitative predictions discussed in the previous sections, but gives a quantitative figure of the coherence of the beam, a  coherent fraction of CF=83\% that measures the ``coherence purity''.
 
@@
     \caption{Coherent modes of the focused beam.
-Top: lower mode ($m$=0), with 83\% occupation and FWHM of 7.3$\times$6.7 nm$^2$.
-Botton: mode $m$=1, with 13\% occupation and FWHM of 19.0$\times$6.7 nm$^2$}
+Top: lower mode ($m$=0), with 83\% occupation and FWHM of 7.3$\times$6.7~nm$^2$.
+Botton: mode $m$=1, with 13\% occupation and FWHM of 19.0$\times$6.7~nm$^2$}
 \end{figure}
@@
 
-After the complete analysis done in the previous section on how the ID16A optics modify the characteristics of the x-ray beam, one can verify that the analytical previsions (Section~\ref{level0}) were a good estimator of the values then obtained using more sophisticated methods.
-The size of the focused beam (e.g, 10x5 for EBS) and the transmission (3\%) calculated analytically are in good agreement with geometric ray tracing (8.4x4.8, and 2.28\%).
-When characterising beam sizes, one usually assume that the intensity distributions are Gaussian and therefore they are well characterised by the FWHM value.
-From Fig.~\ref{fig:ray tracing} one can see that the horizontal intensity distribution is not Gaussian for the ESRF, as a result of cropping the beam with the VSS opened to 50$\mu$m.
-Although the electron beam statistics is in a very good approximations Gaussian, the undulator emission is not.
-In our ray tracing simulations we always suppose that we work at the resonance, then we applied the Gaussian approximation used in Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas} to compute the geometrical characteristics of the source.
+After the complete analysis done in the previous section on how the ID16A optics modify the characteristics of the X-ray beam, one can verify that the analytical previsions (Section~\ref{level0}) were a good estimator of the values obtained using more sophisticated methods.
+The size of the focused beam (e.g, 10x5 for EBS) and the transmission (3\%) calculated analytically are in good agreement with geometric ray-tracing (8.4x4.8, and 2.28\%).
+When characterizing beam sizes, one usually assumes that the intensity distributions are Gaussian and therefore they are well characterized by the FWHM value.
+From Fig.~\ref{fig:ray-tracing} one can see that the horizontal intensity distribution is not Gaussian for the ESRF-1, as a result of cropping the beam with the VSS opened to 50$\mu$m.
+Although the electron beam statistics is in a very good approximation Gaussian, the undulator emission is not.
+In our ray-tracing simulations, we always suppose that we work at the resonance, then we applied the Gaussian approximation used in Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas} to compute the geometrical characteristics of the source.
 This is certainly acceptable at the resonance but will fail out of resonance.
-Indeed, these lack of Gaussianity is relevant in the calculations of the coherent fraction.
-We approximate the coherence fraction by Eq.~\ref{eq:coherent fraction}, which for zero emittance ($\sigma_{x,y}=\sigma_{x',y'}$=0) becomes CF=1, as expected because a zero emittance electron beam (a single electron or a filament beam) that travel in an magnetic field emits a fully coherent wavefront.
+Indeed, this lack of Gaussianity is relevant in the calculations of the coherent fraction.
+We approximate the coherence fraction by Eq.~\ref{eq:coherent fraction}, which for zero emittance ($\sigma_{x,y}=\sigma_{x',y'}$=0) becomes CF=1, as expected because a zero emittance electron beam (a single-electron or a filament beam) that travel in a magnetic field emits a fully coherent wavefront.
 However, the photon source for a filament beam would have an emittance of $\sigma_u \sigma_{u'} \approx \lambda / 2 \pi$ (using Eqs.~\ref{eq:photon small sigmas}), which is in contradiction with theory, that says that a pure coherent state has emittance of $\lambda / 4 \pi$, as happens for a Gaussian photon beam.
@@
 However, its calculation requires to perform the coherent mode decomposition, as done in Section~\ref{comsyl} which is very costly.
-It is therefore the case to find the compromise between accuracy and cost, or where in the hierarchical series of methodologies proposed one should stop.
-
-The combination of ray tracing and wave optics methods used indicate that the focalized beam size due to  demagnification with mirrors (giving for EBS 10$\times$5 nm$^2$ analytically, 8.4$\times$4.8 nm$^2$ with ray tracing) has to be corrected with diffraction effect (8.4$\times$7.8 nm$^2$ hybrid, 12.7$\times$9.2 nm$^2$ SRW) and then add the slope error effect.
-The slope errors do not to blow up the size and kit it in the specifications range (10.8$\times$8.2 nm$^2$ hybrid 13.6$\times$9.9 nm$^2$ SRW).
-It is important to treat the the slope errors correctly.
-Slope errors can be included in the ray tracing using the data in Appendix~\ref{appendix:metrology} and applying specular reflection on the optical surface with errors.
-This is the standard mode of including slope errors in SHADOW, using measured profiles or synthetic ones.
-It works well for the case of incoherent beams, but in our case, because of the high coherence of the beam the ray tracing method (specular reflection) overestimates the blowing of the focal spot conducting to wrong results.
+It is, therefore, the case to find the compromise between accuracy and cost, or where in the hierarchical series of methodologies proposed one should stop.
+
+The combination of ray-tracing and wave optics methods used indicate that the focalized beam size due to  demagnification with mirrors (giving for EBS 10$\times$5~nm$^2$ analytically, 8.4$\times$4.8~nm$^2$ with ray-tracing) has to be corrected with diffraction effect (8.4$\times$7.8~nm$^2$ hybrid, 12.7$\times$9.2~nm$^2$ SRW) and then add the slope error effect.
+The slope errors do not blow up the size and kit it in the specifications range (10.8$\times$8.2~nm$^2$ hybrid 13.6$\times$9.9~nm$^2$ SRW).
+It is important to treat the slope errors correctly.
+Slope errors can be included in the ray-tracing using the data in Appendix~\ref{appendix:metrology} and applying specular reflection on the optical surface with errors.
+This is the standard model of including slope errors in SHADOW, using measured profiles or synthetic ones.
+It works well for the case of incoherent beams, but in our case, because of the high coherence of the beam the ray-tracing method (specular reflection) overestimates the blowing of the focal spot conducting to wrong results.
 Therefore the slope errors of the KB system have been considered using a ``wave optics approach'', as explained in Ref.~\cite{hybrid}.
@@
 
-% The first step in the wave optics simulation is to assume a full coherent source and simulate the propagation of such wavefront through the beamline.
-
-The wavefront simulation of the beamline include propagation of the beam from element to element (drift spaces) using different possible propagators.
+% The first step in the wave optics simulation is to assume a full coherent source and simulate the propagation of such a wavefront through the beamline.
+
+The wavefront simulation of the beamline includes the propagation of the beam from element to element (drift spaces) using different possible propagators.
 A propagator based on the Freskel-Kirchhoff integral is used.
 The applications of this (and other) operators imply the calculation of an integral (or sum) at each pixel of the image plane.
-This is costly from the computer point of view, so Fourier Transforms (Fourier Optics) and its Fast Fourier Transform implementation are used to reduce amount of operations and the calculation time.
+This is costly from the computer point of view, so Fourier Transforms (Fourier Optics) and its Fast Fourier Transform implementation are used to reducing the number of operations and the calculation time.
 This is practically the only choice for propagating 2D wavefronts.
-The use of Fourier Transforms adds another problem: the result is very dependent on the sampling of the wavefront (i.e., pixel size, number of points, dimension of the window where the wavefront is defined).
+The use of Fourier Transforms adds another problem: the result is very dependent on the sampling of the wavefront (i.e., pixel size, number of points, the dimension of the window where the wavefront is defined).
 This means the usually one has to test many configurations before finding one that works reasonably well (even though one can never guarantee the complete accuracy of the sampling used).
 A full wavefront optics simulation is an iterative process, where the user has to refine the sampling parameters element by element to be sure the wavefront at each position is sampled correctly.
-This is a time consuming work in particular for 2D simulation.
+This is a time-consuming work in particular for 2D simulation.
 We found very useful to start a wavefront simulation in 1D, separating the horizontal and vertical simulations, using a simplified system with ideal sources and ideal focusing elements.
-The use of 1D propagation is justified because in in synchrotron beamlines the horizontal and vertical coupling is usually small.
-This 1D simulations help in refining the choice of propagator and its parameters (resampling, zoom factors) that can be reused in a full 2D simulations.
-
-Regarding computational effort, simulations concerning ray tracing (Section~\ref{level1}) and coherent optics (Sections~\ref{wofry} and \ref{srw_se}) can be done interactively in a normal laptop.
+The use of 1D propagation is justified because in synchrotron beamlines the horizontal and vertical coupling is usually small.
+These 1D simulations help in refining the choice of propagator and its parameters (resampling, zoom factors) that can be reused in full 2D simulations.
+
+Regarding computational effort, simulations concerning ray-tracing (Section~\ref{level1}) and coherent optics (Sections~\ref{wofry} and \ref{srw_se}) can be done interactively in a normal laptop.
 However, for those concerning partial coherence, very long calculations are needed.
-For example, Monte Carlo calculations in Section~\ref{srw_me} for the multi electron case spent a few hours on a 56-core computer cluster for simulating \inred{NN} electrons.
-For performing the COMSYL coherent mode decomposition the calculation time depends very much on the dimensions on the problem.
+For example, Monte Carlo calculations in Section~\ref{srw_me} for the multi-electron case spent a few hours on a 56-core computer cluster for simulating 22k electrons.
+For performing the COMSYL coherent mode decomposition the calculation time depends very much on the dimensions of the problem.
 For EBS the wavefronts were sampled using 1007$\times$335 pixels, and 1103 modes were computed representing 98\% of the spectral density.
 This represents 5 MB per wavefront and a matrix of 1695 GB to diagonalize.
-It took less that four hours in the a cluster of 28 cores.
+It took less than four hours in a cluster of 28 cores.
 The calculation is more expensive if one wants to decompose a source that is pretty incoherent like our ESRF-High-$\beta$.
@@
 We obtained 4016 modes totalizing only 86\% of the spectral density.
-The calculation took more than 4 days in a 84 cores cluster.
+The calculation took more than 4 days in an 84 core cluster.
 This was done as a test case, to demonstrate the feasibility of the calculations.
-It is not recommended to do similar calculations that imply a the cost because a source that is quite incoherent was treated with partial coherence methods.
+It is not recommended to do similar calculations that imply a cost because a source that is quite incoherent was treated with partial coherence methods.
 %\inred{In Appendix B we present for the ESRF High$\beta$ similar results than those in Section~\ref{comsyl}}.
@@
 Some effects that are typically studied when designing and simulating a beamline have not been discussed here.
-For example the study of the power emitted by the source and its heat load effect in the first (or firsts) slits and optical elements (here the ML monochromator).
-Also, the reflectivity of the elements are not considered.
-The simulation of this was not included in the simulations, because it can be estimated analytically considering an average reflectivity about 0.7 for multilayers and 0.9 for mirrors, therefore a transmission factor of 0.44 for the beamline flux due to reflectivity of mirror coatings.
+For example, the study of the power emitted by the source and its heat load effect in the first (or firsts) slits and optical elements (here the ML monochromator).
+Also, the reflectivity of the elements is not considered.
+The simulation of this was not included in the simulations, because it can be estimated analytically considering an average reflectivity about 0.7 for multilayers and 0.9 for mirrors, therefore a transmission factor of 0.44 for the beamline flux due to the reflectivity of mirror coatings.
 
@@
 
-We have obtained quantitative values of flux, transmission, size of the focused beam, and coherent fraction for ID16A, a modern beamline with already exceptional performances in terms of nanofucusing and coherence.
-We predicted advances the improvements in these parameters when the EBS will start in 2020.
-We obtained these parameters using a hierarchical flowchart applying methods with increasing complexity and requiring higher computational effort.
-
-The analytical estimation of the main parameters is mandatory first step in the beamline conception and analysis.
+We have obtained quantitative values of flux, transmission, size of the focused beam, and coherent fraction for ID16A, a modern beamline with already exceptional performances in terms of nanofocusing and coherence.
+We predicted advances in the improvements in these parameters when the EBS will start in 2020.
+We obtained these parameters using a hierarchical flowchart applying methods with increasing complexity and requiring a higher computational effort.
+
+The analytical estimation of the main parameters is the mandatory first step in the beamline conception and analysis.
 We applied concepts based on optical magnification and numerical aperture acceptance and calculated approximately the coherent fraction at the source.
-These values help to perform sanity check of the computer simulations.
-A ray tracing simulation is inexpensive and provides important information.
+These values help to perform a sanity check of the computer simulations.
+A ray-tracing simulation is inexpensive and provides important information.
 The comparison of the CF with the beamline transmitivity would give an idea of the beam coherence, reflected in the number modes present in the final beam.
 For high coherence, like for ID16A, the wave optics calculations are necessary.
-If the final beam has a very high coherence coherent optics (1D simplifications, and single electron calculations) may give good results.
+If the final beam has a very high coherence coherent optics (1D simplifications, and single-electron calculations) may give good results.
 Full calculations based on coherent mode decomposition Were introduced here.
-It is a new way of quantitatively assess the quality of the partial coherent beam via the calculation of the coherent modes of the source, propagation of the modes and recalculation the coherent mode expansion on the final beam.
-The CF or occupations of the first mode is the parameter used to measure the coherence quality.
-
-% It is interesting to remark that although Monte Carlo multielectron method can in principle calculate Eq.~\ref{CSD} the storage at every point is unmanagleable, so usually only some 'cuts' of the CSD are stored, such as $W(x_1,0,x_2,0;z,\omega)$ (horizontal) or $W(0,y_1,0,y_2;z,\omega)$ (vertical) \todo{move this to the discussion?}.
+It is a new way of quantitatively assess the quality of the partially coherent beam via the calculation of the coherent modes of the source, propagation of the modes and recalculation of the coherent mode expansion on the final beam.
+The CF or occupation of the first mode is the parameter used to measure the coherence quality.
+f
+
+% It is interesting to remark that although Monte Carlo multi-electron method can in principle calculate Eq.~\ref{CSD} the storage at every point is unmanagleable, so usually only some 'cuts' of the CSD are stored, such as $W(x_1,0,x_2,0;z,\omega)$ (horizontal) or $W(0,y_1,0,y_2;z,\omega)$ (vertical) \todo{move this to the discussion?}.
 %
@@
 The residual height error of the multilayer monochromator and the KB mirrors, after removing the main ellitical shape, are shown in Fig.~\ref{fig:metrology}.
-These profiles are used along the text in ray tracing and wave optics simulations.
+These profiles are used along the text in ray-tracing and wave optics simulations.
 

### 2019-06-14 GRAPHICS/figure3.sh
preparing script for arxiv

@@ -1,0 +1,2 @@
+montage -mode concatenate idealelementsESRF.png idealelementsEBS.png hybridnoerrorsESRF.png hybridnoerrorsEBS.png hybriderrorsESRF.png hybriderrorsEBS.png -tile 2x3 figure3.png
+

### 2019-06-14 GRAPHICS/figure9b.sh

@@ -1,0 +1,1 @@
+montage -mode concatenate ebs_mode0.png ebs_mode1.png ebs_mode2.png ebs_mode3.png -tile 2x2 figure9b.png 

### 2019-06-14 mkarxiv.sh

@@ -1,0 +1,117 @@
+#
+# script to create the ArXiv submission
+#
+# please be sure that you run pdflatex paper-hierarchical.tex in the main directory
+# before running the script, so paper-hierarchical.bbl and the *pdf files are created
+
+cd /users/srio/Working/paper-hierarchical
+rm -rf ARXIV2 ARXIV
+mkdir ARXIV
+cd ARXIV
+cp ../paper-hierarchical.tex .
+cp ../paper-hierarchical.bbl .
+cp ../iucr.cls .
+cp ../harvard.sty .
+cp ../siunitx.sty .
+cp ../siunitx-abbreviations.cfg .
+cp ../tableUndulator.txt .
+cp ../tableElectron.txt .
+
+cp -r ../GRAPHICS .
+
+cp paper-hierarchical.tex paper-hierarchical.tex.orig
+
+# biblio
+# sed -e "s.^\\\\bibliography{references}.\\\\input{paper-hierarchical\.bbl}." paper-hierarchical.tex -i
+sed -e "s.\\\\bibliographystyle.\\% \\\\bibliographystyle.g" paper-hierarchical.tex -i
+sed -e "s.^\\\\bibliography{references}.\\% \\\\bibliography{references}." paper-hierarchical.tex -i
+sed -e "s.^\\\\end{document}.\\% \\\\end{document}." paper-hierarchical.tex -i
+cat ../paper-hierarchical.bbl >> paper-hierarchical.tex
+echo "\end{document}" >> paper-hierarchical.tex
+
+
+# figure 1
+cp GRAPHICS/ID16A.pdf ./figure1.pdf
+sed -e "s/GRAPHICS\/ID16A.pdf/figure1.pdf/" paper-hierarchical.tex -i
+
+# figure 2
+cp GRAPHICS/fluxU18.png ./figure2.png
+sed -e "s/GRAPHICS\/fluxU18.png/figure2.png/" paper-hierarchical.tex -i
+
+# figure 3
+cp GRAPHICS/figure3.png ./figure3.png
+sed -e "s/GRAPHICS\/idealelementsESRF/figure3/" paper-hierarchical.tex -i
+sed -e "/idealelementsEBS/d" paper-hierarchical.tex -i
[…]

### 2019-06-14 compile.sh
update

@@ -1,0 +1,5 @@
+pdflatex paper-hierarchical.tex
+bibtex paper-hierarchical
+pdflatex paper-hierarchical.tex
+pdflatex paper-hierarchical.tex
+okular paper-hierarchical.pdf

### 2019-06-14 mkarxiv.sh

@@ -23,6 +23,8 @@ cp paper-hierarchical.tex paper-hierarchical.tex.orig
 
 # biblio
 # sed -e "s.^\\\\bibliography{references}.\\\\input{paper-hierarchical\.bbl}." paper-hierarchical.tex -i
+pdflatex paper-hierarchical.tex
+bibtex paper-hierarchical
 sed -e "s.\\\\bibliographystyle.\\% \\\\bibliographystyle.g" paper-hierarchical.tex -i
 sed -e "s.^\\\\bibliography{references}.\\% \\\\bibliography{references}." paper-hierarchical.tex -i
 sed -e "s.^\\\\end{document}.\\% \\\\end{document}." paper-hierarchical.tex -i

### 2019-06-14 mkarxiv.sh
update

@@ -42,13 +42,7 @@ sed -e "s/GRAPHICS\/fluxU18.png/figure2.png/" paper-hierarchical.tex -i
 
 # figure 3
 cp GRAPHICS/figure3.png ./figure3.png
-sed -e "s/GRAPHICS\/idealelementsESRF/figure3/" paper-hierarchical.tex -i
-sed -e "/idealelementsEBS/d" paper-hierarchical.tex -i
-sed -e "/hybridnoerrorsESRF/d" paper-hierarchical.tex -i
-sed -e "/hybridnoerrorsEBS/d" paper-hierarchical.tex -i
-sed -e "/hybriderrorsESRF/d" paper-hierarchical.tex -i
-sed -e "/hybriderrorsEBS/d" paper-hierarchical.tex -i
-
+sed -e "s/GRAPHICS\/figure3/figure3/" paper-hierarchical.tex -i
 
 # figure 4
 cp GRAPHICS/wofry1D.png ./figure4.png
@@ -82,13 +76,9 @@ sed -e "s/GRAPHICS\/up_to_mode_id16a.eps/figure8b.eps/" paper-hierarchical.tex -
 
 # figure 9
 cp GRAPHICS/ebs_spectral_density.png ./figure9a.png
-sed -e "s/GRAPHICS\/ebs_spectral_density.png/figure9a.png/" paper-hierarchical.tex -i
 cp GRAPHICS/figure9b.png ./figure9b.png
-sed -e "s/GRAPHICS\/ebs_mode0/figure9b/" paper-hierarchical.tex -i
-sed -e "/ebs_mode1.png/d" paper-hierarchical.tex -i
-sed -e "/ebs_mode2.png/d" paper-hierarchical.tex -i
-sed -e "/ebs_mode3.png/d" paper-hierarchical.tex -i
-
+sed -e "s/GRAPHICS\/ebs_spectral_density.png/figure9a.png/" paper-hierarchical.tex -i
+sed -e "s/GRAPHICS\/figure9b/figure9b/" paper-hierarchical.tex -i
 
 # figure 10
 cp GRAPHICS/propagated_modes.eps ./figure10.eps

### 2019-06-14 mkarxiv.sh
update

@@ -1,15 +1,17 @@
 #
 # script to create the ArXiv submission
 #
-# please be sure that you run pdflatex paper-hierarchical.tex in the main directory
-# before running the script, so paper-hierarchical.bbl and the *pdf files are created
+# please be sure that you run compile.sh in the main directory before running 
+# this script, so paper-hierarchical.bbl and the GRAPHICS/*eps-converted-to.pdf files 
+# are created
+#
 
 cd /users/srio/Working/paper-hierarchical
 rm -rf ARXIV2 ARXIV
 mkdir ARXIV
 cd ARXIV
 cp ../paper-hierarchical.tex .
-cp ../paper-hierarchical.bbl .
+#cp ../paper-hierarchical.bbl .
 cp ../iucr.cls .
 cp ../harvard.sty .
 cp ../siunitx.sty .
@@ -68,10 +70,10 @@ cp GRAPHICS/ebs_slopeTE_slopeOE_h.pdf ./figure7b.pdf
 sed -e "s/GRAPHICS\/ebs_slopeTE_slopeOE_h.pdf/figure7b.pdf/" paper-hierarchical.tex -i
 
 # figure 8
-cp GRAPHICS/cumulated_occupation.eps ./figure8a.eps
-sed -e "s/GRAPHICS\/cumulated_occupation.eps/figure8a.eps/" paper-hierarchical.tex -i
-cp GRAPHICS/up_to_mode_id16a.eps ./figure8b.eps
-sed -e "s/GRAPHICS\/up_to_mode_id16a.eps/figure8b.eps/" paper-hierarchical.tex -i
+cp GRAPHICS/cumulated_occupation-eps-converted-to.pdf ./figure8a.pdf
+sed -e "s/GRAPHICS\/cumulated_occupation.eps/figure8a.pdf/" paper-hierarchical.tex -i
+cp GRAPHICS/up_to_mode_id16a-eps-converted-to.pdf ./figure8b.pdf
+sed -e "s/GRAPHICS\/up_to_mode_id16a.eps/figure8b.pdf/" paper-hierarchical.tex -i
 
 
 # figure 9
@@ -81,12 +83,12 @@ sed -e "s/GRAPHICS\/ebs_spectral_density.png/figure9a.png/" paper-hierarchical.t
 sed -e "s/GRAPHICS\/figure9b/figure9b/" paper-hierarchical.tex -i
 
 # figure 10
[…]

### 2019-06-17 paper-hierarchical.tex
update
arxiv almost ready

@@
 \usepackage{color,xcolor}
-\usepackage{siunitx}
+%\usepackage{siunitx}
 \usepackage[normalem]{ulem}
@@
 \begin{abstract}
-We consider different approaches to simulating a modern X-ray beamline.
+We consider different approaches to simulate a modern X-ray beamline.
 Several methodologies with increasing complexity are applied to discuss the relevant parameters that quantify the beamline performance.
@@
 The different algorithms and methodologies are implemented in the software suite OASYS \cite{codeOASYS}.
-Those are discussed with emphasis placed upon the their benefits and limitations.
+Those are discussed with emphasis placed upon the their benefits and limitations of each.
 \end{abstract}
@@
 Every modern beamline follows a procedure of conceptual design plus technical design where detailed simulations are essential.
-These calculations allowing testing of the design parameters in a virtual computer environment where advantages and limitations can be studied quantitatively.
-Selection of the optimal optical layout requires the detailed simulation of the optics imperfections, that in many cases are the limiting factor of the beamline optical configuration.
+These calculations allow testing of the design parameters in a virtual computer environment where advantages and limitations can be studied quantitatively.
+The selection of the optimal optical layout requires a detailed simulation of the optics imperfections, that in many cases are the limiting factor of the beamline optics.
 Deformations of the optical elements due to heat load need to be controlled, therefore optics simulations have to include results from engineering modeling of the thermal deformations usually done using finite element analysis.
@@
 A fully coherent hard X-ray beam is still a dream for the new generation storage rings.
-Despite a considerable increase in the coherent fraction of the new beams, the coherent fraction at typically hard X-ray photon energies is still a few percents for the new sources.
-Ray-tracing analysis does not become obsolete but has to be complemented with an analysis based on wave optics.
+Despite a considerable increase in the coherent fraction of the new beams, the coherent fraction at typically hard X-ray photon energies is still a few percent for the new sources.
+Consequently, the ray-tracing analysis does not become obsolete but has to be complemented with an analysis based on wave optics.
 Because of the partial coherence, the wave optics analysis becomes nontrivial and can be done with different levels of approximation, as discussed in this work.
@@
 It displays characteristics increasingly aimed for in modern beamlines: extremely high demagnification (sub-micrometer focused beam size) and high coherence.
-It combines coherent imaging techniques and X-ray fluorescence microscopy to perform quantitative 3D characterization of the morphology and the elemental composition of specimens in their native state~\ingreen{\cite{XRFID16A}}.
+It combines coherent imaging techniques and X-ray fluorescence microscopy to perform quantitative 3D characterization of the morphology and the elemental composition of specimens in their native state~\cite{XRFID16A}.
 
@@
 
- In this paper, some quantitative results on how the storage ring upgrade will affect a particular beamline are presented and it is structured as follows: we first introduce in Section~\ref{Beamline description} the main parameters of the beamline and explain the particular choice of the parameters used in the calculations.
+ In this paper, some quantitative results on how the storage ring upgrade will affect a particular beamline are presented.
+The paper is structured as follows: we first introduce in Section~\ref{Beamline description} the main parameters of the beamline and explain the particular choice of the parameters used in the calculations.
 Section~\ref{Simulations} is the central part of the paper, and describes the different methods used to analyze the beamline with increasing complexity, setting a hierarchy of methodologies that can be exported to the simulations of any beamline.
@@
 The photon source size and divergence are approximately given by the convolution of the electron beam sizes and the undulator natural sizes.
-If we consider Gaussian distributions, the photon source is:
+If we consider Gaussian distributions, the photon source has sizes $\Sigma$ and divergences $\Sigma'$ given by:
 \begin{align}
@@
 
-Using the values for U18.3, $n=1$, and  $E=17225$~eV (Table \ref{tab:Undulators}) and the optical configuration described in Table \ref{tab:Positions}, one obtains a focal size 139$\times$5~nm$^2$ (H$\times$V) for ESRF-1 and $10\times5$~nm$^2$ for EBS (Table~\ref{tab:HandCalculations}).
+Using the values for U18.3, $n$=1, and  $E$=17225~eV (Table \ref{tab:Undulators}) and the optical configuration described in Table \ref{tab:Positions}, one obtains a focal size 139$\times$5~nm$^2$ (H$\times$V) for ESRF-1 and $10\times5$~nm$^2$ for EBS (Table~\ref{tab:HandCalculations}).
 The horizontal focal spot for the ESRF-1 lattice is larger than the specified target value for the beamline, therefore the VSS could be closed to 50~$\mu$m thus reducing the beam size at the VSS plane by a factor ca.
@@
 Values are FWHM.
-Values in parentheses correspond to closing the VSS slit to 50~$\mu$m (21.3 if approximating by the Gaussian $\sigma$).}
+Values in parentheses correspond to closing the VSS slit to 50~$\mu$m.
+    %(21.3 if approximating by the Gaussian $\sigma$).
+    }
     \resizebox{\textwidth}{!}{%
     \begin{tabular}{r|ccc|ccc}
-                            &               & \textbf{ESRF (high-$\beta$)} &   &         & \textbf{EBS}  &    \\ 
+                            &               & \textbf{ESRF-1 (high-$\beta$)} &   &         & \textbf{EBS}  &    \\ 
     Element:                & Undulator     & VSS          & Sample            & Undulator & VVS   & Sample       \\
@@
 Figure~\ref{fig:FluxU18} shows the spectra calculated using SRW over an aperture of $1 \times 1$~mm$^2$ placed at $d=27.1$~m.
-This aperture opening is set to fully accept the central cone at the resonance peak of the harmonic in use ($\max(\Sigma') \times d = 0.73$~mm), with $\Sigma'$ values from in Table~\ref{tab:HandCalculations}.
+This aperture opening is set to fully accept the central cone at the resonance peak of the harmonic in use.
+%($\max(\Sigma') \times d = 0.73$~mm), with $\Sigma'$ values from in Table~\ref{tab:HandCalculations}. 
 The peak intensity is about  $2.9\cdot10^{14}$ photons/s/0.1{\%}bw for EBS and $1.9\cdot10^{14}$ for ESRF-1 (see \ref{fig:FluxU18}).
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{GRAPHICS/fluxU18.png}
+    \includegraphics[width=0.95\textwidth]{GRAPHICS/fluxU18.png}
     \caption{Flux of the U18.3 undulator integrated over a square slit placed at 27.1~m from the ID centre with aperture of 1 and 4 mm.}
@@
 The multilayer monochromator accepts the whole beam.
-We consider a VSS closed to $S_h=50~\mu$m in horizontal and fully opened in vertical.
+We consider a VSS closed to $S_h$=50~$\mu$m in horizontal and fully opened in vertical.
 For ESRF-1 its transmission is 1/8 (as discussed before) and 1 for EBS.
 In a similar way, the KB mirrors will crop the beam to a size that is the projection of their useful optical length onto a direction perpendicular to the optical axis.
-Using a mirror length of $L_v=60$~mm and $L_h=26$~mm for VFM and HFM, respectively, and incident angle of 15~mrad, the KB behaves as an aperture of $900 \times 390~\mu$m$^2$.
+Using a mirror length of $L_v$=60~mm and $L_h$=26~mm for VFM and HFM, respectively, and incident angle of 15~mrad, the KB behaves as an aperture of 900 $\times$ 390~$\mu$m$^2$.
 We get N.A.
-values of $6.2\times21.6~\mu$rad$^2$ (aperture size divided by the distance from the mirror to the VSS in H and to the source in V).
+values of 6.2 $\times$ 21.6~$\mu$rad$^2$ (aperture size divided by the distance from the mirror to the VSS in H and to the source in V).
 The transmission coefficient is the ratio of the accepted N.A.
@@
 To assess the focused beam size given by a coherent beam, one can consider the beamline in two sections: First, the ML monochromator and VSS that produce a coherent plane wave as the source (vertical) and VSS (horizontal) are both very far from the entrance aperture of the focusing optics.
-Second, the KB-system playing a dual role, acting as a focusing element of focal distance $f=5$~cm, but also cropping the beam by acting as a square aperture of $D \times D=$ 400$\times$400~$\mu$m$^2$ (the length of the mirror projected in a plane perpendicular to the beam propagation).
+Second, the KB-system playing a dual role, acting as a focusing element of focal distance $f=5$~cm, but also cropping the beam by acting as a square aperture of roughly D$\times$D=400$\times$400~$\mu$m$^2$ (the length of the mirror projected in a plane perpendicular to the beam propagation).
 %\todo{REMOVE: we obtained 900 in H, that will be reduced by 0.5 as propagated from 10 cm to 5 cm upstream of the sample, and 390 in V approximated by 400}.
@@
 In this section, we perform ray-tracing calculations to study the effect of real optics (including aberrations and slope errors) upon the beam size and beamline transmission.
-The first subsection deals with the effect of aberrations and transmission due to geometrical considerations.
-The second subsection shows how the effect of slope errors of the KB mirrors degrade the focal spot.
+The first part deals with the effect of aberrations and transmission due to geometrical considerations.
+The second part shows how the effect of slope errors of the KB mirrors degrade the focal spot.
 
@@
 Ray-tracing is a simple and extremely powerful technique for calculating the main characteristics of the photon beam (size, divergence, photon distribution) at every point of the beamline.
-Although very powerful commercial ray-tracing packages are available in the market, they are generally ill-adapted for synchrotron simulations because they do not cover the particularities of X-ray devices and sources.
-For that purposes, several packages created in the synchrotron community and are available, like SHADOW \cite{codeSHADOW}, RAY \cite{codeRAY}, McXtrace \cite{codeMCXTRACE} or XRT \cite{codeXRT}.
+%Although very powerful commercial ray-tracing packages are available in the market, they are generally ill-adapted for synchrotron simulations because they do not cover the particularities of X-ray devices and sources.
+For that purposes,
+Several packages created in the synchrotron community and are available, like SHADOW \cite{codeSHADOW}, RAY \cite{codeRAY}, McXtrace \cite{codeMCXTRACE} or XRT \cite{codeXRT}.
 For this study, we use SHADOW and its interface ShadowOUI \cite{codeSHADOWOUI}, available as a module of the OASYS suite.
 
-We performed ray-tracing of the ID16A beamline using an undulator source tuned to have its resonance at 17225~eV, considering the ML as a cylindrical mirror, and using elliptical cylinder reflectors with finite dimensions as the KB focusing elements.
-The reflector shape is chosen to be elliptical as this is the theoretical shape producing a point-to-point focusing.
+We performed ray-tracing of the ID16A beamline using an undulator source tuned to have its resonance at 17225~eV and approximated using Gaussian distributions (Eqs.~\ref{eq:photon big sigmas}), considering the ML as a cylindrical mirror, and using elliptical cylinder reflectors with finite dimensions as the KB focusing elements.
+The KB reflector shape are elliptical as required for a point-to-point focusing.
 For a first calculation, the reflectors (ML and KBs) are considered perfect and have no slope errors.
@@
 Secondly, one can observe a much larger luminosity in the EBS case.
-The transmission values are 0.162\% for ESRF-1 and 2.22\% for EBS.
+The transmission values are 0.16\% for ESRF-1 and 2.22\% for EBS.
 These values confirm the analytical estimations (0.2\% for ESRF-1 and 2.95\% for EBS).
-As mentioned before, these values are similar to the coherent fraction \cite{arxivCF} (0.12 for ESRF-1 and 2.73\% for EBS), meaning that a very high coherence of the beam is expected, therefore it is expected some broadening of the beam due to diffraction effects originated by the mirrors acceptance.
-
-\begin{figure}
-\label{fig:ray-tracing}
-\centering
-~~~~~ESRF-1~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~EBS
-\newline
-\includegraphics[width=0.45\textwidth]{GRAPHICS/idealelementsESRF.png}
-\includegraphics[width=0.45\textwidth]{GRAPHICS/idealelementsEBS.png}
-
-\includegraphics[width=0.45\textwidth]{GRAPHICS/hybridnoerrorsESRF.png}
-\includegraphics[width=0.45\textwidth]{GRAPHICS/hybridnoerrorsEBS.png}
-
-\includegraphics[width=0.45\textwidth]{GRAPHICS/hybriderrorsESRF.png}
-\includegraphics[width=0.45\textwidth]{GRAPHICS/hybriderrorsEBS.png}
-\caption{Image produced by different types of ideal focusing for the ESRF (left column) and EBS (right column). The source with parameters in Table~\ref{tab:HandCalculations}. 
-The first row has been calculated using the ideal reflectors (cylindrical for ML, elliptical for the BK mirrors) without slope errors. 
-The second row shows the image calculated the hybrid option (ray-tracing plus coherent diffraction) also for ideal reflectors with no slope errors.
-It shows the focal spot degradation due to the diffraction arising from the beam aperturing by the KB mirrors.
-The calculations in the bottom row also include slope errors using the real mirror profiles measures at the ESRF metrology lab (see Appendix~\ref{appendix:metrology}).
-Further beam broadening and the growth of satellite structures in the vertical intensity distribution can be observed.
-The focal beam dimensions (FWHM of the histograms) are written in the plot title.
-}
-\end{figure}
+As mentioned before, these values are similar to the coherent fraction \cite{arxivCF} (0.12 for ESRF-1 and 2.73\% for EBS), meaning that a very high coherence of the beam is expected.
+It is therefore expected some broadening of the beam due to diffraction effects originated by the mirrors acceptance.
 
@@
 
+\begin{figure}
+\label{fig:ray-tracing}
+\centering
+~~~~~ESRF-1~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~EBS
+\newline
+\includegraphics[width=0.95\textwidth]{GRAPHICS/figure3.png}
+\caption{Image produced by different types of ideal focusing for the ESRF (left column) and EBS (right column).
+The source follows Gaussian distributions with parameters in Table~\ref{tab:HandCalculations}.
+a) Simulations using the ideal reflectors (cylindrical for ML, elliptical for the BK mirrors) without slope errors. 
+b) Images calculated using the hybrid method (ray-tracing plus coherent diffraction) with ideal reflectors with no slope errors.
+It shows the focal spot degradation due to the diffraction arising from the cropping of the beam by KB mirrors.
+c) Simulations using the hybrid method including slope errors from the real mirror profiles measured at the ESRF metrology lab (see Appendix~\ref{appendix:metrology}).
+It can be observed further beam broadening and the growth of satellite structures.
+The focal beam dimensions (FWHM of the histograms) are written in the plot title.
+}
+\end{figure}
+
 The surface figure of the optical elements is another limiting effect for determining the size and intensity of the beam at the image position.
@@
 These experimental data are used in all the simulations in this paper.
-Simulations using the hybrid model including slope errors are shown in Fig.~\ref{fig:ray-tracing}c.
-A small broadening of the peak due to slope errors can be observed, but, more importantly, a new satellite structure close to the vertical intensity peak is produced.
+Simulations using the hybrid method including slope errors are shown in Fig.~\ref{fig:ray-tracing}c.
+A small broadening of the peak due to slope errors can be observed, but, more importantly, new satellite structures close to the main peak appear.
 
@@
 We first analyze the beamline supposing a fully coherent beam.
-Subection~\ref{wofry} shows a simplified 1D model.
+Subsection~\ref{wofry} shows a simplified 1D model.
 Subsection~\ref{srw_se} uses a full 2D model with accurate simulation of the undulator emission but without considering storage ring emittance.
-These simplifications allow the simulation of diffraction effects by apertures and slope/figure errors.
+These simplifications allow the estimation of diffraction effects by apertures and slope or figure errors.
 Subsequently, we introduce methods to treat the partial coherence originated by the storage ring emittance.
@@
 We perform here a simplified wavefront simulation in 1D, thus decoupling the simulations of the horizontal and vertical planes.
-The source can be simplified by using a point source (spherical wavefront) and applying a Gaussian intensity profile with $\sigma=\sigma'_u d$ with $\sigma'_u$ given by Eq.~\ref{eq:photon small sigmas} and $d$ the distance from the point source to the observation plane.
+The source can be simplified by using a point source (spherical wavefront) and applying a Gaussian intensity profile with $\sigma$=$\sigma'_u d$ with $\sigma'_u$ given by Eq.~\ref{eq:photon small sigmas} and $d$ the distance from the point source to the observation plane.
 Each optical element can be simplified by separating its action into two parts: the focusing behavior implemented as an ideal lens, and the finite size because of the boundaries, implemented as a slit.
-The WOFRY package \cite{codeWOFRY} in OASYS permits such simulations, and it is possible to achieve satisfactory results in a few calculation iterations.
+The WOFRY package \cite{codeWOFRY} in OASYS was used for these simulations.
 The beam size at different positions of the beamline is shown in Table~\ref{tablewofry1D}.
@@
 
-A next step in the wave optics simulation is, assuming full coherence, to generate a 2D source wavefront that matches in detail the real source (an undulator in a storage ring) and then simulate the propagation of such a wavefront through the optical elements with models that match well the characteristics of the real elements.
-The emission by a filament beam is fully coherent.
+A next step in the wave optics simulation is, assuming full coherence, to generate a 2D source wavefront that matches in detail the real source (an undulator in a storage ring) and then simulate the propagation of such a wavefront through the optical elements with models that reproduce well the characteristics of the real elements.
 Once the source is defined, an approach of how best to simulate the beamline optical elements.
-Increasing the complexity of the simulations one can: a) use ideal thin lenses preceded by slits to emulate finite apertures (as done in the 1D model presented in Section~\ref{wofry}); or b) use grazing incidence focusing optics models \cite{Canestrari2014}.
-Furthermore, one can introduce the effect of imperfect optics on the simulations, i.e.
-slope/figure errors to both ideal lenses and mirrors.
-
+Once the source is defined, several approaches of increasing complexity are used: a) use ideal thin lenses preceded by slits to emulate finite apertures (as done in the 1D model presented in Section~\ref{wofry}) b) use grazing incidence focusing optics models \cite{Canestrari2014}; and c) introduce the effect of imperfect optics on the simulations, i.e.
+slope and figure errors to both ideal lenses and mirrors.
 The computer code chosen for these calculations is ``Synchrotron Radiation Workshop'' (SRW) \cite{codeSRW}, which over the past two decades has been extensively benchmarked throughout different synchrotron radiation facilities and has become widely accepted within the X-ray community.
-Figure \ref{fig:SingleElectron} shows the horizontal integrated intensity at the focal position for the ID16A.
+
+%One of the difficulties with wave-optics simulations is the correct sampling and range of the input wavefront.
+If chosen poorly, the simulation will be affected by replicas and aliasing, compromising, then the quality of the obtained data.
+Choosing the range of the wavefront is straightforward: one can base the choice either on the first optical element aperture or on the beam footprint at its entrance.
+Usually one chooses the smaller.
+Choosing a very small range for the wavefront leads to cropping and diffraction effects that may not come from the beamline, but only a numerical consequence of truncating the wavefront.
+For the ID16A, the first optical element is the multi-layer monochromator and its aperture accepts the full beam of both the ESRF-1 and EBS lattices.
+%Choosing the correct sampling is a more difficult task: good sampling implies choosing an appropriate number of pixels such as the phase of the electric field being propagated varies less than $\pi$ between two adjacent pixels.
+This rule may lead to very big arrays: SRW keeps in memory both $\sigma-$ and $\pi-$polarizations.
+Doubling the number of pixels on both directions of a 2D array implies an increase in memory consumption of about 8 times.
+
+%All simulations using SRW were done using the same grid ($3.5\times1.5$mm$^2$) and initial sampling ($740\times150$) - the initial wavefront was resampled by a factor of 8 in order to mitigate aliasing and replicas throughout the beamline.
+
+Figure \ref{fig:SingleElectron} shows the horizontal intensity profile for a filament beam (no emittances).
 It can be seen that there is little difference between using thin ideal lenses or the elliptical mirror model.
-The elliptical mirror presents a slight reduction in intensity and asymmetry but no significant differences are observed, which is in accordance with the findings of \cite{Canestrari2014}.
+The elliptical mirror presents a slight reduction in intensity and asymmetry but no significant differences are observed, which is in agreement with the findings of \cite{Canestrari2014}.
 Notice that the differences are minimal and to accentuate them we have chosen logarithmically scaled plots to represent the intensity distributions at the focal position.
@@
 Fig.~5 in \cite{Canestrari2014}.
-
-It is also possible to add slope errors to the simulations to study the resulting impact upon the focal spot quality (Fig.~\ref{fig:SingleElectron}).
-For that, the available surface metrology data of the reflective elements (see Appendix~\ref{appendix:metrology})  are used.
-For both cases (i.e., simulating the beamline using thin optical elements or elliptical mirrors), the beam sizes calculated as the FWHM at the focal position are little affected by the presence of the slope errors (see Table \ref{tab:SRW_results}).
-There is a significant reduction and blurring of the interference fringes compared to observations with no slope error.
-They are almost completely removed and associated with a large increase of the background level: the radiation found on the main lobe (focus) shifts to the side lobes when the slope errors are considered.
-This is accompanied by an obvious reduction in the peak intensity.
-Table \ref{tab:SRW_results} compiles the beam sizes at focal position for mentioned simulations.
+Increasing the complexity of the model increments considerably the simulation time, which went from typically $\sim16$~s to $\sim24$~s.
+
+% Table \ref{tab:SRW_results} compiles the beam sizes at focal position for mentioned simulations.
 
@@
 
+It is also possible to add slope errors to the simulations to study the resulting impact upon the focal spot quality (Fig.~\ref{fig:SingleElectron}b).
+For that, the available surface metrology data of the reflective elements (see Appendix~\ref{appendix:metrology})  are used.
+For both cases (i.e., simulating the beamline using thin optical elements or elliptical mirrors), the beam sizes calculated as the FWHM at the focal position are little affected by the presence of the slope errors (see Table \ref{tab:SRW_results}).
+There is, however, a significant reduction and blurring of the interference fringes compared to observations with no slope error.
+They are almost completely removed and associated with a large increase of the background level: part of the radiation in the main peak shifts to the side lobes when the slope errors are considered.
+This is accompanied by an obvious reduction in the peak intensity.
+Adding the mirrors figure errors to the model slightly increases the simulation time: a simulation of a filament beam using elliptical mirrors with figure errors runs in less than 30s.
+
 \begin{table}\label{tab:SRW_results}
 \centering
-\caption{Beam dimensions (FWHM, in nm) at the focal position for different models of focusing elements.}
+\caption{Calculated beam dimensions (FWHM, in nm) at the focal position for different models of focusing elements using SRW.}
 \resizebox{\textwidth}{!}{%
 \begin{tabular}{ccccc}
-Focusing elements                 & Plane      &  Single electron & EBS         & ESRF-1       \\
+Focusing elements                 & Plane      &  Single electron & EBS         & ESRF-1       \\ \\
 ideal lenses/ elliptical mirrors  & Horizontal &  7.94/7.97       & 12.20/12.70 & 17.50/16.80 \\
@@
 \label{srw_me}
+
 The wavefront propagation methodology is based on the sampling of a single wavefront (monochromatic and coherent) in a 2D space (in horizontal and vertical spatial dimensions) at a given point of the beamline.
-This initial point considered as source cannot be the source physical position, because i) when considering point or spherical waves all the intensity would be concentrated in a single pixel so the wavefront is ill-defined, ii) when using synchrotron sources (i.e., undulators) theory \cite{jackson} permits the calculation of the radiation in positions far from the electron trajectory.
-The limitation of the method is that one wavefront is not representative for a beam that, in storage rings, is produced by the emission of many electrons circulating through the corresponding magnet structure.
+This initial point considered as source cannot be the source physical position, because i) a point or spherical wave would concentrate all its intensity in a single pixel so the wavefront is ill-defined, and ii) the theory (e.g., \cite{jackson}) used for calculating of  synchrotron emission (i.e., undulators) permits the calculation of the radiation in positions far from the electron trajectory.
+The limitation of the wavefront propagation method is that a single wavefront is not representative for a beam that: in storage rings it is produced by the incoherent supperposition of the emission of many electrons circulating through the corresponding magnet structure.
 
 Diffraction limited sources would emit radiation that can be described in good approximation by a single wavefront.
-But even for new ultra low-emittance storage rings (improperly called diffraction-limited sources, see discussion in \cite{arxivCF}) the contribution of the electron dimensions ($\sigma_{e(h,v)}$ and $\sigma'_{e(h,v)}$) is not negligeable as compared with the natural emission of radiation ($\sigma_u$ and $\sigma'_u$ for undulators as in Eq.~\ref{eq:photon small sigmas}).
+But even for new ultra low-emittance storage rings (improperly called diffraction-limited sources, see discussion in \cite{arxivCF}) the contribution of the electron dimensions ($\sigma_{e(h,v)}$ and $\sigma'_{e(h,v)}$) is not negligible as compared with the natural emission of radiation ($\sigma_u$ and $\sigma'_u$ for undulators as in Eq.~\ref{eq:photon small sigmas}).
 To account for this effect one should complete or extend the wavefront simulations to take into account that in fact, not only a single wavefront contributes to the properties of the beam.
@@
 
+There are three approaches currently used for including the electron emittance effects in the wave optics simulations: i) use of convolutions, ii) Monte Carlo multi-electron sampling, and iii) Propagation of coherent modes.
+The convolution method consists of calculating the emission characteristics and then convolving the resulting intensity with Gaussians describing the electron beam.
+This method is simple and fast, as only requires simulation of a single wavefront.
+However, it presents two important limitations: it can only be applied to the intensity calculations and not to other parameters of interest for describing coherence, such as the degree of coherence, etc.; and it can only be applied at a plane right after the source position (it is usually unknown how the different optical elements will affect the source contribution to the phase space).
+The next method, Monte Carlo multi-electron propagation, is a statistical method based on the idea that one can exactly calculate the emission and its propagation for one electron that is deterministically described by its initial conditions at the entrance of the magnetic structure.
+It is then possible to perform many simulations for different initial conditions of the traveling electron and make the propagation along the beamline until the observation plane.
+Then, the different wavefronts (one for each electron) are combined to obtain a statistical map of the intensity and other magnitudes useful to describe partially coherent beams.
+This method presents two difficulties.
+The first one is convergence: it is not possible to know a priori how many electrons must be sampled to obtain a given level of accuracy.
+The second is computing resources.
+The complex task of calculating the emission of one electron, propagated it through every element of the beamline socoring the results must be repeated thousand of times.
+The third method, the propagation of coherent modes, will be discussed in the next section.
+
 \begin{figure}
@@
 
-There are three approaches currently used for including the electron emittance effects in the wave optics simulations: 1) use of convolutions, 2) Monte Carlo multi-electron sampling, and 3) Propagation of coherent modes.
-The convolution method consists of calculating the emission characteristics and then convolving the resulting intensity with Gaussians describing the electron beam.
-This method is simple and fast, as only requires simulation of a single wavefront.
-However, it presents two important limitations: i) it can only be applied to the intensity calculations and not to other parameters of interest for describing coherence, such as the degree of coherence, etc., and ii) it can be applied only at the source position: it is usually unknown how the different optical elements will affect the source contribution to the phase space.
-The next method, the Monte Carlo multi-electron propagation,  is a statistical method based on the idea that one can exactly calculate the emission and its propagation for one electron that is deterministically described by its initial conditions at the entrance of the magnetic structure.
-It is, therefore, possible to perform many simulations for different initial conditions of the traveling electron and make the propagation along the beamline until the observation plane.
-Then one can statistically combine the different wavefronts (one for each electron) to obtain a sampling of the intensity and other magnitudes useful to describe partially coherent beams.
-This method presents two difficulties.
-One is convergence.
-One cannot know a priori how many electrons must be sampled to obtain a given level of accuracy.
-The second is computing resources.
-The emission of each electron must be calculated at the source level, propagated through every element of the beamline, and the resulting wavefront must be analyzed to extract the information to be scored and stored.
-In the third method, the propagation of coherent modes will be discussed in the next section.
-The Monte Carlo multi-electron method has been proposed by O. Chubar and it is fully implemented in SRW \cite{multiE}. We have used this code for the calculations presented in this section.
-
-The source used for simulations is the U18.3 undulator tuned to its first harmonic at $E=17225$~eV with a finite emittance electron beam (see Table \ref{tab:eBeam}).
-Two set of parameters were used for the electron beam: the machine parameters of the ESRF-1 lattice (high-$\beta$) and the new EBS straight sections (Tables \ref{tab:TableSources} and \ref{tab:eBeam}).
-Simulations for both storage ring lattices were carried out modeling the beamline using thin ideal lenses and elliptical mirror models with and without the addition of the slope errors.
-Results are presented in Figures \ref{fig:FiniteEmittanceA} and \ref{fig:FiniteEmittanceB}; and Table \ref{tab:SRW_results}.
-
-The same observations made in the previous section (fully coherent simulations) are pertinent here: there is a reduction of the peak intensity when comparing the simple lens model with the elliptical mirror.
-The beam sizes and shapes are similar at the focal position, only.
+The results presented in this section are calculated using the Monte Carlo multi-electron method.
+It has been proposed by O.
+Chubar and it is fully implemented in SRW \cite{multiE}.
+The simulations presented here are built upon the ones described in the previous section (\ref{srw_se}).
+The different initial conditions were sampled from the 6D-electron phase space: its position on the cartesian space $(x_e, y_e, z_e)$; its direction in both vertical and horizontal planes  $(x'_e, y'_e)$ and its energy (as the electron energy spread is taken into account).
+Although simulations for highly coherent beamline converge rapidly (a few electrons are sufficient), this is in no way the case of the ID16A at ESRF-1: the VSS and the KB aperture reject a very high number of photons.
+To ensure good statistics when averaging the intensities of the different wavefronts from the sampled electrons, one has to go to a very high number of iterations.
+For the simulations presented here, 50000 wavefronts were used, amounting to a $\sim7$~h simulation time on a 56-core computer cluster.
+Convergence is faster for the EBS lattice as the electron-beam phase space is smaller.
+Results are presented in Figures \ref{fig:FiniteEmittanceA} and \ref{fig:FiniteEmittanceB} and Table \ref{tab:SRW_results}.
+
+The most important feature of the partially-coherent simulations, when compared to one of a filament-beam, is that the refraction fringes from the KB system aperture are smoothed out, although still present.
+The trends shown for the zero-emittance case are also found here, such as the reduction of the peak intensity and asymmetries when comparing the simple lens model with the elliptical mirror (although here it is more subtle).
+The major difference found when comparing the ESRF-1 and the EBS is, once more, the beam intensity at the sample position, thus permitting the beamline after the upgrade to exploit a more brilliant beam, that is, more photons on the same phase space.
+
 %{\todo{more detailed comments on comparisons with filament beam results? For comparison also information on calculation time, number of electrons, ...}
@@
 \begin{equation}\label{CSD}
-W(x_1,y_1,x_2,y_2;z,\omega) = <E*(x_1,y_1; z,\omega) E(x_2,y_2;z,\omega)>
+W(x_1,y_1,x_2,y_2;z,\omega) = <E^{*}(x_1,y_1; z,\omega) E(x_2,y_2;z,\omega)>
 \end{equation}
@@
 that measures the coherence of the beam between two points, and ranges from fully incoherent ($\mu=0$) to fully coherent ($\mu$=1). 
-Eq.~\ref{CSD} can in principle be calculated using Kim's convolution theorem \cite{kim1986} (see \cite{geloni2008} for a full discussion on the validity) but its storage is prohibitive in present computers: supposing sampling one spatial dimension by $N\approx1000$ pixels, a wavefront will contain $N^2$ pixels, and the CSD will be sampled over $N^4$ points.
+Eq.~\ref{CSD} can in principle be calculated using Kim's convolution theorem \cite{kim1986} (see \cite{geloni2008} for a full discussion on its validity) but its storage is prohibitive in present computers: supposing sampling one spatial dimension by $N\approx1000$ pixels, a wavefront will contain $N^2$ pixels, and the CSD will be sampled over $N^4$ points.
 They are complex numbers (16 bits) so the only storage is of the order of GB or TB.
@@
 where $\lambda_m$ are the eigenvalues or weights, and $\Phi_m$ are the eigenfunctions or (orthonormal) coherent modes.
-
 A theoretical study and a numerical algorithm for the coherent mode decomposition of undulator emission in storage rings has been recently proposed  \cite{GlassThesis,GlassEPL} and is implemented in the computer package COMSYL \cite{codeCOMSYL}.
@@
 Fig.~\ref{fig:histomodes}a shows the cumulated occupation of the coherent modes for the full central cone undulator emission.
-It shows that the EBS cumulated occupation grows very steeply with the number of coherent modes, whereas for ESRF-1 the number of coherent modes needed to attain a given value of cumulated occupation much higher.
-This can also be shown in Fig.\ref{fig:histomodes}b, where the number of modes to fill 50\%, 75\% and 95\% of spectral density is compared for EBS and ESRF-1.
+It can be noticed that the EBS cumulated occupation grows very steeply with the number of coherent modes, whereas for ESRF-1 the number of coherent modes needed to attain a given value of cumulated occupation much higher.
+This can also be shown in Fig.~\ref{fig:histomodes}b, where the number of modes to fill 50\%, 75\% and 95\% of spectral density is compared for EBS and ESRF-1.
 
@@
         \includegraphics[width=0.95\textwidth]{GRAPHICS/cumulated_occupation.eps}
-        \includegraphics[width=0.95\textwidth]{GRAPHICS/up_to_mode_id16a.eps}
+        \includegraphics[width=0.9\textwidth]{GRAPHICS/up_to_mode_id16a.eps}
     \caption{Cumulated occupation of the coherent modes for the EBS and ESRF-High-$\beta$ (top), and number of modes needed to reach a given percent of CSD occupation (bottom). }
@@
 \end{equation}
-Figs.~\ref{fig:spectraldensity}b-e represent the coherent modes with $m$=0 to $m$=3.
+Figsures~\ref{fig:spectraldensity}b-e represent the coherent modes with $m$=0 to $m$=3.
 It can be shown that the first mode is centered on the axis but extends over a spatial region smaller than the spectral density.
@@
         \includegraphics[width=\textwidth]{GRAPHICS/ebs_spectral_density.png}
-        \includegraphics[width=0.45\textwidth]{GRAPHICS/ebs_mode0.png}
-        \includegraphics[width=0.45\textwidth]{GRAPHICS/ebs_mode1.png}
-        \includegraphics[width=0.45\textwidth]{GRAPHICS/ebs_mode2.png}
-        \includegraphics[width=0.45\textwidth]{GRAPHICS/ebs_mode3.png}
+        \includegraphics[width=\textwidth]{GRAPHICS/figure9b.png}
     \caption{a) Spectral density $S(x,y)$ at the source position for EBS. b) Intensity distribution of the first coherent modes ($m$=0 to $m$=3) at the source position.
@@
 
-The coherent modes calculated by COMSYL at the EBS source have been propagated using the WOFRY package implementing a simplified beamline with ideal focusing elements and slits to determine their numerical aperture (like in Section~\ref{wofry}).
+The coherent modes calculated by COMSYL at the EBS source have been propagated in OASYS using the WOFRY package implementing a simplified beamline with ideal focusing elements and slits to determine their numerical aperture (like in Section~\ref{wofry}).
 Fig.~\ref{fig:mode transmission} shows how the intensity of the first modes is reduced by geometrical considerations.
@@
 Therefore, the coherent modes of the source after propagation and cropping by the beamline are no longer coherent modes of the resulting beam.
-Fig.~\ref{fig:mode transmission} shows that the most important transfer ratio ($M_{mm}$)transfer for the lowest mode is about $M_{00}$=25\%, then $M_{22}$=10\% and $M_{44}$=6\%.
+Fig.~\ref{fig:mode transmission} shows that the most important transfer ratio ($M_{mm}$) transfer for the lowest mode is about $M_{00}$=25\%, then $M_{22}$=10\% and $M_{44}$=6\%.
 The other modes shown transfer less than 5\% transfer, and modes higher than $m$=10 practically do not contribute to the final spectral density.
@@
 After the complete analysis done in the previous section on how the ID16A optics modify the characteristics of the X-ray beam, one can verify that the analytical previsions (Section~\ref{level0}) were a good estimator of the values obtained using more sophisticated methods.
-The size of the focused beam (e.g, 10x5 for EBS) and the transmission (3\%) calculated analytically are in good agreement with geometric ray-tracing (8.4x4.8, and 2.28\%).
+The size of the focused beam (e.g, 10$\times$5~nm$^2$ for EBS) and the transmission (3\%) calculated analytically are in good agreement with geometric ray-tracing (8.4$\times$4.8~nm$^2$, and 2.28\%).
 When characterizing beam sizes, one usually assumes that the intensity distributions are Gaussian and therefore they are well characterized by the FWHM value.
@@
 Indeed, this lack of Gaussianity is relevant in the calculations of the coherent fraction.
-We approximate the coherence fraction by Eq.~\ref{eq:coherent fraction}, which for zero emittance ($\sigma_{x,y}=\sigma_{x',y'}$=0) becomes CF=1, as expected because a zero emittance electron beam (a single-electron or a filament beam) that travel in a magnetic field emits a fully coherent wavefront.
+We approximate the coherence fraction by Eq.~\ref{eq:coherent fraction}, which for zero emittance ($\sigma_{x,y}=\sigma_{x',y'}$=0) becomes CF=1, as expected because a filament electron beam in a magnetic field emits a fully coherent wavefront.
 However, the photon source for a filament beam would have an emittance of $\sigma_u \sigma_{u'} \approx \lambda / 2 \pi$ (using Eqs.~\ref{eq:photon small sigmas}), which is in contradiction with theory, that says that a pure coherent state has emittance of $\lambda / 4 \pi$, as happens for a Gaussian photon beam.
@@
 
-The combination of ray-tracing and wave optics methods used indicate that the focalized beam size due to  demagnification with mirrors (giving for EBS 10$\times$5~nm$^2$ analytically, 8.4$\times$4.8~nm$^2$ with ray-tracing) has to be corrected with diffraction effect (8.4$\times$7.8~nm$^2$ hybrid, 12.7$\times$9.2~nm$^2$ SRW) and then add the slope error effect.
-The slope errors do not blow up the size and kit it in the specifications range (10.8$\times$8.2~nm$^2$ hybrid 13.6$\times$9.9~nm$^2$ SRW).
+The combination of ray-tracing and wave optics methods used indicate that the focalized beam size due to  demagnification with mirrors (giving for EBS 10$\times$5~nm$^2$ analytically, 8.8$\times$4.5~nm$^2$ with ray-tracing) has to be corrected with diffraction effect (11.2$\times$8~nm$^2$ hybrid, inred{12.7$\times$9.2~nm$^2$ SRW}) and then add the slope error effect.
+The slope errors do not blow up the size and kit it in the specifications range (11.4$\times$8.5~nm$^2$ hybrid \inred{13.6$\times$9.9~nm$^2$ SRW}).
 It is important to treat the slope errors correctly.
@@
 However, for those concerning partial coherence, very long calculations are needed.
-For example, Monte Carlo calculations in Section~\ref{srw_me} for the multi-electron case spent a few hours on a 56-core computer cluster for simulating 22k electrons.
+For example, Monte Carlo calculations in Section~\ref{srw_me} for the multi-electron case spent a few hours on a 56-core computer cluster for simulating 22000 electrons.
 For performing the COMSYL coherent mode decomposition the calculation time depends very much on the dimensions of the problem.
@@
 The CF or occupation of the first mode is the parameter used to measure the coherence quality.
-f
 
@@
 \label{appendix:metrology}
-The residual height error of the multilayer monochromator and the KB mirrors, after removing the main ellitical shape, are shown in Fig.~\ref{fig:metrology}.
+The residual height error of the multilayer monochromator and the KB mirrors, after removing the main ellitical shape, together with their power spectral densities are shown in Fig.~\ref{fig:metrology}.
 These profiles are used along the text in ray-tracing and wave optics simulations.
@@
 \includegraphics[width=0.95\textwidth]{GRAPHICS/metrology.eps}
-\caption{Error height profiles used for the simulations (left) and Power Spectral Density (right) for the ML monochromator (top row), the KB vertical focusing mirror (central row), and the KB horizontal focusing mirror (bottom row).}
+\caption{Error height profiles used for the simulations (left) and power spectral density (right) for the ML monochromator (top row), the KB vertical focusing mirror (central row), and the KB horizontal focusing mirror (bottom row).}
 \end{figure}

### 2019-06-17 paper-hierarchical.tex
Updates from Overleaf

@@
      %-------------------------------------------------------------------------
-     \journalcode{K}              % Indicate the journal to which submitted
+     \journalcode{S}              % Indicate the journal to which submitted
                                   %   A - Acta Crystallographica Section A
@@
 
+% emails:
+%  srio@esrf.eu
+%  celestre@esrf.eu
+%  mark.glass@gmx.de
+%  giogio.92@gmail.com
+% juan.reyes-herrera@esrf.eu
+% barrett@esrf.eu
+% jdasilva@esrf.eu
+% cloetens@esrf.eu
+% xshi@anl.gov
+% lrebuffi@anl.gov
+%
+%
+%
      % Use \shortauthor to indicate an abbreviated author list for use in
@@
 \caption{Parameters of the undulators available at the ID16A beamline: period ($\lambda_u$), length ($L$), number of identical undulators ($N$) and deflection parameter ($K$) values for the three main photon energies in use.
-The valuse are shown for the two ESRF magnetic lattices: the old ESRF-1 (labelled ESRF, with $E_e = 6.04$~GeV) and the new ESRF-EBS (labelled EBS, with $E_e = 6.00$~GeV).
+The valuse are shown for the two ESRF magnetic lattices: the old ESRF-1, with $E_e = 6.04$~GeV), and the new ESRF-EBS, with $E_e = 6.00$~GeV.
 $n$ is the emission harmonic in use for each particular photon energy.}
@@
 \begin{tabular}{lcccccc}
-name  & $\lambda_u$ [mm] & $L$ [m] & N & $K$ (11200~eV) & $K$ (17050~eV) & $K$ (33600~eV) \\ 
-\input{tableUndulator.txt}
+name  & $\lambda_u$ [mm] & $L$ [m] & N & $K$ (11200~eV) & $K$ (17050~eV) & $K$ (33600~eV) \\ \\
+%\input{tableUndulator.txt}
+ ESRF-1 U18.3  & 18.30  & 1.40  & 2 & 1.175(n=1) & 0.470(n=1) & 1.175(n=3)\\ 
+ ESRF-1 U22.4  & 22.40  & 1.40  & 2 & 0.873(n=1) & 1.855(n=3) & 0.873(n=3)\\ \\
+ EBS U18.3  & 18.30  & 1.40  & 2 & 1.156(n=1) & 0.437(n=1) & 1.156(n=3)\\ 
+ EBS U22.4  & 22.40  & 1.40  & 2 & 0.852(n=1) & 1.836(n=3) & 0.852(n=3)\\ 
 \end{tabular}%
@@
 Values taken at the symmetry point of the straight section, where the insertion devices are placed.
-Values for ESRF-1 are taken from \cite{ESRF2014}.
-For EBS, two (very close) design values are used, from Ref.~\cite{ESRF2014} labelled EBS-S28A, and those coming from further lattice refinement, labelled EBS-S28D.
-\todo{Remove ESRF-AT and EBS-S28A} }
+Values are taken from \cite{ESRF2014}, updated for EBS with
+    the lattice labelled S28D. 
+    %For EBS, two (very close) design values are used, from Ref.~\cite{ESRF2014} labelled EBS-S28A, and those coming from further lattice refinement, labelled EBS-S28D. 
+    %\todo{Remove ESRF-AT and EBS-S28A} 
+    }
         \resizebox{\textwidth}{!}{%
         \begin{tabular}{ccccc}
-         Lattice & $\sigma_e$ horizontal {[}$\mu$m{]} &  $\sigma_e'$ horizontal {[}$\mu$rad{]} & $\sigma_e$ vertical {[}$\mu$m{]} & $\sigma_e'$ vertical {[}$\mu$rad{]} \\
-         \input{tableElectron.txt}
-
-%    \begin{tabular}{rcc}
-%                                                & \textbf{ESRF (high-$\beta$)} & \textbf{EBS}     \\ 
-%         $\sigma_e$ horizontal {[}$\mu$m{]}:    & 415.00                      & 30.20             \\
-%         $\sigma_e'$ horizontal {[}$\mu$rad{]}: & 10.30                       & 4.40              \\
-%         $\sigma_e$ vertical {[}$\mu$m{]}:      & 3.40                        & 3.60              \\
-%         $\sigma_e'$ vertical {[}$\mu$rad{]}:   & 1.20                        & 1.40              \\ 
+         Lattice & $\sigma_e$ horizontal {[}$\mu$m{]} &  $\sigma_e'$ horizontal {[}$\mu$rad{]} & $\sigma_e$ vertical {[}$\mu$m{]} & $\sigma_e'$ vertical {[}$\mu$rad{]} \\ \\
+         %\input{tableElectron.txt}
+         ESRF-1  & 387.80  & 3.50  & 10.30 & 1.20 \\ 
+         %ESRF-AT  & 414.97  & 3.43  & 10.31 & 1.16 \\ 
+         %EBS-S28A  & 27.20  & 3.40  & 5.20 & 1.40 \\ 
+         EBS     & 30.18  & 3.64  & 4.37 & 1.37 \\ 
     \end{tabular}
@@
     \centering
-        \includegraphics[width=0.95\textwidth]{GRAPHICS/se_idealTE_idealOE_h.pdf}
-        \includegraphics[width=0.95\textwidth]{GRAPHICS/ebs_slopeTE_slopeOE_h.pdf}
+        \includegraphics[width=0.95\textwidth]{GRAPHICS/filament_beam_ideal.png}
+        \includegraphics[width=0.95\textwidth]{GRAPHICS/filament_beam_real.png}
     \label{fig:SingleElectron}
     \caption{Intensity profile at the sample position produced by the single-electron emission propagated through the ID16A beamline.
-Top: the perfect focusing with ideal lenses (black) is compared with the focusing using elliptical mirrors (red).
+Top: the perfect focusing with ideal lenses (green) is compared with the focusing using elliptical mirrors (blue).
 Bottom: the same as in the previous plot, but adding the effect of the mirrors slope errors in both cases.}
@@
 Focusing elements                 & Plane      &  Single electron & EBS         & ESRF-1       \\ \\
-ideal lenses/ elliptical mirrors  & Horizontal &  7.94/7.97       & 12.20/12.70 & 17.50/16.80 \\
-the same, with errors             & Horizontal &  9.35/10.50       & 12.90/13.65 & 17.70/17.03 \\
+ideal lenses/ elliptical mirrors  & Horizontal &  8.53/9.54       & 12.71/11.56 & 17.75/12.41 \\
+the same, with errors             & Horizontal &  9.36/10.61      & 13.20/12.52 & 17.78/15.71 \\
 ideal lenses/ elliptical mirrors  & Vertical   &  7.11/7.26       & 8.40/9.02   & 8.24/8.77    \\
@@
     \centering
-        \includegraphics[width=0.95\textwidth]{GRAPHICS/esrf_idealTE_idealOE_h.pdf}
-        \includegraphics[width=0.95\textwidth]{GRAPHICS/esrf_slopeTE_slopeOE_h.pdf}
+        \includegraphics[width=0.95\textwidth]{GRAPHICS/ESRF_ideal.png}
+        \includegraphics[width=0.95\textwidth]{GRAPHICS/ESRF_real.png}
     \label{fig:FiniteEmittanceA}
@@
     \centering
-        \includegraphics[width=0.95\textwidth]{GRAPHICS/ebs_idealTE_idealOE_h.pdf}
-        \includegraphics[width=0.95\textwidth]{GRAPHICS/ebs_slopeTE_slopeOE_h.pdf}
+        \includegraphics[width=0.95\textwidth]{GRAPHICS/EBS_ideal.png}
+        \includegraphics[width=0.95\textwidth]{GRAPHICS/EBS_real.png}
     \label{fig:FiniteEmittanceB}
@@
 It has been proposed by O.
-Chubar and it is fully implemented in SRW \cite{multiE}.
+Chubar and it is fully implemented in SRW \cite{Chubar2011b}.
 The simulations presented here are built upon the ones described in the previous section (\ref{srw_se}).
@@
 
-The combination of ray-tracing and wave optics methods used indicate that the focalized beam size due to  demagnification with mirrors (giving for EBS 10$\times$5~nm$^2$ analytically, 8.8$\times$4.5~nm$^2$ with ray-tracing) has to be corrected with diffraction effect (11.2$\times$8~nm$^2$ hybrid, inred{12.7$\times$9.2~nm$^2$ SRW}) and then add the slope error effect.
+The combination of ray-tracing and wave optics methods used indicate that the focalized beam size due to  demagnification with mirrors (giving for EBS 10$\times$5~nm$^2$ analytically, 8.8$\times$4.5~nm$^2$ with ray-tracing) has to be corrected with diffraction effect (11.2$\times$8~nm$^2$ hybrid, 12.7$\times$9.2~nm$^2$ SRW) and then add the slope error effect.
 The slope errors do not blow up the size and kit it in the specifications range (11.4$\times$8.5~nm$^2$ hybrid \inred{13.6$\times$9.9~nm$^2$ SRW}).

### 2019-06-17 paper-hierarchical.tex
rafael

@@
 the same, with errors             & Horizontal &  9.36/10.61      & 13.20/12.52 & 17.78/15.71 \\
-ideal lenses/ elliptical mirrors  & Vertical   &  7.11/7.26       & 8.40/9.02   & 8.24/8.77    \\
-the same, with errors             & Vertical   &  7.44/8.46       & 8.66/9.87   & 8.52/9.66    \\                                            
+ideal lenses/ elliptical mirrors  & Vertical   &  7.22/7.94       & 8.66/9.38   & 8.90/10.11    \\
+the same, with errors             & Vertical   &  7.70/8.90       & 8.42/9.14   & 8.90/9.86    \\                                            
 \end{tabular}
@@
 
-The combination of ray-tracing and wave optics methods used indicate that the focalized beam size due to  demagnification with mirrors (giving for EBS 10$\times$5~nm$^2$ analytically, 8.8$\times$4.5~nm$^2$ with ray-tracing) has to be corrected with diffraction effect (11.2$\times$8~nm$^2$ hybrid, 12.7$\times$9.2~nm$^2$ SRW) and then add the slope error effect.
-The slope errors do not blow up the size and kit it in the specifications range (11.4$\times$8.5~nm$^2$ hybrid \inred{13.6$\times$9.9~nm$^2$ SRW}).
+The combination of ray-tracing and wave optics methods used indicate that the focalized beam size due to  demagnification with mirrors (giving for EBS 10$\times$5~nm$^2$ analytically, 8.8$\times$4.5~nm$^2$ with ray-tracing) has to be corrected with diffraction effect (11.2$\times$8~nm$^2$ hybrid, 12.56$\times$9.38~nm$^2$ SRW) and then add the slope error effect.
+The slope errors do not blow up the size and kit it in the specifications range (11.4$\times$8.5~nm$^2$ hybrid 13.6$\times$9.14~nm$^2$ SRW).
 It is important to treat the slope errors correctly.

### 2019-06-19 paper-hierarchical.tex
changes xianbo

@@
 In a similar way, the KB mirrors will crop the beam to a size that is the projection of their useful optical length onto a direction perpendicular to the optical axis.
-Using a mirror length of $L_v$=60~mm and $L_h$=26~mm for VFM and HFM, respectively, and incident angle of 15~mrad, the KB behaves as an aperture of 900 $\times$ 390~$\mu$m$^2$.
+Using a mirror length of $L_v$=60~mm and $L_h$=26~mm for VFM and HFM, respectively, and incident angle of 15~mrad, the KB behaves as an aperture of  390 $\times$ 900 (H $\times$ V) $\mu$m$^2$.
 We get N.A.
-values of 6.2 $\times$ 21.6~$\mu$rad$^2$ (aperture size divided by the distance from the mirror to the VSS in H and to the source in V).
+values of 2.7 $\times$ 4.9~$\mu$rad$^2$ (aperture size divided by the distance from the mirror to the VSS in H and to the source in V).
 The transmission coefficient is the ratio of the accepted N.A.
-over the beam divergence, resulting in 9.5\% (H) and 0.18\% (V)for ESRF-1 and 16.3\% and 0.18\% for EBS.
-The overall transmission is the product of H and V ratios, therefore 0.2\% for ESRF-1 and 2.95\% for EBS, therefore $3.8\cdot10^{12}$ photons/s for ESRF-1 and $8.5\cdot10^{13}$ photons/s for EBS.
+over the beam divergence, resulting in
+4.1\% (H) and 40\% (V)for ESRF-1 and 7.1\% and 40\% for EBS. The overall transmission considers the product of H and V ratios, therefore 
+0.21\% for ESRF-1 and 2.85\% for EBS, therefore 
+$4\cdot10^{12}$ photons/s for ESRF-1 and $8.3\cdot10^{13}$ photons/s for EBS. 
 %\todo{check with references $10^9$ at 33.6 keV.

### 2019-06-21 paper-hierarchical.tex
some fixes

@@
 A fully coherent hard X-ray beam is still a dream for the new generation storage rings.
-Despite a considerable increase in the coherent fraction of the new beams, the coherent fraction at typically hard X-ray photon energies is still a few percent for the new sources.
+Despite a considerable increase in the coherent fraction for the upgraded sources, at hard X-ray photon energies it is still a few percent.
 Consequently, the ray-tracing analysis does not become obsolete but has to be complemented with an analysis based on wave optics.
@@
 In Section \ref{level0} we calculate with simple methods approximate values of focal size, flux and coherent fraction.
-Accuracy can be improved using geometrical methods in Section \ref{level1} that allow a more complete simulation of the beamline optics within certain limits which are related, in this example, to high demagnification and high coherence of the studied beamline.
+Accuracy can be improved using ray-tracing methods in Section \ref{level1} that allow a more complete simulation of the beamline optics within certain limits which are related, in this example, to high demagnification and high coherence of the studied beamline.
 The wave optics section~\ref{level2} includes different models of increasing complexity as the required accuracy of the results becomes more demanding.
@@
 Values taken at the symmetry point of the straight section, where the insertion devices are placed.
-Values are taken from \cite{ESRF2014}, updated for EBS with
+Values are taken from \cite{orangebook}, and updated for EBS with
     the lattice labelled S28D. 
@@
 
-The source size and divergence are usually approximated supposing a Gaussian distribution of the electron positions and divergences (Table \ref{tab:eBeam}) and single-electron photon source size and divergence.
+The source size and divergence of the photon source are usually calculated approximately supposing Gaussian distribution of both the electron positions and divergences (Table \ref{tab:eBeam}) and the single-electron emission.
 In some cases, such as for undulator emission, this Gaussian approximation may not be accurate and have some implications that will be discussed later.
@@
 
-Using the values for U18.3, $n$=1, and  $E$=17225~eV (Table \ref{tab:Undulators}) and the optical configuration described in Table \ref{tab:Positions}, one obtains a focal size 139$\times$5~nm$^2$ (H$\times$V) for ESRF-1 and $10\times5$~nm$^2$ for EBS (Table~\ref{tab:HandCalculations}).
+Using the values for U18.3, $n$=1, and  $E$=17225~eV and the optical configuration described in Table \ref{tab:Positions}, one obtains a focal size 139$\times$5~nm$^2$ (H$\times$V) for ESRF-1 and $10\times5$~nm$^2$ for EBS (Table~\ref{tab:HandCalculations}).
 The horizontal focal spot for the ESRF-1 lattice is larger than the specified target value for the beamline, therefore the VSS could be closed to 50~$\mu$m thus reducing the beam size at the VSS plane by a factor ca.
@@
 over the beam divergence, resulting in
-4.1\% (H) and 40\% (V)for ESRF-1 and 7.1\% and 40\% for EBS. The overall transmission considers the product of H and V ratios, therefore 
-0.21\% for ESRF-1 and 2.85\% for EBS, therefore 
+4.1\% (H) and 40\% (V) for ESRF-1 and 7.1\% and 40\% for EBS. The overall transmission considers the product of H and V ratios, therefore 
+0.21\% for ESRF-1 and 2.85\% for EBS, thus giving 
 $4\cdot10^{12}$ photons/s for ESRF-1 and $8.3\cdot10^{13}$ photons/s for EBS. 
@@
 
-In summary, we have shown in this section how simple analytical calculations that can be done mostly ``by hand'' help to estimate the geometrical beam size at focal position (17$\times$5~nm$^2$ for ESRF-1, 10$\times$5~nm$^2$ for the EBS) the flux ($3.8\cdot10^{12}$ photons/s for ESRF-1 and $8.5\cdot10^{13}$ photons/s for EBS) and also anticipate that the beam is highly coherent producing a diffraction-limited focal spot of  7.75$\times$7.75~nm$^2$.
+In summary, we have shown in this section how simple analytical calculations that can be done mostly ``by hand'' help to estimate the geometrical beam size at focal position (17$\times$5~nm$^2$ for ESRF-1, 10$\times$5~nm$^2$ for the EBS) the flux ($4.0\cdot10^{12}$ photons/s for ESRF-1 and $8.3\cdot10^{13}$ photons/s for EBS) and also anticipate that the beam is highly coherent producing a diffraction-limited focal spot of  7.75$\times$7.75~nm$^2$.
 Thus far we ignored the effect of real optics exhibiting surface errors, the reflectivity of the elements, and other aspects that will be treated in depth in the next sections.
@@
 The transmission values are 0.16\% for ESRF-1 and 2.22\% for EBS.
-These values confirm the analytical estimations (0.2\% for ESRF-1 and 2.95\% for EBS).
+These values confirm the analytical estimations (0.21\% for ESRF-1 and 2.85\% for EBS).
 As mentioned before, these values are similar to the coherent fraction \cite{arxivCF} (0.12 for ESRF-1 and 2.73\% for EBS), meaning that a very high coherence of the beam is expected.
@@
 This is more evident in the vertical direction, because of the smaller focal size.
-The spot size is about 7-8~nm in V for both storage ring lattices, and in horizontal is about 15~nm for ESRF-1 High$\beta$ and 11~nm for EBS.
+The spot size is 8~nm in V for both storage ring lattices, and in horizontal is about 15~nm for ESRF-1 High$\beta$ and 11~nm for EBS.
 It is interesting to note the almost round spot that will be obtained with the new EBS lattice.
@@
 The source follows Gaussian distributions with parameters in Table~\ref{tab:HandCalculations}.
-a) Simulations using the ideal reflectors (cylindrical for ML, elliptical for the BK mirrors) without slope errors. 
+a) Simulations using the ideal reflectors (cylindrical for ML, elliptical for the KB mirrors) without slope errors. 
 b) Images calculated using the hybrid method (ray-tracing plus coherent diffraction) with ideal reflectors with no slope errors.
@@
 A next step in the wave optics simulation is, assuming full coherence, to generate a 2D source wavefront that matches in detail the real source (an undulator in a storage ring) and then simulate the propagation of such a wavefront through the optical elements with models that reproduce well the characteristics of the real elements.
-Once the source is defined, an approach of how best to simulate the beamline optical elements.
 Once the source is defined, several approaches of increasing complexity are used: a) use ideal thin lenses preceded by slits to emulate finite apertures (as done in the 1D model presented in Section~\ref{wofry}) b) use grazing incidence focusing optics models \cite{Canestrari2014}; and c) introduce the effect of imperfect optics on the simulations, i.e.
@@
 The elliptical mirror presents a slight reduction in intensity and asymmetry but no significant differences are observed, which is in agreement with the findings of \cite{Canestrari2014}.
-Notice that the differences are minimal and to accentuate them we have chosen logarithmically scaled plots to represent the intensity distributions at the focal position.
 It is also important to mention that this equivalency between ideal lenses and elliptical mirrors is only valid close to the focal position: down- or upstream the focal position, differences are no longer negligible - c.f.
@@
     \caption{
-    Intensity profile at the sample position produced by multi-electron method sampled with parameters of the ESRF-1 lattice, using ideal focusing elements (top), and including the effect of the mirrors slope errors (bottom).
+    Intensity profile at the sample position produced by multi-electron method sampled with parameters of the ESRF-1 lattice.
+Top: the perfect focusing with ideal lenses (green) is compared with the focusing using elliptical mirrors (blue).
+Bottom: the same as in the previous plot, but adding the effect of the mirrors slope errors in both cases.
     }
@@
     \caption{
-    Intensity profile at the sample position produced by multi-electron method sampled with parameters of the EBS lattice, using ideal focusing elements (top), and including the effect of the mirrors slope errors (bottom).
+    Intensity profile at the sample position produced by multi-electron method sampled with parameters of the EBS lattice. 
+    Top: the perfect focusing with ideal lenses (green) is compared with the focusing using elliptical mirrors (blue).
+Bottom: the same as in the previous plot, but adding the effect of the mirrors slope errors in both cases.
     }
@@
 This analysis of the mode transmission is more dramatic if one looks at the total intensity in the mode, which is given by its eigenvalue.
-Neglecting modes higher than 10, the lowest coherent mode ($m$=0) carries the 49.5\% of the beam intensity, 18.4\% for the mode $m$=2, and less than 5\% for the others.
+Neglecting modes higher than 10, the lowest coherent mode at the source ($m$=0) carries the 49.5\% of the beam intensity after being cropped by the beamline, 18.4\% for the mode $m$=2, and less than 5\% for the others.
 
@@
 After the complete analysis done in the previous section on how the ID16A optics modify the characteristics of the X-ray beam, one can verify that the analytical previsions (Section~\ref{level0}) were a good estimator of the values obtained using more sophisticated methods.
-The size of the focused beam (e.g, 10$\times$5~nm$^2$ for EBS) and the transmission (3\%) calculated analytically are in good agreement with geometric ray-tracing (8.4$\times$4.8~nm$^2$, and 2.28\%).
+The size of the focused beam (e.g, 10$\times$5~nm$^2$ for EBS) and the transmission (2.85\%) calculated analytically are in good agreement with geometric ray-tracing (8.4$\times$4.8~nm$^2$, and 2.28\%).
 When characterizing beam sizes, one usually assumes that the intensity distributions are Gaussian and therefore they are well characterized by the FWHM value.
@@
 
-The combination of ray-tracing and wave optics methods used indicate that the focalized beam size due to  demagnification with mirrors (giving for EBS 10$\times$5~nm$^2$ analytically, 8.8$\times$4.5~nm$^2$ with ray-tracing) has to be corrected with diffraction effect (11.2$\times$8~nm$^2$ hybrid, 12.56$\times$9.38~nm$^2$ SRW) and then add the slope error effect.
-The slope errors do not blow up the size and kit it in the specifications range (11.4$\times$8.5~nm$^2$ hybrid 13.6$\times$9.14~nm$^2$ SRW).
+The combination of ray-tracing and wave optics methods used indicate that the focalized beam size due to  demagnification with mirrors (giving for EBS 10$\times$5~nm$^2$ analytically, 8.8$\times$4.5~nm$^2$ with ray-tracing) has to be corrected with diffraction effect (11.2$\times$8~nm$^2$ hybrid, 11.6$\times$9.4~nm$^2$ SRW) and then add the slope error effect.
+The slope errors do not blow up the size and kit it in the specifications range (11.4$\times$8.5~nm$^2$ hybrid 12.5$\times$9.1~nm$^2$ SRW).
 It is important to treat the slope errors correctly.
@@
 If the final beam has a very high coherence coherent optics (1D simplifications, and single-electron calculations) may give good results.
-Full calculations based on coherent mode decomposition Were introduced here.
+Full calculations based on coherent mode decomposition were introduced here.
 It is a new way of quantitatively assess the quality of the partially coherent beam via the calculation of the coherent modes of the source, propagation of the modes and recalculation of the coherent mode expansion on the final beam.
@@
 \begin{figure}\label{fig:metrology}
-\includegraphics[width=0.95\textwidth]{GRAPHICS/metrology.eps}
+\includegraphics[width=0.9\textwidth]{GRAPHICS/metrology.eps}
 \caption{Error height profiles used for the simulations (left) and power spectral density (right) for the ML monochromator (top row), the KB vertical focusing mirror (central row), and the KB horizontal focusing mirror (bottom row).}

### 2019-06-21 paper-hierarchical.tex
last fixes

@@
 However, apertures centered at the optical axis would clean the beam from these undesired photons.
-If we compare the values of the beamline transmission due to the geometrical aperture  (0.2\% for ESRF-1 and 2.95\% for EBS) with the CF values, we remark that they are very close.
+If we compare the values of the beamline transmission due to the geometrical aperture  (0.21\% for ESRF-1 and 2.85\% for EBS) with the CF values, we remark that they are very close.
 This suggests that the beam at the focal position will be highly coherent because the fraction of the photons removed from the beam matches the CF value that quantifies the fraction of ``good'' or ``coherent'' photons.
@@
 \begin{figure}\label{fig:metrology}
-\includegraphics[width=0.9\textwidth]{GRAPHICS/metrology.eps}
+\includegraphics[width=0.9\textwidth]{GRAPHICS/metrology.pdf}
 \caption{Error height profiles used for the simulations (left) and power spectral density (right) for the ML monochromator (top row), the KB vertical focusing mirror (central row), and the KB horizontal focusing mirror (bottom row).}

### 2019-06-21 mkJSR.sh

@@ -7,9 +7,9 @@
 #
 
 cd /users/srio/Working/paper-hierarchical
-rm -rf ARXIV2 ARXIV
-mkdir ARXIV
-cd ARXIV
+rm -rf JSR JSR
+mkdir JSR
+cd JSR
 cp ../paper-hierarchical.tex .
 #cp ../paper-hierarchical.bbl .
 cp ../iucr.cls .
@@ -97,8 +97,8 @@ sed -e "s/GRAPHICS\/final_mode0.png/figure12a.png/" paper-hierarchical.tex -i
 sed -e "s/GRAPHICS\/final_mode1.png/figure12b.png/" paper-hierarchical.tex -i
 
 # figure 13
-cp GRAPHICS/metrology-eps-converted-to.pdf ./figure13.pdf
-sed -e "s/GRAPHICS\/metrology.eps/figure13.pdf/" paper-hierarchical.tex -i
+cp GRAPHICS/metrology.pdf ./figure13.pdf
+sed -e "s/GRAPHICS\/metrology.pdf/figure13.pdf/" paper-hierarchical.tex -i
 
 rm -rf GRAPHICS
 

### 2019-08-09 paper-hierarchical.tex
revised version

@@
 
+ %-------------------------------------------------------------------------
+ %
+ %-------------------------------------------------------------------------
+
 \section{Introduction}
@@
 It uses one or several undulators as the source.
-During the period 2011-2018, it operated exploiting the ESRF-1 storage ring’s double-bend achromat (DBA) lattice with horizontal emittance about 4000~pm.
+During the period 2011-2018, it operated exploiting the ESRF-1 storage ring double-bend achromat (DBA) lattice with horizontal emittance about 4000~pm.
 In the second phase of the ESRF Upgrade, the original storage ring is being replaced by a new hybrid multi-bend achromat (HMBA) lattice.
 This new lattice will drastically improve the performances of the ESRF source.
-The resulting Extreme Brilliant Source (EBS) \cite{orangebook} will increase the brightness and coherent fraction by a factor of more than 100, by strongly reducing the horizontal emittance.
+The resulting Extreme Brilliant Source (EBS) \cite{orangebook} will increase the brightness and coherent fraction by \inred{one or two orders of magnitude}, by strongly reducing the horizontal emittance.
 This is possible using almost the same ring circumference and electron energy $E_e$.
@@
 It is dedicated to nanofocusing applications.
-The beamline can operate at two fixed photon energies: 17050~eV and 33600~eV.
+The beamline can operate at two fixed photon energies \inred{(at ca.
+17 and 33 keV) }.
 The straight section in the magnet lattice houses four undulators, two 18~mm and another two of 22.4~mm period.
@@
 \centering
-\caption{Parameters of the undulators available at the ID16A beamline: period ($\lambda_u$), length ($L$), number of identical undulators ($N$) and deflection parameter ($K$) values for the three main photon energies in use.
-The valuse are shown for the two ESRF magnetic lattices: the old ESRF-1, with $E_e = 6.04$~GeV), and the new ESRF-EBS, with $E_e = 6.00$~GeV.
+\caption{Parameters of the undulators available at the ID16A beamline: period ($\lambda_u$), length ($L$), number of identical undulators ($N$) and deflection parameter ($K$) values for \inred{two main photon energies}.
+The valuse are shown for the two ESRF magnetic lattices: the old ESRF-1, with $E_e = 6.04$~GeV), and the new EBS, with $E_e = 6.00$~GeV.
 $n$ is the emission harmonic in use for each particular photon energy.}
@@
 \begin{tabular}{lcccccc}
-name  & $\lambda_u$ [mm] & $L$ [m] & N & $K$ (11200~eV) & $K$ (17050~eV) & $K$ (33600~eV) \\ \\
+name  & $\lambda_u$ [mm] & $L$ [m] & N &  $K$ (17050~eV) & $K$ (33600~eV) \\ \\
 %\input{tableUndulator.txt}
- ESRF-1 U18.3  & 18.30  & 1.40  & 2 & 1.175(n=1) & 0.470(n=1) & 1.175(n=3)\\ 
- ESRF-1 U22.4  & 22.40  & 1.40  & 2 & 0.873(n=1) & 1.855(n=3) & 0.873(n=3)\\ \\
- EBS U18.3  & 18.30  & 1.40  & 2 & 1.156(n=1) & 0.437(n=1) & 1.156(n=3)\\ 
- EBS U22.4  & 22.40  & 1.40  & 2 & 0.852(n=1) & 1.836(n=3) & 0.852(n=3)\\ 
+ ESRF-1 U18.3  & 18.30  & 1.40  & 2 & 0.470(n=1) & 1.175(n=3)\\ 
+ ESRF-1 U22.4  & 22.40  & 1.40  & 2 & 1.855(n=3) & 0.873(n=3)\\ \\
+ EBS U18.3       & 18.30  & 1.40  & 2 &  0.437(n=1) & 1.156(n=3)\\ 
+ EBS U22.4       & 22.40  & 1.40  & 2 &  1.836(n=3) & 0.852(n=3)\\ 
 \end{tabular}%
@@
 In the sketch ML stands for the focusing multilayer monochromator; VSS for virtual source slit; and KB for the Kirkpatrick-Baez set of mirrors.
-The sketch is drawn out of scale.}
+The sketch is drawn \inred{not to} scale.}
 \end{figure}
@@
     \centering
-    \caption{ESRF-1 and EBS main electron beam parameters.
-Electron beam parameters for the high-$\beta$ (labelled ESRF) and the EBS straight sections.
-Values taken at the symmetry point of the straight section, where the insertion devices are placed.
-Values are taken from \cite{orangebook}, and updated for EBS with
-    the lattice labelled S28D. 
+    \caption{ \inred{Electron beam sizes for the ESRF-1 high-$\beta$ straight sections and the EBS straight sections.
+Values taken at the symmetry point of the straight section, where the insertion devices are placed}.
+Values are  from \cite{orangebook}, and updated for EBS with the lattice labelled S28D.
     %For EBS, two (very close) design values are used, from Ref.~\cite{ESRF2014} labelled EBS-S28A, and those coming from further lattice refinement, labelled EBS-S28D. 
@@
 
+ %-------------------------------------------------------------------------
+ %
+ %-------------------------------------------------------------------------
+
 \section{Simulations}
@@
 
-For all simulations presented hereafter, we selected the configuration using a single undulator U18.3 placed in the center of the straight section and tuned to its first harmonic at a photon energy of $E=17225$~eV: a value close to the usual operation value shown in Table \ref{tab:Undulators}.
+For all simulations presented hereafter, we selected the configuration using a single undulator U18.3 placed in the center of the straight section and tuned to its first harmonic at a photon energy of $E=17225$~eV (Table \ref{tab:Undulators}).
 Calculations were done using the two different storage ring lattices mentioned before: the (already dismantled) ESRF-1 where the undulators are in a high-$\beta$ straight section (larger horizontal emittance) and the new EBS (lower horizontal emittance) straight sections.
@@
 
+ %-------------------------------------------------------------------------
 \subsection{Analytical calculations}
@@
 The electron beam usually presents different horizontal and vertical beam emittances, but the undulator natural source size or divergence presents radial symmetry at the central cone (excluding polarization).
+\inred{Eqs.~\ref{eq:photon small sigmas} are approximations, valid at the resonance, and under the assumptions that the undulator emission has Gaussian distribution in both angular and spatial coordinates, which is not the case as discussed below and in Ref.~\cite{arxivCF}} .
 The photon source size and divergence are approximately given by the convolution of the electron beam sizes and the undulator natural sizes.
@@
 The beam sizes at different positions in the beamline are estimated in different ways depending upon whether the beam is focused or not.
-If the beam is focused, its dimensions, to a first approximation, are the photon source size multiplied by the corresponding magnification factor.
+If the beam is focused, its dimensions, to a first \inred{geometrical} approximation \inred{also neglecting diffraction effects}, are the photon source size multiplied by the corresponding magnification factor.
 For predicting the photon beam sizes in a position out of the focus the beam divergence is used.
 In an ideal {\'{e}}tendue-conserving optical system, where the Smith-Helmholz invariant applies, the resulting beam divergence transforms inversely as the beam size: it can be calculated as the product between the beam divergence at the source with the demagnification factor $M^{-1}$.
+\inred{It is, however, limited by the numerical aperture accepted by the optical elements due to finite size.}
 
@@
 
-%Effects of slope errors and diffraction limits can be simply estimated by hand calculations as well...
-%The smearing of the focal size because of diffractive effects by the mirror clipping will also be discussed.
-
 \begin{table}\label{tab:HandCalculations}
@@
     \caption{Photon beam size and divergence for selected ID16A beamline positions.
-Values below were obtained considering undulator U18.3 tuned to its first harmonic at the photon energy of $E=17225$~eV for both the high-$\beta$ (ESRF-1) and the EBS straight sections.
+Values below were obtained considering undulator U18.3 tuned to its first harmonic at the photon energy of $E=17225$~eV for both the \inred{ESRF-1 (high-$\beta$)} and the EBS straight sections.
 Values are FWHM.
@@
     Position [m]:                          & 0.00                & 40.00                   & 185.00                  & 0.00    & 40.00   & 185.00       \\
-    FWHM$_h$ [$\mu$m]:      & 977.19        & 404.00 (50.0)& 0.139 (0.017)     & 71.26   & 29.46   & 0.0102       \\
-    FWHM$_v$ [$\mu$m]:      & 9.60          & 482          & 0.0052            & 10.02   & 486.5   & 0.0054       \\
+    FWHM$_h$ [$\mu$m]:      & 977.19            & 404.00 (50.0)    & 0.139 (0.017)     & 71.26   & 29.46   & 0.0102       \\
+    FWHM$_v$ [$\mu$m]:      & 9.60                 & 478.7                  & 0.0052                  & 10.00   & 483.6   & 0.0054       \\
     \newline
-    FWHM'$_h$ [$\mu$rad]:   & 26.96         & 65.22        & 189062   & 15.60   & 37.74   & 109404       \\
-    FWHM'$_v$ [$\mu$rad]:   & 12.04         & 12.04                 & 22264    & 12.16   & 12.16   & 22489        \\                                                   
+    FWHM'$_h$ [$\mu$rad]:   & 26.93            & 65.14                   & 7800                     & 15.54   & 37.60   & 7800       \\
+    FWHM'$_v$ [$\mu$rad]:   & 11.97            & 11.97                   & 9000                     & 12.09   & 12.09   & 9000        \\                                                   
     \end{tabular}%
@@
 
-Next, we want to estimate the number of photons in the sample.
+Next, we want to estimate the number of photons at the sample position.
 For that, we need the spectrum at the source.
@@
 This aperture opening is set to fully accept the central cone at the resonance peak of the harmonic in use.
-%($\max(\Sigma') \times d = 0.73$~mm), with $\Sigma'$ values from in Table~\ref{tab:HandCalculations}. 
-The peak intensity is about  $2.9\cdot10^{14}$ photons/s/0.1{\%}bw for EBS and $1.9\cdot10^{14}$ for ESRF-1 (see \ref{fig:FluxU18}).
+The peak intensity is about  $2.9\cdot10^{14}$ photons/s/0.1{\%}bw for EBS and $1.9\cdot10^{14}$ for ESRF-1 (see \inred{Fig.~} \ref{fig:FluxU18}).
 One can appreciate how the emittance plays an important role in both peak width and especially in peak value.
@@
     \includegraphics[width=0.95\textwidth]{GRAPHICS/fluxU18.png}
-    \caption{Flux of the U18.3 undulator integrated over a square slit placed at 27.1~m from the ID centre with aperture of 1 and 4 mm.}
+    \caption{Flux of the U18.3 undulator integrated over a square slit \inred{(1$\times$1~mm$^2$)} placed at 27.1~m from the ID centre.}
 \end{figure}
@@
 The multilayer monochromator accepts the whole beam.
-We consider a VSS closed to $S_h$=50~$\mu$m in horizontal and fully opened in vertical.
+We consider a \inred{slit (}VSS\inred{)} closed to $S_h$=50~$\mu$m in horizontal and fully opened in vertical.
 For ESRF-1 its transmission is 1/8 (as discussed before) and 1 for EBS.
@@
 $4\cdot10^{12}$ photons/s for ESRF-1 and $8.3\cdot10^{13}$ photons/s for EBS.
-%\todo{check with references $10^9$ at 33.6 keV.
-I get: 1.6e+15 * 0.00295 * 0.7**3 = 1.62e+12.
-How closed are the slits?
-We used the undulator peak value over the whole ML bandwidth, that overestimates the integrated flux.}
-%\todo{need to consider ML reflectivity for the 3 reflections, also undulator is not perfect, the horizontal slit may be closed to less than $50 \mu m$, also KB mirrors may not be fully illuminated and slope error losses due to initial ML mono are neglected}
 
 One can also make some simple inferences about the beam coherence.
-The coherent fraction (CF) \cite{arxivCF} of the undulator beam is the occupation of the lowest coherent mode and will be discussed further in Section~\ref{level2}.
+The coherent fraction (CF) of the undulator beam is the occupation of the lowest coherent mode \cite{arxivCF} and will be discussed further in Section~\ref{level2}.
 An approximated estimation of CF can be obtained using Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas}.
@@
 It is a good indicator of the quality of a coherent beam (a fully coherent beam has CF=1 and an incoherent beam CF=0).
-The beamline optics act as a ``coherence filter'' by removing the ``bad'' or ``incoherent'' photons and in consequence increasing the CF.
+The beamline optics act as a ``coherence filter'' by removing the ``incoherent'' photons and in consequence increasing the CF.
 A complete cleaning is impossible because of the spatial overlapping of the different coherence modes.
@@
 If we compare the values of the beamline transmission due to the geometrical aperture  (0.21\% for ESRF-1 and 2.85\% for EBS) with the CF values, we remark that they are very close.
-This suggests that the beam at the focal position will be highly coherent because the fraction of the photons removed from the beam matches the CF value that quantifies the fraction of ``good'' or ``coherent'' photons.
+This suggests that the beam at the focal position will be highly coherent because the fraction of the photons removed from the beam matches the CF value that quantifies the fraction of  ``coherent'' photons.
 
 To assess the focused beam size given by a coherent beam, one can consider the beamline in two sections: First, the ML monochromator and VSS that produce a coherent plane wave as the source (vertical) and VSS (horizontal) are both very far from the entrance aperture of the focusing optics.
-Second, the KB-system playing a dual role, acting as a focusing element of focal distance $f=5$~cm, but also cropping the beam by acting as a square aperture of roughly D$\times$D=400$\times$400~$\mu$m$^2$ (the length of the mirror projected in a plane perpendicular to the beam propagation).
-%\todo{REMOVE: we obtained 900 in H, that will be reduced by 0.5 as propagated from 10 cm to 5 cm upstream of the sample, and 390 in V approximated by 400}.
+Second, the KB-system playing a dual role, acting as a focusing element of focal distance $f=5$~cm, but also cropping the beam by acting as a square aperture of roughly D$\times$D=400$\times$400~$\mu$m$^2$ (the length of the mirror projected in a plane perpendicular to the beam propagation
+\footnote{\inred{we obtained 900 $\mu$m in horizontal that is reduced by a factor 0.5 as propagated from 10 cm to 5 cm upstream of the sample, and 390 $\mu$m in vertical.
+These numbers are approximated by 400$\times$400$\mu$m$^2$ }}
+).
 This aperture has two effects: i) as discussed before, it removes a lot of off-axis photons increasing the CF to a value close to one, and ii) broadens of the focal spot because of coherent diffraction.
 The diffraction by a square aperture of a collapsing spherical wavefront has an intensity distribution proportional to $\sinc^2(k D x/2 f) \sinc^2(k D y / 2 f)$,
-%\todo{include citation?}
 where $k=2\pi/\lambda$, and $x$ ($y$) is the horizontal (vertical) coordinate.
@@
 
-%-------------------------------------------------------------------------
-% 
 %-------------------------------------------------------------------------
@@
 This will permit us to calculate the refractive index that should be used for each ray.
-The characteristics of each ray can also be extended to include the electric field amplitude.
-This allows extending the pure geometrical tracing to include physical models that take into account the optical element reflectivity or transmission.
-This combination of using a geometrical model plus a physical model allows to simulate every element used in a synchrotron beamline, such as mirrors and combinations of them (e.g., KBs); lenses, compound refractive lenses, and transfocators; gratings of any type and shape, crystal systems, etc.
+The characteristics of each ray can also be extended to include the electric field amplitude\inred{, thus complementing} the pure geometrical tracing to include physical models that take into account the optical element reflectivity or transmission.
+This combination of using a geometrical model plus a physical model \inred{permits to simulate} every element used in a synchrotron beamline, such as mirrors and combinations of them (e.g., KBs); lenses, compound refractive lenses, and transfocators; gratings of any type and shape, crystal systems, etc.
 Ray-tracing is a simple and extremely powerful technique for calculating the main characteristics of the photon beam (size, divergence, photon distribution) at every point of the beamline.
-%Although very powerful commercial ray-tracing packages are available in the market, they are generally ill-adapted for synchrotron simulations because they do not cover the particularities of X-ray devices and sources.
-For that purposes,
 Several packages created in the synchrotron community and are available, like SHADOW \cite{codeSHADOW}, RAY \cite{codeRAY}, McXtrace \cite{codeMCXTRACE} or XRT \cite{codeXRT}.
@@
 \includegraphics[width=0.95\textwidth]{GRAPHICS/figure3.png}
-\caption{Image produced by different types of ideal focusing for the ESRF (left column) and EBS (right column).
+\caption{Image produced by different types of ideal focusing for the \inred{ESRF-1} (left column) and EBS (right column).
 The source follows Gaussian distributions with parameters in Table~\ref{tab:HandCalculations}.
@@
 c) Simulations using the hybrid method including slope errors from the real mirror profiles measured at the ESRF metrology lab (see Appendix~\ref{appendix:metrology}).
-It can be observed further beam broadening and the growth of satellite structures.
 The focal beam dimensions (FWHM of the histograms) are written in the plot title.
@@
  %-------------------------------------------------------------------------
-% 
-%-------------------------------------------------------------------------
-
 \subsection{Wave optics simulations}
@@
 
-As discussed in Section~\ref{level0}, the clipping of the beam, mainly due to the acceptance of the KB mirrors produces a significant broadening of the focal spot.
+As discussed in Section~\ref{level0}, the clipping of the beam, mainly due to the acceptance of the KB mirrors\inred{,} produces a significant broadening of the focal spot.
 Very simple calculation of the diffraction limit can be done using a convergent spherical beam clipped by a slit of 400$\times$400$~\mu$m$^2$ placed 5~cm upstream of the focus (at the second KB mirror position), producing a spot of about 7.75$\times$7.75~nm$^2$, therefore qualifying this beamline to be a diffraction-limited beamline.
@@
 slope and figure errors to both ideal lenses and mirrors.
-The computer code chosen for these calculations is ``Synchrotron Radiation Workshop'' (SRW) \cite{codeSRW}, which over the past two decades has been extensively benchmarked throughout different synchrotron radiation facilities and has become widely accepted within the X-ray community.
-
-%One of the difficulties with wave-optics simulations is the correct sampling and range of the input wavefront.
-If chosen poorly, the simulation will be affected by replicas and aliasing, compromising, then the quality of the obtained data.
-Choosing the range of the wavefront is straightforward: one can base the choice either on the first optical element aperture or on the beam footprint at its entrance.
-Usually one chooses the smaller.
-Choosing a very small range for the wavefront leads to cropping and diffraction effects that may not come from the beamline, but only a numerical consequence of truncating the wavefront.
-For the ID16A, the first optical element is the multi-layer monochromator and its aperture accepts the full beam of both the ESRF-1 and EBS lattices.
-%Choosing the correct sampling is a more difficult task: good sampling implies choosing an appropriate number of pixels such as the phase of the electric field being propagated varies less than $\pi$ between two adjacent pixels.
-This rule may lead to very big arrays: SRW keeps in memory both $\sigma-$ and $\pi-$polarizations.
-Doubling the number of pixels on both directions of a 2D array implies an increase in memory consumption of about 8 times.
-
-%All simulations using SRW were done using the same grid ($3.5\times1.5$mm$^2$) and initial sampling ($740\times150$) - the initial wavefront was resampled by a factor of 8 in order to mitigate aliasing and replicas throughout the beamline.
+The computer code chosen for these calculations is ``Synchrotron Radiation Workshop'' (SRW) \cite{codeSRW}, which over the past two decades has been extensively benchmarked \inred{in} different synchrotron radiation facilities and has become widely accepted within the X-ray community.
 
@@
 Increasing the complexity of the model increments considerably the simulation time, which went from typically $\sim16$~s to $\sim24$~s.
-
-% Table \ref{tab:SRW_results} compiles the beam sizes at focal position for mentioned simulations.
 
@@
 This initial point considered as source cannot be the source physical position, because i) a point or spherical wave would concentrate all its intensity in a single pixel so the wavefront is ill-defined, and ii) the theory (e.g., \cite{jackson}) used for calculating of  synchrotron emission (i.e., undulators) permits the calculation of the radiation in positions far from the electron trajectory.
-The limitation of the wavefront propagation method is that a single wavefront is not representative for a beam that: in storage rings it is produced by the incoherent supperposition of the emission of many electrons circulating through the corresponding magnet structure.
+\inred{The the wavefront propagation method is limited by the fact that a single wavefront is not representative for a beam.
+Multiple wavefronts must ba taken into account to include effects of partial coherence.
+In storage rings, partial coherent due to the supperposition of the emission of many individual non-identical electrons traversing a magnet structure.
+}
 
 Diffraction limited sources would emit radiation that can be described in good approximation by a single wavefront.
-But even for new ultra low-emittance storage rings (improperly called diffraction-limited sources, see discussion in \cite{arxivCF}) the contribution of the electron dimensions ($\sigma_{e(h,v)}$ and $\sigma'_{e(h,v)}$) is not negligible as compared with the natural emission of radiation ($\sigma_u$ and $\sigma'_u$ for undulators as in Eq.~\ref{eq:photon small sigmas}).
+But even for new ultra low-emittance storage rings (improperly called diffraction-limited sources, see discussion in \cite{arxivCF}) the contribution of the electron dimensions ($\sigma_{e(h,v)}$ and $\sigma'_{e(h,v)}$) is not negligible as compared with the natural emission of radiation \inred{at photon energies of interest} ($\sigma_u$ and $\sigma'_u$ for undulators as in Eq.~\ref{eq:photon small sigmas}).
 To account for this effect one should complete or extend the wavefront simulations to take into account that in fact, not only a single wavefront contributes to the properties of the beam.
@@
 
-There are three approaches currently used for including the electron emittance effects in the wave optics simulations: i) use of convolutions, ii) Monte Carlo multi-electron sampling, and iii) Propagation of coherent modes.
+There are three approaches currently used for including the electron emittance effects in the wave optics simulations: i) use of convolutions, ii) Monte Carlo multi-electron sampling, and iii) \inred{p}ropagation of coherent modes.
 The convolution method consists of calculating the emission characteristics and then convolving the resulting intensity with Gaussians describing the electron beam.
-This method is simple and fast, as only requires simulation of a single wavefront.
+This method is simple and fast, as \inred{it} only requires simulation of a single wavefront.
 However, it presents two important limitations: it can only be applied to the intensity calculations and not to other parameters of interest for describing coherence, such as the degree of coherence, etc.; and it can only be applied at a plane right after the source position (it is usually unknown how the different optical elements will affect the source contribution to the phase space).
@@
 The second is computing resources.
-The complex task of calculating the emission of one electron, propagated it through every element of the beamline socoring the results must be repeated thousand of times.
+The complex task of calculating the emission of one electron, propagated it through every element of the beamline \inred{and scoring the results} must be repeated thousand of times.
 The third method, the propagation of coherent modes, will be discussed in the next section.
@@
 
-%{\todo{more detailed comments on comparisons with filament beam results? For comparison also information on calculation time, number of electrons, ...}
-
-% 
-% Synchrotron radiation on the hard X-rays regime is hardly full coherent even for the so-called diffraction limited sources (see Table \ref{tab:Coherence}).
-To better describe the nature of the emitted radiation, one makes use of the partial coherent simulations using the \textsc{srw}, i.e.
-incoherent addition of the intensity from the emission from electrons sampled from the source's 6D phase-space \cite{Chubar2011b}.
-This is a natural extension from the method presented on the previous section.
-
-% 
-% \newpage
-% 
-%  %-------------------------------------------------------------------------
-%  % Manuel (?)
-%  %-------------------------------------------------------------------------
-%  
 \subsubsection{Full Wave optics simulation: partial coherence case by means of coherent mode decomposition}
 \label{comsyl}
-% 
-% One of the mentioned drawbacks of the Monte Carlo multi-electron methods is the a-priori unability to determine the size of sampling space (i.e., how many electrons) for accurate calculations. 
-% Another important limitation is the impossibility with available computer resources to store (although can be computed) the full 4D matrices fully describing the partial coherence properties (cross spectral density) at any point of the beamline.
 
@@
 Fig.~\ref{fig:histomodes}a shows the cumulated occupation of the coherent modes for the full central cone undulator emission.
-It can be noticed that the EBS cumulated occupation grows very steeply with the number of coherent modes, whereas for ESRF-1 the number of coherent modes needed to attain a given value of cumulated occupation much higher.
-This can also be shown in Fig.~\ref{fig:histomodes}b, where the number of modes to fill 50\%, 75\% and 95\% of spectral density is compared for EBS and ESRF-1.
+The EBS cumulated occupation grows very steeply with the number of coherent modes, whereas for ESRF-1 the number of coherent modes needed to attain a given value of cumulated occupation much higher.
+This can also be shown in Fig.~\ref{fig:histomodes}b, where the number of modes to fill 50\%, 75\% and \inred{85\%} of spectral density is compared for EBS and ESRF-1.
 
@@
 \end{equation}
-Figsures~\ref{fig:spectraldensity}b-e represent the coherent modes with $m$=0 to $m$=3.
+Figure~\ref{fig:spectraldensity}b represents the coherent modes with $m$=0 to $m$=3.
 It can be shown that the first mode is centered on the axis but extends over a spatial region smaller than the spectral density.
-The next mode ($m$=1) contributes to extend the spectral density in horizontal but presents zero intensity at the origin.
+The next mode ($m$=1) \inred{extends the spectral density in the horizontal direction} but presents zero intensity at the origin.
 Successive modes extend the spectral density in horizontal, sometimes presenting a lobe center at zero and sometimes not (e.g., $m$=3).
-Modes with several lobes along the vertical will also appear, but not so often than in the horizontal, because of the elliptical shape of the spectral density.
-The coherent modes are orthonormal, thus verifying
+Modes with several lobes along the vertical \inred{direction} will also appear, but not so often than in the horizontal, because of the elliptical shape of the spectral density.
+The coherent modes are orthonormal,
 \begin{equation}\label{eq:mode normalization}
@@
 Fig.~\ref{fig:mode transmission} shows how the intensity of the first modes is reduced by geometrical considerations.
-%We recall that in the coherent mode expansion, the modes are orthonormal, therefore the integrated intensity for any mode is always one at the source. 
 As discussed, coherent modes will be ``cut'' in a different manner, depending on the effect of the slit.
@@
 The other modes shown transfer less than 5\% transfer, and modes higher than $m$=10 practically do not contribute to the final spectral density.
-Note that, as discussed before, the transmission of mode $m$=2 is higher than mode $m$=1, because this one does not display a central lobe (Fig.~\ref{fig:spectraldensity}c).
+Note that, as discussed before, the transmission of mode $m$=2 is higher than mode $m$=1, because this mode does not display a central lobe (Fig.~\ref{fig:spectraldensity}c).
 
@@
 Making this numerical calculation with COMSYL, one obtains a mode occupation spectrum (see Fig.~\ref{fig:rediagonalization}) where the lowest mode accounts for 83\% of the total spectral density.
-The next mode accounts for only 13\% and the rest is practically negligible.
+The next mode accounts for only 13\% and the \inred{the next ones are} practically negligible.
 The intensity of these two coherent modes is shown in Fig.~\ref{fig:final modes}.
@@
 \end{figure}
-%\todo{
-%\begin{itemize}
-% \item introduce slope errors in WOFRY propagation
-% \item propagate modes with SRW
-% \item create some plots with cuts of W
-%\end{itemize}
-%}
+
+ %-------------------------------------------------------------------------
+ %
+ %-------------------------------------------------------------------------
+\vspace{1cm}
 
@@
 From Fig.~\ref{fig:ray-tracing} one can see that the horizontal intensity distribution is not Gaussian for the ESRF-1, as a result of cropping the beam with the VSS opened to 50$\mu$m.
-Although the electron beam statistics is in a very good approximation Gaussian, the undulator emission is not.
+Although the electron beam statistics inred{is Gaussian in good approximation}, the undulator emission is not.
 In our ray-tracing simulations, we always suppose that we work at the resonance, then we applied the Gaussian approximation used in Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas} to compute the geometrical characteristics of the source.
 This is certainly acceptable at the resonance but will fail out of resonance.
-Indeed, this lack of Gaussianity is relevant in the calculations of the coherent fraction.
-We approximate the coherence fraction by Eq.~\ref{eq:coherent fraction}, which for zero emittance ($\sigma_{x,y}=\sigma_{x',y'}$=0) becomes CF=1, as expected because a filament electron beam in a magnetic field emits a fully coherent wavefront.
-However, the photon source for a filament beam would have an emittance of $\sigma_u \sigma_{u'} \approx \lambda / 2 \pi$ (using Eqs.~\ref{eq:photon small sigmas}), which is in contradiction with theory, that says that a pure coherent state has emittance of $\lambda / 4 \pi$, as happens for a Gaussian photon beam.
-This is justified considering that Eqs.~\ref{eq:photon small sigmas} are approximations obtained by fitting with Gaussians the theoretical non-Gaussian distributions (see Ref.~\cite{elleaume}).
+Indeed, this \inred{separation from of Gaussian behaviour} is relevant in the calculations of the coherent fraction.
+We approximate the coherence fraction by Eq.~\ref{eq:coherent fraction}, which for zero emittance ($\sigma_{e,(x,y)}=\sigma'_{e,(x,y)}$=0) becomes CF=1, as expected because a filament electron beam in a magnetic field emits a fully coherent wavefront.
+However, the photon source for a filament beam would have an emittance of $\sigma_u \sigma'_{u} \approx \lambda / 2 \pi$ (using Eqs.~\ref{eq:photon small sigmas}), which is in contradiction with theory, that says that a pure coherent state has emittance of $\lambda / 4 \pi$, as happens for a Gaussian photon beam.
+\inred{
+This apparent contradiction is due to the fact that we deal with approximations: consider that Eqs.~\ref{eq:photon small sigmas} are approximations obtained by fitting with Gaussians the theoretical non-Gaussian distributions (see Ref.~\cite{elleaume}), but the resulting Gaussians are not Fourier transform pairs.
+If instead using \sigma'_{u}$ from Eq,~\ref{eq:photon small sigmas}, we force it to verify   $\sigma_u \sigma'_{u} = \lambda / (4 \pi)$, it underestimates the source size (this can be numerically verified using e.g., SRW).
+These inconsistencies are also translated to the CF if calculated by Eq.~\ref{eq:coherent fraction}.
+} 
 The correct definition of CF as the occupation of the lowest coherent mode does not present these incongruences.
-However, its calculation requires to perform the coherent mode decomposition, as done in Section~\ref{comsyl} which is very costly.
+However, its calculation requires to perform the coherent mode decomposition, as done in Section~\ref{comsyl} which is costly.
 It is, therefore, the case to find the compromise between accuracy and cost, or where in the hierarchical series of methodologies proposed one should stop.
@@
 The combination of ray-tracing and wave optics methods used indicate that the focalized beam size due to  demagnification with mirrors (giving for EBS 10$\times$5~nm$^2$ analytically, 8.8$\times$4.5~nm$^2$ with ray-tracing) has to be corrected with diffraction effect (11.2$\times$8~nm$^2$ hybrid, 11.6$\times$9.4~nm$^2$ SRW) and then add the slope error effect.
-The slope errors do not blow up the size and kit it in the specifications range (11.4$\times$8.5~nm$^2$ hybrid 12.5$\times$9.1~nm$^2$ SRW).
+The slope errors do not blow up the size and \inred{keep} it in the specifications range (11.4$\times$8.5~nm$^2$ hybrid 12.5$\times$9.1~nm$^2$ SRW).
 It is important to treat the slope errors correctly.
@@
 This is the standard model of including slope errors in SHADOW, using measured profiles or synthetic ones.
-It works well for the case of incoherent beams, but in our case, because of the high coherence of the beam the ray-tracing method (specular reflection) overestimates the blowing of the focal spot conducting to wrong results.
+It works well for the case of incoherent beams, but in our case, because of the high coherence of the beam the ray-tracing method (specular reflection) overestimates the blowing of the focal spot.
 Therefore the slope errors of the KB system have been considered using a ``wave optics approach'', as explained in Ref.~\cite{hybrid}.
@@
 This is the algorithm used in the simulations in Sections~\ref{level1}, \ref{srw_se} and \ref{srw_me}.
-
-% The first step in the wave optics simulation is to assume a full coherent source and simulate the propagation of such a wavefront through the beamline.
 
@@
 However, for those concerning partial coherence, very long calculations are needed.
-For example, Monte Carlo calculations in Section~\ref{srw_me} for the multi-electron case spent a few hours on a 56-core computer cluster for simulating 22000 electrons.
+For example, Monte Carlo calculations in Section~\ref{srw_me} for the multi-electron case spent a few hours on a 56-core computer cluster for simulating 50000 electrons.
 For performing the COMSYL coherent mode decomposition the calculation time depends very much on the dimensions of the problem.
@@
 It took less than four hours in a cluster of 28 cores.
-The calculation is more expensive if one wants to decompose a source that is pretty incoherent like our ESRF-High-$\beta$.
-This is a limit case, it was used wavefields of 10075$\times$201 pixels to accommodate the large spatial extension of the source.
-We obtained 4016 modes totalizing only 86\% of the spectral density.
+The calculation is more expensive if one wants to decompose a source that is \inred{mostly} incoherent like our ESRF-High-$\beta$.
+This is a limit case, it used wavefields of 10075$\times$201 pixels to accommodate the large spatial extension of the source.
+We obtained 4016 modes \inred{containing} only 86\% of the spectral density.
 The calculation took more than 4 days in an 84 core cluster.
@@
 It is not recommended to do similar calculations that imply a cost because a source that is quite incoherent was treated with partial coherence methods.
-%\inred{In Appendix B we present for the ESRF High$\beta$ similar results than those in Section~\ref{comsyl}}.
 
@@
 
-%\todo{
-%Complete discussion. Some ideas:
-%\begin{itemize}
-% \item DISCUSS EXPERIMENTAL RESULTS?
-% \item Discuss coherence length?
-% \item Comment on the heat-load deformation 
-% \item Extend the effect of reflectivity by ML and mirror %coating
-% \item Talk a bit more on Oasys?
-% \item comment on general strategy for beamline %simulations and the application of the shown hierarchical %model. Make a flowchart
-%\end{itemize}
-% }
+ %-------------------------------------------------------------------------
+ %
+ %-------------------------------------------------------------------------
 
@@
 
+\inred{
 The analytical estimation of the main parameters is the mandatory first step in the beamline conception and analysis.
 We applied concepts based on optical magnification and numerical aperture acceptance and calculated approximately the coherent fraction at the source.
-These values help to perform a sanity check of the computer simulations.
+\inred{It is shown that  that simple analytical calculations give a reasonable estimation of basic performances that measure beamline performance (acceptance of the beamline optics, energy resolution, dimensions of the focused beam).}
+
+The analytical values help to perform a sanity check of the computer simulations.
 A ray-tracing simulation is inexpensive and provides important information.
-The comparison of the CF with the beamline transmitivity would give an idea of the beam coherence, reflected in the number modes present in the final beam.
-For high coherence, like for ID16A, the wave optics calculations are necessary.
-If the final beam has a very high coherence coherent optics (1D simplifications, and single-electron calculations) may give good results.
-Full calculations based on coherent mode decomposition were introduced here.
-It is a new way of quantitatively assess the quality of the partially coherent beam via the calculation of the coherent modes of the source, propagation of the modes and recalculation of the coherent mode expansion on the final beam.
-The CF or occupation of the first mode is the parameter used to measure the coherence quality.
-
-% It is interesting to remark that although Monte Carlo multi-electron method can in principle calculate Eq.~\ref{CSD} the storage at every point is unmanagleable, so usually only some 'cuts' of the CSD are stored, such as $W(x_1,0,x_2,0;z,\omega)$ (horizontal) or $W(0,y_1,0,y_2;z,\omega)$ (vertical) \todo{move this to the discussion?}.
+It is therefore recommended to always perform a basic ray tracing although its limitation for beamlines exploiting highly coherent X-ray beams.
+Ray tracing does not take into account the
+phase changes induced by the beamline optics, which has a crucial impact on the wavefront characteristics and determine the (diffraction limited) focused beam size.
+Wave optical methods combined with ray tracing (hybrid) help to evaluate the possible diffraction effects produced by beam clipping and slope errros for a highly coherent beam.
+
+The comparison of the analytical CF with the beamline transmitivity gives an indication of the beam coherence.
+When these values are similar, as for ID16A, the resulting beam will be highly coherent, thus wave optics calculations are required.
+Simplified coherent optics (1D calculations) and also single-electron calculations may give good results if the final beam has a very high coherence like in our case.
+All these simulations can be done using the OASYS integrated environment.
+
+Wave optical simulations considering partial coherence implement complex methodologies that require considerable computer ressources.
+This is needed in cases when the beam coherence is large but cannot be considered fully coherent, which will be the case of several beamlines in new storage rings.
+Two approaches are presented:  Monte Carlo method and coherent mode decomposition.
+For ID16A we have studied in detail the effect of the slope errors using Monte Carlo.
+Full calculations based on coherent mode decomposition were introduced here to study the coherence transfer.
+This new technique quantitatively assesses the quality of the partially coherent beam via the calculation of the coherent modes of the source, the propagation of the modes and the recalculation of the coherent mode expansion on the final beam.
+The Coherent fraction or occupation of the first coherent mode is the parameter used to measure the coherence quality of the propagated beam (83\% for the ID16A at EBS at 17 keV).
+
+}
+
+ %-------------------------------------------------------------------------
  %
-
-%\ack{\todo{someone to thank?
-- RC: before the PHANGS in Trieste I was contacting Oleg about the elliptical mirrors.
-There was a bug there and he corrected it.
-%Only after fixed it we were able to simulate the ID16A with the mirror elements.
-So maybe something mentioning some disucssions on the early stages of the SRW %simulations.}}
+ %-------------------------------------------------------------------------
 
@@
 
-%\section{COMSYL results for the ESRF-H$\beta$}
-%\todo{Under development}
+ %-------------------------------------------------------------------------
+ %
+ %-------------------------------------------------------------------------
 

### 2019-08-09 mkJSR.sh

@@ -102,6 +102,8 @@ sed -e "s/GRAPHICS\/metrology.pdf/figure13.pdf/" paper-hierarchical.tex -i
 
 rm -rf GRAPHICS
 
+sed -e "s/documentclass/documentclass[preprint]/" paper-hierarchical.tex -i
+
 pdflatex paper-hierarchical.tex
 pdflatex paper-hierarchical.tex
 okular paper-hierarchical.pdf
